### 2018-01-23 ChIP-seq_ER/qrime/02.Supplementary.Rmd
first commit

---
title: "Supplementary"
author: "Andrew Holding & Matt Eldridge"
date: "9 January 2017"
output:
  pdf_document:
    number_sections: yes
    toc: yes
    fig_caption: yes
header-includes:
  - \usepackage{longtable}
  - \usepackage{booktabs}
---

****************************

```{r configuration, include = FALSE, cache = FALSE}
knitr::opts_chunk$set(cache = FALSE)
```

```{r load_libraries, include = FALSE, cache = FALSE}
library(tidyr)
library(dplyr)
library(tibble)
library(MSnbase)
library(RColorBrewer)
library(ggplot2)
library(kfigr)
library(gridExtra)
library(xtable)
library(limma)
```

```{r load_functions, include = FALSE, cache = FALSE}
source("Functions/plots.R")
source("Functions/normalization.R")
```

```{r xtable_options, include=FALSE, cache = FALSE}

options(xtable.comment = FALSE)
options(xtable.booktabs = TRUE)

addtorow          <- list()
addtorow$pos      <- list()
addtorow$pos[[1]] <- c(0)
addtorow$command  <- c(paste("\\hline \n",
                             "\\endhead \n",
                             "\\hline \n",
                             "{\\footnotesize Continued on next page} \n",
                             "\\endfoot \n",
                             "\\endlastfoot \n",sep=""))
```

```{r read_protein_information, include = FALSE}
proteinInfo <- read.delim("ReferenceData/protein_gene_description.txt", stringsAsFactors = FALSE)
```

```{r read_metadata, include = FALSE}
metadata <- read.delim("SampleData/metadata.txt", check.names = FALSE)
runs <- levels(metadata$Run)
groups <- levels(metadata$Group)
controlGroup <- "IgG"
```

```{r colours, include = FALSE}
groupColours <- c(brewer.pal(9, "Blues")[5:7], brewer.pal(9, "RdPu")[5:7], brewer.pal(9, "Greens")[6])
names(groupColours) <- groups
sampleColours <- groupColours[metadata$Group]
names(sampleColours) <- metadata$Sample
```

```{r read_intensities, include = FALSE}

peptideData <- list()

for (run in runs)
{
  cat(run, "\n", sep = "")

  data <- read.delim(paste("ProcessedData/", run, ".txt", sep = ""), stringsAsFactors = FALSE, check.names = FALSE)

  peptideCount <- data %>% select(Sequence, Modifications) %>% distinct %>% nrow
  cat("Peptides:", peptideCount, "\n", sep = "")
  if (peptideCount != nrow(data))
    stop("Error: multiple rows for at least one peptide/modification (is this PSM-level data?)")

  proteinCount <- data %>% select(Protein) %>% distinct %>% nrow
  cat("Proteins:", proteinCount, "\n", sep = "")

  sampleColumns <- metadata %>%
    filter(Run == run) %>%
    select(Group) %>%
    unlist(use.names = FALSE) %>%
    as.character

  data <- MSnSet(
    exprs = data %>% select(one_of(sampleColumns)) %>% as.matrix,
    fData = data %>% select(Sequence, Modifications, Protein),
    pData = metadata %>% filter(Run == run) %>% mutate(RowName = Group) %>% column_to_rownames("RowName")
  )

  peptideData[[run]] <- data
}
```

# Introduction

This report is from an analysis of the RIME-TMT proteomics data for an experiment carried out by Andrew Holding to quantify specific interactors of estrogen receptor (ER) and forkhead box protein A1 (FOXA1) in MCF-7 cells.

The RIME (Rapid Immunoprecipitation of Endogenous Proteins) technique was used to pull down ER and FOXA1 and their interacting proteins, which were subsequently assessed and quantitated using mass spectrometry with tandem mass tags (TMT). TMT runs were carried out for 3 biological replicates, estrogen-starved MCF-7 cells from the same cell line at different times/passages, before and after addition of estrogen with samples taken at 3 different time points -- 0, 45 and 90 minutes after estrogen addition. A pull down of Immunoglobulin G (IgG) was also run as a control.

The digests from two of the biological replicates (PR622 and PR650) were analyzed with 3 separate runs of the mass spectrometer to increase coverage; data from the repeated runs were combined prior to the analysis carried out here. The other sample (PR590) was run once only.

`r figr("samples_table", TRUE, type="Table")` shows the isobaric tags used for each sample within each of the runs.

```{r samples_table, echo=FALSE, warning=FALSE, message=FALSE, results='asis'}
print(
  xtable(
    metadata %>% select(Run, Tag, Group) %>% spread(Run, Tag),
    caption = "Isobaric tags used for each sample (group) and run."
  ),
  include.rownames = FALSE,
  size = "\\setlength{\\tabcolsep}{12pt}"
)
```

# Peptide intensity data

Raw spectra were processed using Proteome Discover 2.1 to produce peptide-level intensity data with a single set of intensity values per distinct peptide. Only peptides with unique high-confidence protein matches were included.

Multiple peptide-spectrum matches (PSMs) for the same peptide were combined using Proteome Discoverer by summing the PSM-level intensities. Peptide sequences with different modifications are treated as distinct peptides and the data provided contain intensities values for each modification identified for a peptide sequence. Several distinct peptide sequences (with modifications) may have been identified for any given protein.

`r figr("peptide_protein_count_table", TRUE, type="Table")` shows the numbers of distinct peptides and proteins obtained in each TMT run. Also given are the numbers of peptides and proteins after filtering peptides with missing intensity values in one or more TMT channels.

```{r filter_missing_intensities, include=FALSE}
peptideDataExcludingMissingValues <- list()
for (run in runs)
{
  cat(run, "\n", sep = "")
  data <- peptideData[[run]]
  data <- data[which(complete.cases(exprs(data))),]
  peptideDataExcludingMissingValues[[run]] <- data
}
```

```{r combine_peptide_intensities, include = FALSE}

# function for combining data for runs into single MSnSet object
combineRuns <- function(runData, runs)
{
  combinedData <- NULL

  for (run in runs)
  {
    data <- runData[[run]]

    sampleNames(data) <- pData(data) %>%
      transmute(Sample = paste(Run, Group, sep = ":")) %>%
      unlist %>%
      as.character

    featureNames(data) <- fData(data) %>%
      transmute(ID = paste(Sequence, Modifications, sep = "|")) %>%
      unlist %>%
      as.character()

    if (is.null(combinedData))
    {
      combinedData <- data
    } else
    {
      combinedData <- BiocGenerics::combine(combinedData, data)
    }
  }

  return(combinedData)
}

combinedPeptideData <- combineRuns(peptideData, runs)
combinedPeptideDataExcludingMissingValues <- combineRuns(peptideDataExcludingMissingValues, runs)
```

```{r peptide_protein_count_table, echo=FALSE, warning=FALSE, message=FALSE, results='asis'}

countProteins <- function(data) data %>% fData %>% select(Protein) %>% distinct %>% nrow

peptideData[["All"]] <- combinedPeptideData
peptideDataExcludingMissingValues[["All"]] <- combinedPeptideDataExcludingMissingValues

counts <- bind_rows(
  peptideData %>% sapply(nrow) %>% t %>% as.data.frame %>% mutate(label = "Peptides"),
  peptideDataExcludingMissingValues %>% sapply(nrow) %>% t %>% as.data.frame %>% mutate(label = "Peptides with no missing values"),
  peptideData %>% sapply(countProteins) %>% t %>% as.data.frame %>% mutate(label = "Proteins"),
  peptideDataExcludingMissingValues %>% sapply(countProteins) %>% t %>% as.data.frame %>% mutate(label = "Proteins with no missing values")
)
rownames(counts) <- counts$label
counts <- counts %>% select(-label)

peptideData[["All"]] <- NULL
peptideDataExcludingMissingValues[["All"]] <- NULL

print(
  xtable(
    counts,
    align = "lrrrr",
    caption = "Numbers of peptides and proteins observed in each run."
  ),
  include.rownames = TRUE,
  size = "\\setlength{\\tabcolsep}{12pt}"
)
```

`r figr("intensity_plots", TRUE, type="Figure")` shows the distribution of intensities for each sample within each run.

```{r intensity_plots, echo=FALSE, warning=FALSE, message=FALSE, fig.cap="Density plots of intensities from each run.", fig.height=8, fig.width=5}

minIntensity <- -2
maxIntensity <- 15

plots <- vector("list")

for (i in 1:length(runs))
{
  run <- runs[i]
  intensities <- peptideData[[run]] %>% exprs %>% log2 %>% as.data.frame
  plots[[run]] <- intensityDistributionPlot(
    intensities,
    groups,
    groupColours,
    title = run,
    xlab=expression(log[2](intensity)),
    minIntensity = minIntensity,
    maxIntensity = maxIntensity,
    showLegend = (i == 1))
}

do.call("grid.arrange", c(plots, ncol = 1))
```

## Missing intensity values

Missing intensity values can result even though ions are present at detectable concentrations due to the stochastic nature of the mass spectrometry acquisition method. It is reasonable to expect that these missing values are randomly distributed in the data. Alternatively, missing values may occur when there is a low abundance of ions, below the limit of detection of the instrument. These biologically relevant missing values are not randomly distributed, affecting only those proteins that are expressed at low levels in the samples analysed. The R/Bioconductor package `MSnBase` provides imputation methods for both types of missing value.

In this analysis, missing values have been handled in 3 different ways. The first approach is to exclude all peptide-level mesasurements where there is a missing value for one or more of the samples within a run. In addition, two imputation methods have been performed: one in which the missing values are set to the smallest non-missing value in the data for that run and the other in which k-nearest neighbour (KNN) averaging is applied.

```{r imputation, include = FALSE}
peptideDataImputedMin <- list()
peptideDataImputedKNN <- list()
for (run in runs)
{
  peptideDataImputedMin[[run]] <- impute(peptideData[[run]], "min")
  peptideDataImputedKNN[[run]] <- impute(peptideData[[run]], "knn")
}
```

`r figr("intensity_plots_impute_min", TRUE, type="Figure")` shows the distribution of intensities for each sample within each run following imputation using the smallest non-missing values. A small hump just below zero on the log~2~ scale is clearly visible for each run.

```{r intensity_plots_impute_min, echo=FALSE, warning=FALSE, message=FALSE, fig.cap="Density plots of intensities from each run following imputation of missing values using the smallest non-missing value in the data for each run.", fig.height=8, fig.width=5}

minIntensity <- -2
maxIntensity <- 15

plots <- vector("list")

for (i in 1:length(runs))
{
  run <- runs[i]
  intensities <- peptideDataImputedMin[[run]] %>% exprs %>% log2 %>% as.data.frame
  plots[[run]] <- intensityDistributionPlot(
    intensities,
    groups,
    groupColours,
    title = run,
    xlab=expression(log[2](intensity)),
    minIntensity = minIntensity,
    maxIntensity = maxIntensity,
    showLegend = (i == 1))
}

do.call("grid.arrange", c(plots, ncol = 1))
```

# Normalization of intensity data

The intensity distribution for PR622 suggests that the labeled samples have been pooled with differing protein concentrations and that some normalization is required to properly assess differences between groups. Normalization techniques that are commonly applied to microarray expression data and mRNA sequencing data assume that only a few genes, e.g. a few hundred out of tens of thousands, are expressed at very different levels between samples. This assumption may not hold for a RIME experiment where a specific set of interacting proteins is targeted. For example, if the majority of interacting proteins were to bind less strongly at one time point relative to another, quantile normalization or scale normalization approaches would have the effect of removing the general trend. In this case, following normalization, those proteins that bind less strongly may appear to be largely unchanged, while proteins with very similar levels of binding at the different time points would have artificially increased binding levels. Therefore some care is required when interpreting the results of a differential binding analysis for RIME TMT experiments.

In this analysis, quantile normalization and scale normalization were applied for each run separately.

The IgG control was used to assess non-specific binding. The binding of a protein was considered to be non-specific if the binding level of that protein in the condition of interest, e.g. ER at 45 minutes, was not significantly above the level observed for the IgG control. Proteins detected with the IgG control are expressed at lower levels in the PR590 and PR650 runs. Normalization of the IgG control along with the other samples would make it more difficult to distinguish between specific and non-specific binding since the intensities of the IgG control would be scaled to a similar level of those of the other samples.

Among a number of normalization approaches tested here, normalizations were carried out that both include and exclude the IgG control. `r figr("normalized_intensity_plots", TRUE, type="Figure")` shows the normalized intensity distributions following scale normalization of all peptide intensities including those with missing values. `r figr("normalizedExcludingIgG_intensity_plots", TRUE, type="Figure")` shows the normalized intensity distributions following scale normalization for all samples except the IgG controls.

```{r normalization, include = FALSE}

peptideDataNormalized <- list()

peptideDataNormalized[["IncludeMissingValues_NoNormalization"]] <- peptideData
peptideDataNormalized[["ExcludeMissingValues_NoNormalization"]] <- peptideDataExcludingMissingValues
peptideDataNormalized[["ImputeMin_NoNormalization"]] <- peptideDataImputedMin
peptideDataNormalized[["ImputeKNN_NoNormalization"]] <- peptideDataImputedKNN

for (missingValueMethod in c("IncludeMissingValues", "ExcludeMissingValues", "ImputeMin", "ImputeKNN"))
{
  peptideDataForNormalization <- peptideDataNormalized[[paste(missingValueMethod, "NoNormalization", sep = "_")]]

  quantileNormalizationAnalysis <- paste(missingValueMethod, "QuantileNormalization", sep = "_")
  quantileNormalizationExcludingIgGAnalysis <-
    paste(missingValueMethod, "QuantileNormalizationExcludingIgG", sep = "_")
  scaleNormalizationAnalysis <- paste(missingValueMethod, "ScaleNormalization", sep = "_")
  scaleNormalizationExcludingIgGAnalysis <-
    paste(missingValueMethod, "ScaleNormalizationExcludingIgG", sep = "_")
  scaleNormalizationTopIgGAnalysis <- paste(missingValueMethod, "ScaleNormalizationTopIgG", sep = "_")

  peptideDataNormalized[[quantileNormalizationAnalysis]] <- list()
  peptideDataNormalized[[quantileNormalizationExcludingIgGAnalysis]] <- list()
  peptideDataNormalized[[scaleNormalizationAnalysis]] <- list()
  peptideDataNormalized[[scaleNormalizationExcludingIgGAnalysis]] <- list()
  peptideDataNormalized[[scaleNormalizationTopIgGAnalysis]] <- list()

  numberOfPeptidesWithHighestIgG <- 10

  for (run in runs)
  {
    data <- peptideDataForNormalization[[run]]

    # quantile normalization
    peptideDataNormalized[[quantileNormalizationAnalysis]][[run]] <- normalize(data, method = "quantiles")

    # scale normalization
    peptideDataNormalized[[scaleNormalizationAnalysis]][[run]] <- normalizeMedianScaling(data)

    samples <- setdiff(sampleNames(data), controlGroup)

    # quantile normalization
    normalized <- normalize(data[,samples], method = "quantiles")
    peptideDataNormalized[[quantileNormalizationExcludingIgGAnalysis]][[run]] <- data
    exprs(peptideDataNormalized[[quantileNormalizationExcludingIgGAnalysis]][[run]])[,samples] <-
      exprs(normalized)

    # scale normalization
    peptideDataNormalized[[scaleNormalizationExcludingIgGAnalysis]][[run]] <-
      normalizeMedianScaling(data, samples)

    # scale normalization using peptides with highest IgG intensities for computing sample medians
    controlIntensities <- data %>%
      exprs %>%
      as.data.frame %>%
      select(one_of(controlGroup)) %>%
      unlist %>%
      as.numeric
    controlIntensities[is.na(controlIntensities)] <- -Inf
    controlIntensityOrder <- rev(order(controlIntensities))
    peptideDataNormalized[[scaleNormalizationTopIgGAnalysis]][[run]] <-
      normalizeMedianScaling(data, rowsForCalculatingMedian = controlIntensityOrder[1:numberOfPeptidesWithHighestIgG])
  }
}
```

```{r normalized_intensity_plots, echo=FALSE, warning=FALSE, message=FALSE, fig.cap="Density plots of normalized intensities from each TMT run where scale normalization was applied to peptide intensities that include misssing values.", fig.height=8, fig.width=5}

analysis <- "IncludeMissingValues_ScaleNormalization"

minIntensity <- 0
maxIntensity <- 15

plots <- vector("list")

for (i in 1:length(runs))
{
  run <- runs[i]
  intensities <- peptideDataNormalized[[analysis]][[run]] %>% exprs %>% log2 %>% as.data.frame
  plots[[run]] <- intensityDistributionPlot(
    intensities,
    groups,
    groupColours,
    title = run,
    xlab=expression(log[2](intensity)),
    minIntensity = minIntensity,
    maxIntensity = maxIntensity,
    showLegend = (i == 1))
}

do.call("grid.arrange", c(plots, ncol = 1))
```

```{r normalizedExcludingIgG_intensity_plots, echo=FALSE, warning=FALSE, message=FALSE, fig.cap="Density plots of normalized intensities from each TMT run where scale normalization was applied to peptide intensities for all samples except the IgG controls.", fig.height=8, fig.width=5}

analysis <- "IncludeMissingValues_ScaleNormalizationExcludingIgG"

minIntensity <- 0
maxIntensity <- 15

plots <- vector("list")

for (i in 1:length(runs))
{
  run <- runs[i]
  intensities <- peptideDataNormalized[[analysis]][[run]] %>% exprs %>% log2 %>% as.data.frame
  plots[[run]] <- intensityDistributionPlot(
    intensities,
    groups,
    groupColours,
    title = run,
    xlab=expression(log[2](intensity)),
    minIntensity = minIntensity,
    maxIntensity = maxIntensity,
    showLegend = (i == 1))
}

do.call("grid.arrange", c(plots, ncol = 1))
```

A further normalization approach was attempted that scales the data from all samples within a run based on a subset of peptides that are considered most likely to be at similar concentrations between the samples. The assumption is that the peptides with the highest intensity values in the IgG control are the result of non-specific binding and that that binding is consistent across all samples. Similar to the scale normalization applied in the first approach, the median intensity within each sample is computed to determine a scaling factor for each sample within a run but the median computed only uses the `r numberOfPeptidesWithHighestIgG` peptides with the highest IgG measurements. As before, normalization is carried out within each run separately.

`r figr("normalizedIgG_intensity_plots", TRUE, type="Figure")` shows the resulting normalized intensity distributions. Strikingly, the IgG intensity distribution is shifted to lower values in run PR622, making it more consistent with the other 2 runs.

```{r normalizedIgG_intensity_plots, echo=FALSE, warning=FALSE, message=FALSE, fig.cap=paste("Density plots of normalized intensities from each TMT run where scale normalization is applied based on", numberOfPeptidesWithHighestIgG, "peptides with the highest IgG intensities in each run."), fig.height=8, fig.width=5}

analysis <- "IncludeMissingValues_ScaleNormalizationTopIgG"

minIntensity <- 0
maxIntensity <- 15

plots <- vector("list")

for (i in 1:length(runs))
{
  run <- runs[i]
  intensities <- peptideDataNormalized[[analysis]][[run]] %>% exprs %>% log2 %>% as.data.frame
  plots[[run]] <- intensityDistributionPlot(
    intensities,
    groups,
    groupColours,
    title = run,
    xlab=expression(log[2](intensity)),
    minIntensity = minIntensity,
    maxIntensity = maxIntensity,
    showLegend = (i == 1))
}

do.call("grid.arrange", c(plots, ncol = 1))
```

# Protein-level quantification

Protein-level quantification was carried out for each run separately by summing normalized intensity values for all peptides matching to a particular protein. This is similar in principle to the gene-level quantification that is carried out during microarray analysis or the read-counting approach in RNA-Seq. Essentially, all signal attributable to a particular protein is assigned to that protein. Intensity values vary in magnitude depending on the elution profile for the peptide and when the peptide is sampled. Relative intensities for a peptide between the tags or samples within a run should be consistent but the signal may include some contribution from a contaminating, co-eluting peptide, the extent of which may differ for different PSMs. In summing intensities from multiple measurements for the same peptide, those peptides that have higher levels of intensity values will have a greater weight in determining the overall intensity; higher intensity measurements are likely to more accurately reflect differences between samples so this may be advantageous.

When no imputation of missing values was carried out, peptides with missing values in one or more samples were excluded from the summation. Missing values that arise for technical reasons instead of being at undetectable levels in a sample would otherwise reduce the overall intensity in that sample relative to other samples.

```{r summarize_protein_intensities, include=FALSE}

proteinData <- list()

for (analysis in names(peptideDataNormalized))
{
  peptideDataForSummarization <- peptideDataNormalized[[analysis]]

  proteinData[[analysis]] <- list()

  for (run in runs)
  {
    data <- peptideDataForSummarization[[run]]

    # exclude rows with missing values (will have no effect if imputation carried out)
    data <- data[which(complete.cases(exprs(data))),]

    data <- combineFeatures(data, groupBy = fData(data)$Protein, fun = "sum", na.rm = FALSE)
    fData(data) <- fData(data) %>% select(Protein)

    proteinData[[analysis]][[run]] <- data
  }
}
```

```{r combine_protein_intensities, include = FALSE}

combinedProteinData <- list()

for (analysis in names(proteinData))
{
  for (run in runs)
  {
    data <- proteinData[[analysis]][[run]]

    sampleNames(data) <- pData(data) %>%
      transmute(Sample = paste(Run, Group, sep = ":")) %>%
      unlist %>%
      as.character

    if (is.null(combinedProteinData[[analysis]]))
    {
      combinedProteinData[[analysis]] <- data
    } else
    {
      combinedProteinData[[analysis]] <- BiocGenerics::combine(combinedProteinData[[analysis]], data)
    }
  }
}
```

```{r replicate_counts, include=FALSE}

isRepresented <- function(x) any(!is.na(x))

getReplicateCounts <- function(data)
{
  metadata <- pData(data) %>% rownames_to_column("Sample")

  intensities <- data %>%
    exprs %>%
    as.data.frame %>%
    mutate(N = 0)

  for (run in runs)
  {
    samples <- metadata %>%
      filter(Run == run) %>%
      select(Sample) %>%
      unlist %>%
      as.character

    intensities <- intensities %>%
      mutate(N = N + apply(select(., one_of(samples)), 1, isRepresented))
  }

  replicateCounts <- intensities %>%
    select(N) %>%
    mutate(Count = 0) %>%
    group_by(N) %>%
    summarize_each(funs(length)) %>%
    ungroup %>%
    mutate(N = as.character(N))

  replicateCounts <- replicateCounts %>%
    bind_rows(data_frame(N = "total", Count = sum(replicateCounts$Count)))

  return(replicateCounts)
}

replicateCounts <- getReplicateCounts(combinedProteinData[["ImputeKNN_ScaleNormalization"]])

replicateCountTableData <- inner_join(
  getReplicateCounts(combinedProteinData[["ImputeKNN_ScaleNormalization"]]) %>%
    rename(`Including missing values` = Count),
  getReplicateCounts(combinedProteinData[["ExcludeMissingValues_ScaleNormalization"]]) %>%
    rename(`Excluding missing values` = Count),
  by = "N"
)

replicateCountTableData <- replicateCountTableData %>%
  as.data.frame %>%
  column_to_rownames(var = "N") %>%
  t %>%
  as.data.frame %>%
  rownames_to_column(var = "Runs/replicates")
```

The statistical analysis of differential binding between groups that follows requires at least 3 observations per group. `r figr("replicate_count_table", TRUE, type="Table")` gives the number of protein identified in just a single run, two runs or all three runs. A total of `r (replicateCounts %>% filter(N == "all"))$Count` proteins were identified across all runs of which `r (replicateCounts %>% filter(N == "3"))$Count` were observed in all three runs, allowing for a statistical analysis. log~2~ fold changes are still computed for proteins identified only in one or two of the runs, but no measure of the statistical significance is given in the differential binding analysis.

```{r replicate_count_table, echo=FALSE, warning=FALSE, message=FALSE, results='asis'}
# colnames(replicateCounts) <- c("Runs/replicates", "Proteins (includes missing values)", "Proteins")
print(
  xtable(
    replicateCountTableData,
    # align = "llr",
    caption = "Numbers of proteins identified in differing numbers of runs and total number of proteins identified in all runs."
  ),
  include.rownames = FALSE,
  size = "\\setlength{\\tabcolsep}{12pt}"
)
```

# Principal Component Analysis

Protein-level intensities were scaled to sum to 1.0 for each protein within a run, allowing for comparison across runs in a principal component analysis (PCA).

`r figr("pca_plot", TRUE, type="Figure")` shows a PCA plot for the first two principal components using all proteins that were sampled in all three runs. The plot shows a clear separation of the ER, FOXA1 and IgG control pull-downs. The 0, 45 and 90 minute timepoints within each of the ER and FOXA1 groups are not completely separated in the first two principal components. `r figr("pca_plot_er", TRUE, type="Figure")` shows the PCA plot for just the ER samples.

```{r pca_plot, echo=FALSE, warning=FALSE, message=FALSE, fig.cap="Principal Componenent Analysis for proteins sampled in all 3 runs. The PCA was based on protein-level data resulting from summation of quantile normalized peptide intensities in which missing values were imputed using KNN-based nearest neighbour averaging. The first two principal components are displayed.", fig.height=5}

data <- combinedProteinData[["ImputeKNN_QuantileNormalization"]]

metadata <- pData(data) %>% rownames_to_column("Sample")

scaledIntensities <- data %>%
  exprs %>%
  as.data.frame %>%
  filter(complete.cases(.))

for (run in runs)
{
  samples <- metadata %>%
    filter(Run == run) %>%
    select(Sample) %>%
    unlist %>%
    as.character

  summedIntensities <- scaledIntensities %>%
    select(one_of(samples)) %>%
    rowSums

  scaledIntensities <- scaledIntensities %>%
    mutate_each(funs(. / summedIntensities), one_of(samples))
}

print(pcaPlot(scaledIntensities, metadata$Sample, metadata$Group, groupColours, labels = metadata$Run, legend = TRUE))
```

```{r pca_plot_er, echo=FALSE, warning=FALSE, message=FALSE, fig.cap="Principal Componenent Analysis for ER-interactng proteins sampled in all 3 runs. The first two principal components are displayed.", fig.height=5}

metadata <- metadata %>% filter(grepl("^ER", Group))

print(pcaPlot(scaledIntensities, metadata$Sample, metadata$Group, groupColours, labels = metadata$Run, legend = TRUE))
```

# Differential Binding

```{r differential_expression_paramters, include=FALSE}
significanceLevel <- 0.05
controlLogFoldChangeThreshold <- 1.0
```

A statistical analysis of differentially-expressed peptides was carried out using **limma**, a R/Bioconductor package commonly used in the analysis of microarray and RNA-seq data, but applicable to data from any quantitative expression technology.

Limma uses linear models to assess differential expression in the context of multifactor experimental designs. It is able to analyze comparisons between many RNA targets simultaneously (or in this case many proteins) and has features that make these analyses stable even for experiments with small numbers of samples; this is achieved by borrowing information across genes (proteins).

In this analysis, limma was used to estimate log~2~ fold changes and standard errors by fitting a linear model for each protein where the model includes variables for the group (the ER and FOXA1 pull-downs at 0, 45 and 90 minute time points and the IgG control at 45 minutes) and the run. This is essentially a two-way ANOVA, a generalization of a paired analysis, in which comparisons between pull-downs at different time points are made within each run.

Limma employs an empirical Bayes method to moderate the standard errors of the estimated log~2~ fold changes; this results in more stable inference and improved power, especially for experiments with small numbers of replicates.

Multiple testing correction of p-values was applied using the Benjamini-Hochberg method to control the false discovery rate (FDR). The adjusted p-value (also known as a q-value) can be understood as follows. If all proteins with q-value below a threshold of `r significanceLevel` are selected as differentially expressed, then the expected proportion of false discoveries in the selected group is controlled to be less than the threshold value, in this case 5%.

The B-statistic (lods or B) is the log odds that the protein is differentially expressed. For example, if B = 1.5, the odds of differential expression is $e^{1.5} = 4.48$, i.e, about four and a half to one. The probability that the protein is differentially expressed is $4.48 / (1 + 4.48) = 0.82$. A B-statistic of zero corresponds to a 50-50 chance that the gene is differentially expressed. The B-statistic has been automatically adjusted for multiple testing by assuming that 1% of the proteins are expected to be differentially expressed.

```{r fit_linear_model, include = FALSE}

factors <- c("Group", "Run")

model <- as.formula(paste(c("~ 0", factors), collapse = " + "))

fittedLinearModels <- list()
fittedLinearModelsExcludingControls <- list()

for (analysis in names(combinedProteinData))
{
  cat(analysis, "\n", sep = "")
  data <- combinedProteinData[[analysis]]

  # log2 transform
  exprs(data) <- log2(exprs(data))

  design <- model.matrix(model, data = pData(data))

  colnames(design) <- colnames(design) %>%
    sub(pattern = "^Group", replacement = "") %>%
    gsub(pattern = " ", replacement = "_")

  fittedLinearModels[[analysis]] <- lmFit(data, design = design)

  samples <- pData(data) %>%
    rownames_to_column("Sample") %>%
    filter(Group != controlGroup) %>%
    select(Sample) %>%
    unlist(use.names = FALSE)

  data <- data[, sampleNames(data) %in% samples]

  design <- model.matrix(model, data = pData(data))

  colnames(design) <- colnames(design) %>%
    sub(pattern = "^Group", replacement = "") %>%
    gsub(pattern = " ", replacement = "_")

  fittedLinearModelsExcludingControls[[analysis]] <- lmFit(data, design = design)
}
```

```{r contrasts, include = FALSE}

contrasts <- c(
  "ER 45min vs ER 0min",
  "ER 90min vs ER 0min",
  "ER 90min vs ER 45min",
  "FOXA1 45min vs FOXA1 0min",
  "FOXA1 90min vs FOXA1 0min",
  "FOXA1 90min vs FOXA1 45min"
)

reformatGroup <- function(x) gsub(x, pattern = " ", replacement = "_")

reformatContrasts <- function(x)
{
  x %>%
    gsub(pattern = " ", replacement = "_") %>%
    sub(pattern = "_-_", replacement = " - ") %>%
    sub(pattern = "_vs_", replacement = " - ")
}

reformattedContrasts <- reformatContrasts(contrasts)

fittedContrasts <- list()

for (analysis in names(combinedProteinData))
{
  cat(analysis, "\n", sep = "")
  fittedLinearModel <- fittedLinearModels[[analysis]]
  # fittedLinearModel <- fittedLinearModelsExcludingControls[[analysis]]
  contrastMatrix <- makeContrasts(contrasts = reformattedContrasts, levels = fittedLinearModel$design)
  fittedContrasts[[analysis]] <- contrasts.fit(fittedLinearModel, contrastMatrix)
  fittedContrasts[[analysis]] <- eBayes(fittedContrasts[[analysis]], trend = TRUE, robust = TRUE)
}
```

```{r create_contrast_results, include = FALSE}

createContrastResults <- function(
  msnset,
  fittedLinearModel,
  fittedLinearModelExcludingControls,
  fittedContrasts,
  runs,
  contrastGroups,
  controlGroup,
  proteinInfo)
{
  reformattedContrast <- reformatContrasts(paste(contrastGroups, collapse = " - "))
  reformattedContrastGroups <- reformattedContrast %>% strsplit(" - ") %>% unlist
  reformatedControlGroup <- reformatGroup(controlGroup)

  results <- topTable(fittedContrasts, coef = reformattedContrast, number = Inf, sort.by = "none")

  fittedIntensities <- as.data.frame(fittedLinearModel$coefficients)
  # fittedIntensities <- as.data.frame(fittedLinearModelExcludingControls$coefficients)
  contrastFittedIntensities <- select(fittedIntensities, one_of(reformattedContrastGroups))
  results$logFC <- contrastFittedIntensities[, 1] - contrastFittedIntensities[, 2]

  fittedIntensities <- as.data.frame(fittedLinearModel$coefficients)
  contrastFittedIntensities <- select(fittedIntensities, one_of(reformattedContrastGroups))
  controlFittedIntensities <- fittedIntensities[, reformatedControlGroup]
  results$logFCcontrol = apply(contrastFittedIntensities - controlFittedIntensities, 1, max)

  samples <- pData(msnset) %>%
    mutate(Sample = rownames(.)) %>%
    filter(Group %in% c(contrastGroups, controlGroup)) %>%
    select(Sample) %>%
    unlist(use.names = FALSE)

  intensities <- exprs(msnset) %>%
    as.data.frame %>%
    select(one_of(samples))

  # add N column for the number of replicates, i.e. runs where intensity
  # is not NA for both groups
  n <- rep(0, nrow(intensities))
  for (run in runs)
  {
    samples <- pData(msnset) %>%
      mutate(Sample = rownames(.)) %>%
      filter(Group %in% contrastGroups) %>%
      filter(Run == run) %>%
      select(Sample) %>%
      unlist(use.names = FALSE)
    n <- n +
      intensities %>%
      select(one_of(samples)) %>%
      transmute(n = ifelse(complete.cases(.), 1, 0)) %>%
      unlist(use.names = FALSE)
  }

  results <- results %>%
    mutate(N = n) %>%
    bind_cols(fittedIntensities %>% select(one_of(rev(reformattedContrastGroups), controlGroup))) %>%
    bind_cols(intensities) %>%
    left_join(proteinInfo, by = "Protein")

  results <- bind_cols(
    results %>% select(Protein, Gene, Description, N),
    results %>% select(-Protein, -Gene, -Description, -N)
  )

  results 
}

```

The IgG control was used to assess non-specific binding. The binding of a protein is considered to be non-specific if the log~2~ fold change relative to the IgG control is less than `r controlLogFoldChangeThreshold`. In the differential binding analysis, protein intensity values are fitted to a linear model and the fitted values for each condition being compared are also compared to the IgG control. For each comparison of two groups, the maximum log~2~ fold change from each of the two groups above the IgG control is used to determine if the binding is specific.

\newpage

```{r contrast_1, include = FALSE}
contrastGroups <- c("ER 45min", "ER 0min")
contrastName <- paste(contrastGroups, collapse = " vs ")
processingDescription <- "excluding peptides with missing intensities, no normalization"
analysis <- "ExcludeMissingValues_NoNormalization"
```

## `r paste(contrastName, processingDescription, sep = ", ")`

```{r results_1, include = FALSE}

data <- combinedProteinData[[analysis]]
exprs(data) <- log2(exprs(data))

results <- createContrastResults(
  data,
  fittedLinearModels[[analysis]],
  fittedLinearModelsExcludingControls[[analysis]],
  fittedContrasts[[analysis]],
  runs,
  contrastGroups,
  controlGroup,
  proteinInfo)
```

```{r ma_plot_1, echo=FALSE, warning=FALSE, message=FALSE, fig.cap=paste("MA plot of the average intensity against log~2~ fold change for the ", contrastName, " comparison (", processingDescription, "). Top ranking differentially-expressed proteins with false discovery rate below 0.05 are highlighted in pink. Open circles indicate that the protein is non-specific from the IgG control comparison.", sep = ""), fig.height=6}
print(maPlot(results,
             significanceLevel = significanceLevel,
             controlLogFoldChangeThreshold = controlLogFoldChangeThreshold,
             minLogFoldChangeForLabelling = 1.5,
             maxNumberLabelledProteins = 20,
             pointSize = 1.5))
```

```{r volcano_plot_1, echo=FALSE, warning=FALSE, message=FALSE, fig.cap=paste("Volcano plot of the average intensity against log~2~ fold change for the ", contrastName, " comparison (", processingDescription, "). Top ranking differentially-expressed proteins with false discovery rate below 0.05 are highlighted in pink. Open circles indicate that the protein is non-specific from the IgG control comparison.", sep = ""), fig.height=6}
print(volcanoPlot(results,
                  significanceLevel = significanceLevel,
                  minLogFoldChange = -2.0,
                  maxLogFoldChange = 2.75,
                  controlLogFoldChangeThreshold = controlLogFoldChangeThreshold,
                  minLogFoldChangeForLabelling = 1.5,
                  maxNumberLabelledProteins = 20,
                  pointSize = 1.5))
```

```{r pvalue_histogram_1, echo=FALSE, warning=FALSE, message=FALSE, fig.cap=paste("Histogram of p-values for the ", contrastName, " comparison (", processingDescription, ")", sep = ""), fig.height=3.5, fig.width=4.5}
print(histogram(results, "P.Value", xlab = "p-value"))
```

```{r qq_plot_1, echo=FALSE, warning=FALSE, message=FALSE, fig.cap=paste("QQ plot of the adjusted p-values for the ", contrastName, " comparison (", processingDescription, ")", sep = ""), fig.height=5, fig.width=5}
q <- results %>%
  select(adj.P.Val) %>%
  filter(!is.na(adj.P.Val)) %>%
  arrange(adj.P.Val) %>%
  unlist %>%
  as.numeric
qqplot(-log10(ppoints(length(q))),
       -log10(q),
       xlab=expression(Expected~~-log[10](italic(q))),
       ylab=expression(Observed~~-log[10](italic(q))))
abline(0, 1, col = "red")
```

\newpage

```{r contrast_2, include = FALSE}
contrastGroups <- c("ER 45min", "ER 0min")
contrastName <- paste(contrastGroups, collapse = " vs ")
processingDescription <- "excluding peptides with missing intensities, quantile normalization"
analysis <- "ExcludeMissingValues_QuantileNormalization"
```

## `r paste(contrastName, processingDescription, sep = ", ")`

```{r results_2, include = FALSE}

data <- combinedProteinData[[analysis]]
exprs(data) <- log2(exprs(data))

results <- createContrastResults(
  data,
  fittedLinearModels[[analysis]],
  fittedLinearModelsExcludingControls[[analysis]],
  fittedContrasts[[analysis]],
  runs,
  contrastGroups,
  controlGroup,
  proteinInfo)
```

```{r ma_plot_2, echo=FALSE, warning=FALSE, message=FALSE, fig.cap=paste("MA plot of the average intensity against log~2~ fold change for the ", contrastName, " comparison (", processingDescription, "). Top ranking differentially-expressed proteins with false discovery rate below 0.05 are highlighted in pink. Open circles indicate that the protein is non-specific from the IgG control comparison.", sep = ""), fig.height=6}
print(maPlot(results,
             significanceLevel = significanceLevel,
             controlLogFoldChangeThreshold = controlLogFoldChangeThreshold,
             minLogFoldChangeForLabelling = 1.5,
             maxNumberLabelledProteins = 50,
             pointSize = 1.5))
```

```{r volcano_plot_2, echo=FALSE, warning=FALSE, message=FALSE, fig.cap=paste("Volcano plot of the average intensity against log~2~ fold change for the ", contrastName, " comparison (", processingDescription, "). Top ranking differentially-expressed proteins with false discovery rate below 0.05 are highlighted in pink. Open circles indicate that the protein is non-specific from the IgG control comparison.", sep = ""), fig.height=6}
print(volcanoPlot(results,
                  significanceLevel = significanceLevel,
                  minLogFoldChange = -2.0,
                  maxLogFoldChange = 2.75,
                  controlLogFoldChangeThreshold = controlLogFoldChangeThreshold,
                  minLogFoldChangeForLabelling = 1.5,
                  maxNumberLabelledProteins = 50,
                  pointSize = 1.5))
```

```{r pvalue_histogram_2, echo=FALSE, warning=FALSE, message=FALSE, fig.cap=paste("Histogram of p-values for the ", contrastName, " comparison (", processingDescription, ")", sep = ""), fig.height=3.5, fig.width=4.5}
print(histogram(results, "P.Value", xlab = "p-value"))
```

```{r qq_plot_2, echo=FALSE, warning=FALSE, message=FALSE, fig.cap=paste("QQ plot of the adjusted p-values for the ", contrastName, " comparison (", processingDescription, ")", sep = ""), fig.height=5, fig.width=5}
q <- results %>%
  select(adj.P.Val) %>%
  filter(!is.na(adj.P.Val)) %>%
  arrange(adj.P.Val) %>%
  unlist %>%
  as.numeric
qqplot(-log10(ppoints(length(q))),
       -log10(q),
       xlab=expression(Expected~~-log[10](italic(q))),
       ylab=expression(Observed~~-log[10](italic(q))))
abline(0, 1, col = "red")
```

\newpage

```{r contrast_3, include = FALSE}
contrastGroups <- c("ER 45min", "ER 0min")
contrastName <- paste(contrastGroups, collapse = " vs ")
processingDescription <- "excluding peptides with missing intensities, quantile normalization excluding IgG control"
analysis <- "ExcludeMissingValues_QuantileNormalizationExcludingIgG"
```

## `r paste(contrastName, processingDescription, sep = ", ")`

```{r results_3, include = FALSE}

data <- combinedProteinData[[analysis]]
exprs(data) <- log2(exprs(data))

results <- createContrastResults(
  data,
  fittedLinearModels[[analysis]],
  fittedLinearModelsExcludingControls[[analysis]],
  fittedContrasts[[analysis]],
  runs,
  contrastGroups,
  controlGroup,
  proteinInfo)
```

```{r ma_plot_3, echo=FALSE, warning=FALSE, message=FALSE, fig.cap=paste("MA plot of the average intensity against log~2~ fold change for the ", contrastName, " comparison (", processingDescription, "). Top ranking differentially-expressed proteins with false discovery rate below 0.05 are highlighted in pink. Open circles indicate that the protein is non-specific from the IgG control comparison.", sep = ""), fig.height=6}
print(maPlot(results,
             significanceLevel = significanceLevel,
             controlLogFoldChangeThreshold = controlLogFoldChangeThreshold,
             minLogFoldChangeForLabelling = 1.5,
             maxNumberLabelledProteins = 50,
             pointSize = 1.5))
```

```{r volcano_plot_3, echo=FALSE, warning=FALSE, message=FALSE, fig.cap=paste("Volcano plot of the average intensity against log~2~ fold change for the ", contrastName, " comparison (", processingDescription, "). Top ranking differentially-expressed proteins with false discovery rate below 0.05 are highlighted in pink. Open circles indicate that the protein is non-specific from the IgG control comparison.", sep = ""), fig.height=6}
print(volcanoPlot(results,
                  significanceLevel = significanceLevel,
                  minLogFoldChange = -2.0,
                  maxLogFoldChange = 2.75,
                  controlLogFoldChangeThreshold = controlLogFoldChangeThreshold,
                  minLogFoldChangeForLabelling = 1.5,
                  maxNumberLabelledProteins = 50,
                  pointSize = 1.5))
```

```{r pvalue_histogram_3, echo=FALSE, warning=FALSE, message=FALSE, fig.cap=paste("Histogram of p-values for the ", contrastName, " comparison (", processingDescription, ")", sep = ""), fig.height=3.5, fig.width=4.5}
print(histogram(results, "P.Value", xlab = "p-value"))
```

```{r qq_plot_3, echo=FALSE, warning=FALSE, message=FALSE, fig.cap=paste("QQ plot of the adjusted p-values for the ", contrastName, " comparison (", processingDescription, ")", sep = ""), fig.height=5, fig.width=5}
q <- results %>%
  select(adj.P.Val) %>%
  filter(!is.na(adj.P.Val)) %>%
  arrange(adj.P.Val) %>%
  unlist %>%
  as.numeric
qqplot(-log10(ppoints(length(q))),
       -log10(q),
       xlab=expression(Expected~~-log[10](italic(q))),
       ylab=expression(Observed~~-log[10](italic(q))))
abline(0, 1, col = "red")
```

\newpage

```{r contrast_4, include = FALSE}
contrastGroups <- c("ER 45min", "ER 0min")
contrastName <- paste(contrastGroups, collapse = " vs ")
processingDescription <- "excluding peptides with missing intensities, scale normalization"
analysis <- "ExcludeMissingValues_ScaleNormalization"
```

## `r paste(contrastName, processingDescription, sep = ", ")`

```{r results_4, include = FALSE}

data <- combinedProteinData[[analysis]]
exprs(data) <- log2(exprs(data))

results <- createContrastResults(
  data,
  fittedLinearModels[[analysis]],
  fittedLinearModelsExcludingControls[[analysis]],
  fittedContrasts[[analysis]],
  runs,
  contrastGroups,
  controlGroup,
  proteinInfo)
```

```{r ma_plot_4, echo=FALSE, warning=FALSE, message=FALSE, fig.cap=paste("MA plot of the average intensity against log~2~ fold change for the ", contrastName, " comparison (", processingDescription, "). Top ranking differentially-expressed proteins with false discovery rate below 0.05 are highlighted in pink. Open circles indicate that the protein is non-specific from the IgG control comparison.", sep = ""), fig.height=6}
print(maPlot(results,
             significanceLevel = significanceLevel,
             controlLogFoldChangeThreshold = controlLogFoldChangeThreshold,
             minLogFoldChangeForLabelling = 1.5,
             maxNumberLabelledProteins = 50,
             pointSize = 1.5))
```

```{r volcano_plot_4, echo=FALSE, warning=FALSE, message=FALSE, fig.cap=paste("Volcano plot of the average intensity against log~2~ fold change for the ", contrastName, " comparison (", processingDescription, "). Top ranking differentially-expressed proteins with false discovery rate below 0.05 are highlighted in pink. Open circles indicate that the protein is non-specific from the IgG control comparison.", sep = ""), fig.height=6}
print(volcanoPlot(results,
                  significanceLevel = significanceLevel,
                  minLogFoldChange = -2.0,
                  maxLogFoldChange = 2.75,
                  controlLogFoldChangeThreshold = controlLogFoldChangeThreshold,
                  minLogFoldChangeForLabelling = 1.5,
                  maxNumberLabelledProteins = 50,
                  pointSize = 1.5))
```

```{r pvalue_histogram_4, echo=FALSE, warning=FALSE, message=FALSE, fig.cap=paste("Histogram of p-values for the ", contrastName, " comparison (", processingDescription, ")", sep = ""), fig.height=3.5, fig.width=4.5}
print(histogram(results, "P.Value", xlab = "p-value"))
```

```{r qq_plot_4, echo=FALSE, warning=FALSE, message=FALSE, fig.cap=paste("QQ plot of the adjusted p-values for the ", contrastName, " comparison (", processingDescription, ")", sep = ""), fig.height=5, fig.width=5}
q <- results %>%
  select(adj.P.Val) %>%
  filter(!is.na(adj.P.Val)) %>%
  arrange(adj.P.Val) %>%
  unlist %>%
  as.numeric
qqplot(-log10(ppoints(length(q))),
       -log10(q),
       xlab=expression(Expected~~-log[10](italic(q))),
       ylab=expression(Observed~~-log[10](italic(q))))
abline(0, 1, col = "red")
```

\newpage

```{r contrast_5, include = FALSE}
contrastGroups <- c("ER 45min", "ER 0min")
contrastName <- paste(contrastGroups, collapse = " vs ")
processingDescription <- paste("excluding peptides with missing intensities, scale normalization using top", numberOfPeptidesWithHighestIgG, "IgG peptides")
analysis <- "ExcludeMissingValues_ScaleNormalizationTopIgG"
```

## `r paste(contrastName, processingDescription, sep = ", ")`

```{r results_5, include = FALSE}

data <- combinedProteinData[[analysis]]
exprs(data) <- log2(exprs(data))

results <- createContrastResults(
  data,
  fittedLinearModels[[analysis]],
  fittedLinearModelsExcludingControls[[analysis]],
  fittedContrasts[[analysis]],
  runs,
  contrastGroups,
  controlGroup,
  proteinInfo)
```

```{r ma_plot_5, echo=FALSE, warning=FALSE, message=FALSE, fig.cap=paste("MA plot of the average intensity against log~2~ fold change for the ", contrastName, " comparison (", processingDescription, "). Top ranking differentially-expressed proteins with false discovery rate below 0.05 are highlighted in pink. Open circles indicate that the protein is non-specific from the IgG control comparison.", sep = ""), fig.height=6}
print(maPlot(results,
             significanceLevel = significanceLevel,
             controlLogFoldChangeThreshold = controlLogFoldChangeThreshold,
             minLogFoldChangeForLabelling = 1.5,
             maxNumberLabelledProteins = 50,
             pointSize = 1.5))
```

```{r volcano_plot_5, echo=FALSE, warning=FALSE, message=FALSE, fig.cap=paste("Volcano plot of the average intensity against log~2~ fold change for the ", contrastName, " comparison (", processingDescription, "). Top ranking differentially-expressed proteins with false discovery rate below 0.05 are highlighted in pink. Open circles indicate that the protein is non-specific from the IgG control comparison.", sep = ""), fig.height=6}
print(volcanoPlot(results,
                  significanceLevel = significanceLevel,
                  minLogFoldChange = -2.0,
                  maxLogFoldChange = 2.75,
                  controlLogFoldChangeThreshold = controlLogFoldChangeThreshold,
                  minLogFoldChangeForLabelling = 1.5,
                  maxNumberLabelledProteins = 50,
                  pointSize = 1.5))
```

```{r pvalue_histogram_5, echo=FALSE, warning=FALSE, message=FALSE, fig.cap=paste("Histogram of p-values for the ", contrastName, " comparison (", processingDescription, ")", sep = ""), fig.height=3.5, fig.width=4.5}
print(histogram(results, "P.Value", xlab = "p-value"))
```

```{r qq_plot_5, echo=FALSE, warning=FALSE, message=FALSE, fig.cap=paste("QQ plot of the adjusted p-values for the ", contrastName, " comparison (", processingDescription, ")", sep = ""), fig.height=5, fig.width=5}
q <- results %>%
  select(adj.P.Val) %>%
  filter(!is.na(adj.P.Val)) %>%
  arrange(adj.P.Val) %>%
  unlist %>%
  as.numeric
qqplot(-log10(ppoints(length(q))),
       -log10(q),
       xlab=expression(Expected~~-log[10](italic(q))),
       ylab=expression(Observed~~-log[10](italic(q))))
abline(0, 1, col = "red")
```

\newpage

```{r contrast_6, include = FALSE}
contrastGroups <- c("ER 45min", "ER 0min")
contrastName <- paste(contrastGroups, collapse = " vs ")
processingDescription <- "KNN nearest neighbour averaging imputation of missing values, quantile normalization"
analysis <- "ImputeKNN_QuantileNormalization"
```

## `r paste(contrastName, processingDescription, sep = ", ")`

```{r results_6, include = FALSE}

data <- combinedProteinData[[analysis]]
exprs(data) <- log2(exprs(data))

results <- createContrastResults(
  data,
  fittedLinearModels[[analysis]],
  fittedLinearModelsExcludingControls[[analysis]],
  fittedContrasts[[analysis]],
  runs,
  contrastGroups,
  controlGroup,
  proteinInfo)
```

```{r ma_plot_6, echo=FALSE, warning=FALSE, message=FALSE, fig.cap=paste("MA plot of the average intensity against log~2~ fold change for the ", contrastName, " comparison (", processingDescription, "). Top ranking differentially-expressed proteins with false discovery rate below 0.05 are highlighted in pink. Open circles indicate that the protein is non-specific from the IgG control comparison.", sep = ""), fig.height=6}
print(maPlot(results,
             significanceLevel = significanceLevel,
             controlLogFoldChangeThreshold = controlLogFoldChangeThreshold,
             minLogFoldChangeForLabelling = 1.5,
             maxNumberLabelledProteins = 50,
             pointSize = 1.5))
```

```{r volcano_plot_6, echo=FALSE, warning=FALSE, message=FALSE, fig.cap=paste("Volcano plot of the average intensity against log~2~ fold change for the ", contrastName, " comparison (", processingDescription, "). Top ranking differentially-expressed proteins with false discovery rate below 0.05 are highlighted in pink. Open circles indicate that the protein is non-specific from the IgG control comparison.", sep = ""), fig.height=6}
print(volcanoPlot(results,
                  significanceLevel = significanceLevel,
                  minLogFoldChange = -2.0,
                  maxLogFoldChange = 2.75,
                  controlLogFoldChangeThreshold = controlLogFoldChangeThreshold,
                  minLogFoldChangeForLabelling = 1.5,
                  maxNumberLabelledProteins = 50,
                  pointSize = 1.5))
```

```{r pvalue_histogram_6, echo=FALSE, warning=FALSE, message=FALSE, fig.cap=paste("Histogram of p-values for the ", contrastName, " comparison (", processingDescription, ")", sep = ""), fig.height=3.5, fig.width=4.5}
print(histogram(results, "P.Value", xlab = "p-value"))
```

```{r qq_plot_6, echo=FALSE, warning=FALSE, message=FALSE, fig.cap=paste("QQ plot of the adjusted p-values for the ", contrastName, " comparison (", processingDescription, ")", sep = ""), fig.height=5, fig.width=5}
q <- results %>%
  select(adj.P.Val) %>%
  filter(!is.na(adj.P.Val)) %>%
  arrange(adj.P.Val) %>%
  unlist %>%
  as.numeric
qqplot(-log10(ppoints(length(q))),
       -log10(q),
       xlab=expression(Expected~~-log[10](italic(q))),
       ylab=expression(Observed~~-log[10](italic(q))))
abline(0, 1, col = "red")
```

\newpage

# Comparison of differential binding analysis approaches

In this section, comparisons of the differential binding results obtained using differing methods for handling missing values and varying normalization techniques are presented. The B-statistic (log odds) is used to rank the proteins from most significantly differentially expressed to least significant. Scatter plots of both the B-statistic generated for the two methods being compared and the rank are presented. The protein with the highest B-statistic, i.e. the most significantly differentially expressed protein, is given a rank of 1.

Differing imputation methods have relatively little effect on the results for most proteins. This is largely because only a small proportion of the proteins have peptide-level measurements containing missing values. There are a few proteins that are only represented in the differential binding results when imputation is carried out, for which all peptide-level measurements contain missing values. These proteins are not represented in the comparions with the analysis in which measurements with missing values are removed.

Different normalization approaches have a more substantial impact on the differential binding analysis.

\newpage

## Removal of missing values vs KNN imputation (quantile normalization in both)

```{r exclude_missing_vs_knn_imputation, include = FALSE}

contrastGroups <- c("ER 45min", "ER 0min")
contrastName <- paste(contrastGroups, collapse = " vs ")

analysis <- "ExcludeMissingValues_QuantileNormalization"
data <- combinedProteinData[[analysis]]
exprs(data) <- log2(exprs(data))
results1 <- createContrastResults(
  data,
  fittedLinearModels[[analysis]],
  fittedLinearModelsExcludingControls[[analysis]],
  fittedContrasts[[analysis]],
  runs,
  contrastGroups,
  controlGroup,
  proteinInfo)

analysis <- "ImputeKNN_QuantileNormalization"
data <- combinedProteinData[[analysis]]
exprs(data) <- log2(exprs(data))
results2 <- createContrastResults(
  data,
  fittedLinearModels[[analysis]],
  fittedLinearModelsExcludingControls[[analysis]],
  fittedContrasts[[analysis]],
  runs,
  contrastGroups,
  controlGroup,
  proteinInfo)

results <- inner_join(
  results1 %>% select(Protein, B1 = B),
  results2 %>% select(Protein, B2 = B),
  by = "Protein"
  ) %>%
  filter(complete.cases(.)) %>%
  mutate(rank1 = rank(-B1), rank2 = rank(-B2))
```

```{r exclude_missing_vs_knn_imputation_B_plot, echo=FALSE, warning=FALSE, message=FALSE, fig.cap=paste("Plot of B-statistic (log odds) for differential expression of proteins for the ", contrastName, " contrast, comparing results following exclusion of missing values and KNN imputation.", sep = ""), fig.height=2.9, fig.width=3.5}
print(scatterPlot(results, "B1", "B2", "B-statistic (excluding missing values)", "B-statistic (KNN imputation)"))
```

```{r exclude_missing_vs_knn_imputation_rank_plot, echo=FALSE, warning=FALSE, message=FALSE, fig.cap=paste("Plot of the rank of differentially-expressed proteins for the ", contrastName, " contrast, comparing results following exclusion of missing values and KNN imputation.", sep = ""), fig.height=2.9, fig.width=3.5}
print(scatterPlot(results, "rank1", "rank2", "Rank (excluding missing values)", "Rank (KNN imputation)"))
```

\newpage

## Removal of missing values vs lowest value imputation (quantile normalization in both)

```{r exclude_missing_vs_min_imputation, include = FALSE}

contrastGroups <- c("ER 45min", "ER 0min")
contrastName <- paste(contrastGroups, collapse = " vs ")

analysis <- "ExcludeMissingValues_QuantileNormalization"
data <- combinedProteinData[[analysis]]
exprs(data) <- log2(exprs(data))
results1 <- createContrastResults(
  data,
  fittedLinearModels[[analysis]],
  fittedLinearModelsExcludingControls[[analysis]],
  fittedContrasts[[analysis]],
  runs,
  contrastGroups,
  controlGroup,
  proteinInfo)

analysis <- "ImputeMin_QuantileNormalization"
data <- combinedProteinData[[analysis]]
exprs(data) <- log2(exprs(data))
results2 <- createContrastResults(
  data,
  fittedLinearModels[[analysis]],
  fittedLinearModelsExcludingControls[[analysis]],
  fittedContrasts[[analysis]],
  runs,
  contrastGroups,
  controlGroup,
  proteinInfo)

results <- inner_join(
  results1 %>% select(Protein, B1 = B),
  results2 %>% select(Protein, B2 = B),
  by = "Protein"
  ) %>%
  filter(complete.cases(.)) %>%
  mutate(rank1 = rank(-B1), rank2 = rank(-B2))
```

```{r exclude_missing_vs_min_imputation_B_plot, echo=FALSE, warning=FALSE, message=FALSE, fig.cap=paste("Plot of B-statistic (log odds) for differential expression of proteins for the ", contrastName, " contrast, comparing results following exclusion of missing values and lowest value imputation.", sep = ""), fig.height=2.9, fig.width=3.5}
print(scatterPlot(results, "B1", "B2", "B-statistic (excluding missing values)", "B-statistic (lowest value imputation)"))
```

```{r exclude_missing_vs_min_imputation_rank_plot, echo=FALSE, warning=FALSE, message=FALSE, fig.cap=paste("Plot of the rank of differentially-expressed proteins for the ", contrastName, " contrast, comparing results following exclusion of missing values and lowest value imputation.", sep = ""), fig.height=2.9, fig.width=3.5}
print(scatterPlot(results, "rank1", "rank2", "Rank (excluding missing values)", "Rank (lowest value imputation)"))
```

\newpage

## Quantile vs Scale normalization (missing values excluded in both)

```{r quantile_vs_scale_normalization, include = FALSE}

contrastGroups <- c("ER 45min", "ER 0min")
contrastName <- paste(contrastGroups, collapse = " vs ")

analysis <- "ExcludeMissingValues_QuantileNormalization"
data <- combinedProteinData[[analysis]]
exprs(data) <- log2(exprs(data))
results1 <- createContrastResults(
  data,
  fittedLinearModels[[analysis]],
  fittedLinearModelsExcludingControls[[analysis]],
  fittedContrasts[[analysis]],
  runs,
  contrastGroups,
  controlGroup,
  proteinInfo)

analysis <- "ExcludeMissingValues_ScaleNormalization"
data <- combinedProteinData[[analysis]]
exprs(data) <- log2(exprs(data))
results2 <- createContrastResults(
  data,
  fittedLinearModels[[analysis]],
  fittedLinearModelsExcludingControls[[analysis]],
  fittedContrasts[[analysis]],
  runs,
  contrastGroups,
  controlGroup,
  proteinInfo)

results <- inner_join(
  results1 %>% select(Protein, B1 = B),
  results2 %>% select(Protein, B2 = B),
  by = "Protein"
  ) %>%
  filter(complete.cases(.)) %>%
  mutate(rank1 = rank(-B1), rank2 = rank(-B2))
```

```{r scale_vs_quantile_normalization_B_plot, echo=FALSE, warning=FALSE, message=FALSE, fig.cap=paste("Plot of B-statistic (log odds) for differential expression of proteins for the ", contrastName, " contrast, comparing results following quantile and scale normalization.", sep = ""), fig.height=2.9, fig.width=3.5}
print(scatterPlot(results, "B1", "B2", "B-statistic (quantile normalization)", "B-statistic (scale normalization)"))
```

```{r scale_vs_quantile_normalization_rank_plot, echo=FALSE, warning=FALSE, message=FALSE, fig.cap=paste("Plot of the rank of differentially-expressed proteins for the ", contrastName, " contrast, comparing results following quantile and scale normalization.", sep = ""), fig.height=2.9, fig.width=3.5}
print(scatterPlot(results, "rank1", "rank2", "Rank (quantile normalization)", "Rank (scale normalization)"))
```

\newpage

## Scale normalization based on all peptides vs peptides with highest IgG intensities (missing values excluded in both)

```{r scale_normalization_all_vs_topIgG, include = FALSE}

contrastGroups <- c("ER 45min", "ER 0min")
contrastName <- paste(contrastGroups, collapse = " vs ")

analysis <- "ExcludeMissingValues_ScaleNormalization"
data <- combinedProteinData[[analysis]]
exprs(data) <- log2(exprs(data))
results1 <- createContrastResults(
  data,
  fittedLinearModels[[analysis]],
  fittedLinearModelsExcludingControls[[analysis]],
  fittedContrasts[[analysis]],
  runs,
  contrastGroups,
  controlGroup,
  proteinInfo)

analysis <- "ExcludeMissingValues_ScaleNormalizationTopIgG"
data <- combinedProteinData[[analysis]]
exprs(data) <- log2(exprs(data))
results2 <- createContrastResults(
  data,
  fittedLinearModels[[analysis]],
  fittedLinearModelsExcludingControls[[analysis]],
  fittedContrasts[[analysis]],
  runs,
  contrastGroups,
  controlGroup,
  proteinInfo)

results <- inner_join(
  results1 %>% select(Protein, B1 = B),
  results2 %>% select(Protein, B2 = B),
  by = "Protein"
  ) %>%
  filter(complete.cases(.)) %>%
  mutate(rank1 = rank(-B1), rank2 = rank(-B2))
```

```{r scale_normalization_all_vs_topIgG_B_plot, echo=FALSE, warning=FALSE, message=FALSE, fig.cap=paste("Plot of B-statistic (log odds) for differential expression of proteins for the ", contrastName, " contrast, comparing results following scale normalization based on all peptides and those with the highest IgG intensities.", sep = ""), fig.height=2.5, fig.width=3.5}
print(scatterPlot(results, "B1", "B2", "B-statistic (scale normalization based on all peptides)", "B-statistic (scale normalization, peptides with highest IgG)"))
```

```{r scale_normalization_all_vs_topIgG_rank_plot, echo=FALSE, warning=FALSE, message=FALSE, fig.cap=paste("Plot of the rank of differentially-expressed proteins for the ", contrastName, " contrast, comparing results following scale normalization based on all peptides and those with the highest IgG intensities.", sep = ""), fig.height=2.5, fig.width=3.5}
print(scatterPlot(results, "rank1", "rank2", "Rank (scale normalization based on all peptides)", "Rank (scale normalization, peptides with highest IgG)"))
```

\newpage

# Comparisoin of variance between specfic and non-specific binding

```{r  echo=FALSE, warning=FALSE, message=FALSE}

maPlot <- function(differentialExpressionResults, selectedGenes = NULL,
                   xlab = expression(average~log[2](intensity)),
                   ylab = expression(log[2](fold~change)),
                   significanceLevel = 0.05,
                   minLogFoldChangeForLabelling = 1.5,
                   maxNumberLabelledProteins = 50,
                   controlLogFoldChangeThreshold = -Inf,
                   pointSize = 2.5) {
  if (!"logFCcontrol" %in% colnames(differentialExpressionResults))
    differentialExpressionResults$logFCcontrol <- Inf

  differentialExpressionResults$group <-
    ifelse(differentialExpressionResults$Gene %in% selectedGenes, 0,
    ifelse(!is.na(differentialExpressionResults$adj.P.Val) & differentialExpressionResults$adj.P.Val <= significanceLevel & differentialExpressionResults$logFCcontrol >= controlLogFoldChangeThreshold, 1,
    ifelse(!is.na(differentialExpressionResults$adj.P.Val) & differentialExpressionResults$adj.P.Val <= significanceLevel, 2,
    ifelse(abs(differentialExpressionResults$logFC) >= minLogFoldChangeForLabelling & differentialExpressionResults$logFCcontrol >= controlLogFoldChangeThreshold, 3,
    ifelse(abs(differentialExpressionResults$logFC) >= minLogFoldChangeForLabelling, 4,
    ifelse(differentialExpressionResults$logFCcontrol >= controlLogFoldChangeThreshold, 5, 6))))))

  # 0 - selected
  # 1 - significant, specific
  # 2 - significant, non-specific
  # 3 - non-significant, large logfc, specific
  # 4 - non-significant, large logfc, non-specific
  # 5 - specific
  # 6 - non-specific

  differentialExpressionResults <- differentialExpressionResults %>% arrange(desc(group))

  differentialExpressionResults$group <- factor(differentialExpressionResults$group, 0:6)

  differentialExpressionResults <- differentialExpressionResults %>%
    mutate(label = paste(" ", Gene)) %>%
    mutate(logFCcontrolAlpha = pmin(logFCcontrol, 2.0))

  labelSubset <- bind_rows(
    differentialExpressionResults %>%
      filter(group != 0 & adj.P.Val <= significanceLevel) %>%
      arrange(desc(abs(logFC))) %>%
      slice(1:(maxNumberLabelledProteins / 2)),
    differentialExpressionResults %>%
      filter(group != 0 & abs(logFC) >= minLogFoldChangeForLabelling) %>%
      arrange(desc(abs(logFC))) %>%
      slice(1:maxNumberLabelledProteins)
  ) %>%
    unique %>%
    slice(1:maxNumberLabelledProteins)

    if (maxNumberLabelledProteins == 0) labelSubset <- labelSubset %>% slice(0)

  # plot <- ggplot(differentialExpressionResults, aes(x = AveExpr, y = logFC, colour = group, size = group, shape = group, alpha = logFCcontrolAlpha))
  plot <- ggplot(differentialExpressionResults, aes(x = AveExpr, y = logFC, colour = group))+ coord_cartesian(ylim = c(-3.5,3.5))
  plot <- plot + geom_hline(yintercept = 0, color="gray50", size = 0.5)
  plot <- plot + geom_point(alpha=0.75)
  plot <- plot + geom_text(data = labelSubset, aes(label = label), hjust = 0, size = 3.5)
  plot <- plot + geom_point(data = subset(differentialExpressionResults, group == 0), alpha = 0.7)
  plot <- plot + geom_text(data = subset(differentialExpressionResults, group == 0), aes(label = label), hjust = 0, size = 4.0, alpha = 1)
  plot <- plot + scale_colour_manual(values = c("green", "red", "blue", "red", "blue", "red", "blue"), drop = FALSE)
  # plot <- plot + scale_colour_manual(values = c("blue", "deeppink3", "deeppink3", "gray50", "gray50", "gray50", "gray50"), drop = FALSE)
  # plot <- plot + scale_size_manual(values = c(1.2 * pointSize, pointSize, pointSize, 0.8 * pointSize, 0.8 * pointSize, 0.6 * pointSize, 0.6 * pointSize), drop = FALSE)
  # plot <- plot + scale_shape_manual(values = c(16, 16, 1, 16, 1, 16, 1), drop = FALSE)
  #plot <- plot + scale_alpha(range = c(0.2, 1.0))
  plot <- plot + xlab(xlab)
  plot <- plot + ylab(ylab)
  plot <- plot + theme_bw()
  plot <- plot + theme(
    text = element_text(size = 14),
    legend.position = "none"
  )

  save(differentialExpressionResults,file="Results/QRIME.RData") 
  return(plot)
}

contrastGroups <- c("ER 45min", "ER 0min")
contrastName <- paste(contrastGroups, collapse = " vs ")

analysis <- "ExcludeMissingValues_QuantileNormalization"

data <- combinedProteinData[[analysis]]
exprs(data) <- log2(exprs(data))

results <- createContrastResults(
  data,
  fittedLinearModels[[analysis]],
  fittedLinearModelsExcludingControls[[analysis]],
  fittedContrasts[[analysis]],
  runs,
  c("ER 45min", "ER 0min"),
  controlGroup,
  proteinInfo)

```

```{r MA_specificVsNonSpecfic, echo=FALSE, warning=FALSE, message=FALSE, fig.cap="MA plot showing the distribution of non-specific binding. Non-specific binding was identified as proteins that showed less than a log2 enrichment over the IgG channel at either 0 or 45 minutes. The majority of non-specific binding (blue) is found to show less change between time points whereas the specific interactions (red) shows a broader distribution.", fig.height=2.5, fig.width=3.5}

print(maPlot(results,
             significanceLevel = 0.5,
             controlLogFoldChangeThreshold = 1.0,
             minLogFoldChangeForLabelling = 1.5,
             maxNumberLabelledProteins = 0,
             pointSize = 1.5))
```

```{r  echo=FALSE, warning=FALSE, message=FALSE}

load("Results/QRIME.RData")
as.numeric(differentialExpressionResults$group)%%2->differentialExpressionResults$group

tmp<-differentialExpressionResults

```

```{r density_specificVsNonSpecfici, echo=FALSE, warning=FALSE, message=FALSE, fig.cap="Density plot right showing the distribution of non-specific binding. Non-specific binding was identified as proteins that showed less than a log2 enrichment over the IgG channel at either 0 or 45 minutes. The majority of non-specific binding (blue) is found to show less change between time points whereas the specific interactions (red) shows a broader distribution.", fig.height=3.5, fig.width=4.5}

plot(density(tmp$logFC[tmp$group==1]),col="blue",ylab="Density", xlab="Fold Change", main="", xlim=c(3,-3), lwd=2)
lines(density(tmp$logFC[tmp$group==0]),col="red",lwd=2)

```

```{r box_specificVsNonSpecfici, echo=FALSE, warning=FALSE, message=FALSE, fig.cap=" Box plot showing the variance of non-specific vs non-specific binding. Non-specific binding was identified as proteins that showed less than a log2 enrichment over the IgG channel at either 0 or 45 minutes. The the variance of the log fold changes for the non-specific binding interactions was significantly less than that of the specific interactions (F-test, one-sided, p-value < 2.2e*-16).", fig.height=3.5, fig.width=4.5}

boxplot(tmp$logFC[tmp$group==1],tmp$logFC[tmp$group==0],ylab="Fold Change", names=c("Non-specific","Specific"), col=c("#0000FFAA","#FF0000AA"))

```

```{r  echo=FALSE, warning=FALSE, message=FALSE}

var.test(tmp$logFC[tmp$group==1],tmp$logFC[tmp$group==0], alternative="less")

```

\newpage

# Differential binding results tables

```{r contrast_results_thresholds, include = FALSE}
significanceLevel <- 0.05
logFoldChangeThreshold <- 1.0
controlLogFoldChangeThreshold <- 1.0
```

The following tables contain the top ranking differentially expressed proteins for each comparison. Included are all proteins that reach a statistical signficance of `r significanceLevel` in terms of the adjusted p-value and those with an absolute log~2~ fold change of `r logFoldChangeThreshold` or above.

The IgG column gives the larger of the log~2~ fold changes for the two groups against the IgG control and an asterisk indicates specific binding where this log~2~ fold change is above a threshold of `r controlLogFoldChangeThreshold`. N is the number of replicates in which the protein was observed.

In all cases, peptide intensities were quantile normalized and measurements with missing values were removed prior to summarization at the protein level.

```{r contrast_results_table, include = FALSE}

contrastResultsTable <- function(results, contrastName,
                                 significanceLevel = 0.05,
                                 logFoldChangeThreshold = 1.0,
                                 controlLogFoldChangeThreshold = 1.0)
{
  resultsTable <- results %>%
    select(Protein, Gene, N, logFC, AveExpr, adj.P.Val, B, logFCcontrol)

  resultsTable$Specific <- ifelse(resultsTable$logFCcontrol >= controlLogFoldChangeThreshold, "*", "")

  resultsTable <- resultsTable %>%
    filter(adj.P.Val < 0.05 | abs(logFC) >= 1.0) %>%
    arrange(desc(abs(logFC)))

  colnames(resultsTable) <- c("Protein", "Gene", "N", "log2FC", "Avg Expr", "p-value", "B", "IgG", "")

  print(
    xtable(
      resultsTable,
      align = "lllrrrrrrl",
      display = c("d", "s", "s", "d", "f", "f", "g", "f", "f", "s"),
      caption = paste("Top ranking differentially expressed proteins from the", contrastName, "comparison,  sorted by log2 fold change.")
    ),
    include.rownames = FALSE,
    tabular.environment = "longtable",
    floating = FALSE,
    add.to.row = addtorow
  )
}
```

\newpage

```{r results_table_contrast_1_config, include = FALSE}
contrastGroups <- c("ER 45min", "ER 0min")
contrastName <- paste(contrastGroups, collapse = " vs ")
processingDescription <- "excluding peptides with missing intensities, quantile normalization"
analysis <- "ExcludeMissingValues_QuantileNormalization"
```

## `r contrastName`

```{r results_table_contrast_1, echo=FALSE, warning=FALSE, message=FALSE, results='asis'}

data <- combinedProteinData[[analysis]]
exprs(data) <- log2(exprs(data))

results <- createContrastResults(
  data,
  fittedLinearModels[[analysis]],
  fittedLinearModelsExcludingControls[[analysis]],
  fittedContrasts[[analysis]],
  runs,
  contrastGroups,
  controlGroup,
  proteinInfo)

contrastResultsTable(results, contrastName,
                     significanceLevel = 0.05,
                     logFoldChangeThreshold = 1.0,
                     controlLogFoldChangeThreshold = 1.0)

resultsFile <- paste("Results/", paste(gsub(" ", "_", contrastGroups), collapse = "_vs_"), analysis, ".txt", sep = "")
write.table(results, resultsFile, sep = "\t", row.names = FALSE, quote = FALSE)
```

\newpage

```{r results_table_contrast_2_config, include = FALSE}
contrastGroups <- c("ER 90min", "ER 0min")
contrastName <- paste(contrastGroups, collapse = " vs ")
processingDescription <- "excluding peptides with missing intensities, quantile normalization"
analysis <- "ExcludeMissingValues_QuantileNormalization"
```

## `r contrastName`

```{r results_table_contrast_2, echo=FALSE, warning=FALSE, message=FALSE, results='asis'}

data <- combinedProteinData[[analysis]]
exprs(data) <- log2(exprs(data))

results <- createContrastResults(
  data,
  fittedLinearModels[[analysis]],
  fittedLinearModelsExcludingControls[[analysis]],
  fittedContrasts[[analysis]],
  runs,
  contrastGroups,
  controlGroup,
  proteinInfo)

contrastResultsTable(results, contrastName,
                     significanceLevel = 0.05,
                     logFoldChangeThreshold = 1.0,
                     controlLogFoldChangeThreshold = 1.0)

resultsFile <- paste("Results/",paste(gsub(" ", "_", contrastGroups), collapse = "_vs_"), analysis, ".txt", sep = "")
write.table(results, resultsFile, sep = "\t", row.names = FALSE, quote = FALSE)
```

\newpage

```{r results_table_contrast_3_config, include = FALSE}
contrastGroups <- c("ER 90min", "ER 45min")
contrastName <- paste(contrastGroups, collapse = " vs ")
processingDescription <- "excluding peptides with missing intensities, quantile normalization"
analysis <- "ExcludeMissingValues_QuantileNormalization"
```

## `r contrastName`

```{r results_table_contrast_3, echo=FALSE, warning=FALSE, message=FALSE, results='asis'}

data <- combinedProteinData[[analysis]]
exprs(data) <- log2(exprs(data))

results <- createContrastResults(
  data,
  fittedLinearModels[[analysis]],
  fittedLinearModelsExcludingControls[[analysis]],
  fittedContrasts[[analysis]],
  runs,
  contrastGroups,
  controlGroup,
  proteinInfo)

contrastResultsTable(results, contrastName,
                     significanceLevel = 0.05,
                     logFoldChangeThreshold = 1.0,
                     controlLogFoldChangeThreshold = 1.0)

resultsFile <- paste("Results/",paste(gsub(" ", "_", contrastGroups), collapse = "_vs_"), analysis, ".txt", sep = "")
write.table(results, resultsFile, sep = "\t", row.names = FALSE, quote = FALSE)
```

\newpage

```{r results_table_contrast_4_config, include = FALSE}
contrastGroups <- c("FOXA1 45min", "FOXA1 0min")
contrastName <- paste(contrastGroups, collapse = " vs ")
processingDescription <- "excluding peptides with missing intensities, quantile normalization"
analysis <- "ExcludeMissingValues_QuantileNormalization"
```

## `r contrastName`

```{r results_table_contrast_4, echo=FALSE, warning=FALSE, message=FALSE, results='asis'}

data <- combinedProteinData[[analysis]]
exprs(data) <- log2(exprs(data))

results <- createContrastResults(
  data,
  fittedLinearModels[[analysis]],
  fittedLinearModelsExcludingControls[[analysis]],
  fittedContrasts[[analysis]],
  runs,
  contrastGroups,
  controlGroup,
  proteinInfo)

contrastResultsTable(results, contrastName,
                     significanceLevel = 0.05,
                     logFoldChangeThreshold = 1.0,
                     controlLogFoldChangeThreshold = 1.0)

resultsFile <- paste("Results/",paste(gsub(" ", "_", contrastGroups), collapse = "_vs_"), analysis, ".txt", sep = "")
write.table(results, resultsFile, sep = "\t", row.names = FALSE, quote = FALSE)
```

\newpage

```{r results_table_contrast_5_config, include = FALSE}
contrastGroups <- c("FOXA1 90min", "FOXA1 0min")
contrastName <- paste(contrastGroups, collapse = " vs ")
processingDescription <- "excluding peptides with missing intensities, quantile normalization"
analysis <- "ExcludeMissingValues_QuantileNormalization"
```

## `r contrastName`

```{r results_table_contrast_5, echo=FALSE, warning=FALSE, message=FALSE, results='asis'}

data <- combinedProteinData[[analysis]]
exprs(data) <- log2(exprs(data))

results <- createContrastResults(
  data,
  fittedLinearModels[[analysis]],
  fittedLinearModelsExcludingControls[[analysis]],
  fittedContrasts[[analysis]],
  runs,
  contrastGroups,
  controlGroup,
  proteinInfo)

contrastResultsTable(results, contrastName,
                     significanceLevel = 0.05,
                     logFoldChangeThreshold = 1.0,
                     controlLogFoldChangeThreshold = 1.0)

resultsFile <- paste("Results/",paste(gsub(" ", "_", contrastGroups), collapse = "_vs_"), analysis, ".txt", sep = "")
write.table(results, resultsFile, sep = "\t", row.names = FALSE, quote = FALSE)
```

\newpage

```{r results_table_contrast_6_config, include = FALSE}
contrastGroups <- c("FOXA1 90min", "FOXA1 45min")
contrastName <- paste(contrastGroups, collapse = " vs ")
processingDescription <- "excluding peptides with missing intensities, quantile normalization"
analysis <- "ExcludeMissingValues_QuantileNormalization"
```

## `r contrastName`

```{r results_table_contrast_6, echo=FALSE, warning=FALSE, message=FALSE, results='asis'}

data <- combinedProteinData[[analysis]]
exprs(data) <- log2(exprs(data))

results <- createContrastResults(
  data,
  fittedLinearModels[[analysis]],
  fittedLinearModelsExcludingControls[[analysis]],
  fittedContrasts[[analysis]],
  runs,
  contrastGroups,
  controlGroup,
  proteinInfo)

contrastResultsTable(results, contrastName,
                     significanceLevel = 0.05,
                     logFoldChangeThreshold = 1.0,
                     controlLogFoldChangeThreshold = 1.0)

resultsFile <- paste("Results/", paste(gsub(" ", "_", contrastGroups), collapse = "_vs_"), analysis, ".txt", sep = "")
write.table(results, resultsFile, sep = "\t", row.names = FALSE, quote = FALSE)
```

```{r save, eval = TRUE, include = FALSE, cache = FALSE}
save.image(file = "Results/QRIME.RData")
```

### 2018-01-23 ChIP-seq_ER/qrime/04.QRIMESceneInvestigatonOverview.Rmd
first commit

--- 2018-01-23 ChIP-seq_ER/qrime/02.Supplementary.Rmd
@@
 ---
-title: "Supplementary"
+title: "QRIME Scene Investigation - Overview"
 author: "Andrew Holding & Matt Eldridge"
@@
   pdf_document:
+    fig_caption: yes
     number_sections: yes
     toc: yes
-    fig_caption: yes
+  html_document:
+    toc: yes
 header-includes:
@@
 metadata <- read.delim("SampleData/metadata.txt", check.names = FALSE)
+contrastsTable <- read.delim("SampleData/contrasts.txt", check.names = FALSE)
 runs <- levels(metadata$Run)
@@
 
-# Introduction
+\newpage
 
-This report is from an analysis of the RIME-TMT proteomics data for an experiment carried out by Andrew Holding to quantify specific interactors of estrogen receptor (ER) and forkhead box protein A1 (FOXA1) in MCF-7 cells.
+# SampleSheet
 
-The RIME (Rapid Immunoprecipitation of Endogenous Proteins) technique was used to pull down ER and FOXA1 and their interacting proteins, which were subsequently assessed and quantitated using mass spectrometry with tandem mass tags (TMT).
-TMT runs were carried out for 3 biological replicates, estrogen-starved MCF-7 cells from the same cell line at different times/passages, before and after addition of estrogen with samples taken at 3 different time points -- 0, 45 and 90 minutes after estrogen addition.
-A pull down of Immunoglobulin G (IgG) was also run as a control.
-
-The digests from two of the biological replicates (PR622 and PR650) were analyzed with 3 separate runs of the mass spectrometer to increase coverage; data from the repeated runs were combined prior to the analysis carried out here.
-The other sample (PR590) was run once only.
-
-`r figr("samples_table", TRUE, type="Table")` shows the isobaric tags used for each sample within each of the runs.
+SampleSheet and ContrastTable configurations are loaded from tab-delimited format text files in the 'SampleData' folder.
+Reports are generated where requested and output to the 'Reports' folder.
+Plain text results are also provided in the 'Results' folder.
 
@@
 )
+
+print(
+  xtable(
+    contrastsTable %>% select(Contrast, Report),
+    caption = "Configured Constrasts."
+  ),
+  include.rownames = FALSE,
+  size = "\\setlength{\\tabcolsep}{12pt}"
+)
 ```
@@
 # Peptide intensity data
-
-Raw spectra were processed using Proteome Discover 2.1 to produce peptide-level intensity data with a single set of intensity values per distinct peptide.
-Only peptides with unique high-confidence protein matches were included.
-
-Multiple peptide-spectrum matches (PSMs) for the same peptide were combined using Proteome Discoverer by summing the PSM-level intensities.
-Peptide sequences with different modifications are treated as distinct peptides and the data provided contain intensities values for each modification identified for a peptide sequence.
-Several distinct peptide sequences (with modifications) may have been identified for any given protein.
-
-`r figr("peptide_protein_count_table", TRUE, type="Table")` shows the numbers of distinct peptides and proteins obtained in each TMT run.
-Also given are the numbers of peptides and proteins after filtering peptides with missing intensity values in one or more TMT channels.
 
@@
 
-## Missing intensity values
+\newpage
 
-Missing intensity values can result even though ions are present at detectable concentrations due to the stochastic nature of the mass spectrometry acquisition method.
-It is reasonable to expect that these missing values are randomly distributed in the data.
-Alternatively, missing values may occur when there is a low abundance of ions, below the limit of detection of the instrument.
-These biologically relevant missing values are not randomly distributed, affecting only those proteins that are expressed at low levels in the samples analysed.
-The R/Bioconductor package `MSnBase` provides imputation methods for both types of missing value.
-
-In this analysis, missing values have been handled in 3 different ways.
-The first approach is to exclude all peptide-level mesasurements where there is a missing value for one or more of the samples within a run.
-In addition, two imputation methods have been performed: one in which the missing values are set to the smallest non-missing value in the data for that run and the other in which k-nearest neighbour (KNN) averaging is applied.
+# Normalization of intensity data
 
@@
 ```
-
-`r figr("intensity_plots_impute_min", TRUE, type="Figure")` shows the distribution of intensities for each sample within each run following imputation using the smallest non-missing values.
-A small hump just below zero on the log~2~ scale is clearly visible for each run.
-
-```{r intensity_plots_impute_min, echo=FALSE, warning=FALSE, message=FALSE, fig.cap="Density plots of intensities from each run following imputation of missing values using the smallest non-missing value in the data for each run.", fig.height=8, fig.width=5}
-
-minIntensity <- -2
-maxIntensity <- 15
-
-plots <- vector("list")
-
-for (i in 1:length(runs))
-{
-  run <- runs[i]
-  intensities <- peptideDataImputedMin[[run]] %>% exprs %>% log2 %>% as.data.frame
-  plots[[run]] <- intensityDistributionPlot(
-    intensities,
-    groups,
-    groupColours,
-    title = run,
-    xlab=expression(log[2](intensity)),
-    minIntensity = minIntensity,
-    maxIntensity = maxIntensity,
-    showLegend = (i == 1))
-}
-
-do.call("grid.arrange", c(plots, ncol = 1))
-```
-
-# Normalization of intensity data
-
-The intensity distribution for PR622 suggests that the labeled samples have been pooled with differing protein concentrations and that some normalization is required to properly assess differences between groups.
-Normalization techniques that are commonly applied to microarray expression data and mRNA sequencing data assume that only a few genes, e.g.
-a few hundred out of tens of thousands, are expressed at very different levels between samples.
-This assumption may not hold for a RIME experiment where a specific set of interacting proteins is targeted.
-For example, if the majority of interacting proteins were to bind less strongly at one time point relative to another, quantile normalization or scale normalization approaches would have the effect of removing the general trend.
-In this case, following normalization, those proteins that bind less strongly may appear to be largely unchanged, while proteins with very similar levels of binding at the different time points would have artificially increased binding levels.
-Therefore some care is required when interpreting the results of a differential binding analysis for RIME TMT experiments.
-
-In this analysis, quantile normalization and scale normalization were applied for each run separately.
-
-The IgG control was used to assess non-specific binding.
-The binding of a protein was considered to be non-specific if the binding level of that protein in the condition of interest, e.g.
-ER at 45 minutes, was not significantly above the level observed for the IgG control.
-Proteins detected with the IgG control are expressed at lower levels in the PR590 and PR650 runs.
-Normalization of the IgG control along with the other samples would make it more difficult to distinguish between specific and non-specific binding since the intensities of the IgG control would be scaled to a similar level of those of the other samples.
-
-Among a number of normalization approaches tested here, normalizations were carried out that both include and exclude the IgG control.
-`r figr("normalized_intensity_plots", TRUE, type="Figure")` shows the normalized intensity distributions following scale normalization of all peptide intensities including those with missing values.
-`r figr("normalizedExcludingIgG_intensity_plots", TRUE, type="Figure")` shows the normalized intensity distributions following scale normalization for all samples except the IgG controls.
 
@@
 
-```{r normalized_intensity_plots, echo=FALSE, warning=FALSE, message=FALSE, fig.cap="Density plots of normalized intensities from each TMT run where scale normalization was applied to peptide intensities that include misssing values.", fig.height=8, fig.width=5}
+```{r normalized_intensity_plots, echo=FALSE, warning=FALSE, message=FALSE, fig.cap="Density plots of normalized intensities from each TMT run where quantile normalization was applied to peptide intensities that include misssing values.", fig.height=6.5, fig.width=5}
 
-analysis <- "IncludeMissingValues_ScaleNormalization"
+analysis <- "IncludeMissingValues_QuantileNormalization"
 
@@
 
-```{r normalizedExcludingIgG_intensity_plots, echo=FALSE, warning=FALSE, message=FALSE, fig.cap="Density plots of normalized intensities from each TMT run where scale normalization was applied to peptide intensities for all samples except the IgG controls.", fig.height=8, fig.width=5}
-
-analysis <- "IncludeMissingValues_ScaleNormalizationExcludingIgG"
-
-minIntensity <- 0
-maxIntensity <- 15
-
-plots <- vector("list")
-
-for (i in 1:length(runs))
-{
-  run <- runs[i]
-  intensities <- peptideDataNormalized[[analysis]][[run]] %>% exprs %>% log2 %>% as.data.frame
-  plots[[run]] <- intensityDistributionPlot(
-    intensities,
-    groups,
-    groupColours,
-    title = run,
-    xlab=expression(log[2](intensity)),
-    minIntensity = minIntensity,
-    maxIntensity = maxIntensity,
-    showLegend = (i == 1))
-}
-
-do.call("grid.arrange", c(plots, ncol = 1))
-```
-
-A further normalization approach was attempted that scales the data from all samples within a run based on a subset of peptides that are considered most likely to be at similar concentrations between the samples.
-The assumption is that the peptides with the highest intensity values in the IgG control are the result of non-specific binding and that that binding is consistent across all samples.
-Similar to the scale normalization applied in the first approach, the median intensity within each sample is computed to determine a scaling factor for each sample within a run but the median computed only uses the `r numberOfPeptidesWithHighestIgG` peptides with the highest IgG measurements.
-As before, normalization is carried out within each run separately.
-
-`r figr("normalizedIgG_intensity_plots", TRUE, type="Figure")` shows the resulting normalized intensity distributions.
-Strikingly, the IgG intensity distribution is shifted to lower values in run PR622, making it more consistent with the other 2 runs.
-
-```{r normalizedIgG_intensity_plots, echo=FALSE, warning=FALSE, message=FALSE, fig.cap=paste("Density plots of normalized intensities from each TMT run where scale normalization is applied based on", numberOfPeptidesWithHighestIgG, "peptides with the highest IgG intensities in each run."), fig.height=8, fig.width=5}
-
-analysis <- "IncludeMissingValues_ScaleNormalizationTopIgG"
-
-minIntensity <- 0
-maxIntensity <- 15
-
-plots <- vector("list")
-
-for (i in 1:length(runs))
-{
-  run <- runs[i]
-  intensities <- peptideDataNormalized[[analysis]][[run]] %>% exprs %>% log2 %>% as.data.frame
-  plots[[run]] <- intensityDistributionPlot(
-    intensities,
-    groups,
-    groupColours,
-    title = run,
-    xlab=expression(log[2](intensity)),
-    minIntensity = minIntensity,
-    maxIntensity = maxIntensity,
-    showLegend = (i == 1))
-}
-
-do.call("grid.arrange", c(plots, ncol = 1))
-```
+\newpage
 
 # Protein-level quantification
-
-Protein-level quantification was carried out for each run separately by summing normalized intensity values for all peptides matching to a particular protein.
-This is similar in principle to the gene-level quantification that is carried out during microarray analysis or the read-counting approach in RNA-Seq.
-Essentially, all signal attributable to a particular protein is assigned to that protein.
-Intensity values vary in magnitude depending on the elution profile for the peptide and when the peptide is sampled.
-Relative intensities for a peptide between the tags or samples within a run should be consistent but the signal may include some contribution from a contaminating, co-eluting peptide, the extent of which may differ for different PSMs.
-In summing intensities from multiple measurements for the same peptide, those peptides that have higher levels of intensity values will have a greater weight in determining the overall intensity; higher intensity measurements are likely to more accurately reflect differences between samples so this may be advantageous.
-
-When no imputation of missing values was carried out, peptides with missing values in one or more samples were excluded from the summation.
-Missing values that arise for technical reasons instead of being at undetectable levels in a sample would otherwise reduce the overall intensity in that sample relative to other samples.
 
@@
 
-The statistical analysis of differential binding between groups that follows requires at least 3 observations per group.
-`r figr("replicate_count_table", TRUE, type="Table")` gives the number of protein identified in just a single run, two runs or all three runs.
-A total of `r (replicateCounts %>% filter(N == "all"))$Count` proteins were identified across all runs of which `r (replicateCounts %>% filter(N == "3"))$Count` were observed in all three runs, allowing for a statistical analysis.
-log~2~ fold changes are still computed for proteins identified only in one or two of the runs, but no measure of the statistical significance is given in the differential binding analysis.
-
 ```{r replicate_count_table, echo=FALSE, warning=FALSE, message=FALSE, results='asis'}
@@
 # Principal Component Analysis
-
-Protein-level intensities were scaled to sum to 1.0 for each protein within a run, allowing for comparison across runs in a principal component analysis (PCA).
-
-`r figr("pca_plot", TRUE, type="Figure")` shows a PCA plot for the first two principal components using all proteins that were sampled in all three runs.
-The plot shows a clear separation of the ER, FOXA1 and IgG control pull-downs.
-The 0, 45 and 90 minute timepoints within each of the ER and FOXA1 groups are not completely separated in the first two principal components.
-`r figr("pca_plot_er", TRUE, type="Figure")` shows the PCA plot for just the ER samples.
 
@@
 
-```{r pca_plot_er, echo=FALSE, warning=FALSE, message=FALSE, fig.cap="Principal Componenent Analysis for ER-interactng proteins sampled in all 3 runs.
-The first two principal components are displayed.", fig.height=5}
+```{r knitreports, include = FALSE, cache = FALSE}
 
-metadata <- metadata %>% filter(grepl("^ER", Group))
+library("rmarkdown")
 
-print(pcaPlot(scaledIntensities, metadata$Sample, metadata$Group, groupColours, labels = metadata$Run, legend = TRUE))
-```
-
-# Differential Binding
-
-```{r differential_expression_paramters, include=FALSE}
-significanceLevel <- 0.05
-controlLogFoldChangeThreshold <- 1.0
-```
-
-A statistical analysis of differentially-expressed peptides was carried out using **limma**, a R/Bioconductor package commonly used in the analysis of microarray and RNA-seq data, but applicable to data from any quantitative expression technology.
-
-Limma uses linear models to assess differential expression in the context of multifactor experimental designs.
-It is able to analyze comparisons between many RNA targets simultaneously (or in this case many proteins) and has features that make these analyses stable even for experiments with small numbers of samples; this is achieved by borrowing information across genes (proteins).
-
-In this analysis, limma was used to estimate log~2~ fold changes and standard errors by fitting a linear model for each protein where the model includes variables for the group (the ER and FOXA1 pull-downs at 0, 45 and 90 minute time points and the IgG control at 45 minutes) and the run.
-This is essentially a two-way ANOVA, a generalization of a paired analysis, in which comparisons between pull-downs at different time points are made within each run.
-
-Limma employs an empirical Bayes method to moderate the standard errors of the estimated log~2~ fold changes; this results in more stable inference and improved power, especially for experiments with small numbers of replicates.
-
-Multiple testing correction of p-values was applied using the Benjamini-Hochberg method to control the false discovery rate (FDR).
-The adjusted p-value (also known as a q-value) can be understood as follows.
-If all proteins with q-value below a threshold of `r significanceLevel` are selected as differentially expressed, then the expected proportion of false discoveries in the selected group is controlled to be less than the threshold value, in this case 5%.
-
-The B-statistic (lods or B) is the log odds that the protein is differentially expressed.
-For example, if B = 1.5, the odds of differential expression is $e^{1.5} = 4.48$, i.e, about four and a half to one.
-The probability that the protein is differentially expressed is $4.48 / (1 + 4.48) = 0.82$.
-A B-statistic of zero corresponds to a 50-50 chance that the gene is differentially expressed.
-The B-statistic has been automatically adjusted for multiple testing by assuming that 1% of the proteins are expected to be differentially expressed.
-
-```{r fit_linear_model, include = FALSE}
-
-factors <- c("Group", "Run")
-
-model <- as.formula(paste(c("~ 0", factors), collapse = " + "))
-
-fittedLinearModels <- list()
-fittedLinearModelsExcludingControls <- list()
-
-for (analysis in names(combinedProteinData))
+for (contrast in contrastsTable$Contrast[contrastsTable$Report=="Y"])
 {
-  cat(analysis, "\n", sep = "")
-  data <- combinedProteinData[[analysis]]
-
-  # log2 transform
-  exprs(data) <- log2(exprs(data))
-
-  design <- model.matrix(model, data = pData(data))
-
-  colnames(design) <- colnames(design) %>%
-    sub(pattern = "^Group", replacement = "") %>%
-    gsub(pattern = " ", replacement = "_")
-
-  fittedLinearModels[[analysis]] <- lmFit(data, design = design)
-
-  samples <- pData(data) %>%
-    rownames_to_column("Sample") %>%
-    filter(Group != controlGroup) %>%
-    select(Sample) %>%
-    unlist(use.names = FALSE)
-
-  data <- data[, sampleNames(data) %in% samples]
-
-  design <- model.matrix(model, data = pData(data))
-
-  colnames(design) <- colnames(design) %>%
-    sub(pattern = "^Group", replacement = "") %>%
-    gsub(pattern = " ", replacement = "_")
-
-  fittedLinearModelsExcludingControls[[analysis]] <- lmFit(data, design = design)
-}
-```
-
-```{r contrasts, include = FALSE}
-
-contrasts <- c(
-  "ER 45min vs ER 0min",
-  "ER 90min vs ER 0min",
-  "ER 90min vs ER 45min",
-  "FOXA1 45min vs FOXA1 0min",
-  "FOXA1 90min vs FOXA1 0min",
-  "FOXA1 90min vs FOXA1 45min"
-)
-
-reformatGroup <- function(x) gsub(x, pattern = " ", replacement = "_")
-
-reformatContrasts <- function(x)
-{
-  x %>%
-    gsub(pattern = " ", replacement = "_") %>%
-    sub(pattern = "_-_", replacement = " - ") %>%
-    sub(pattern = "_vs_", replacement = " - ")
+  render('05.QRIMESceneInvestigatonReport.Rmd', output_file = paste('Reports/05.QRIMEReport_', contrast, '.pdf', sep = ''))
 }
 
-reformattedContrasts <- reformatContrasts(contrasts)
-
-fittedContrasts <- list()
-
-for (analysis in names(combinedProteinData))
-{
-  cat(analysis, "\n", sep = "")
-  fittedLinearModel <- fittedLinearModels[[analysis]]
-  # fittedLinearModel <- fittedLinearModelsExcludingControls[[analysis]]
-  contrastMatrix <- makeContrasts(contrasts = reformattedContrasts, levels = fittedLinearModel$design)
-  fittedContrasts[[analysis]] <- contrasts.fit(fittedLinearModel, contrastMatrix)
-  fittedContrasts[[analysis]] <- eBayes(fittedContrasts[[analysis]], trend = TRUE, robust = TRUE)
-}
-```
-
-```{r create_contrast_results, include = FALSE}
-
-createContrastResults <- function(
-  msnset,
-  fittedLinearModel,
-  fittedLinearModelExcludingControls,
-  fittedContrasts,
-  runs,
-  contrastGroups,
-  controlGroup,
-  proteinInfo)
-{
-  reformattedContrast <- reformatContrasts(paste(contrastGroups, collapse = " - "))
-  reformattedContrastGroups <- reformattedContrast %>% strsplit(" - ") %>% unlist
-  reformatedControlGroup <- reformatGroup(controlGroup)
-
-  results <- topTable(fittedContrasts, coef = reformattedContrast, number = Inf, sort.by = "none")
-
-  fittedIntensities <- as.data.frame(fittedLinearModel$coefficients)
-  # fittedIntensities <- as.data.frame(fittedLinearModelExcludingControls$coefficients)
-  contrastFittedIntensities <- select(fittedIntensities, one_of(reformattedContrastGroups))
-  results$logFC <- contrastFittedIntensities[, 1] - contrastFittedIntensities[, 2]
-
-  fittedIntensities <- as.data.frame(fittedLinearModel$coefficients)
-  contrastFittedIntensities <- select(fittedIntensities, one_of(reformattedContrastGroups))
-  controlFittedIntensities <- fittedIntensities[, reformatedControlGroup]
-  results$logFCcontrol = apply(contrastFittedIntensities - controlFittedIntensities, 1, max)
-
-  samples <- pData(msnset) %>%
-    mutate(Sample = rownames(.)) %>%
-    filter(Group %in% c(contrastGroups, controlGroup)) %>%
-    select(Sample) %>%
-    unlist(use.names = FALSE)
-
-  intensities <- exprs(msnset) %>%
-    as.data.frame %>%
-    select(one_of(samples))
-
-  # add N column for the number of replicates, i.e. runs where intensity
-  # is not NA for both groups
-  n <- rep(0, nrow(intensities))
-  for (run in runs)
-  {
-    samples <- pData(msnset) %>%
-      mutate(Sample = rownames(.)) %>%
-      filter(Group %in% contrastGroups) %>%
-      filter(Run == run) %>%
-      select(Sample) %>%
-      unlist(use.names = FALSE)
-    n <- n +
-      intensities %>%
-      select(one_of(samples)) %>%
-      transmute(n = ifelse(complete.cases(.), 1, 0)) %>%
-      unlist(use.names = FALSE)
-  }
-
-  results <- results %>%
-    mutate(N = n) %>%
-    bind_cols(fittedIntensities %>% select(one_of(rev(reformattedContrastGroups), controlGroup))) %>%
-    bind_cols(intensities) %>%
-    left_join(proteinInfo, by = "Protein")
-
-  results <- bind_cols(
-    results %>% select(Protein, Gene, Description, N),
-    results %>% select(-Protein, -Gene, -Description, -N)
-  )
-
-  results 
-}
-
-```
-
-The IgG control was used to assess non-specific binding.
-The binding of a protein is considered to be non-specific if the log~2~ fold change relative to the IgG control is less than `r controlLogFoldChangeThreshold`.
-In the differential binding analysis, protein intensity values are fitted to a linear model and the fitted values for each condition being compared are also compared to the IgG control.
-For each comparison of two groups, the maximum log~2~ fold change from each of the two groups above the IgG control is used to determine if the binding is specific.
-
-\newpage
-
-```{r contrast_1, include = FALSE}
-contrastGroups <- c("ER 45min", "ER 0min")
-contrastName <- paste(contrastGroups, collapse = " vs ")
-processingDescription <- "excluding peptides with missing intensities, no normalization"
-analysis <- "ExcludeMissingValues_NoNormalization"
-```
-
-## `r paste(contrastName, processingDescription, sep = ", ")`
-
-```{r results_1, include = FALSE}
-
-data <- combinedProteinData[[analysis]]
-exprs(data) <- log2(exprs(data))
-
-results <- createContrastResults(
-  data,
-  fittedLinearModels[[analysis]],
-  fittedLinearModelsExcludingControls[[analysis]],
-  fittedContrasts[[analysis]],
-  runs,
-  contrastGroups,
-  controlGroup,
-  proteinInfo)
-```
-
-```{r ma_plot_1, echo=FALSE, warning=FALSE, message=FALSE, fig.cap=paste("MA plot of the average intensity against log~2~ fold change for the ", contrastName, " comparison (", processingDescription, ").
-Top ranking differentially-expressed proteins with false discovery rate below 0.05 are highlighted in pink.
-Open circles indicate that the protein is non-specific from the IgG control comparison.", sep = ""), fig.height=6}
-print(maPlot(results,
-             significanceLevel = significanceLevel,
-             controlLogFoldChangeThreshold = controlLogFoldChangeThreshold,
-             minLogFoldChangeForLabelling = 1.5,
-             maxNumberLabelledProteins = 20,
-             pointSize = 1.5))
-```
-
-```{r volcano_plot_1, echo=FALSE, warning=FALSE, message=FALSE, fig.cap=paste("Volcano plot of the average intensity against log~2~ fold change for the ", contrastName, " comparison (", processingDescription, ").
-Top ranking differentially-expressed proteins with false discovery rate below 0.05 are highlighted in pink.
-Open circles indicate that the protein is non-specific from the IgG control comparison.", sep = ""), fig.height=6}
-print(volcanoPlot(results,
-                  significanceLevel = significanceLevel,
-                  minLogFoldChange = -2.0,
-                  maxLogFoldChange = 2.75,
-                  controlLogFoldChangeThreshold = controlLogFoldChangeThreshold,
-                  minLogFoldChangeForLabelling = 1.5,
-                  maxNumberLabelledProteins = 20,
-                  pointSize = 1.5))
-```
-
-```{r pvalue_histogram_1, echo=FALSE, warning=FALSE, message=FALSE, fig.cap=paste("Histogram of p-values for the ", contrastName, " comparison (", processingDescription, ")", sep = ""), fig.height=3.5, fig.width=4.5}
-print(histogram(results, "P.Value", xlab = "p-value"))
-```
-
-```{r qq_plot_1, echo=FALSE, warning=FALSE, message=FALSE, fig.cap=paste("QQ plot of the adjusted p-values for the ", contrastName, " comparison (", processingDescription, ")", sep = ""), fig.height=5, fig.width=5}
-q <- results %>%
-  select(adj.P.Val) %>%
-  filter(!is.na(adj.P.Val)) %>%
-  arrange(adj.P.Val) %>%
-  unlist %>%
-  as.numeric
-qqplot(-log10(ppoints(length(q))),
-       -log10(q),
-       xlab=expression(Expected~~-log[10](italic(q))),
-       ylab=expression(Observed~~-log[10](italic(q))))
-abline(0, 1, col = "red")
-```
-
-\newpage
-
-```{r contrast_2, include = FALSE}
-contrastGroups <- c("ER 45min", "ER 0min")
-contrastName <- paste(contrastGroups, collapse = " vs ")
-processingDescription <- "excluding peptides with missing intensities, quantile normalization"
-analysis <- "ExcludeMissingValues_QuantileNormalization"
-```
-
-## `r paste(contrastName, processingDescription, sep = ", ")`
-
-```{r results_2, include = FALSE}
-
-data <- combinedProteinData[[analysis]]
-exprs(data) <- log2(exprs(data))
-
-results <- createContrastResults(
-  data,
-  fittedLinearModels[[analysis]],
-  fittedLinearModelsExcludingControls[[analysis]],
-  fittedContrasts[[analysis]],
-  runs,
-  contrastGroups,
-  controlGroup,
-  proteinInfo)
-```
-
-```{r ma_plot_2, echo=FALSE, warning=FALSE, message=FALSE, fig.cap=paste("MA plot of the average intensity against log~2~ fold change for the ", contrastName, " comparison (", processingDescription, ").
-Top ranking differentially-expressed proteins with false discovery rate below 0.05 are highlighted in pink.
-Open circles indicate that the protein is non-specific from the IgG control comparison.", sep = ""), fig.height=6}
-print(maPlot(results,
-             significanceLevel = significanceLevel,
-             controlLogFoldChangeThreshold = controlLogFoldChangeThreshold,
-             minLogFoldChangeForLabelling = 1.5,
-             maxNumberLabelledProteins = 50,
-             pointSize = 1.5))
-```
-
-```{r volcano_plot_2, echo=FALSE, warning=FALSE, message=FALSE, fig.cap=paste("Volcano plot of the average intensity against log~2~ fold change for the ", contrastName, " comparison (", processingDescription, ").
-Top ranking differentially-expressed proteins with false discovery rate below 0.05 are highlighted in pink.
-Open circles indicate that the protein is non-specific from the IgG control comparison.", sep = ""), fig.height=6}
-print(volcanoPlot(results,
-                  significanceLevel = significanceLevel,
-                  minLogFoldChange = -2.0,
-                  maxLogFoldChange = 2.75,
-                  controlLogFoldChangeThreshold = controlLogFoldChangeThreshold,
-                  minLogFoldChangeForLabelling = 1.5,
-                  maxNumberLabelledProteins = 50,
-                  pointSize = 1.5))
-```
-
-```{r pvalue_histogram_2, echo=FALSE, warning=FALSE, message=FALSE, fig.cap=paste("Histogram of p-values for the ", contrastName, " comparison (", processingDescription, ")", sep = ""), fig.height=3.5, fig.width=4.5}
-print(histogram(results, "P.Value", xlab = "p-value"))
-```
-
-```{r qq_plot_2, echo=FALSE, warning=FALSE, message=FALSE, fig.cap=paste("QQ plot of the adjusted p-values for the ", contrastName, " comparison (", processingDescription, ")", sep = ""), fig.height=5, fig.width=5}
-q <- results %>%
-  select(adj.P.Val) %>%
-  filter(!is.na(adj.P.Val)) %>%
-  arrange(adj.P.Val) %>%
-  unlist %>%
-  as.numeric
-qqplot(-log10(ppoints(length(q))),
-       -log10(q),
-       xlab=expression(Expected~~-log[10](italic(q))),
-       ylab=expression(Observed~~-log[10](italic(q))))
-abline(0, 1, col = "red")
-```
-
-\newpage
-
-```{r contrast_3, include = FALSE}
-contrastGroups <- c("ER 45min", "ER 0min")
-contrastName <- paste(contrastGroups, collapse = " vs ")
-processingDescription <- "excluding peptides with missing intensities, quantile normalization excluding IgG control"
-analysis <- "ExcludeMissingValues_QuantileNormalizationExcludingIgG"
-```
-
-## `r paste(contrastName, processingDescription, sep = ", ")`
-
-```{r results_3, include = FALSE}
-
-data <- combinedProteinData[[analysis]]
-exprs(data) <- log2(exprs(data))
-
-results <- createContrastResults(
-  data,
-  fittedLinearModels[[analysis]],
-  fittedLinearModelsExcludingControls[[analysis]],
-  fittedContrasts[[analysis]],
-  runs,
-  contrastGroups,
-  controlGroup,
-  proteinInfo)
-```
-
-```{r ma_plot_3, echo=FALSE, warning=FALSE, message=FALSE, fig.cap=paste("MA plot of the average intensity against log~2~ fold change for the ", contrastName, " comparison (", processingDescription, ").
-Top ranking differentially-expressed proteins with false discovery rate below 0.05 are highlighted in pink.
-Open circles indicate that the protein is non-specific from the IgG control comparison.", sep = ""), fig.height=6}
-print(maPlot(results,
-             significanceLevel = significanceLevel,
-             controlLogFoldChangeThreshold = controlLogFoldChangeThreshold,
-             minLogFoldChangeForLabelling = 1.5,
-             maxNumberLabelledProteins = 50,
-             pointSize = 1.5))
-```
-
-```{r volcano_plot_3, echo=FALSE, warning=FALSE, message=FALSE, fig.cap=paste("Volcano plot of the average intensity against log~2~ fold change for the ", contrastName, " comparison (", processingDescription, ").
-Top ranking differentially-expressed proteins with false discovery rate below 0.05 are highlighted in pink.
-Open circles indicate that the protein is non-specific from the IgG control comparison.", sep = ""), fig.height=6}
-print(volcanoPlot(results,
-                  significanceLevel = significanceLevel,
-                  minLogFoldChange = -2.0,
-                  maxLogFoldChange = 2.75,
-                  controlLogFoldChangeThreshold = controlLogFoldChangeThreshold,
-                  minLogFoldChangeForLabelling = 1.5,
-                  maxNumberLabelledProteins = 50,
-                  pointSize = 1.5))
-```
-
-```{r pvalue_histogram_3, echo=FALSE, warning=FALSE, message=FALSE, fig.cap=paste("Histogram of p-values for the ", contrastName, " comparison (", processingDescription, ")", sep = ""), fig.height=3.5, fig.width=4.5}
-print(histogram(results, "P.Value", xlab = "p-value"))
-```
-
-```{r qq_plot_3, echo=FALSE, warning=FALSE, message=FALSE, fig.cap=paste("QQ plot of the adjusted p-values for the ", contrastName, " comparison (", processingDescription, ")", sep = ""), fig.height=5, fig.width=5}
-q <- results %>%
-  select(adj.P.Val) %>%
-  filter(!is.na(adj.P.Val)) %>%
-  arrange(adj.P.Val) %>%
-  unlist %>%
-  as.numeric
-qqplot(-log10(ppoints(length(q))),
-       -log10(q),
-       xlab=expression(Expected~~-log[10](italic(q))),
-       ylab=expression(Observed~~-log[10](italic(q))))
-abline(0, 1, col = "red")
-```
-
-\newpage
-
-```{r contrast_4, include = FALSE}
-contrastGroups <- c("ER 45min", "ER 0min")
-contrastName <- paste(contrastGroups, collapse = " vs ")
-processingDescription <- "excluding peptides with missing intensities, scale normalization"
-analysis <- "ExcludeMissingValues_ScaleNormalization"
-```
-
-## `r paste(contrastName, processingDescription, sep = ", ")`
-
-```{r results_4, include = FALSE}
-
-data <- combinedProteinData[[analysis]]
-exprs(data) <- log2(exprs(data))
-
-results <- createContrastResults(
-  data,
-  fittedLinearModels[[analysis]],
-  fittedLinearModelsExcludingControls[[analysis]],
-  fittedContrasts[[analysis]],
-  runs,
-  contrastGroups,
-  controlGroup,
-  proteinInfo)
-```
-
-```{r ma_plot_4, echo=FALSE, warning=FALSE, message=FALSE, fig.cap=paste("MA plot of the average intensity against log~2~ fold change for the ", contrastName, " comparison (", processingDescription, ").
-Top ranking differentially-expressed proteins with false discovery rate below 0.05 are highlighted in pink.
-Open circles indicate that the protein is non-specific from the IgG control comparison.", sep = ""), fig.height=6}
-print(maPlot(results,
-             significanceLevel = significanceLevel,
-             controlLogFoldChangeThreshold = controlLogFoldChangeThreshold,
-             minLogFoldChangeForLabelling = 1.5,
-             maxNumberLabelledProteins = 50,
-             pointSize = 1.5))
-```
-
-```{r volcano_plot_4, echo=FALSE, warning=FALSE, message=FALSE, fig.cap=paste("Volcano plot of the average intensity against log~2~ fold change for the ", contrastName, " comparison (", processingDescription, ").
-Top ranking differentially-expressed proteins with false discovery rate below 0.05 are highlighted in pink.
-Open circles indicate that the protein is non-specific from the IgG control comparison.", sep = ""), fig.height=6}
-print(volcanoPlot(results,
-                  significanceLevel = significanceLevel,
-                  minLogFoldChange = -2.0,
-                  maxLogFoldChange = 2.75,
-                  controlLogFoldChangeThreshold = controlLogFoldChangeThreshold,
-                  minLogFoldChangeForLabelling = 1.5,
-                  maxNumberLabelledProteins = 50,
-                  pointSize = 1.5))
-```
-
-```{r pvalue_histogram_4, echo=FALSE, warning=FALSE, message=FALSE, fig.cap=paste("Histogram of p-values for the ", contrastName, " comparison (", processingDescription, ")", sep = ""), fig.height=3.5, fig.width=4.5}
-print(histogram(results, "P.Value", xlab = "p-value"))
-```
-
-```{r qq_plot_4, echo=FALSE, warning=FALSE, message=FALSE, fig.cap=paste("QQ plot of the adjusted p-values for the ", contrastName, " comparison (", processingDescription, ")", sep = ""), fig.height=5, fig.width=5}
-q <- results %>%
-  select(adj.P.Val) %>%
-  filter(!is.na(adj.P.Val)) %>%
-  arrange(adj.P.Val) %>%
-  unlist %>%
-  as.numeric
-qqplot(-log10(ppoints(length(q))),
-       -log10(q),
-       xlab=expression(Expected~~-log[10](italic(q))),
-       ylab=expression(Observed~~-log[10](italic(q))))
-abline(0, 1, col = "red")
-```
-
-\newpage
-
-```{r contrast_5, include = FALSE}
-contrastGroups <- c("ER 45min", "ER 0min")
-contrastName <- paste(contrastGroups, collapse = " vs ")
-processingDescription <- paste("excluding peptides with missing intensities, scale normalization using top", numberOfPeptidesWithHighestIgG, "IgG peptides")
-analysis <- "ExcludeMissingValues_ScaleNormalizationTopIgG"
-```
-
-## `r paste(contrastName, processingDescription, sep = ", ")`
-
-```{r results_5, include = FALSE}
-
-data <- combinedProteinData[[analysis]]
-exprs(data) <- log2(exprs(data))
-
-results <- createContrastResults(
-  data,
-  fittedLinearModels[[analysis]],
-  fittedLinearModelsExcludingControls[[analysis]],
-  fittedContrasts[[analysis]],
-  runs,
-  contrastGroups,
-  controlGroup,
-  proteinInfo)
-```
-
-```{r ma_plot_5, echo=FALSE, warning=FALSE, message=FALSE, fig.cap=paste("MA plot of the average intensity against log~2~ fold change for the ", contrastName, " comparison (", processingDescription, ").
-Top ranking differentially-expressed proteins with false discovery rate below 0.05 are highlighted in pink.
-Open circles indicate that the protein is non-specific from the IgG control comparison.", sep = ""), fig.height=6}
-print(maPlot(results,
-             significanceLevel = significanceLevel,
-             controlLogFoldChangeThreshold = controlLogFoldChangeThreshold,
-             minLogFoldChangeForLabelling = 1.5,
-             maxNumberLabelledProteins = 50,
-             pointSize = 1.5))
-```
-
-```{r volcano_plot_5, echo=FALSE, warning=FALSE, message=FALSE, fig.cap=paste("Volcano plot of the average intensity against log~2~ fold change for the ", contrastName, " comparison (", processingDescription, ").
-Top ranking differentially-expressed proteins with false discovery rate below 0.05 are highlighted in pink.
-Open circles indicate that the protein is non-specific from the IgG control comparison.", sep = ""), fig.height=6}
-print(volcanoPlot(results,
-                  significanceLevel = significanceLevel,
-                  minLogFoldChange = -2.0,
-                  maxLogFoldChange = 2.75,
-                  controlLogFoldChangeThreshold = controlLogFoldChangeThreshold,
-                  minLogFoldChangeForLabelling = 1.5,
-                  maxNumberLabelledProteins = 50,
-                  pointSize = 1.5))
-```
-
-```{r pvalue_histogram_5, echo=FALSE, warning=FALSE, message=FALSE, fig.cap=paste("Histogram of p-values for the ", contrastName, " comparison (", processingDescription, ")", sep = ""), fig.height=3.5, fig.width=4.5}
-print(histogram(results, "P.Value", xlab = "p-value"))
-```
-
-```{r qq_plot_5, echo=FALSE, warning=FALSE, message=FALSE, fig.cap=paste("QQ plot of the adjusted p-values for the ", contrastName, " comparison (", processingDescription, ")", sep = ""), fig.height=5, fig.width=5}
-q <- results %>%
-  select(adj.P.Val) %>%
-  filter(!is.na(adj.P.Val)) %>%
-  arrange(adj.P.Val) %>%
-  unlist %>%
-  as.numeric
-qqplot(-log10(ppoints(length(q))),
-       -log10(q),
-       xlab=expression(Expected~~-log[10](italic(q))),
-       ylab=expression(Observed~~-log[10](italic(q))))
-abline(0, 1, col = "red")
-```
-
-\newpage
-
-```{r contrast_6, include = FALSE}
-contrastGroups <- c("ER 45min", "ER 0min")
-contrastName <- paste(contrastGroups, collapse = " vs ")
-processingDescription <- "KNN nearest neighbour averaging imputation of missing values, quantile normalization"
-analysis <- "ImputeKNN_QuantileNormalization"
-```
-
-## `r paste(contrastName, processingDescription, sep = ", ")`
-
-```{r results_6, include = FALSE}
-
-data <- combinedProteinData[[analysis]]
-exprs(data) <- log2(exprs(data))
-
-results <- createContrastResults(
-  data,
-  fittedLinearModels[[analysis]],
-  fittedLinearModelsExcludingControls[[analysis]],
-  fittedContrasts[[analysis]],
-  runs,
-  contrastGroups,
-  controlGroup,
-  proteinInfo)
-```
-
-```{r ma_plot_6, echo=FALSE, warning=FALSE, message=FALSE, fig.cap=paste("MA plot of the average intensity against log~2~ fold change for the ", contrastName, " comparison (", processingDescription, ").
-Top ranking differentially-expressed proteins with false discovery rate below 0.05 are highlighted in pink.
-Open circles indicate that the protein is non-specific from the IgG control comparison.", sep = ""), fig.height=6}
-print(maPlot(results,
-             significanceLevel = significanceLevel,
-             controlLogFoldChangeThreshold = controlLogFoldChangeThreshold,
-             minLogFoldChangeForLabelling = 1.5,
-             maxNumberLabelledProteins = 50,
-             pointSize = 1.5))
-```
-
-```{r volcano_plot_6, echo=FALSE, warning=FALSE, message=FALSE, fig.cap=paste("Volcano plot of the average intensity against log~2~ fold change for the ", contrastName, " comparison (", processingDescription, ").
-Top ranking differentially-expressed proteins with false discovery rate below 0.05 are highlighted in pink.
-Open circles indicate that the protein is non-specific from the IgG control comparison.", sep = ""), fig.height=6}
-print(volcanoPlot(results,
-                  significanceLevel = significanceLevel,
-                  minLogFoldChange = -2.0,
-                  maxLogFoldChange = 2.75,
-                  controlLogFoldChangeThreshold = controlLogFoldChangeThreshold,
-                  minLogFoldChangeForLabelling = 1.5,
-                  maxNumberLabelledProteins = 50,
-                  pointSize = 1.5))
-```
-
-```{r pvalue_histogram_6, echo=FALSE, warning=FALSE, message=FALSE, fig.cap=paste("Histogram of p-values for the ", contrastName, " comparison (", processingDescription, ")", sep = ""), fig.height=3.5, fig.width=4.5}
-print(histogram(results, "P.Value", xlab = "p-value"))
-```
-
-```{r qq_plot_6, echo=FALSE, warning=FALSE, message=FALSE, fig.cap=paste("QQ plot of the adjusted p-values for the ", contrastName, " comparison (", processingDescription, ")", sep = ""), fig.height=5, fig.width=5}
-q <- results %>%
-  select(adj.P.Val) %>%
-  filter(!is.na(adj.P.Val)) %>%
-  arrange(adj.P.Val) %>%
-  unlist %>%
-  as.numeric
-qqplot(-log10(ppoints(length(q))),
-       -log10(q),
-       xlab=expression(Expected~~-log[10](italic(q))),
-       ylab=expression(Observed~~-log[10](italic(q))))
-abline(0, 1, col = "red")
-```
-
-\newpage
-
-# Comparison of differential binding analysis approaches
-
-In this section, comparisons of the differential binding results obtained using differing methods for handling missing values and varying normalization techniques are presented.
-The B-statistic (log odds) is used to rank the proteins from most significantly differentially expressed to least significant.
-Scatter plots of both the B-statistic generated for the two methods being compared and the rank are presented.
-The protein with the highest B-statistic, i.e.
-the most significantly differentially expressed protein, is given a rank of 1.
-
-Differing imputation methods have relatively little effect on the results for most proteins.
-This is largely because only a small proportion of the proteins have peptide-level measurements containing missing values.
-There are a few proteins that are only represented in the differential binding results when imputation is carried out, for which all peptide-level measurements contain missing values.
-These proteins are not represented in the comparions with the analysis in which measurements with missing values are removed.
-
-Different normalization approaches have a more substantial impact on the differential binding analysis.
-
-\newpage
-
-## Removal of missing values vs KNN imputation (quantile normalization in both)
-
-```{r exclude_missing_vs_knn_imputation, include = FALSE}
-
-contrastGroups <- c("ER 45min", "ER 0min")
-contrastName <- paste(contrastGroups, collapse = " vs ")
-
-analysis <- "ExcludeMissingValues_QuantileNormalization"
-data <- combinedProteinData[[analysis]]
-exprs(data) <- log2(exprs(data))
-results1 <- createContrastResults(
-  data,
-  fittedLinearModels[[analysis]],
-  fittedLinearModelsExcludingControls[[analysis]],
-  fittedContrasts[[analysis]],
-  runs,
-  contrastGroups,
-  controlGroup,
-  proteinInfo)
-
-analysis <- "ImputeKNN_QuantileNormalization"
-data <- combinedProteinData[[analysis]]
-exprs(data) <- log2(exprs(data))
-results2 <- createContrastResults(
-  data,
-  fittedLinearModels[[analysis]],
-  fittedLinearModelsExcludingControls[[analysis]],
-  fittedContrasts[[analysis]],
-  runs,
-  contrastGroups,
-  controlGroup,
-  proteinInfo)
-
-results <- inner_join(
-  results1 %>% select(Protein, B1 = B),
-  results2 %>% select(Protein, B2 = B),
-  by = "Protein"
-  ) %>%
-  filter(complete.cases(.)) %>%
-  mutate(rank1 = rank(-B1), rank2 = rank(-B2))
-```
-
-```{r exclude_missing_vs_knn_imputation_B_plot, echo=FALSE, warning=FALSE, message=FALSE, fig.cap=paste("Plot of B-statistic (log odds) for differential expression of proteins for the ", contrastName, " contrast, comparing results following exclusion of missing values and KNN imputation.", sep = ""), fig.height=2.9, fig.width=3.5}
-print(scatterPlot(results, "B1", "B2", "B-statistic (excluding missing values)", "B-statistic (KNN imputation)"))
-```
-
-```{r exclude_missing_vs_knn_imputation_rank_plot, echo=FALSE, warning=FALSE, message=FALSE, fig.cap=paste("Plot of the rank of differentially-expressed proteins for the ", contrastName, " contrast, comparing results following exclusion of missing values and KNN imputation.", sep = ""), fig.height=2.9, fig.width=3.5}
-print(scatterPlot(results, "rank1", "rank2", "Rank (excluding missing values)", "Rank (KNN imputation)"))
-```
-
-\newpage
-
-## Removal of missing values vs lowest value imputation (quantile normalization in both)
-
-```{r exclude_missing_vs_min_imputation, include = FALSE}
-
-contrastGroups <- c("ER 45min", "ER 0min")
-contrastName <- paste(contrastGroups, collapse = " vs ")
-
-analysis <- "ExcludeMissingValues_QuantileNormalization"
-data <- combinedProteinData[[analysis]]
-exprs(data) <- log2(exprs(data))
-results1 <- createContrastResults(
-  data,
-  fittedLinearModels[[analysis]],
-  fittedLinearModelsExcludingControls[[analysis]],
-  fittedContrasts[[analysis]],
-  runs,
-  contrastGroups,
-  controlGroup,
-  proteinInfo)
-
-analysis <- "ImputeMin_QuantileNormalization"
-data <- combinedProteinData[[analysis]]
-exprs(data) <- log2(exprs(data))
-results2 <- createContrastResults(
-  data,
-  fittedLinearModels[[analysis]],
-  fittedLinearModelsExcludingControls[[analysis]],
-  fittedContrasts[[analysis]],
-  runs,
-  contrastGroups,
-  controlGroup,
-  proteinInfo)
-
-results <- inner_join(
-  results1 %>% select(Protein, B1 = B),
-  results2 %>% select(Protein, B2 = B),
-  by = "Protein"
-  ) %>%
-  filter(complete.cases(.)) %>%
-  mutate(rank1 = rank(-B1), rank2 = rank(-B2))
-```
-
-```{r exclude_missing_vs_min_imputation_B_plot, echo=FALSE, warning=FALSE, message=FALSE, fig.cap=paste("Plot of B-statistic (log odds) for differential expression of proteins for the ", contrastName, " contrast, comparing results following exclusion of missing values and lowest value imputation.", sep = ""), fig.height=2.9, fig.width=3.5}
-print(scatterPlot(results, "B1", "B2", "B-statistic (excluding missing values)", "B-statistic (lowest value imputation)"))
-```
-
-```{r exclude_missing_vs_min_imputation_rank_plot, echo=FALSE, warning=FALSE, message=FALSE, fig.cap=paste("Plot of the rank of differentially-expressed proteins for the ", contrastName, " contrast, comparing results following exclusion of missing values and lowest value imputation.", sep = ""), fig.height=2.9, fig.width=3.5}
-print(scatterPlot(results, "rank1", "rank2", "Rank (excluding missing values)", "Rank (lowest value imputation)"))
-```
-
-\newpage
-
-## Quantile vs Scale normalization (missing values excluded in both)
-
-```{r quantile_vs_scale_normalization, include = FALSE}
-
-contrastGroups <- c("ER 45min", "ER 0min")
-contrastName <- paste(contrastGroups, collapse = " vs ")
-
-analysis <- "ExcludeMissingValues_QuantileNormalization"
-data <- combinedProteinData[[analysis]]
-exprs(data) <- log2(exprs(data))
-results1 <- createContrastResults(
-  data,
-  fittedLinearModels[[analysis]],
-  fittedLinearModelsExcludingControls[[analysis]],
-  fittedContrasts[[analysis]],
-  runs,
-  contrastGroups,
-  controlGroup,
-  proteinInfo)
-
-analysis <- "ExcludeMissingValues_ScaleNormalization"
-data <- combinedProteinData[[analysis]]
-exprs(data) <- log2(exprs(data))
-results2 <- createContrastResults(
-  data,
-  fittedLinearModels[[analysis]],
-  fittedLinearModelsExcludingControls[[analysis]],
-  fittedContrasts[[analysis]],
-  runs,
-  contrastGroups,
-  controlGroup,
-  proteinInfo)
-
-results <- inner_join(
-  results1 %>% select(Protein, B1 = B),
-  results2 %>% select(Protein, B2 = B),
-  by = "Protein"
-  ) %>%
-  filter(complete.cases(.)) %>%
-  mutate(rank1 = rank(-B1), rank2 = rank(-B2))
-```
-
-```{r scale_vs_quantile_normalization_B_plot, echo=FALSE, warning=FALSE, message=FALSE, fig.cap=paste("Plot of B-statistic (log odds) for differential expression of proteins for the ", contrastName, " contrast, comparing results following quantile and scale normalization.", sep = ""), fig.height=2.9, fig.width=3.5}
-print(scatterPlot(results, "B1", "B2", "B-statistic (quantile normalization)", "B-statistic (scale normalization)"))
-```
-
-```{r scale_vs_quantile_normalization_rank_plot, echo=FALSE, warning=FALSE, message=FALSE, fig.cap=paste("Plot of the rank of differentially-expressed proteins for the ", contrastName, " contrast, comparing results following quantile and scale normalization.", sep = ""), fig.height=2.9, fig.width=3.5}
-print(scatterPlot(results, "rank1", "rank2", "Rank (quantile normalization)", "Rank (scale normalization)"))
-```
-
-\newpage
-
-## Scale normalization based on all peptides vs peptides with highest IgG intensities (missing values excluded in both)
-
-```{r scale_normalization_all_vs_topIgG, include = FALSE}
-
-contrastGroups <- c("ER 45min", "ER 0min")
-contrastName <- paste(contrastGroups, collapse = " vs ")
-
-analysis <- "ExcludeMissingValues_ScaleNormalization"
-data <- combinedProteinData[[analysis]]
-exprs(data) <- log2(exprs(data))
-results1 <- createContrastResults(
-  data,
-  fittedLinearModels[[analysis]],
-  fittedLinearModelsExcludingControls[[analysis]],
-  fittedContrasts[[analysis]],
-  runs,
-  contrastGroups,
-  controlGroup,
-  proteinInfo)
-
-analysis <- "ExcludeMissingValues_ScaleNormalizationTopIgG"
-data <- combinedProteinData[[analysis]]
-exprs(data) <- log2(exprs(data))
-results2 <- createContrastResults(
-  data,
-  fittedLinearModels[[analysis]],
-  fittedLinearModelsExcludingControls[[analysis]],
-  fittedContrasts[[analysis]],
-  runs,
-  contrastGroups,
-  controlGroup,
-  proteinInfo)
-
-results <- inner_join(
-  results1 %>% select(Protein, B1 = B),
-  results2 %>% select(Protein, B2 = B),
-  by = "Protein"
-  ) %>%
-  filter(complete.cases(.)) %>%
-  mutate(rank1 = rank(-B1), rank2 = rank(-B2))
-```
-
-```{r scale_normalization_all_vs_topIgG_B_plot, echo=FALSE, warning=FALSE, message=FALSE, fig.cap=paste("Plot of B-statistic (log odds) for differential expression of proteins for the ", contrastName, " contrast, comparing results following scale normalization based on all peptides and those with the highest IgG intensities.", sep = ""), fig.height=2.5, fig.width=3.5}
-print(scatterPlot(results, "B1", "B2", "B-statistic (scale normalization based on all peptides)", "B-statistic (scale normalization, peptides with highest IgG)"))
-```
-
-```{r scale_normalization_all_vs_topIgG_rank_plot, echo=FALSE, warning=FALSE, message=FALSE, fig.cap=paste("Plot of the rank of differentially-expressed proteins for the ", contrastName, " contrast, comparing results following scale normalization based on all peptides and those with the highest IgG intensities.", sep = ""), fig.height=2.5, fig.width=3.5}
-print(scatterPlot(results, "rank1", "rank2", "Rank (scale normalization based on all peptides)", "Rank (scale normalization, peptides with highest IgG)"))
-```
-
-\newpage
-
-# Comparisoin of variance between specfic and non-specific binding
-
-```{r  echo=FALSE, warning=FALSE, message=FALSE}
-
-maPlot <- function(differentialExpressionResults, selectedGenes = NULL,
-                   xlab = expression(average~log[2](intensity)),
-                   ylab = expression(log[2](fold~change)),
-                   significanceLevel = 0.05,
-                   minLogFoldChangeForLabelling = 1.5,
-                   maxNumberLabelledProteins = 50,
-                   controlLogFoldChangeThreshold = -Inf,
-                   pointSize = 2.5) {
-  if (!"logFCcontrol" %in% colnames(differentialExpressionResults))
-    differentialExpressionResults$logFCcontrol <- Inf
-
-  differentialExpressionResults$group <-
-    ifelse(differentialExpressionResults$Gene %in% selectedGenes, 0,
-    ifelse(!is.na(differentialExpressionResults$adj.P.Val) & differentialExpressionResults$adj.P.Val <= significanceLevel & differentialExpressionResults$logFCcontrol >= controlLogFoldChangeThreshold, 1,
-    ifelse(!is.na(differentialExpressionResults$adj.P.Val) & differentialExpressionResults$adj.P.Val <= significanceLevel, 2,
-    ifelse(abs(differentialExpressionResults$logFC) >= minLogFoldChangeForLabelling & differentialExpressionResults$logFCcontrol >= controlLogFoldChangeThreshold, 3,
-    ifelse(abs(differentialExpressionResults$logFC) >= minLogFoldChangeForLabelling, 4,
-    ifelse(differentialExpressionResults$logFCcontrol >= controlLogFoldChangeThreshold, 5, 6))))))
-
-  # 0 - selected
-  # 1 - significant, specific
-  # 2 - significant, non-specific
-  # 3 - non-significant, large logfc, specific
-  # 4 - non-significant, large logfc, non-specific
-  # 5 - specific
-  # 6 - non-specific
-
-  differentialExpressionResults <- differentialExpressionResults %>% arrange(desc(group))
-
-  differentialExpressionResults$group <- factor(differentialExpressionResults$group, 0:6)
-
-  differentialExpressionResults <- differentialExpressionResults %>%
-    mutate(label = paste(" ", Gene)) %>%
-    mutate(logFCcontrolAlpha = pmin(logFCcontrol, 2.0))
-
-  labelSubset <- bind_rows(
-    differentialExpressionResults %>%
-      filter(group != 0 & adj.P.Val <= significanceLevel) %>%
-      arrange(desc(abs(logFC))) %>%
-      slice(1:(maxNumberLabelledProteins / 2)),
-    differentialExpressionResults %>%
-      filter(group != 0 & abs(logFC) >= minLogFoldChangeForLabelling) %>%
-      arrange(desc(abs(logFC))) %>%
-      slice(1:maxNumberLabelledProteins)
-  ) %>%
-    unique %>%
-    slice(1:maxNumberLabelledProteins)
-
-    if (maxNumberLabelledProteins == 0) labelSubset <- labelSubset %>% slice(0)
-
-  # plot <- ggplot(differentialExpressionResults, aes(x = AveExpr, y = logFC, colour = group, size = group, shape = group, alpha = logFCcontrolAlpha))
-  plot <- ggplot(differentialExpressionResults, aes(x = AveExpr, y = logFC, colour = group))+ coord_cartesian(ylim = c(-3.5,3.5))
-  plot <- plot + geom_hline(yintercept = 0, color="gray50", size = 0.5)
-  plot <- plot + geom_point(alpha=0.75)
-  plot <- plot + geom_text(data = labelSubset, aes(label = label), hjust = 0, size = 3.5)
-  plot <- plot + geom_point(data = subset(differentialExpressionResults, group == 0), alpha = 0.7)
-  plot <- plot + geom_text(data = subset(differentialExpressionResults, group == 0), aes(label = label), hjust = 0, size = 4.0, alpha = 1)
-  plot <- plot + scale_colour_manual(values = c("green", "red", "blue", "red", "blue", "red", "blue"), drop = FALSE)
-  # plot <- plot + scale_colour_manual(values = c("blue", "deeppink3", "deeppink3", "gray50", "gray50", "gray50", "gray50"), drop = FALSE)
-  # plot <- plot + scale_size_manual(values = c(1.2 * pointSize, pointSize, pointSize, 0.8 * pointSize, 0.8 * pointSize, 0.6 * pointSize, 0.6 * pointSize), drop = FALSE)
-  # plot <- plot + scale_shape_manual(values = c(16, 16, 1, 16, 1, 16, 1), drop = FALSE)
-  #plot <- plot + scale_alpha(range = c(0.2, 1.0))
-  plot <- plot + xlab(xlab)
-  plot <- plot + ylab(ylab)
-  plot <- plot + theme_bw()
-  plot <- plot + theme(
-    text = element_text(size = 14),
-    legend.position = "none"
-  )
-
-  save(differentialExpressionResults,file="Results/QRIME.RData") 
-  return(plot)
-}
-
-contrastGroups <- c("ER 45min", "ER 0min")
-contrastName <- paste(contrastGroups, collapse = " vs ")
-
-analysis <- "ExcludeMissingValues_QuantileNormalization"
-
-data <- combinedProteinData[[analysis]]
-exprs(data) <- log2(exprs(data))
-
-results <- createContrastResults(
-  data,
-  fittedLinearModels[[analysis]],
-  fittedLinearModelsExcludingControls[[analysis]],
-  fittedContrasts[[analysis]],
-  runs,
-  c("ER 45min", "ER 0min"),
-  controlGroup,
-  proteinInfo)
-
-```
-
-```{r MA_specificVsNonSpecfic, echo=FALSE, warning=FALSE, message=FALSE, fig.cap="MA plot showing the distribution of non-specific binding.
-Non-specific binding was identified as proteins that showed less than a log2 enrichment over the IgG channel at either 0 or 45 minutes.
-The majority of non-specific binding (blue) is found to show less change between time points whereas the specific interactions (red) shows a broader distribution.", fig.height=2.5, fig.width=3.5}
-
-print(maPlot(results,
-             significanceLevel = 0.5,
-             controlLogFoldChangeThreshold = 1.0,
-             minLogFoldChangeForLabelling = 1.5,
-             maxNumberLabelledProteins = 0,
-             pointSize = 1.5))
-```
-
-```{r  echo=FALSE, warning=FALSE, message=FALSE}
-
-load("Results/QRIME.RData")
-as.numeric(differentialExpressionResults$group)%%2->differentialExpressionResults$group
-
-tmp<-differentialExpressionResults
-
-```
-
-```{r density_specificVsNonSpecfici, echo=FALSE, warning=FALSE, message=FALSE, fig.cap="Density plot right showing the distribution of non-specific binding.
-Non-specific binding was identified as proteins that showed less than a log2 enrichment over the IgG channel at either 0 or 45 minutes.
-The majority of non-specific binding (blue) is found to show less change between time points whereas the specific interactions (red) shows a broader distribution.", fig.height=3.5, fig.width=4.5}
-
-plot(density(tmp$logFC[tmp$group==1]),col="blue",ylab="Density", xlab="Fold Change", main="", xlim=c(3,-3), lwd=2)
-lines(density(tmp$logFC[tmp$group==0]),col="red",lwd=2)
-
-```
-
-```{r box_specificVsNonSpecfici, echo=FALSE, warning=FALSE, message=FALSE, fig.cap=" Box plot showing the variance of non-specific vs non-specific binding.
-Non-specific binding was identified as proteins that showed less than a log2 enrichment over the IgG channel at either 0 or 45 minutes.
-The the variance of the log fold changes for the non-specific binding interactions was significantly less than that of the specific interactions (F-test, one-sided, p-value < 2.2e*-16).", fig.height=3.5, fig.width=4.5}
-
-boxplot(tmp$logFC[tmp$group==1],tmp$logFC[tmp$group==0],ylab="Fold Change", names=c("Non-specific","Specific"), col=c("#0000FFAA","#FF0000AA"))
-
-```
-
-```{r  echo=FALSE, warning=FALSE, message=FALSE}
-
-var.test(tmp$logFC[tmp$group==1],tmp$logFC[tmp$group==0], alternative="less")
-
-```
-
-\newpage
-
-# Differential binding results tables
-
-```{r contrast_results_thresholds, include = FALSE}
-significanceLevel <- 0.05
-logFoldChangeThreshold <- 1.0
-controlLogFoldChangeThreshold <- 1.0
-```
-
-The following tables contain the top ranking differentially expressed proteins for each comparison.
-Included are all proteins that reach a statistical signficance of `r significanceLevel` in terms of the adjusted p-value and those with an absolute log~2~ fold change of `r logFoldChangeThreshold` or above.
-
-The IgG column gives the larger of the log~2~ fold changes for the two groups against the IgG control and an asterisk indicates specific binding where this log~2~ fold change is above a threshold of `r controlLogFoldChangeThreshold`.
-N is the number of replicates in which the protein was observed.
-
-In all cases, peptide intensities were quantile normalized and measurements with missing values were removed prior to summarization at the protein level.
-
-```{r contrast_results_table, include = FALSE}
-
-contrastResultsTable <- function(results, contrastName,
-                                 significanceLevel = 0.05,
-                                 logFoldChangeThreshold = 1.0,
-                                 controlLogFoldChangeThreshold = 1.0)
-{
-  resultsTable <- results %>%
-    select(Protein, Gene, N, logFC, AveExpr, adj.P.Val, B, logFCcontrol)
-
-  resultsTable$Specific <- ifelse(resultsTable$logFCcontrol >= controlLogFoldChangeThreshold, "*", "")
-
-  resultsTable <- resultsTable %>%
-    filter(adj.P.Val < 0.05 | abs(logFC) >= 1.0) %>%
-    arrange(desc(abs(logFC)))
-
-  colnames(resultsTable) <- c("Protein", "Gene", "N", "log2FC", "Avg Expr", "p-value", "B", "IgG", "")
-
-  print(
-    xtable(
-      resultsTable,
-      align = "lllrrrrrrl",
-      display = c("d", "s", "s", "d", "f", "f", "g", "f", "f", "s"),
-      caption = paste("Top ranking differentially expressed proteins from the", contrastName, "comparison,  sorted by log2 fold change.")
-    ),
-    include.rownames = FALSE,
-    tabular.environment = "longtable",
-    floating = FALSE,
-    add.to.row = addtorow
-  )
-}
-```
-
-\newpage
-
-```{r results_table_contrast_1_config, include = FALSE}
-contrastGroups <- c("ER 45min", "ER 0min")
-contrastName <- paste(contrastGroups, collapse = " vs ")
-processingDescription <- "excluding peptides with missing intensities, quantile normalization"
-analysis <- "ExcludeMissingValues_QuantileNormalization"
-```
-
-## `r contrastName`
-
-```{r results_table_contrast_1, echo=FALSE, warning=FALSE, message=FALSE, results='asis'}
-
-data <- combinedProteinData[[analysis]]
-exprs(data) <- log2(exprs(data))
-
-results <- createContrastResults(
-  data,
-  fittedLinearModels[[analysis]],
-  fittedLinearModelsExcludingControls[[analysis]],
-  fittedContrasts[[analysis]],
-  runs,
-  contrastGroups,
-  controlGroup,
-  proteinInfo)
-
-contrastResultsTable(results, contrastName,
-                     significanceLevel = 0.05,
-                     logFoldChangeThreshold = 1.0,
-                     controlLogFoldChangeThreshold = 1.0)
-
-resultsFile <- paste("Results/", paste(gsub(" ", "_", contrastGroups), collapse = "_vs_"), analysis, ".txt", sep = "")
-write.table(results, resultsFile, sep = "\t", row.names = FALSE, quote = FALSE)
-```
-
-\newpage
-
-```{r results_table_contrast_2_config, include = FALSE}
-contrastGroups <- c("ER 90min", "ER 0min")
-contrastName <- paste(contrastGroups, collapse = " vs ")
-processingDescription <- "excluding peptides with missing intensities, quantile normalization"
-analysis <- "ExcludeMissingValues_QuantileNormalization"
-```
-
-## `r contrastName`
-
-```{r results_table_contrast_2, echo=FALSE, warning=FALSE, message=FALSE, results='asis'}
-
-data <- combinedProteinData[[analysis]]
-exprs(data) <- log2(exprs(data))
-
-results <- createContrastResults(
-  data,
-  fittedLinearModels[[analysis]],
-  fittedLinearModelsExcludingControls[[analysis]],
-  fittedContrasts[[analysis]],
-  runs,
-  contrastGroups,
-  controlGroup,
-  proteinInfo)
-
-contrastResultsTable(results, contrastName,
-                     significanceLevel = 0.05,
-                     logFoldChangeThreshold = 1.0,
-                     controlLogFoldChangeThreshold = 1.0)
-
-resultsFile <- paste("Results/",paste(gsub(" ", "_", contrastGroups), collapse = "_vs_"), analysis, ".txt", sep = "")
-write.table(results, resultsFile, sep = "\t", row.names = FALSE, quote = FALSE)
-```
-
-\newpage
-
-```{r results_table_contrast_3_config, include = FALSE}
-contrastGroups <- c("ER 90min", "ER 45min")
-contrastName <- paste(contrastGroups, collapse = " vs ")
-processingDescription <- "excluding peptides with missing intensities, quantile normalization"
-analysis <- "ExcludeMissingValues_QuantileNormalization"
-```
-
-## `r contrastName`
-
-```{r results_table_contrast_3, echo=FALSE, warning=FALSE, message=FALSE, results='asis'}
-
-data <- combinedProteinData[[analysis]]
-exprs(data) <- log2(exprs(data))
-
-results <- createContrastResults(
-  data,
-  fittedLinearModels[[analysis]],
-  fittedLinearModelsExcludingControls[[analysis]],
-  fittedContrasts[[analysis]],
-  runs,
-  contrastGroups,
-  controlGroup,
-  proteinInfo)
-
-contrastResultsTable(results, contrastName,
-                     significanceLevel = 0.05,
-                     logFoldChangeThreshold = 1.0,
-                     controlLogFoldChangeThreshold = 1.0)
-
-resultsFile <- paste("Results/",paste(gsub(" ", "_", contrastGroups), collapse = "_vs_"), analysis, ".txt", sep = "")
-write.table(results, resultsFile, sep = "\t", row.names = FALSE, quote = FALSE)
-```
-
-\newpage
-
-```{r results_table_contrast_4_config, include = FALSE}
-contrastGroups <- c("FOXA1 45min", "FOXA1 0min")
-contrastName <- paste(contrastGroups, collapse = " vs ")
-processingDescription <- "excluding peptides with missing intensities, quantile normalization"
-analysis <- "ExcludeMissingValues_QuantileNormalization"
-```
-
-## `r contrastName`
-
-```{r results_table_contrast_4, echo=FALSE, warning=FALSE, message=FALSE, results='asis'}
-
-data <- combinedProteinData[[analysis]]
-exprs(data) <- log2(exprs(data))
-
-results <- createContrastResults(
-  data,
-  fittedLinearModels[[analysis]],
-  fittedLinearModelsExcludingControls[[analysis]],
-  fittedContrasts[[analysis]],
-  runs,
-  contrastGroups,
-  controlGroup,
-  proteinInfo)
-
-contrastResultsTable(results, contrastName,
-                     significanceLevel = 0.05,
-                     logFoldChangeThreshold = 1.0,
-                     controlLogFoldChangeThreshold = 1.0)
-
-resultsFile <- paste("Results/",paste(gsub(" ", "_", contrastGroups), collapse = "_vs_"), analysis, ".txt", sep = "")
-write.table(results, resultsFile, sep = "\t", row.names = FALSE, quote = FALSE)
-```
-
-\newpage
-
-```{r results_table_contrast_5_config, include = FALSE}
-contrastGroups <- c("FOXA1 90min", "FOXA1 0min")
-contrastName <- paste(contrastGroups, collapse = " vs ")
-processingDescription <- "excluding peptides with missing intensities, quantile normalization"
-analysis <- "ExcludeMissingValues_QuantileNormalization"
-```
-
-## `r contrastName`
-
-```{r results_table_contrast_5, echo=FALSE, warning=FALSE, message=FALSE, results='asis'}
-
-data <- combinedProteinData[[analysis]]
-exprs(data) <- log2(exprs(data))
-
-results <- createContrastResults(
-  data,
-  fittedLinearModels[[analysis]],
-  fittedLinearModelsExcludingControls[[analysis]],
-  fittedContrasts[[analysis]],
-  runs,
-  contrastGroups,
-  controlGroup,
-  proteinInfo)
-
-contrastResultsTable(results, contrastName,
-                     significanceLevel = 0.05,
-                     logFoldChangeThreshold = 1.0,
-                     controlLogFoldChangeThreshold = 1.0)
-
-resultsFile <- paste("Results/",paste(gsub(" ", "_", contrastGroups), collapse = "_vs_"), analysis, ".txt", sep = "")
-write.table(results, resultsFile, sep = "\t", row.names = FALSE, quote = FALSE)
-```
-
-\newpage
-
-```{r results_table_contrast_6_config, include = FALSE}
-contrastGroups <- c("FOXA1 90min", "FOXA1 45min")
-contrastName <- paste(contrastGroups, collapse = " vs ")
-processingDescription <- "excluding peptides with missing intensities, quantile normalization"
-analysis <- "ExcludeMissingValues_QuantileNormalization"
-```
-
-## `r contrastName`
-
-```{r results_table_contrast_6, echo=FALSE, warning=FALSE, message=FALSE, results='asis'}
-
-data <- combinedProteinData[[analysis]]
-exprs(data) <- log2(exprs(data))
-
-results <- createContrastResults(
-  data,
-  fittedLinearModels[[analysis]],
-  fittedLinearModelsExcludingControls[[analysis]],
-  fittedContrasts[[analysis]],
-  runs,
-  contrastGroups,
-  controlGroup,
-  proteinInfo)
-
-contrastResultsTable(results, contrastName,
-                     significanceLevel = 0.05,
-                     logFoldChangeThreshold = 1.0,
-                     controlLogFoldChangeThreshold = 1.0)
-
-resultsFile <- paste("Results/", paste(gsub(" ", "_", contrastGroups), collapse = "_vs_"), analysis, ".txt", sep = "")
-write.table(results, resultsFile, sep = "\t", row.names = FALSE, quote = FALSE)
 ```

### 2018-01-23 ChIP-seq_ER/qrime/05.QRIMESceneInvestigatonReport.Rmd
first commit

--- 2018-01-23 ChIP-seq_ER/qrime/02.Supplementary.Rmd
@@
 ---
-title: "Supplementary"
+title: "QRIME Scene Investigation - Report"
 author: "Andrew Holding & Matt Eldridge"
@@
 
-****************************
-
-```{r configuration, include = FALSE, cache = FALSE}
+```{r configuration_report, include = FALSE, cache = FALSE}
 knitr::opts_chunk$set(cache = FALSE)
 ```
+\newpage
 
-```{r load_libraries, include = FALSE, cache = FALSE}
-library(tidyr)
-library(dplyr)
-library(tibble)
-library(MSnbase)
-library(RColorBrewer)
-library(ggplot2)
-library(kfigr)
-library(gridExtra)
-library(xtable)
-library(limma)
-```
-
-```{r load_functions, include = FALSE, cache = FALSE}
-source("Functions/plots.R")
-source("Functions/normalization.R")
-```
-
-```{r xtable_options, include=FALSE, cache = FALSE}
-
-options(xtable.comment = FALSE)
-options(xtable.booktabs = TRUE)
-
-addtorow          <- list()
-addtorow$pos      <- list()
-addtorow$pos[[1]] <- c(0)
-addtorow$command  <- c(paste("\\hline \n",
-                             "\\endhead \n",
-                             "\\hline \n",
-                             "{\\footnotesize Continued on next page} \n",
-                             "\\endfoot \n",
-                             "\\endlastfoot \n",sep=""))
-```
-
-```{r read_protein_information, include = FALSE}
-proteinInfo <- read.delim("ReferenceData/protein_gene_description.txt", stringsAsFactors = FALSE)
-```
-
-```{r read_metadata, include = FALSE}
-metadata <- read.delim("SampleData/metadata.txt", check.names = FALSE)
-runs <- levels(metadata$Run)
-groups <- levels(metadata$Group)
-controlGroup <- "IgG"
-```
-
-```{r colours, include = FALSE}
-groupColours <- c(brewer.pal(9, "Blues")[5:7], brewer.pal(9, "RdPu")[5:7], brewer.pal(9, "Greens")[6])
-names(groupColours) <- groups
-sampleColours <- groupColours[metadata$Group]
-names(sampleColours) <- metadata$Sample
-```
-
-```{r read_intensities, include = FALSE}
-
-peptideData <- list()
-
-for (run in runs)
-{
-  cat(run, "\n", sep = "")
-
-  data <- read.delim(paste("ProcessedData/", run, ".txt", sep = ""), stringsAsFactors = FALSE, check.names = FALSE)
-
-  peptideCount <- data %>% select(Sequence, Modifications) %>% distinct %>% nrow
-  cat("Peptides:", peptideCount, "\n", sep = "")
-  if (peptideCount != nrow(data))
-    stop("Error: multiple rows for at least one peptide/modification (is this PSM-level data?)")
-
-  proteinCount <- data %>% select(Protein) %>% distinct %>% nrow
-  cat("Proteins:", proteinCount, "\n", sep = "")
-
-  sampleColumns <- metadata %>%
-    filter(Run == run) %>%
-    select(Group) %>%
-    unlist(use.names = FALSE) %>%
-    as.character
-
-  data <- MSnSet(
-    exprs = data %>% select(one_of(sampleColumns)) %>% as.matrix,
-    fData = data %>% select(Sequence, Modifications, Protein),
-    pData = metadata %>% filter(Run == run) %>% mutate(RowName = Group) %>% column_to_rownames("RowName")
-  )
-
-  peptideData[[run]] <- data
-}
-```
-
-# Introduction
-
-This report is from an analysis of the RIME-TMT proteomics data for an experiment carried out by Andrew Holding to quantify specific interactors of estrogen receptor (ER) and forkhead box protein A1 (FOXA1) in MCF-7 cells.
-
-The RIME (Rapid Immunoprecipitation of Endogenous Proteins) technique was used to pull down ER and FOXA1 and their interacting proteins, which were subsequently assessed and quantitated using mass spectrometry with tandem mass tags (TMT).
-TMT runs were carried out for 3 biological replicates, estrogen-starved MCF-7 cells from the same cell line at different times/passages, before and after addition of estrogen with samples taken at 3 different time points -- 0, 45 and 90 minutes after estrogen addition.
-A pull down of Immunoglobulin G (IgG) was also run as a control.
-
-The digests from two of the biological replicates (PR622 and PR650) were analyzed with 3 separate runs of the mass spectrometer to increase coverage; data from the repeated runs were combined prior to the analysis carried out here.
-The other sample (PR590) was run once only.
-
-`r figr("samples_table", TRUE, type="Table")` shows the isobaric tags used for each sample within each of the runs.
-
-```{r samples_table, echo=FALSE, warning=FALSE, message=FALSE, results='asis'}
-print(
-  xtable(
-    metadata %>% select(Run, Tag, Group) %>% spread(Run, Tag),
-    caption = "Isobaric tags used for each sample (group) and run."
-  ),
-  include.rownames = FALSE,
-  size = "\\setlength{\\tabcolsep}{12pt}"
-)
-```
-
-# Peptide intensity data
-
-Raw spectra were processed using Proteome Discover 2.1 to produce peptide-level intensity data with a single set of intensity values per distinct peptide.
-Only peptides with unique high-confidence protein matches were included.
-
-Multiple peptide-spectrum matches (PSMs) for the same peptide were combined using Proteome Discoverer by summing the PSM-level intensities.
-Peptide sequences with different modifications are treated as distinct peptides and the data provided contain intensities values for each modification identified for a peptide sequence.
-Several distinct peptide sequences (with modifications) may have been identified for any given protein.
-
-`r figr("peptide_protein_count_table", TRUE, type="Table")` shows the numbers of distinct peptides and proteins obtained in each TMT run.
-Also given are the numbers of peptides and proteins after filtering peptides with missing intensity values in one or more TMT channels.
-
-```{r filter_missing_intensities, include=FALSE}
-peptideDataExcludingMissingValues <- list()
-for (run in runs)
-{
-  cat(run, "\n", sep = "")
-  data <- peptideData[[run]]
-  data <- data[which(complete.cases(exprs(data))),]
-  peptideDataExcludingMissingValues[[run]] <- data
-}
-```
-
-```{r combine_peptide_intensities, include = FALSE}
-
-# function for combining data for runs into single MSnSet object
-combineRuns <- function(runData, runs)
-{
-  combinedData <- NULL
-
-  for (run in runs)
-  {
-    data <- runData[[run]]
-
-    sampleNames(data) <- pData(data) %>%
-      transmute(Sample = paste(Run, Group, sep = ":")) %>%
-      unlist %>%
-      as.character
-
-    featureNames(data) <- fData(data) %>%
-      transmute(ID = paste(Sequence, Modifications, sep = "|")) %>%
-      unlist %>%
-      as.character()
-
-    if (is.null(combinedData))
-    {
-      combinedData <- data
-    } else
-    {
-      combinedData <- BiocGenerics::combine(combinedData, data)
-    }
-  }
-
-  return(combinedData)
-}
-
-combinedPeptideData <- combineRuns(peptideData, runs)
-combinedPeptideDataExcludingMissingValues <- combineRuns(peptideDataExcludingMissingValues, runs)
-```
-
-```{r peptide_protein_count_table, echo=FALSE, warning=FALSE, message=FALSE, results='asis'}
-
-countProteins <- function(data) data %>% fData %>% select(Protein) %>% distinct %>% nrow
-
-peptideData[["All"]] <- combinedPeptideData
-peptideDataExcludingMissingValues[["All"]] <- combinedPeptideDataExcludingMissingValues
-
-counts <- bind_rows(
-  peptideData %>% sapply(nrow) %>% t %>% as.data.frame %>% mutate(label = "Peptides"),
-  peptideDataExcludingMissingValues %>% sapply(nrow) %>% t %>% as.data.frame %>% mutate(label = "Peptides with no missing values"),
-  peptideData %>% sapply(countProteins) %>% t %>% as.data.frame %>% mutate(label = "Proteins"),
-  peptideDataExcludingMissingValues %>% sapply(countProteins) %>% t %>% as.data.frame %>% mutate(label = "Proteins with no missing values")
-)
-rownames(counts) <- counts$label
-counts <- counts %>% select(-label)
-
-peptideData[["All"]] <- NULL
-peptideDataExcludingMissingValues[["All"]] <- NULL
-
-print(
-  xtable(
-    counts,
-    align = "lrrrr",
-    caption = "Numbers of peptides and proteins observed in each run."
-  ),
-  include.rownames = TRUE,
-  size = "\\setlength{\\tabcolsep}{12pt}"
-)
-```
-
-`r figr("intensity_plots", TRUE, type="Figure")` shows the distribution of intensities for each sample within each run.
-
-```{r intensity_plots, echo=FALSE, warning=FALSE, message=FALSE, fig.cap="Density plots of intensities from each run.", fig.height=8, fig.width=5}
-
-minIntensity <- -2
-maxIntensity <- 15
-
-plots <- vector("list")
-
-for (i in 1:length(runs))
-{
-  run <- runs[i]
-  intensities <- peptideData[[run]] %>% exprs %>% log2 %>% as.data.frame
-  plots[[run]] <- intensityDistributionPlot(
-    intensities,
-    groups,
-    groupColours,
-    title = run,
-    xlab=expression(log[2](intensity)),
-    minIntensity = minIntensity,
-    maxIntensity = maxIntensity,
-    showLegend = (i == 1))
-}
-
-do.call("grid.arrange", c(plots, ncol = 1))
-```
-
-## Missing intensity values
-
-Missing intensity values can result even though ions are present at detectable concentrations due to the stochastic nature of the mass spectrometry acquisition method.
-It is reasonable to expect that these missing values are randomly distributed in the data.
-Alternatively, missing values may occur when there is a low abundance of ions, below the limit of detection of the instrument.
-These biologically relevant missing values are not randomly distributed, affecting only those proteins that are expressed at low levels in the samples analysed.
-The R/Bioconductor package `MSnBase` provides imputation methods for both types of missing value.
-
-In this analysis, missing values have been handled in 3 different ways.
-The first approach is to exclude all peptide-level mesasurements where there is a missing value for one or more of the samples within a run.
-In addition, two imputation methods have been performed: one in which the missing values are set to the smallest non-missing value in the data for that run and the other in which k-nearest neighbour (KNN) averaging is applied.
-
-```{r imputation, include = FALSE}
-peptideDataImputedMin <- list()
-peptideDataImputedKNN <- list()
-for (run in runs)
-{
-  peptideDataImputedMin[[run]] <- impute(peptideData[[run]], "min")
-  peptideDataImputedKNN[[run]] <- impute(peptideData[[run]], "knn")
-}
-```
-
-`r figr("intensity_plots_impute_min", TRUE, type="Figure")` shows the distribution of intensities for each sample within each run following imputation using the smallest non-missing values.
-A small hump just below zero on the log~2~ scale is clearly visible for each run.
-
-```{r intensity_plots_impute_min, echo=FALSE, warning=FALSE, message=FALSE, fig.cap="Density plots of intensities from each run following imputation of missing values using the smallest non-missing value in the data for each run.", fig.height=8, fig.width=5}
-
-minIntensity <- -2
-maxIntensity <- 15
-
-plots <- vector("list")
-
-for (i in 1:length(runs))
-{
-  run <- runs[i]
-  intensities <- peptideDataImputedMin[[run]] %>% exprs %>% log2 %>% as.data.frame
-  plots[[run]] <- intensityDistributionPlot(
-    intensities,
-    groups,
-    groupColours,
-    title = run,
-    xlab=expression(log[2](intensity)),
-    minIntensity = minIntensity,
-    maxIntensity = maxIntensity,
-    showLegend = (i == 1))
-}
-
-do.call("grid.arrange", c(plots, ncol = 1))
-```
-
-# Normalization of intensity data
-
-The intensity distribution for PR622 suggests that the labeled samples have been pooled with differing protein concentrations and that some normalization is required to properly assess differences between groups.
-Normalization techniques that are commonly applied to microarray expression data and mRNA sequencing data assume that only a few genes, e.g.
-a few hundred out of tens of thousands, are expressed at very different levels between samples.
-This assumption may not hold for a RIME experiment where a specific set of interacting proteins is targeted.
-For example, if the majority of interacting proteins were to bind less strongly at one time point relative to another, quantile normalization or scale normalization approaches would have the effect of removing the general trend.
-In this case, following normalization, those proteins that bind less strongly may appear to be largely unchanged, while proteins with very similar levels of binding at the different time points would have artificially increased binding levels.
-Therefore some care is required when interpreting the results of a differential binding analysis for RIME TMT experiments.
-
-In this analysis, quantile normalization and scale normalization were applied for each run separately.
-
-The IgG control was used to assess non-specific binding.
-The binding of a protein was considered to be non-specific if the binding level of that protein in the condition of interest, e.g.
-ER at 45 minutes, was not significantly above the level observed for the IgG control.
-Proteins detected with the IgG control are expressed at lower levels in the PR590 and PR650 runs.
-Normalization of the IgG control along with the other samples would make it more difficult to distinguish between specific and non-specific binding since the intensities of the IgG control would be scaled to a similar level of those of the other samples.
-
-Among a number of normalization approaches tested here, normalizations were carried out that both include and exclude the IgG control.
-`r figr("normalized_intensity_plots", TRUE, type="Figure")` shows the normalized intensity distributions following scale normalization of all peptide intensities including those with missing values.
-`r figr("normalizedExcludingIgG_intensity_plots", TRUE, type="Figure")` shows the normalized intensity distributions following scale normalization for all samples except the IgG controls.
-
-```{r normalization, include = FALSE}
-
-peptideDataNormalized <- list()
-
-peptideDataNormalized[["IncludeMissingValues_NoNormalization"]] <- peptideData
-peptideDataNormalized[["ExcludeMissingValues_NoNormalization"]] <- peptideDataExcludingMissingValues
-peptideDataNormalized[["ImputeMin_NoNormalization"]] <- peptideDataImputedMin
-peptideDataNormalized[["ImputeKNN_NoNormalization"]] <- peptideDataImputedKNN
-
-for (missingValueMethod in c("IncludeMissingValues", "ExcludeMissingValues", "ImputeMin", "ImputeKNN"))
-{
-  peptideDataForNormalization <- peptideDataNormalized[[paste(missingValueMethod, "NoNormalization", sep = "_")]]
-
-  quantileNormalizationAnalysis <- paste(missingValueMethod, "QuantileNormalization", sep = "_")
-  quantileNormalizationExcludingIgGAnalysis <-
-    paste(missingValueMethod, "QuantileNormalizationExcludingIgG", sep = "_")
-  scaleNormalizationAnalysis <- paste(missingValueMethod, "ScaleNormalization", sep = "_")
-  scaleNormalizationExcludingIgGAnalysis <-
-    paste(missingValueMethod, "ScaleNormalizationExcludingIgG", sep = "_")
-  scaleNormalizationTopIgGAnalysis <- paste(missingValueMethod, "ScaleNormalizationTopIgG", sep = "_")
-
-  peptideDataNormalized[[quantileNormalizationAnalysis]] <- list()
-  peptideDataNormalized[[quantileNormalizationExcludingIgGAnalysis]] <- list()
-  peptideDataNormalized[[scaleNormalizationAnalysis]] <- list()
-  peptideDataNormalized[[scaleNormalizationExcludingIgGAnalysis]] <- list()
-  peptideDataNormalized[[scaleNormalizationTopIgGAnalysis]] <- list()
-
-  numberOfPeptidesWithHighestIgG <- 10
-
-  for (run in runs)
-  {
-    data <- peptideDataForNormalization[[run]]
-
-    # quantile normalization
-    peptideDataNormalized[[quantileNormalizationAnalysis]][[run]] <- normalize(data, method = "quantiles")
-
-    # scale normalization
-    peptideDataNormalized[[scaleNormalizationAnalysis]][[run]] <- normalizeMedianScaling(data)
-
-    samples <- setdiff(sampleNames(data), controlGroup)
-
-    # quantile normalization
-    normalized <- normalize(data[,samples], method = "quantiles")
-    peptideDataNormalized[[quantileNormalizationExcludingIgGAnalysis]][[run]] <- data
-    exprs(peptideDataNormalized[[quantileNormalizationExcludingIgGAnalysis]][[run]])[,samples] <-
-      exprs(normalized)
-
-    # scale normalization
-    peptideDataNormalized[[scaleNormalizationExcludingIgGAnalysis]][[run]] <-
-      normalizeMedianScaling(data, samples)
-
-    # scale normalization using peptides with highest IgG intensities for computing sample medians
-    controlIntensities <- data %>%
-      exprs %>%
-      as.data.frame %>%
-      select(one_of(controlGroup)) %>%
-      unlist %>%
-      as.numeric
-    controlIntensities[is.na(controlIntensities)] <- -Inf
-    controlIntensityOrder <- rev(order(controlIntensities))
-    peptideDataNormalized[[scaleNormalizationTopIgGAnalysis]][[run]] <-
-      normalizeMedianScaling(data, rowsForCalculatingMedian = controlIntensityOrder[1:numberOfPeptidesWithHighestIgG])
-  }
-}
-```
-
-```{r normalized_intensity_plots, echo=FALSE, warning=FALSE, message=FALSE, fig.cap="Density plots of normalized intensities from each TMT run where scale normalization was applied to peptide intensities that include misssing values.", fig.height=8, fig.width=5}
-
-analysis <- "IncludeMissingValues_ScaleNormalization"
-
-minIntensity <- 0
-maxIntensity <- 15
-
-plots <- vector("list")
-
-for (i in 1:length(runs))
-{
-  run <- runs[i]
-  intensities <- peptideDataNormalized[[analysis]][[run]] %>% exprs %>% log2 %>% as.data.frame
-  plots[[run]] <- intensityDistributionPlot(
-    intensities,
-    groups,
-    groupColours,
-    title = run,
-    xlab=expression(log[2](intensity)),
-    minIntensity = minIntensity,
-    maxIntensity = maxIntensity,
-    showLegend = (i == 1))
-}
-
-do.call("grid.arrange", c(plots, ncol = 1))
-```
-
-```{r normalizedExcludingIgG_intensity_plots, echo=FALSE, warning=FALSE, message=FALSE, fig.cap="Density plots of normalized intensities from each TMT run where scale normalization was applied to peptide intensities for all samples except the IgG controls.", fig.height=8, fig.width=5}
-
-analysis <- "IncludeMissingValues_ScaleNormalizationExcludingIgG"
-
-minIntensity <- 0
-maxIntensity <- 15
-
-plots <- vector("list")
-
-for (i in 1:length(runs))
-{
-  run <- runs[i]
-  intensities <- peptideDataNormalized[[analysis]][[run]] %>% exprs %>% log2 %>% as.data.frame
-  plots[[run]] <- intensityDistributionPlot(
-    intensities,
-    groups,
-    groupColours,
-    title = run,
-    xlab=expression(log[2](intensity)),
-    minIntensity = minIntensity,
-    maxIntensity = maxIntensity,
-    showLegend = (i == 1))
-}
-
-do.call("grid.arrange", c(plots, ncol = 1))
-```
-
-A further normalization approach was attempted that scales the data from all samples within a run based on a subset of peptides that are considered most likely to be at similar concentrations between the samples.
-The assumption is that the peptides with the highest intensity values in the IgG control are the result of non-specific binding and that that binding is consistent across all samples.
-Similar to the scale normalization applied in the first approach, the median intensity within each sample is computed to determine a scaling factor for each sample within a run but the median computed only uses the `r numberOfPeptidesWithHighestIgG` peptides with the highest IgG measurements.
-As before, normalization is carried out within each run separately.
-
-`r figr("normalizedIgG_intensity_plots", TRUE, type="Figure")` shows the resulting normalized intensity distributions.
-Strikingly, the IgG intensity distribution is shifted to lower values in run PR622, making it more consistent with the other 2 runs.
-
-```{r normalizedIgG_intensity_plots, echo=FALSE, warning=FALSE, message=FALSE, fig.cap=paste("Density plots of normalized intensities from each TMT run where scale normalization is applied based on", numberOfPeptidesWithHighestIgG, "peptides with the highest IgG intensities in each run."), fig.height=8, fig.width=5}
-
-analysis <- "IncludeMissingValues_ScaleNormalizationTopIgG"
-
-minIntensity <- 0
-maxIntensity <- 15
-
-plots <- vector("list")
-
-for (i in 1:length(runs))
-{
-  run <- runs[i]
-  intensities <- peptideDataNormalized[[analysis]][[run]] %>% exprs %>% log2 %>% as.data.frame
-  plots[[run]] <- intensityDistributionPlot(
-    intensities,
-    groups,
-    groupColours,
-    title = run,
-    xlab=expression(log[2](intensity)),
-    minIntensity = minIntensity,
-    maxIntensity = maxIntensity,
-    showLegend = (i == 1))
-}
-
-do.call("grid.arrange", c(plots, ncol = 1))
-```
-
-# Protein-level quantification
-
-Protein-level quantification was carried out for each run separately by summing normalized intensity values for all peptides matching to a particular protein.
-This is similar in principle to the gene-level quantification that is carried out during microarray analysis or the read-counting approach in RNA-Seq.
-Essentially, all signal attributable to a particular protein is assigned to that protein.
-Intensity values vary in magnitude depending on the elution profile for the peptide and when the peptide is sampled.
-Relative intensities for a peptide between the tags or samples within a run should be consistent but the signal may include some contribution from a contaminating, co-eluting peptide, the extent of which may differ for different PSMs.
-In summing intensities from multiple measurements for the same peptide, those peptides that have higher levels of intensity values will have a greater weight in determining the overall intensity; higher intensity measurements are likely to more accurately reflect differences between samples so this may be advantageous.
-
-When no imputation of missing values was carried out, peptides with missing values in one or more samples were excluded from the summation.
-Missing values that arise for technical reasons instead of being at undetectable levels in a sample would otherwise reduce the overall intensity in that sample relative to other samples.
-
-```{r summarize_protein_intensities, include=FALSE}
-
-proteinData <- list()
-
-for (analysis in names(peptideDataNormalized))
-{
-  peptideDataForSummarization <- peptideDataNormalized[[analysis]]
-
-  proteinData[[analysis]] <- list()
-
-  for (run in runs)
-  {
-    data <- peptideDataForSummarization[[run]]
-
-    # exclude rows with missing values (will have no effect if imputation carried out)
-    data <- data[which(complete.cases(exprs(data))),]
-
-    data <- combineFeatures(data, groupBy = fData(data)$Protein, fun = "sum", na.rm = FALSE)
-    fData(data) <- fData(data) %>% select(Protein)
-
-    proteinData[[analysis]][[run]] <- data
-  }
-}
-```
-
-```{r combine_protein_intensities, include = FALSE}
-
-combinedProteinData <- list()
-
-for (analysis in names(proteinData))
-{
-  for (run in runs)
-  {
-    data <- proteinData[[analysis]][[run]]
-
-    sampleNames(data) <- pData(data) %>%
-      transmute(Sample = paste(Run, Group, sep = ":")) %>%
-      unlist %>%
-      as.character
-
-    if (is.null(combinedProteinData[[analysis]]))
-    {
-      combinedProteinData[[analysis]] <- data
-    } else
-    {
-      combinedProteinData[[analysis]] <- BiocGenerics::combine(combinedProteinData[[analysis]], data)
-    }
-  }
-}
-```
-
-```{r replicate_counts, include=FALSE}
-
-isRepresented <- function(x) any(!is.na(x))
-
-getReplicateCounts <- function(data)
-{
-  metadata <- pData(data) %>% rownames_to_column("Sample")
-
-  intensities <- data %>%
-    exprs %>%
-    as.data.frame %>%
-    mutate(N = 0)
-
-  for (run in runs)
-  {
-    samples <- metadata %>%
-      filter(Run == run) %>%
-      select(Sample) %>%
-      unlist %>%
-      as.character
-
-    intensities <- intensities %>%
-      mutate(N = N + apply(select(., one_of(samples)), 1, isRepresented))
-  }
-
-  replicateCounts <- intensities %>%
-    select(N) %>%
-    mutate(Count = 0) %>%
-    group_by(N) %>%
-    summarize_each(funs(length)) %>%
-    ungroup %>%
-    mutate(N = as.character(N))
-
-  replicateCounts <- replicateCounts %>%
-    bind_rows(data_frame(N = "total", Count = sum(replicateCounts$Count)))
-
-  return(replicateCounts)
-}
-
-replicateCounts <- getReplicateCounts(combinedProteinData[["ImputeKNN_ScaleNormalization"]])
-
-replicateCountTableData <- inner_join(
-  getReplicateCounts(combinedProteinData[["ImputeKNN_ScaleNormalization"]]) %>%
-    rename(`Including missing values` = Count),
-  getReplicateCounts(combinedProteinData[["ExcludeMissingValues_ScaleNormalization"]]) %>%
-    rename(`Excluding missing values` = Count),
-  by = "N"
-)
-
-replicateCountTableData <- replicateCountTableData %>%
-  as.data.frame %>%
-  column_to_rownames(var = "N") %>%
-  t %>%
-  as.data.frame %>%
-  rownames_to_column(var = "Runs/replicates")
-```
-
-The statistical analysis of differential binding between groups that follows requires at least 3 observations per group.
-`r figr("replicate_count_table", TRUE, type="Table")` gives the number of protein identified in just a single run, two runs or all three runs.
-A total of `r (replicateCounts %>% filter(N == "all"))$Count` proteins were identified across all runs of which `r (replicateCounts %>% filter(N == "3"))$Count` were observed in all three runs, allowing for a statistical analysis.
-log~2~ fold changes are still computed for proteins identified only in one or two of the runs, but no measure of the statistical significance is given in the differential binding analysis.
-
-```{r replicate_count_table, echo=FALSE, warning=FALSE, message=FALSE, results='asis'}
-# colnames(replicateCounts) <- c("Runs/replicates", "Proteins (includes missing values)", "Proteins")
-print(
-  xtable(
-    replicateCountTableData,
-    # align = "llr",
-    caption = "Numbers of proteins identified in differing numbers of runs and total number of proteins identified in all runs."
-  ),
-  include.rownames = FALSE,
-  size = "\\setlength{\\tabcolsep}{12pt}"
-)
-```
-
-# Principal Component Analysis
-
-Protein-level intensities were scaled to sum to 1.0 for each protein within a run, allowing for comparison across runs in a principal component analysis (PCA).
-
-`r figr("pca_plot", TRUE, type="Figure")` shows a PCA plot for the first two principal components using all proteins that were sampled in all three runs.
-The plot shows a clear separation of the ER, FOXA1 and IgG control pull-downs.
-The 0, 45 and 90 minute timepoints within each of the ER and FOXA1 groups are not completely separated in the first two principal components.
-`r figr("pca_plot_er", TRUE, type="Figure")` shows the PCA plot for just the ER samples.
-
-```{r pca_plot, echo=FALSE, warning=FALSE, message=FALSE, fig.cap="Principal Componenent Analysis for proteins sampled in all 3 runs.
-The PCA was based on protein-level data resulting from summation of quantile normalized peptide intensities in which missing values were imputed using KNN-based nearest neighbour averaging.
-The first two principal components are displayed.", fig.height=5}
-
-data <- combinedProteinData[["ImputeKNN_QuantileNormalization"]]
-
-metadata <- pData(data) %>% rownames_to_column("Sample")
-
-scaledIntensities <- data %>%
-  exprs %>%
-  as.data.frame %>%
-  filter(complete.cases(.))
-
-for (run in runs)
-{
-  samples <- metadata %>%
-    filter(Run == run) %>%
-    select(Sample) %>%
-    unlist %>%
-    as.character
-
-  summedIntensities <- scaledIntensities %>%
-    select(one_of(samples)) %>%
-    rowSums
-
-  scaledIntensities <- scaledIntensities %>%
-    mutate_each(funs(. / summedIntensities), one_of(samples))
-}
-
-print(pcaPlot(scaledIntensities, metadata$Sample, metadata$Group, groupColours, labels = metadata$Run, legend = TRUE))
-```
-
-```{r pca_plot_er, echo=FALSE, warning=FALSE, message=FALSE, fig.cap="Principal Componenent Analysis for ER-interactng proteins sampled in all 3 runs.
-The first two principal components are displayed.", fig.height=5}
-
-metadata <- metadata %>% filter(grepl("^ER", Group))
-
-print(pcaPlot(scaledIntensities, metadata$Sample, metadata$Group, groupColours, labels = metadata$Run, legend = TRUE))
-```
-
-# Differential Binding
+# Differential binding
 
@@
 ```
-
-A statistical analysis of differentially-expressed peptides was carried out using **limma**, a R/Bioconductor package commonly used in the analysis of microarray and RNA-seq data, but applicable to data from any quantitative expression technology.
-
-Limma uses linear models to assess differential expression in the context of multifactor experimental designs.
-It is able to analyze comparisons between many RNA targets simultaneously (or in this case many proteins) and has features that make these analyses stable even for experiments with small numbers of samples; this is achieved by borrowing information across genes (proteins).
-
-In this analysis, limma was used to estimate log~2~ fold changes and standard errors by fitting a linear model for each protein where the model includes variables for the group (the ER and FOXA1 pull-downs at 0, 45 and 90 minute time points and the IgG control at 45 minutes) and the run.
-This is essentially a two-way ANOVA, a generalization of a paired analysis, in which comparisons between pull-downs at different time points are made within each run.
-
-Limma employs an empirical Bayes method to moderate the standard errors of the estimated log~2~ fold changes; this results in more stable inference and improved power, especially for experiments with small numbers of replicates.
-
-Multiple testing correction of p-values was applied using the Benjamini-Hochberg method to control the false discovery rate (FDR).
-The adjusted p-value (also known as a q-value) can be understood as follows.
-If all proteins with q-value below a threshold of `r significanceLevel` are selected as differentially expressed, then the expected proportion of false discoveries in the selected group is controlled to be less than the threshold value, in this case 5%.
-
-The B-statistic (lods or B) is the log odds that the protein is differentially expressed.
-For example, if B = 1.5, the odds of differential expression is $e^{1.5} = 4.48$, i.e, about four and a half to one.
-The probability that the protein is differentially expressed is $4.48 / (1 + 4.48) = 0.82$.
-A B-statistic of zero corresponds to a 50-50 chance that the gene is differentially expressed.
-The B-statistic has been automatically adjusted for multiple testing by assuming that 1% of the proteins are expected to be differentially expressed.
 
@@
 
-contrasts <- c(
-  "ER 45min vs ER 0min",
-  "ER 90min vs ER 0min",
-  "ER 90min vs ER 45min",
-  "FOXA1 45min vs FOXA1 0min",
-  "FOXA1 90min vs FOXA1 0min",
-  "FOXA1 90min vs FOXA1 45min"
-)
+contrasts <- contrastsTable$Contrast
 
@@
 
-The IgG control was used to assess non-specific binding.
-The binding of a protein is considered to be non-specific if the log~2~ fold change relative to the IgG control is less than `r controlLogFoldChangeThreshold`.
-In the differential binding analysis, protein intensity values are fitted to a linear model and the fitted values for each condition being compared are also compared to the IgG control.
-For each comparison of two groups, the maximum log~2~ fold change from each of the two groups above the IgG control is used to determine if the binding is specific.
-
-\newpage
-
-```{r contrast_1, include = FALSE}
-contrastGroups <- c("ER 45min", "ER 0min")
-contrastName <- paste(contrastGroups, collapse = " vs ")
-processingDescription <- "excluding peptides with missing intensities, no normalization"
-analysis <- "ExcludeMissingValues_NoNormalization"
-```
-
-## `r paste(contrastName, processingDescription, sep = ", ")`
-
-```{r results_1, include = FALSE}
-
-data <- combinedProteinData[[analysis]]
-exprs(data) <- log2(exprs(data))
-
-results <- createContrastResults(
-  data,
-  fittedLinearModels[[analysis]],
-  fittedLinearModelsExcludingControls[[analysis]],
-  fittedContrasts[[analysis]],
-  runs,
-  contrastGroups,
-  controlGroup,
-  proteinInfo)
-```
-
-```{r ma_plot_1, echo=FALSE, warning=FALSE, message=FALSE, fig.cap=paste("MA plot of the average intensity against log~2~ fold change for the ", contrastName, " comparison (", processingDescription, ").
-Top ranking differentially-expressed proteins with false discovery rate below 0.05 are highlighted in pink.
-Open circles indicate that the protein is non-specific from the IgG control comparison.", sep = ""), fig.height=6}
-print(maPlot(results,
-             significanceLevel = significanceLevel,
-             controlLogFoldChangeThreshold = controlLogFoldChangeThreshold,
-             minLogFoldChangeForLabelling = 1.5,
-             maxNumberLabelledProteins = 20,
-             pointSize = 1.5))
-```
-
-```{r volcano_plot_1, echo=FALSE, warning=FALSE, message=FALSE, fig.cap=paste("Volcano plot of the average intensity against log~2~ fold change for the ", contrastName, " comparison (", processingDescription, ").
-Top ranking differentially-expressed proteins with false discovery rate below 0.05 are highlighted in pink.
-Open circles indicate that the protein is non-specific from the IgG control comparison.", sep = ""), fig.height=6}
-print(volcanoPlot(results,
-                  significanceLevel = significanceLevel,
-                  minLogFoldChange = -2.0,
-                  maxLogFoldChange = 2.75,
-                  controlLogFoldChangeThreshold = controlLogFoldChangeThreshold,
-                  minLogFoldChangeForLabelling = 1.5,
-                  maxNumberLabelledProteins = 20,
-                  pointSize = 1.5))
-```
-
-```{r pvalue_histogram_1, echo=FALSE, warning=FALSE, message=FALSE, fig.cap=paste("Histogram of p-values for the ", contrastName, " comparison (", processingDescription, ")", sep = ""), fig.height=3.5, fig.width=4.5}
-print(histogram(results, "P.Value", xlab = "p-value"))
-```
-
-```{r qq_plot_1, echo=FALSE, warning=FALSE, message=FALSE, fig.cap=paste("QQ plot of the adjusted p-values for the ", contrastName, " comparison (", processingDescription, ")", sep = ""), fig.height=5, fig.width=5}
-q <- results %>%
-  select(adj.P.Val) %>%
-  filter(!is.na(adj.P.Val)) %>%
-  arrange(adj.P.Val) %>%
-  unlist %>%
-  as.numeric
-qqplot(-log10(ppoints(length(q))),
-       -log10(q),
-       xlab=expression(Expected~~-log[10](italic(q))),
-       ylab=expression(Observed~~-log[10](italic(q))))
-abline(0, 1, col = "red")
-```
-
-\newpage
-
 ```{r contrast_2, include = FALSE}
@@
 
-## `r paste(contrastName, processingDescription, sep = ", ")`
+Here follows the analysis output for `r paste(contrastName, processingDescription, sep = ", ")`.
 
@@
 
+## MA plot
+
 ```{r ma_plot_2, echo=FALSE, warning=FALSE, message=FALSE, fig.cap=paste("MA plot of the average intensity against log~2~ fold change for the ", contrastName, " comparison (", processingDescription, ").
-Top ranking differentially-expressed proteins with false discovery rate below 0.05 are highlighted in pink.
+Top ranking differentially-bound proteins with false discovery rate below 0.05 are highlighted in pink.
 Open circles indicate that the protein is non-specific from the IgG control comparison.", sep = ""), fig.height=6}
@@
 
+\newpage
+
+## Volcano plot
+
 ```{r volcano_plot_2, echo=FALSE, warning=FALSE, message=FALSE, fig.cap=paste("Volcano plot of the average intensity against log~2~ fold change for the ", contrastName, " comparison (", processingDescription, ").
-Top ranking differentially-expressed proteins with false discovery rate below 0.05 are highlighted in pink.
+Top ranking differentially-bound proteins with false discovery rate below 0.05 are highlighted in pink.
 Open circles indicate that the protein is non-specific from the IgG control comparison.", sep = ""), fig.height=6}
@@
 ```
+\newpage
+
+## P-value histogram
 
@@
 ```
+
+## QQ plot
 
@@
 
-```{r contrast_3, include = FALSE}
-contrastGroups <- c("ER 45min", "ER 0min")
-contrastName <- paste(contrastGroups, collapse = " vs ")
-processingDescription <- "excluding peptides with missing intensities, quantile normalization excluding IgG control"
-analysis <- "ExcludeMissingValues_QuantileNormalizationExcludingIgG"
-```
-
-## `r paste(contrastName, processingDescription, sep = ", ")`
-
-```{r results_3, include = FALSE}
-
-data <- combinedProteinData[[analysis]]
-exprs(data) <- log2(exprs(data))
-
-results <- createContrastResults(
-  data,
-  fittedLinearModels[[analysis]],
-  fittedLinearModelsExcludingControls[[analysis]],
-  fittedContrasts[[analysis]],
-  runs,
-  contrastGroups,
-  controlGroup,
-  proteinInfo)
-```
-
-```{r ma_plot_3, echo=FALSE, warning=FALSE, message=FALSE, fig.cap=paste("MA plot of the average intensity against log~2~ fold change for the ", contrastName, " comparison (", processingDescription, ").
-Top ranking differentially-expressed proteins with false discovery rate below 0.05 are highlighted in pink.
-Open circles indicate that the protein is non-specific from the IgG control comparison.", sep = ""), fig.height=6}
-print(maPlot(results,
-             significanceLevel = significanceLevel,
-             controlLogFoldChangeThreshold = controlLogFoldChangeThreshold,
-             minLogFoldChangeForLabelling = 1.5,
-             maxNumberLabelledProteins = 50,
-             pointSize = 1.5))
-```
-
-```{r volcano_plot_3, echo=FALSE, warning=FALSE, message=FALSE, fig.cap=paste("Volcano plot of the average intensity against log~2~ fold change for the ", contrastName, " comparison (", processingDescription, ").
-Top ranking differentially-expressed proteins with false discovery rate below 0.05 are highlighted in pink.
-Open circles indicate that the protein is non-specific from the IgG control comparison.", sep = ""), fig.height=6}
-print(volcanoPlot(results,
-                  significanceLevel = significanceLevel,
-                  minLogFoldChange = -2.0,
-                  maxLogFoldChange = 2.75,
-                  controlLogFoldChangeThreshold = controlLogFoldChangeThreshold,
-                  minLogFoldChangeForLabelling = 1.5,
-                  maxNumberLabelledProteins = 50,
-                  pointSize = 1.5))
-```
-
-```{r pvalue_histogram_3, echo=FALSE, warning=FALSE, message=FALSE, fig.cap=paste("Histogram of p-values for the ", contrastName, " comparison (", processingDescription, ")", sep = ""), fig.height=3.5, fig.width=4.5}
-print(histogram(results, "P.Value", xlab = "p-value"))
-```
-
-```{r qq_plot_3, echo=FALSE, warning=FALSE, message=FALSE, fig.cap=paste("QQ plot of the adjusted p-values for the ", contrastName, " comparison (", processingDescription, ")", sep = ""), fig.height=5, fig.width=5}
-q <- results %>%
-  select(adj.P.Val) %>%
-  filter(!is.na(adj.P.Val)) %>%
-  arrange(adj.P.Val) %>%
-  unlist %>%
-  as.numeric
-qqplot(-log10(ppoints(length(q))),
-       -log10(q),
-       xlab=expression(Expected~~-log[10](italic(q))),
-       ylab=expression(Observed~~-log[10](italic(q))))
-abline(0, 1, col = "red")
-```
-
-\newpage
-
-```{r contrast_4, include = FALSE}
-contrastGroups <- c("ER 45min", "ER 0min")
-contrastName <- paste(contrastGroups, collapse = " vs ")
-processingDescription <- "excluding peptides with missing intensities, scale normalization"
-analysis <- "ExcludeMissingValues_ScaleNormalization"
-```
-
-## `r paste(contrastName, processingDescription, sep = ", ")`
-
-```{r results_4, include = FALSE}
-
-data <- combinedProteinData[[analysis]]
-exprs(data) <- log2(exprs(data))
-
-results <- createContrastResults(
-  data,
-  fittedLinearModels[[analysis]],
-  fittedLinearModelsExcludingControls[[analysis]],
-  fittedContrasts[[analysis]],
-  runs,
-  contrastGroups,
-  controlGroup,
-  proteinInfo)
-```
-
-```{r ma_plot_4, echo=FALSE, warning=FALSE, message=FALSE, fig.cap=paste("MA plot of the average intensity against log~2~ fold change for the ", contrastName, " comparison (", processingDescription, ").
-Top ranking differentially-expressed proteins with false discovery rate below 0.05 are highlighted in pink.
-Open circles indicate that the protein is non-specific from the IgG control comparison.", sep = ""), fig.height=6}
-print(maPlot(results,
-             significanceLevel = significanceLevel,
-             controlLogFoldChangeThreshold = controlLogFoldChangeThreshold,
-             minLogFoldChangeForLabelling = 1.5,
-             maxNumberLabelledProteins = 50,
-             pointSize = 1.5))
-```
-
-```{r volcano_plot_4, echo=FALSE, warning=FALSE, message=FALSE, fig.cap=paste("Volcano plot of the average intensity against log~2~ fold change for the ", contrastName, " comparison (", processingDescription, ").
-Top ranking differentially-expressed proteins with false discovery rate below 0.05 are highlighted in pink.
-Open circles indicate that the protein is non-specific from the IgG control comparison.", sep = ""), fig.height=6}
-print(volcanoPlot(results,
-                  significanceLevel = significanceLevel,
-                  minLogFoldChange = -2.0,
-                  maxLogFoldChange = 2.75,
-                  controlLogFoldChangeThreshold = controlLogFoldChangeThreshold,
-                  minLogFoldChangeForLabelling = 1.5,
-                  maxNumberLabelledProteins = 50,
-                  pointSize = 1.5))
-```
-
-```{r pvalue_histogram_4, echo=FALSE, warning=FALSE, message=FALSE, fig.cap=paste("Histogram of p-values for the ", contrastName, " comparison (", processingDescription, ")", sep = ""), fig.height=3.5, fig.width=4.5}
-print(histogram(results, "P.Value", xlab = "p-value"))
-```
-
-```{r qq_plot_4, echo=FALSE, warning=FALSE, message=FALSE, fig.cap=paste("QQ plot of the adjusted p-values for the ", contrastName, " comparison (", processingDescription, ")", sep = ""), fig.height=5, fig.width=5}
-q <- results %>%
-  select(adj.P.Val) %>%
-  filter(!is.na(adj.P.Val)) %>%
-  arrange(adj.P.Val) %>%
-  unlist %>%
-  as.numeric
-qqplot(-log10(ppoints(length(q))),
-       -log10(q),
-       xlab=expression(Expected~~-log[10](italic(q))),
-       ylab=expression(Observed~~-log[10](italic(q))))
-abline(0, 1, col = "red")
-```
-
-\newpage
-
-```{r contrast_5, include = FALSE}
-contrastGroups <- c("ER 45min", "ER 0min")
-contrastName <- paste(contrastGroups, collapse = " vs ")
-processingDescription <- paste("excluding peptides with missing intensities, scale normalization using top", numberOfPeptidesWithHighestIgG, "IgG peptides")
-analysis <- "ExcludeMissingValues_ScaleNormalizationTopIgG"
-```
-
-## `r paste(contrastName, processingDescription, sep = ", ")`
-
-```{r results_5, include = FALSE}
-
-data <- combinedProteinData[[analysis]]
-exprs(data) <- log2(exprs(data))
-
-results <- createContrastResults(
-  data,
-  fittedLinearModels[[analysis]],
-  fittedLinearModelsExcludingControls[[analysis]],
-  fittedContrasts[[analysis]],
-  runs,
-  contrastGroups,
-  controlGroup,
-  proteinInfo)
-```
-
-```{r ma_plot_5, echo=FALSE, warning=FALSE, message=FALSE, fig.cap=paste("MA plot of the average intensity against log~2~ fold change for the ", contrastName, " comparison (", processingDescription, ").
-Top ranking differentially-expressed proteins with false discovery rate below 0.05 are highlighted in pink.
-Open circles indicate that the protein is non-specific from the IgG control comparison.", sep = ""), fig.height=6}
-print(maPlot(results,
-             significanceLevel = significanceLevel,
-             controlLogFoldChangeThreshold = controlLogFoldChangeThreshold,
-             minLogFoldChangeForLabelling = 1.5,
-             maxNumberLabelledProteins = 50,
-             pointSize = 1.5))
-```
-
-```{r volcano_plot_5, echo=FALSE, warning=FALSE, message=FALSE, fig.cap=paste("Volcano plot of the average intensity against log~2~ fold change for the ", contrastName, " comparison (", processingDescription, ").
-Top ranking differentially-expressed proteins with false discovery rate below 0.05 are highlighted in pink.
-Open circles indicate that the protein is non-specific from the IgG control comparison.", sep = ""), fig.height=6}
-print(volcanoPlot(results,
-                  significanceLevel = significanceLevel,
-                  minLogFoldChange = -2.0,
-                  maxLogFoldChange = 2.75,
-                  controlLogFoldChangeThreshold = controlLogFoldChangeThreshold,
-                  minLogFoldChangeForLabelling = 1.5,
-                  maxNumberLabelledProteins = 50,
-                  pointSize = 1.5))
-```
-
-```{r pvalue_histogram_5, echo=FALSE, warning=FALSE, message=FALSE, fig.cap=paste("Histogram of p-values for the ", contrastName, " comparison (", processingDescription, ")", sep = ""), fig.height=3.5, fig.width=4.5}
-print(histogram(results, "P.Value", xlab = "p-value"))
-```
-
-```{r qq_plot_5, echo=FALSE, warning=FALSE, message=FALSE, fig.cap=paste("QQ plot of the adjusted p-values for the ", contrastName, " comparison (", processingDescription, ")", sep = ""), fig.height=5, fig.width=5}
-q <- results %>%
-  select(adj.P.Val) %>%
-  filter(!is.na(adj.P.Val)) %>%
-  arrange(adj.P.Val) %>%
-  unlist %>%
-  as.numeric
-qqplot(-log10(ppoints(length(q))),
-       -log10(q),
-       xlab=expression(Expected~~-log[10](italic(q))),
-       ylab=expression(Observed~~-log[10](italic(q))))
-abline(0, 1, col = "red")
-```
-
-\newpage
-
-```{r contrast_6, include = FALSE}
-contrastGroups <- c("ER 45min", "ER 0min")
-contrastName <- paste(contrastGroups, collapse = " vs ")
-processingDescription <- "KNN nearest neighbour averaging imputation of missing values, quantile normalization"
-analysis <- "ImputeKNN_QuantileNormalization"
-```
-
-## `r paste(contrastName, processingDescription, sep = ", ")`
-
-```{r results_6, include = FALSE}
-
-data <- combinedProteinData[[analysis]]
-exprs(data) <- log2(exprs(data))
-
-results <- createContrastResults(
-  data,
-  fittedLinearModels[[analysis]],
-  fittedLinearModelsExcludingControls[[analysis]],
-  fittedContrasts[[analysis]],
-  runs,
-  contrastGroups,
-  controlGroup,
-  proteinInfo)
-```
-
-```{r ma_plot_6, echo=FALSE, warning=FALSE, message=FALSE, fig.cap=paste("MA plot of the average intensity against log~2~ fold change for the ", contrastName, " comparison (", processingDescription, ").
-Top ranking differentially-expressed proteins with false discovery rate below 0.05 are highlighted in pink.
-Open circles indicate that the protein is non-specific from the IgG control comparison.", sep = ""), fig.height=6}
-print(maPlot(results,
-             significanceLevel = significanceLevel,
-             controlLogFoldChangeThreshold = controlLogFoldChangeThreshold,
-             minLogFoldChangeForLabelling = 1.5,
-             maxNumberLabelledProteins = 50,
-             pointSize = 1.5))
-```
-
-```{r volcano_plot_6, echo=FALSE, warning=FALSE, message=FALSE, fig.cap=paste("Volcano plot of the average intensity against log~2~ fold change for the ", contrastName, " comparison (", processingDescription, ").
-Top ranking differentially-expressed proteins with false discovery rate below 0.05 are highlighted in pink.
-Open circles indicate that the protein is non-specific from the IgG control comparison.", sep = ""), fig.height=6}
-print(volcanoPlot(results,
-                  significanceLevel = significanceLevel,
-                  minLogFoldChange = -2.0,
-                  maxLogFoldChange = 2.75,
-                  controlLogFoldChangeThreshold = controlLogFoldChangeThreshold,
-                  minLogFoldChangeForLabelling = 1.5,
-                  maxNumberLabelledProteins = 50,
-                  pointSize = 1.5))
-```
-
-```{r pvalue_histogram_6, echo=FALSE, warning=FALSE, message=FALSE, fig.cap=paste("Histogram of p-values for the ", contrastName, " comparison (", processingDescription, ")", sep = ""), fig.height=3.5, fig.width=4.5}
-print(histogram(results, "P.Value", xlab = "p-value"))
-```
-
-```{r qq_plot_6, echo=FALSE, warning=FALSE, message=FALSE, fig.cap=paste("QQ plot of the adjusted p-values for the ", contrastName, " comparison (", processingDescription, ")", sep = ""), fig.height=5, fig.width=5}
-q <- results %>%
-  select(adj.P.Val) %>%
-  filter(!is.na(adj.P.Val)) %>%
-  arrange(adj.P.Val) %>%
-  unlist %>%
-  as.numeric
-qqplot(-log10(ppoints(length(q))),
-       -log10(q),
-       xlab=expression(Expected~~-log[10](italic(q))),
-       ylab=expression(Observed~~-log[10](italic(q))))
-abline(0, 1, col = "red")
-```
-
-\newpage
-
-# Comparison of differential binding analysis approaches
-
-In this section, comparisons of the differential binding results obtained using differing methods for handling missing values and varying normalization techniques are presented.
-The B-statistic (log odds) is used to rank the proteins from most significantly differentially expressed to least significant.
-Scatter plots of both the B-statistic generated for the two methods being compared and the rank are presented.
-The protein with the highest B-statistic, i.e.
-the most significantly differentially expressed protein, is given a rank of 1.
-
-Differing imputation methods have relatively little effect on the results for most proteins.
-This is largely because only a small proportion of the proteins have peptide-level measurements containing missing values.
-There are a few proteins that are only represented in the differential binding results when imputation is carried out, for which all peptide-level measurements contain missing values.
-These proteins are not represented in the comparions with the analysis in which measurements with missing values are removed.
-
-Different normalization approaches have a more substantial impact on the differential binding analysis.
-
-\newpage
-
-## Removal of missing values vs KNN imputation (quantile normalization in both)
-
-```{r exclude_missing_vs_knn_imputation, include = FALSE}
-
-contrastGroups <- c("ER 45min", "ER 0min")
-contrastName <- paste(contrastGroups, collapse = " vs ")
-
-analysis <- "ExcludeMissingValues_QuantileNormalization"
-data <- combinedProteinData[[analysis]]
-exprs(data) <- log2(exprs(data))
-results1 <- createContrastResults(
-  data,
-  fittedLinearModels[[analysis]],
-  fittedLinearModelsExcludingControls[[analysis]],
-  fittedContrasts[[analysis]],
-  runs,
-  contrastGroups,
-  controlGroup,
-  proteinInfo)
-
-analysis <- "ImputeKNN_QuantileNormalization"
-data <- combinedProteinData[[analysis]]
-exprs(data) <- log2(exprs(data))
-results2 <- createContrastResults(
-  data,
-  fittedLinearModels[[analysis]],
-  fittedLinearModelsExcludingControls[[analysis]],
-  fittedContrasts[[analysis]],
-  runs,
-  contrastGroups,
-  controlGroup,
-  proteinInfo)
-
-results <- inner_join(
-  results1 %>% select(Protein, B1 = B),
-  results2 %>% select(Protein, B2 = B),
-  by = "Protein"
-  ) %>%
-  filter(complete.cases(.)) %>%
-  mutate(rank1 = rank(-B1), rank2 = rank(-B2))
-```
-
-```{r exclude_missing_vs_knn_imputation_B_plot, echo=FALSE, warning=FALSE, message=FALSE, fig.cap=paste("Plot of B-statistic (log odds) for differential expression of proteins for the ", contrastName, " contrast, comparing results following exclusion of missing values and KNN imputation.", sep = ""), fig.height=2.9, fig.width=3.5}
-print(scatterPlot(results, "B1", "B2", "B-statistic (excluding missing values)", "B-statistic (KNN imputation)"))
-```
-
-```{r exclude_missing_vs_knn_imputation_rank_plot, echo=FALSE, warning=FALSE, message=FALSE, fig.cap=paste("Plot of the rank of differentially-expressed proteins for the ", contrastName, " contrast, comparing results following exclusion of missing values and KNN imputation.", sep = ""), fig.height=2.9, fig.width=3.5}
-print(scatterPlot(results, "rank1", "rank2", "Rank (excluding missing values)", "Rank (KNN imputation)"))
-```
-
-\newpage
-
-## Removal of missing values vs lowest value imputation (quantile normalization in both)
-
-```{r exclude_missing_vs_min_imputation, include = FALSE}
-
-contrastGroups <- c("ER 45min", "ER 0min")
-contrastName <- paste(contrastGroups, collapse = " vs ")
-
-analysis <- "ExcludeMissingValues_QuantileNormalization"
-data <- combinedProteinData[[analysis]]
-exprs(data) <- log2(exprs(data))
-results1 <- createContrastResults(
-  data,
-  fittedLinearModels[[analysis]],
-  fittedLinearModelsExcludingControls[[analysis]],
-  fittedContrasts[[analysis]],
-  runs,
-  contrastGroups,
-  controlGroup,
-  proteinInfo)
-
-analysis <- "ImputeMin_QuantileNormalization"
-data <- combinedProteinData[[analysis]]
-exprs(data) <- log2(exprs(data))
-results2 <- createContrastResults(
-  data,
-  fittedLinearModels[[analysis]],
-  fittedLinearModelsExcludingControls[[analysis]],
-  fittedContrasts[[analysis]],
-  runs,
-  contrastGroups,
-  controlGroup,
-  proteinInfo)
-
-results <- inner_join(
-  results1 %>% select(Protein, B1 = B),
-  results2 %>% select(Protein, B2 = B),
-  by = "Protein"
-  ) %>%
-  filter(complete.cases(.)) %>%
-  mutate(rank1 = rank(-B1), rank2 = rank(-B2))
-```
-
-```{r exclude_missing_vs_min_imputation_B_plot, echo=FALSE, warning=FALSE, message=FALSE, fig.cap=paste("Plot of B-statistic (log odds) for differential expression of proteins for the ", contrastName, " contrast, comparing results following exclusion of missing values and lowest value imputation.", sep = ""), fig.height=2.9, fig.width=3.5}
-print(scatterPlot(results, "B1", "B2", "B-statistic (excluding missing values)", "B-statistic (lowest value imputation)"))
-```
-
-```{r exclude_missing_vs_min_imputation_rank_plot, echo=FALSE, warning=FALSE, message=FALSE, fig.cap=paste("Plot of the rank of differentially-expressed proteins for the ", contrastName, " contrast, comparing results following exclusion of missing values and lowest value imputation.", sep = ""), fig.height=2.9, fig.width=3.5}
-print(scatterPlot(results, "rank1", "rank2", "Rank (excluding missing values)", "Rank (lowest value imputation)"))
-```
-
-\newpage
-
-## Quantile vs Scale normalization (missing values excluded in both)
-
-```{r quantile_vs_scale_normalization, include = FALSE}
-
-contrastGroups <- c("ER 45min", "ER 0min")
-contrastName <- paste(contrastGroups, collapse = " vs ")
-
-analysis <- "ExcludeMissingValues_QuantileNormalization"
-data <- combinedProteinData[[analysis]]
-exprs(data) <- log2(exprs(data))
-results1 <- createContrastResults(
-  data,
-  fittedLinearModels[[analysis]],
-  fittedLinearModelsExcludingControls[[analysis]],
-  fittedContrasts[[analysis]],
-  runs,
-  contrastGroups,
-  controlGroup,
-  proteinInfo)
-
-analysis <- "ExcludeMissingValues_ScaleNormalization"
-data <- combinedProteinData[[analysis]]
-exprs(data) <- log2(exprs(data))
-results2 <- createContrastResults(
-  data,
-  fittedLinearModels[[analysis]],
-  fittedLinearModelsExcludingControls[[analysis]],
-  fittedContrasts[[analysis]],
-  runs,
-  contrastGroups,
-  controlGroup,
-  proteinInfo)
-
-results <- inner_join(
-  results1 %>% select(Protein, B1 = B),
-  results2 %>% select(Protein, B2 = B),
-  by = "Protein"
-  ) %>%
-  filter(complete.cases(.)) %>%
-  mutate(rank1 = rank(-B1), rank2 = rank(-B2))
-```
-
-```{r scale_vs_quantile_normalization_B_plot, echo=FALSE, warning=FALSE, message=FALSE, fig.cap=paste("Plot of B-statistic (log odds) for differential expression of proteins for the ", contrastName, " contrast, comparing results following quantile and scale normalization.", sep = ""), fig.height=2.9, fig.width=3.5}
-print(scatterPlot(results, "B1", "B2", "B-statistic (quantile normalization)", "B-statistic (scale normalization)"))
-```
-
-```{r scale_vs_quantile_normalization_rank_plot, echo=FALSE, warning=FALSE, message=FALSE, fig.cap=paste("Plot of the rank of differentially-expressed proteins for the ", contrastName, " contrast, comparing results following quantile and scale normalization.", sep = ""), fig.height=2.9, fig.width=3.5}
-print(scatterPlot(results, "rank1", "rank2", "Rank (quantile normalization)", "Rank (scale normalization)"))
-```
-
-\newpage
-
-## Scale normalization based on all peptides vs peptides with highest IgG intensities (missing values excluded in both)
-
-```{r scale_normalization_all_vs_topIgG, include = FALSE}
-
-contrastGroups <- c("ER 45min", "ER 0min")
-contrastName <- paste(contrastGroups, collapse = " vs ")
-
-analysis <- "ExcludeMissingValues_ScaleNormalization"
-data <- combinedProteinData[[analysis]]
-exprs(data) <- log2(exprs(data))
-results1 <- createContrastResults(
-  data,
-  fittedLinearModels[[analysis]],
-  fittedLinearModelsExcludingControls[[analysis]],
-  fittedContrasts[[analysis]],
-  runs,
-  contrastGroups,
-  controlGroup,
-  proteinInfo)
-
-analysis <- "ExcludeMissingValues_ScaleNormalizationTopIgG"
-data <- combinedProteinData[[analysis]]
-exprs(data) <- log2(exprs(data))
-results2 <- createContrastResults(
-  data,
-  fittedLinearModels[[analysis]],
-  fittedLinearModelsExcludingControls[[analysis]],
-  fittedContrasts[[analysis]],
-  runs,
-  contrastGroups,
-  controlGroup,
-  proteinInfo)
-
-results <- inner_join(
-  results1 %>% select(Protein, B1 = B),
-  results2 %>% select(Protein, B2 = B),
-  by = "Protein"
-  ) %>%
-  filter(complete.cases(.)) %>%
-  mutate(rank1 = rank(-B1), rank2 = rank(-B2))
-```
-
-```{r scale_normalization_all_vs_topIgG_B_plot, echo=FALSE, warning=FALSE, message=FALSE, fig.cap=paste("Plot of B-statistic (log odds) for differential expression of proteins for the ", contrastName, " contrast, comparing results following scale normalization based on all peptides and those with the highest IgG intensities.", sep = ""), fig.height=2.5, fig.width=3.5}
-print(scatterPlot(results, "B1", "B2", "B-statistic (scale normalization based on all peptides)", "B-statistic (scale normalization, peptides with highest IgG)"))
-```
-
-```{r scale_normalization_all_vs_topIgG_rank_plot, echo=FALSE, warning=FALSE, message=FALSE, fig.cap=paste("Plot of the rank of differentially-expressed proteins for the ", contrastName, " contrast, comparing results following scale normalization based on all peptides and those with the highest IgG intensities.", sep = ""), fig.height=2.5, fig.width=3.5}
-print(scatterPlot(results, "rank1", "rank2", "Rank (scale normalization based on all peptides)", "Rank (scale normalization, peptides with highest IgG)"))
-```
-
-\newpage
-
-# Comparisoin of variance between specfic and non-specific binding
-
-```{r  echo=FALSE, warning=FALSE, message=FALSE}
-
-maPlot <- function(differentialExpressionResults, selectedGenes = NULL,
-                   xlab = expression(average~log[2](intensity)),
-                   ylab = expression(log[2](fold~change)),
-                   significanceLevel = 0.05,
-                   minLogFoldChangeForLabelling = 1.5,
-                   maxNumberLabelledProteins = 50,
-                   controlLogFoldChangeThreshold = -Inf,
-                   pointSize = 2.5) {
-  if (!"logFCcontrol" %in% colnames(differentialExpressionResults))
-    differentialExpressionResults$logFCcontrol <- Inf
-
-  differentialExpressionResults$group <-
-    ifelse(differentialExpressionResults$Gene %in% selectedGenes, 0,
-    ifelse(!is.na(differentialExpressionResults$adj.P.Val) & differentialExpressionResults$adj.P.Val <= significanceLevel & differentialExpressionResults$logFCcontrol >= controlLogFoldChangeThreshold, 1,
-    ifelse(!is.na(differentialExpressionResults$adj.P.Val) & differentialExpressionResults$adj.P.Val <= significanceLevel, 2,
-    ifelse(abs(differentialExpressionResults$logFC) >= minLogFoldChangeForLabelling & differentialExpressionResults$logFCcontrol >= controlLogFoldChangeThreshold, 3,
-    ifelse(abs(differentialExpressionResults$logFC) >= minLogFoldChangeForLabelling, 4,
-    ifelse(differentialExpressionResults$logFCcontrol >= controlLogFoldChangeThreshold, 5, 6))))))
-
-  # 0 - selected
-  # 1 - significant, specific
-  # 2 - significant, non-specific
-  # 3 - non-significant, large logfc, specific
-  # 4 - non-significant, large logfc, non-specific
-  # 5 - specific
-  # 6 - non-specific
-
-  differentialExpressionResults <- differentialExpressionResults %>% arrange(desc(group))
-
-  differentialExpressionResults$group <- factor(differentialExpressionResults$group, 0:6)
-
-  differentialExpressionResults <- differentialExpressionResults %>%
-    mutate(label = paste(" ", Gene)) %>%
-    mutate(logFCcontrolAlpha = pmin(logFCcontrol, 2.0))
-
-  labelSubset <- bind_rows(
-    differentialExpressionResults %>%
-      filter(group != 0 & adj.P.Val <= significanceLevel) %>%
-      arrange(desc(abs(logFC))) %>%
-      slice(1:(maxNumberLabelledProteins / 2)),
-    differentialExpressionResults %>%
-      filter(group != 0 & abs(logFC) >= minLogFoldChangeForLabelling) %>%
-      arrange(desc(abs(logFC))) %>%
-      slice(1:maxNumberLabelledProteins)
-  ) %>%
-    unique %>%
-    slice(1:maxNumberLabelledProteins)
-
-    if (maxNumberLabelledProteins == 0) labelSubset <- labelSubset %>% slice(0)
-
-  # plot <- ggplot(differentialExpressionResults, aes(x = AveExpr, y = logFC, colour = group, size = group, shape = group, alpha = logFCcontrolAlpha))
-  plot <- ggplot(differentialExpressionResults, aes(x = AveExpr, y = logFC, colour = group))+ coord_cartesian(ylim = c(-3.5,3.5))
-  plot <- plot + geom_hline(yintercept = 0, color="gray50", size = 0.5)
-  plot <- plot + geom_point(alpha=0.75)
-  plot <- plot + geom_text(data = labelSubset, aes(label = label), hjust = 0, size = 3.5)
-  plot <- plot + geom_point(data = subset(differentialExpressionResults, group == 0), alpha = 0.7)
-  plot <- plot + geom_text(data = subset(differentialExpressionResults, group == 0), aes(label = label), hjust = 0, size = 4.0, alpha = 1)
-  plot <- plot + scale_colour_manual(values = c("green", "red", "blue", "red", "blue", "red", "blue"), drop = FALSE)
-  # plot <- plot + scale_colour_manual(values = c("blue", "deeppink3", "deeppink3", "gray50", "gray50", "gray50", "gray50"), drop = FALSE)
-  # plot <- plot + scale_size_manual(values = c(1.2 * pointSize, pointSize, pointSize, 0.8 * pointSize, 0.8 * pointSize, 0.6 * pointSize, 0.6 * pointSize), drop = FALSE)
-  # plot <- plot + scale_shape_manual(values = c(16, 16, 1, 16, 1, 16, 1), drop = FALSE)
-  #plot <- plot + scale_alpha(range = c(0.2, 1.0))
-  plot <- plot + xlab(xlab)
-  plot <- plot + ylab(ylab)
-  plot <- plot + theme_bw()
-  plot <- plot + theme(
-    text = element_text(size = 14),
-    legend.position = "none"
-  )
-
-  save(differentialExpressionResults,file="Results/QRIME.RData") 
-  return(plot)
-}
-
-contrastGroups <- c("ER 45min", "ER 0min")
-contrastName <- paste(contrastGroups, collapse = " vs ")
-
-analysis <- "ExcludeMissingValues_QuantileNormalization"
-
-data <- combinedProteinData[[analysis]]
-exprs(data) <- log2(exprs(data))
-
-results <- createContrastResults(
-  data,
-  fittedLinearModels[[analysis]],
-  fittedLinearModelsExcludingControls[[analysis]],
-  fittedContrasts[[analysis]],
-  runs,
-  c("ER 45min", "ER 0min"),
-  controlGroup,
-  proteinInfo)
-
-```
-
-```{r MA_specificVsNonSpecfic, echo=FALSE, warning=FALSE, message=FALSE, fig.cap="MA plot showing the distribution of non-specific binding.
-Non-specific binding was identified as proteins that showed less than a log2 enrichment over the IgG channel at either 0 or 45 minutes.
-The majority of non-specific binding (blue) is found to show less change between time points whereas the specific interactions (red) shows a broader distribution.", fig.height=2.5, fig.width=3.5}
-
-print(maPlot(results,
-             significanceLevel = 0.5,
-             controlLogFoldChangeThreshold = 1.0,
-             minLogFoldChangeForLabelling = 1.5,
-             maxNumberLabelledProteins = 0,
-             pointSize = 1.5))
-```
-
-```{r  echo=FALSE, warning=FALSE, message=FALSE}
-
-load("Results/QRIME.RData")
-as.numeric(differentialExpressionResults$group)%%2->differentialExpressionResults$group
-
-tmp<-differentialExpressionResults
-
-```
-
-```{r density_specificVsNonSpecfici, echo=FALSE, warning=FALSE, message=FALSE, fig.cap="Density plot right showing the distribution of non-specific binding.
-Non-specific binding was identified as proteins that showed less than a log2 enrichment over the IgG channel at either 0 or 45 minutes.
-The majority of non-specific binding (blue) is found to show less change between time points whereas the specific interactions (red) shows a broader distribution.", fig.height=3.5, fig.width=4.5}
-
-plot(density(tmp$logFC[tmp$group==1]),col="blue",ylab="Density", xlab="Fold Change", main="", xlim=c(3,-3), lwd=2)
-lines(density(tmp$logFC[tmp$group==0]),col="red",lwd=2)
-
-```
-
-```{r box_specificVsNonSpecfici, echo=FALSE, warning=FALSE, message=FALSE, fig.cap=" Box plot showing the variance of non-specific vs non-specific binding.
-Non-specific binding was identified as proteins that showed less than a log2 enrichment over the IgG channel at either 0 or 45 minutes.
-The the variance of the log fold changes for the non-specific binding interactions was significantly less than that of the specific interactions (F-test, one-sided, p-value < 2.2e*-16).", fig.height=3.5, fig.width=4.5}
-
-boxplot(tmp$logFC[tmp$group==1],tmp$logFC[tmp$group==0],ylab="Fold Change", names=c("Non-specific","Specific"), col=c("#0000FFAA","#FF0000AA"))
-
-```
-
-```{r  echo=FALSE, warning=FALSE, message=FALSE}
-
-var.test(tmp$logFC[tmp$group==1],tmp$logFC[tmp$group==0], alternative="less")
-
-```
-
-\newpage
-
 # Differential binding results tables
@@
 
-The following tables contain the top ranking differentially expressed proteins for each comparison.
+The following tables contain the top ranking differentially bound proteins for each comparison.
 Included are all proteins that reach a statistical signficance of `r significanceLevel` in terms of the adjusted p-value and those with an absolute log~2~ fold change of `r logFoldChangeThreshold` or above.
@@
       display = c("d", "s", "s", "d", "f", "f", "g", "f", "f", "s"),
-      caption = paste("Top ranking differentially expressed proteins from the", contrastName, "comparison,  sorted by log2 fold change.")
+      caption = paste("Top ranking differentially bound proteins from the", contrastName, "comparison,  sorted by log2 fold change.")
     ),
@@
 ```{r results_table_contrast_1_config, include = FALSE}
-contrastGroups <- c("ER 45min", "ER 0min")
+
+contrastGroups <- strsplit2(contrast, " vs ")[1,]
 contrastName <- paste(contrastGroups, collapse = " vs ")
@@
 ```
-
-\newpage
-
-```{r results_table_contrast_2_config, include = FALSE}
-contrastGroups <- c("ER 90min", "ER 0min")
-contrastName <- paste(contrastGroups, collapse = " vs ")
-processingDescription <- "excluding peptides with missing intensities, quantile normalization"
-analysis <- "ExcludeMissingValues_QuantileNormalization"
-```
-
-## `r contrastName`
-
-```{r results_table_contrast_2, echo=FALSE, warning=FALSE, message=FALSE, results='asis'}
-
-data <- combinedProteinData[[analysis]]
-exprs(data) <- log2(exprs(data))
-
-results <- createContrastResults(
-  data,
-  fittedLinearModels[[analysis]],
-  fittedLinearModelsExcludingControls[[analysis]],
-  fittedContrasts[[analysis]],
-  runs,
-  contrastGroups,
-  controlGroup,
-  proteinInfo)
-
-contrastResultsTable(results, contrastName,
-                     significanceLevel = 0.05,
-                     logFoldChangeThreshold = 1.0,
-                     controlLogFoldChangeThreshold = 1.0)
-
-resultsFile <- paste("Results/",paste(gsub(" ", "_", contrastGroups), collapse = "_vs_"), analysis, ".txt", sep = "")
-write.table(results, resultsFile, sep = "\t", row.names = FALSE, quote = FALSE)
-```
-
-\newpage
-
-```{r results_table_contrast_3_config, include = FALSE}
-contrastGroups <- c("ER 90min", "ER 45min")
-contrastName <- paste(contrastGroups, collapse = " vs ")
-processingDescription <- "excluding peptides with missing intensities, quantile normalization"
-analysis <- "ExcludeMissingValues_QuantileNormalization"
-```
-
-## `r contrastName`
-
-```{r results_table_contrast_3, echo=FALSE, warning=FALSE, message=FALSE, results='asis'}
-
-data <- combinedProteinData[[analysis]]
-exprs(data) <- log2(exprs(data))
-
-results <- createContrastResults(
-  data,
-  fittedLinearModels[[analysis]],
-  fittedLinearModelsExcludingControls[[analysis]],
-  fittedContrasts[[analysis]],
-  runs,
-  contrastGroups,
-  controlGroup,
-  proteinInfo)
-
-contrastResultsTable(results, contrastName,
-                     significanceLevel = 0.05,
-                     logFoldChangeThreshold = 1.0,
-                     controlLogFoldChangeThreshold = 1.0)
-
-resultsFile <- paste("Results/",paste(gsub(" ", "_", contrastGroups), collapse = "_vs_"), analysis, ".txt", sep = "")
-write.table(results, resultsFile, sep = "\t", row.names = FALSE, quote = FALSE)
-```
-
-\newpage
-
-```{r results_table_contrast_4_config, include = FALSE}
-contrastGroups <- c("FOXA1 45min", "FOXA1 0min")
-contrastName <- paste(contrastGroups, collapse = " vs ")
-processingDescription <- "excluding peptides with missing intensities, quantile normalization"
-analysis <- "ExcludeMissingValues_QuantileNormalization"
-```
-
-## `r contrastName`
-
-```{r results_table_contrast_4, echo=FALSE, warning=FALSE, message=FALSE, results='asis'}
-
-data <- combinedProteinData[[analysis]]
-exprs(data) <- log2(exprs(data))
-
-results <- createContrastResults(
-  data,
-  fittedLinearModels[[analysis]],
-  fittedLinearModelsExcludingControls[[analysis]],
-  fittedContrasts[[analysis]],
-  runs,
-  contrastGroups,
-  controlGroup,
-  proteinInfo)
-
-contrastResultsTable(results, contrastName,
-                     significanceLevel = 0.05,
-                     logFoldChangeThreshold = 1.0,
-                     controlLogFoldChangeThreshold = 1.0)
-
-resultsFile <- paste("Results/",paste(gsub(" ", "_", contrastGroups), collapse = "_vs_"), analysis, ".txt", sep = "")
-write.table(results, resultsFile, sep = "\t", row.names = FALSE, quote = FALSE)
-```
-
-\newpage
-
-```{r results_table_contrast_5_config, include = FALSE}
-contrastGroups <- c("FOXA1 90min", "FOXA1 0min")
-contrastName <- paste(contrastGroups, collapse = " vs ")
-processingDescription <- "excluding peptides with missing intensities, quantile normalization"
-analysis <- "ExcludeMissingValues_QuantileNormalization"
-```
-
-## `r contrastName`
-
-```{r results_table_contrast_5, echo=FALSE, warning=FALSE, message=FALSE, results='asis'}
-
-data <- combinedProteinData[[analysis]]
-exprs(data) <- log2(exprs(data))
-
-results <- createContrastResults(
-  data,
-  fittedLinearModels[[analysis]],
-  fittedLinearModelsExcludingControls[[analysis]],
-  fittedContrasts[[analysis]],
-  runs,
-  contrastGroups,
-  controlGroup,
-  proteinInfo)
-
-contrastResultsTable(results, contrastName,
-                     significanceLevel = 0.05,
-                     logFoldChangeThreshold = 1.0,
-                     controlLogFoldChangeThreshold = 1.0)
-
-resultsFile <- paste("Results/",paste(gsub(" ", "_", contrastGroups), collapse = "_vs_"), analysis, ".txt", sep = "")
-write.table(results, resultsFile, sep = "\t", row.names = FALSE, quote = FALSE)
-```
-
-\newpage
-
-```{r results_table_contrast_6_config, include = FALSE}
-contrastGroups <- c("FOXA1 90min", "FOXA1 45min")
-contrastName <- paste(contrastGroups, collapse = " vs ")
-processingDescription <- "excluding peptides with missing intensities, quantile normalization"
-analysis <- "ExcludeMissingValues_QuantileNormalization"
-```
-
-## `r contrastName`
-
-```{r results_table_contrast_6, echo=FALSE, warning=FALSE, message=FALSE, results='asis'}
-
-data <- combinedProteinData[[analysis]]
-exprs(data) <- log2(exprs(data))
-
-results <- createContrastResults(
-  data,
-  fittedLinearModels[[analysis]],
-  fittedLinearModelsExcludingControls[[analysis]],
-  fittedContrasts[[analysis]],
-  runs,
-  contrastGroups,
-  controlGroup,
-  proteinInfo)
-
-contrastResultsTable(results, contrastName,
-                     significanceLevel = 0.05,
-                     logFoldChangeThreshold = 1.0,
-                     controlLogFoldChangeThreshold = 1.0)
-
-resultsFile <- paste("Results/", paste(gsub(" ", "_", contrastGroups), collapse = "_vs_"), analysis, ".txt", sep = "")
-write.table(results, resultsFile, sep = "\t", row.names = FALSE, quote = FALSE)
-```
-
-```{r save, eval = TRUE, include = FALSE, cache = FALSE}
-save.image(file = "Results/QRIME.RData")
-```

### 2018-01-23 ChIP-seq_ER/rmarkdown.Rmd
first commit

---
title: "Analysis of the time-dependent response to Estradiol through ChIP-Seq data and the VULCAN package"
author: "Federico M. Giorgi"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document: 
    fig_caption: yes
    fig_height: 8
    fig_width: 8
    toc: true # turns on table of contents
    toc_depth: 3  # up to three depths of headings (specified by #, ## and ###)
    number_sections: true  # if you want number sections at each table header
    theme: united
    highlight: tango
---

```{r setup_vulcan, include=FALSE}
knitr::opts_chunk$set(echo=TRUE)
knitr::opts_chunk$set(fig.align="center")
knitr::opts_chunk$set(fig.path = "rmarkdown_images/")
par.backup<-par()
```

```{r captions, include=FALSE}
set.seed(1)
#Figure and Table Caption Numbering, for HTML do it manually
captabNo <- 1; capfigNo <- 1;

#Function to add the Table Number
captab <- function(x){
  x <- paste0("Table S",captabNo,". ",x)
  captabNo <<- captabNo + 1
  x
}

#Function to add the Figure Number
capfig <- function(x){
  #x <- paste0("Figure S",capfigNo,". ",x)
  #capfigNo <<- capfigNo + 1
  x
}
```

## Introduction
Vulcan (VirtUaL ChIP-Seq Analysis through Networks) is a pipeline that combines
ChIP-Seq data and regulatory networks to obtain transcription factors that are
likely affected by a specific stimulus. In order to do so, our package combines
strategies from different BioConductor packages: \emph{DESeq} for data
normalization, \emph{ChIPpeakAnno} and \emph{DiffBind} for annotation and
definition of ChIP-Seq genomic peaks, \emph{csaw} to define optimal peak width
and \emph{viper} for applying a regulatory network over a differential binding
signature.
Usage of gene regulatory networks to analyze biological systems has witnessed
an exponential increase in the last decade, due to the ease of obtaining
genome-wide expression data (Margolin et al., 2005; Giorgi et al., 2013; Castro
et al., 2015). Recently, the VIPER approach to interrogate these network models
has been proposed to infer transcription factor activity using the expression
of a collection of their putative targets, i.e. their regulon (Alvarez et al.,
2016). In the VIPER algorithm, gene-level differential expression signatures
are obtained for either individual samples (relative to the mean of the dataset)
or between groups of samples, and regulons are tested for enrichment. Ideally,
TF regulons can be tested on promoter-level differential binding signatures
generated from ChIP-Seq experiments, in order to ascertain the global change
in promoter occupancy for gene sets. In our study, we propose an extension
of the VIPER algorithm to specifically analyze TF occupancy in ChIP-Seq
experiments. Our VULCAN algorithm uses ChIP-Seq data obtained for a given TF
to provide candidate coregulators of the response to a given stimulus.
The analysis is based on identifying differentially bound genes and testing
their enrichment in the regulon of potential co-regulatory factors.

## Software setup
VULCAN is conveniently provided as an R package available from the Bioconductor repository.
```{r, message=FALSE, warning=FALSE}
#source("http://bioconductor.org/biocLite.R")
#biocLite("vulcan")
library(vulcan)
```
Other packages are required for the execution of the analysis and the visualization of the results. A \textit{results} folder will be also created to store intermediate steps of the analysis.
```{r, message=FALSE, warning=FALSE}
library(gplots) # for heatmap.2
library(org.Hs.eg.db)
library(gridExtra) # for the pathway part
library(GeneNet) # to generate partial correlation networks
if(!file.exists("results")){dir.create("results")}
```
Finally, we will define here some convenience functions, such as those to convert from gene symbols to entrez ids, and viceversa.
```{r, message=FALSE, warning=FALSE}
list_eg2symbol<-as.list(org.Hs.egSYMBOL[mappedkeys(org.Hs.egSYMBOL)])
e2s<-function(ids){
  ids <- as.character(ids)
  outlist <- list_eg2symbol[ids]
  names(outlist) <- ids
  outlist[is.na(outlist)] <- paste("unknown.", ids[is.na(outlist)], sep = "")
  outlist <- gsub("unknown.unknown.", "", outlist)
  return(outlist)
}
list_symbol2eg <- as.character(org.Hs.egSYMBOL2EG[mappedkeys(org.Hs.egSYMBOL2EG)])
s2e<-function(ids){
  ids <- as.character(ids)
  outlist <- list_symbol2eg[ids]
  names(outlist) <- ids
  outlist[is.na(outlist)] <- paste("unknown.", ids[is.na(outlist)], sep = "")
  outlist <- gsub("unknown.unknown.", "", outlist)
  return(outlist)
}
```

## Import ChIP-Seq data using Vulcan
ChIP-Seq data was generated using a cell line model of ER+ breast cancer, MCF7, at 0’, 45’ and 90’ after stimulation with estradiol. The binding profile of ER at each timepoint was then compared between time points using differential binding analysis. From the temporal comparison ER binding we established four classes of binding pattern: early responders, repressed transcription factors, late responders and candidate cyclic genes.

The first part of our analysis highlights how to import ChIP-Seq data using the VULCAN adn DiffBind packages, provided alignment files (BAM format) and peak files (BED format) for each of the samples. In our dataset, we have 4 replicates for each time point.

### Initial import
VULCAN requires an input sheet file in CSV format describing the samples and the location of the individual input files
```{r, message=FALSE, warning=FALSE}
sheetfile<-"chipseq/holding/sheet.csv"

fname<-"results/001_input.rda"
if(!file.exists(fname)){
  vobj<-vulcan.import(sheetfile)
  save(vobj,file=fname)
} else {
  load(fname)
}
```

### Annotation
VULCAN identifies the location of each peak in each sample, and according to the selected method, assigns a score to each gene promoter. There are a few methods available.

* \textbf{\textit{sum}}: when multiple peaks are found, sum their contributions
* \textbf{\textit{closest}}: when multiple peaks are found, keep only the closest to the TSS as the representative one
* \textbf{\textit{strongest}}: when multiple peaks are found, keep the strongest as the representative one
* \textbf{\textit{farthest}}: when multiple peaks are found, keep only the closest to the TSS as the representative one
* \textbf{\textit{topvar}}: when multiple peaks are found, keep the most varying as the representative one
* \textbf{\textit{lowvar}}: when multiple peaks are found, keep the least varying as the representative one

```{r label, message=FALSE, warning=FALSE}
vobj<-vulcan.annotate(vobj,lborder=-10000,rborder=10000,method="sum")
```

### Normalization
At this step, genes are quantified according to the number of reads that could be associated to their promoters. The algorithms within VULCAN (\textit{viper} and \textit{DESeq2}) require however data to be normalized via Variance-Stabilizing Transformation (Anders and Huber, 2010).

```{r, message=FALSE, warning=FALSE}
vobj<-vulcan.normalize(vobj)
save(vobj,file="results/001_vobj.rda")
```

### Dataset Sample Clustering
Here, we show how the samples cluster together using peak raw counts, VST-normalized peak raw counts and peak RPKMs. In the following heatmaps, there are `r nrow(vobj$rawcounts)` rows, one per gene promoter, and 16 columns, one per sample. The sample name indicates the time point in minutes (T0, T45 or T90) and the number of replicate (_1, _2, _3, _4).
```{r, fig.cap=capfig('Dataset clustering with Raw Counts'), message=FALSE, warning=FALSE}
heatmap.2(vobj$rawcounts,scale="row",
          col=colorpanel(1000,"navy","white","red3"),tracecol="black",labRow="")
mtext("Raw Counts")
```

```{r, fig.cap=capfig('Dataset clustering with VST-normalized Counts'), message=FALSE, warning=FALSE}
heatmap.2(vobj$normalized,
          col=colorpanel(1000,"navy","white","red3"),tracecol="black",labRow="")
mtext("VST-normalized")
```

```{r, fig.cap=capfig('Dataset clustering with RPKMs'), message=FALSE, warning=FALSE}
heatmap.2(vobj$rpkms,scale="row",
          col=colorpanel(1000,"navy","white","red3"),tracecol="black",labRow="")
mtext("RPKMs")
```

Principal Component Analysis further confirms the clustering of replicates in two distinct groups: untreated (T0) and treated (T45/T90).
```{r, fig.cap=capfig('Principal Component Analysis of the dataset, highlighting components 1 and 2'), message=FALSE, warning=FALSE,out.width='100%',dev.args=list(pointsize=16)}
topvar<-apply(vobj$normalized[,],1,var)
topvar<-sort(topvar,dec=TRUE)[1:500]
submat<-vobj$normalized[names(topvar),]
pca<-prcomp(t(submat))
vars<-100*(pca$sdev^2)/sum(pca$sdev^2)
vars<-signif(vars,3)

p1<-1
p2<-2
plot(pca$x[,p1],pca$x[,p2],
     xlab=paste0("PC",p1," (Var.Explained: ",vars[p1],"%)"),
     ylab=paste0("PC",p2," (Var.Explained: ",vars[p2],"%)"),
     type="p",main="PC Analysis",pch=20,col="grey",
     xlim=c(min(pca$x[,p1])*1.1,max(pca$x[,p1])*1.1)
)
mtext("VST-Normalized data",cex=0.8)
textplot2(pca$x[,p1],pca$x[,p2],new=FALSE,
          words=gsub("_[0-9]","",rownames(pca$x),perl=TRUE)
)
grid()
```

A clearer distinction of T45 and T90 is highlighted by PC5:
```{r, fig.cap=capfig('Principal Component Analysis of the dataset, highlighting components 1 and 5'), message=FALSE, warning=FALSE,out.width='100%',dev.args=list(pointsize=16)}
p1<-1
p2<-5
plot(pca$x[,p1],pca$x[,p2],
     xlab=paste0("PC",p1," (Var.Explained: ",vars[p1],"%)"),
     ylab=paste0("PC",p2," (Var.Explained: ",vars[p2],"%)"),
     type="p",main="PC Analysis",pch=20,col="grey",
     xlim=c(min(pca$x[,p1])*1.1,max(pca$x[,p1])*1.1)
)
mtext("VST-Normalized data",cex=0.8)
textplot2(pca$x[,p1],pca$x[,p2],new=FALSE,
          words=gsub("_[0-9]","",rownames(pca$x),perl=TRUE)
)
grid()
```

### MA plots
Without changing the format of the input data, we can use the Bioconductor DiffBind package to visualize the amplitude of changes in ER binding between time points. One way to do this is an MA plot, which shows the differences between measurements taken in two groups (e.g. 45' vs 00'), by transforming the promoter peak intensity data onto M (log ratio) and A (mean average) scales.

We will process the data using the DiffBind package and then use the \textit{dba.plotMA} function to visualize the contrasts (which we can extract using the \textit{dba.contrast} function).
```{r, message=FALSE, warning=FALSE}
sheetfile<-"chipseq/holding/sheet.csv"
fname<-"results/001_diffbind.rda"
if(!file.exists(fname)){
  # Load a sample sheet
  chipseqSamples<-read.csv(sheetfile)
  dbaobj<-dba(sampleSheet=chipseqSamples)

  # Count reads 500bp either side of summits, a peak is considered a peak
  # if it is found in 3 samples out of loaded samples.
  dbaobj<-dba.count(dbaobj,minOverlap=3,summits=500)

  # Define contrasts
  dbaobj<-dba.contrast(dbaobj, categories=DBA_CONDITION)

  # Calculate differential binding using the DESeq2 engine within DiffBind
  dbaobj<-dba.analyze(dbaobj,method=DBA_DESEQ2)

  # Save the DiffBind object
  save(dbaobj,file=fname)
} else {
  load(fname)
}

# Visualize each contrast
dba.contrast(dbaobj)
```
```{r, fig.cap=capfig('MA plot for Contrast 90mins vs 45mins, highlighting each individual peak. Significant peaks are highlighted in red'), message=FALSE, warning=FALSE}
dba.plotMA(dbaobj,contrast=1,method=DBA_DESEQ2)
```
```{r, fig.cap=capfig('MA plot for Contrast 90mins vs 00mins, highlighting each individual peak. Significant peaks are highlighted in red'), message=FALSE, warning=FALSE}
dba.plotMA(dbaobj,contrast=2,method=DBA_DESEQ2)
```
```{r, fig.cap=capfig('MA plot for Contrast 45mins vs 00mins, highlighting each individual peak. Significant peaks are highlighted in red'), message=FALSE, warning=FALSE}
dba.plotMA(dbaobj,contrast=3,method=DBA_DESEQ2)
```

## Apply a coregulatory network over a ChIP-Seq profile
Once the data has been loaded, VULCAN applies a regulatory network over differential binding signatures to define Transcription Factors whose networks are most affected by the treatment. In our analysis, we will be using three different networks generated via ARACNe (Margolin et al., 2006): two are breast cancer-specific and are derived from the TCGA and METABRIC data collection, respectively. A third network is used as a negative control, highlighting regulatory mechanisms derived from the Amyloid Leukemia dataset (TCGA).
### Loading the input networks and processed data
First, we will load the output from the previous paragraphs (chipseq data annotated and normalized) and the three networks.
```{r, message=FALSE, warning=FALSE}
# Load imported vulcan object
load("results/001_vobj.rda")

# Vulcan Analysis (multiple networks)
## Networks
load("networks/laml-tf-regulon.rda")
laml_regulon<-regul
rm(regul)

load("networks/brca-tf-regulon.rda")
tcga_regulon<-regul
rm(regul)

load("networks/metabric-regulon-tfs.rda")
metabric_regulon<-regulon
rm(regulon)
```
Accessing the vulcan object /texit{vobj} can give us informations on the sample groups thereby contained.
```{r, warning=FALSE}
names(vobj$samples)
```
### Running the core Vulcan function
The final Vulcan pipeline step requires three input objects:

* The annotated and normalized chipseq data (/texit{vobj})
* A specific binding signature defined by a contrast between two sample groups
* A regulatory network

```{r, message=FALSE, warning=FALSE}
fname<-"results/002_vobj_networks.rda"
list_eg2symbol<-as.list(org.Hs.egSYMBOL[mappedkeys(org.Hs.egSYMBOL)]) 
if(!file.exists(fname)){
  vobj_tcga_90<-vulcan(vobj,network=tcga_regulon,contrast=c("t90","t0"),annotation=list_eg2symbol)
  vobj_tcga_45<-vulcan(vobj,network=tcga_regulon,contrast=c("t45","t0"),annotation=list_eg2symbol)
  vobj_metabric_90<-vulcan(vobj,network=metabric_regulon,contrast=c("t90","t0"),annotation=list_eg2symbol)
  vobj_metabric_45<-vulcan(vobj,network=metabric_regulon,contrast=c("t45","t0"),annotation=list_eg2symbol)
  vobj_negative_90<-vulcan(vobj,network=laml_regulon,contrast=c("t90","t0"),annotation=list_eg2symbol)
  vobj_negative_45<-vulcan(vobj,network=laml_regulon,contrast=c("t45","t0"),annotation=list_eg2symbol)
  save(
    vobj_tcga_90,
    vobj_tcga_45,
    vobj_metabric_90,
    vobj_metabric_45,
    vobj_negative_90,
    vobj_negative_45,
    file=fname
  )
} else {
  load(fname)
}
```
### Visualize the relative TF activity
Vulcan has now generated activity scores for each Transcription Factor, which specify the global ER binding strength on the promoters of their targets.

#### METABRIC results
The following line plot shows the relative network activity for every TF at two time points. On the y-axis, the Normalized Enrichment Score of Vulcan is provided. Dashed lines indicate a p=0.05 significance threshold for up- and down-regulation.
```{r, fig.cap=capfig("Global TF activity after Estradiol Treatment in MCF7 cells, inferred using the METABRIC network, highlighting the ESR1 TF as an example"), message=FALSE, warning=FALSE,dev.args=list(pointsize=14)}
threshold<-p2z(0.05)
metabric_90=vobj_metabric_90$mrs[,"NES"]
metabric_45=vobj_metabric_45$mrs[,"NES"]
tfs<-names(metabric_45)
metabricmat<-cbind(rep(0,length(metabric_45)),metabric_45[tfs],metabric_90[tfs])
colnames(metabricmat)<-c("T0","T45","T90")
## All TFs
matplot(t(metabricmat),type="l",col="grey",ylab="VULCAN NES",xaxt="n",lty=1,main="All TFs",xlim=c(1,3.3),ylim=c(min(metabricmat),max(metabricmat)))
axis(1,at=c(1:3),labels=colnames(metabricmat))
abline(h=c(0,threshold,-threshold),lty=2)
matplot(t(t(metabricmat["ESR1",])),type="l",col="red",lty=1,lwd=2,add=TRUE)
text(3,metabricmat["ESR1",3],label="ESR1",pos=4,cex=0.6,font=2,col="red3")
mtext("METABRIC network")
```
These TFs can be grouped in classes. For example, TFs whose activity is already significant after 45mins and remains significant at 90 minutes after estradiol treatment can be dubbed **early responders**.
```{r, fig.cap=capfig("Global TF activity after Estradiol Treatment in MCF7 cells, inferred using the METABRIC network, highlighting TFs significantly upregulated at 45 minutes and 90 minutes"), message=FALSE, warning=FALSE,dev.args=list(pointsize=14)}
threshold<-p2z(0.05)
tfclass<-tfs[metabricmat[,"T45"]>=threshold&metabricmat[,"T90"]>=threshold]
matplot(t(metabricmat),type="l",col="grey",ylab="VULCAN NES",xaxt="n",lty=1,main="Early responders",xlim=c(1,3.3),ylim=c(0,max(metabricmat)))
axis(1,at=c(1:3),labels=colnames(metabricmat))
abline(h=c(0,threshold,-threshold),lty=2)
matplot(t(metabricmat[tfclass,]),type="l",col="red3",lty=1,lwd=2,add=TRUE)
# Repel a bit
plabels<-tfclass
oricoords<-sort(setNames(metabricmat[plabels,3],plabels))
newcoords<-setNames(seq(min(oricoords),max(oricoords),length.out=length(oricoords)),names(oricoords))
text(3,newcoords,label=names(oricoords),pos=4,cex=0.6,font=2)
text(3,newcoords["ESR1"],label="ESR1",pos=4,cex=0.6,font=2,col="red3")
mtext("METABRIC network")
```
TFs whose *repressed* targets are bound will yield a negative activity score. These **repressed TFs** are symmetrically opposite to early responder TFs.
```{r, fig.cap=capfig("Global TF activity after Estradiol Treatment in MCF7 cells, inferred using the METABRIC network, highlighting TFs significantly downregulated at 45 minutes and 90 minutes"), message=FALSE, warning=FALSE,dev.args=list(pointsize=14)}
threshold<-p2z(0.05)
tfclass<-tfs[metabricmat[,"T45"]<=-threshold&metabricmat[,"T90"]<=-threshold]
matplot(t(metabricmat),type="l",col="grey",ylab="VULCAN NES",xaxt="n",lty=1,main="Repressed TFs",xlim=c(1,3.3),ylim=c(min(metabricmat),0))
axis(1,at=c(1:3),labels=colnames(metabricmat))
abline(h=c(0,threshold,-threshold),lty=2)
matplot(t(metabricmat[tfclass,]),type="l",col="cornflowerblue",lty=1,lwd=2,add=TRUE)
# Repel a bit
plabels<-c(tfclass,"GRHL2")
oricoords<-sort(setNames(metabricmat[plabels,3],plabels))
newcoords<-setNames(seq(min(oricoords),max(oricoords),length.out=length(oricoords)),names(oricoords))
text(3,newcoords,label=names(oricoords),pos=4,cex=0.6,font=2)
mtext("METABRIC network")
```
Some TFs appear to have their targets bound at 45 minutes, but then unoccupied at 90 minutes. This "updown" behavior is consistent to what observed in previous literature about the cyclic properties of certain components of the ER DNA-binding complex, and therefore we dubbed them **candidate cyclic TFs**.
```{r, fig.cap=capfig("Global TF activity after Estradiol Treatment in MCF7 cells, inferred using the METABRIC network, highlighting TFs significantly upregulated at 45 minutes but not at 90 minutes"), message=FALSE, warning=FALSE,dev.args=list(pointsize=14)}
threshold<-p2z(0.05)
tfclass<-tfs[metabricmat[,"T45"]>=threshold&metabricmat[,"T90"]<=threshold]
matplot(t(metabricmat),type="l",col="grey",ylab="VULCAN NES",xaxt="n",lty=1,main="Candidate Cyclic TFs",xlim=c(1,3.3),ylim=c(min(metabricmat),max(metabricmat)))
axis(1,at=c(1:3),labels=colnames(metabricmat))
abline(h=c(0,threshold,-threshold),lty=2)
matplot(t(metabricmat[tfclass,]),type="l",col="seagreen",lty=1,lwd=2,add=TRUE)
# Repel a bit
plabels<-tfclass
oricoords<-sort(setNames(metabricmat[plabels,3],plabels))
newcoords<-setNames(seq(min(oricoords),max(oricoords),length.out=length(oricoords)),names(oricoords))
text(3,newcoords,label=names(oricoords),pos=4,cex=0.6,font=2)
mtext("METABRIC network")
```
Finally, a category of TFs appear to be activated but at 90 minutes only. We call this category of TFs **late responders**.
```{r, fig.cap=capfig("Global TF activity after Estradiol Treatment in MCF7 cells, inferred using the METABRIC network, highlighting TFs significantly upregulated at 90 minutes but not at 45 minutes"), message=FALSE, warning=FALSE,dev.args=list(pointsize=14)}
threshold<-p2z(0.05)
tfclass<-tfs[metabricmat[,"T45"]<=threshold&metabricmat[,"T45"]>0&metabricmat[,"T90"]>=threshold]
matplot(t(metabricmat),type="l",col="grey",ylab="VULCAN NES",xaxt="n",lty=1,main="Late responders",xlim=c(1,3.3),ylim=c(0,max(metabricmat)))
axis(1,at=c(1:3),labels=colnames(metabricmat))
abline(h=c(0,threshold,-threshold),lty=2)
matplot(t(metabricmat[tfclass,]),type="l",col="darkorange",lty=1,lwd=2,add=TRUE)
# Repel a bit
plabels<-tfclass
oricoords<-sort(setNames(metabricmat[plabels,3],plabels))
newcoords<-setNames(seq(min(oricoords),min(oricoords)*1.2,length.out=length(oricoords)),names(oricoords))
text(3,newcoords,label=names(oricoords),pos=4,cex=0.6,font=2)
mtext("METABRIC network")
```

#### TCGA results
All the TFs and the four categories of TFs were inferred also using the TCGA network.

```{r, fig.cap=capfig("Global TF activity after Estradiol Treatment in MCF7 cells, inferred using the TCGA network, highlighting the ESR1 TF as an example"), message=FALSE, warning=FALSE,dev.args=list(pointsize=14)}
tcga_90=vobj_tcga_90$mrs[,"NES"]
tcga_45=vobj_tcga_45$mrs[,"NES"]
tfs<-names(tcga_45)
tcgamat<-cbind(rep(0,length(tcga_45)),tcga_45[tfs],tcga_90[tfs])
colnames(tcgamat)<-c("T0","T45","T90")
## All TFs
matplot(t(tcgamat),type="l",col="grey",ylab="VULCAN NES",xaxt="n",lty=1,main="All TFs",xlim=c(1,3.3),ylim=c(min(tcgamat),max(tcgamat)))
axis(1,at=c(1:3),labels=colnames(tcgamat))
abline(h=c(0,threshold,-threshold),lty=2)
matplot(t(t(tcgamat["ESR1",])),type="l",col="red",lty=1,lwd=2,add=TRUE)
text(3,tcgamat["ESR1",3],label="ESR1",pos=4,cex=0.6,font=2,col="red3")
mtext("TCGA network")
```

```{r, fig.cap=capfig("Global TF activity after Estradiol Treatment in MCF7 cells, inferred using the TCGA network, highlighting TFs significantly upregulated at 45 minutes and 90 minutes"), message=FALSE, warning=FALSE,dev.args=list(pointsize=14)}
threshold<-p2z(0.05)
tfclass<-tfs[tcgamat[,"T45"]>=threshold&tcgamat[,"T90"]>=threshold]
matplot(t(tcgamat),type="l",col="grey",ylab="VULCAN NES",xaxt="n",lty=1,main="Early responders",xlim=c(1,3.3),ylim=c(0,max(tcgamat)))
axis(1,at=c(1:3),labels=colnames(tcgamat))
abline(h=c(0,threshold,-threshold),lty=2)
matplot(t(tcgamat[tfclass,]),type="l",col="red3",lty=1,lwd=2,add=TRUE)
# Repel a bit
plabels<-tfclass
oricoords<-sort(setNames(tcgamat[plabels,3],plabels))
newcoords<-setNames(seq(min(oricoords),max(oricoords),length.out=length(oricoords)),names(oricoords))
text(3,newcoords,label=names(oricoords),pos=4,cex=0.6,font=2)
text(3,newcoords["ESR1"],label="ESR1",pos=4,cex=0.6,font=2,col="red3")
mtext("TCGA network")
```

```{r, fig.cap=capfig("Global TF activity after Estradiol Treatment in MCF7 cells, inferred using the TCGA network, highlighting TFs significantly downregulated at 45 minutes and 90 minutes"), message=FALSE, warning=FALSE,dev.args=list(pointsize=14)}
threshold<-p2z(0.05)
tfclass<-tfs[tcgamat[,"T45"]<=-threshold&tcgamat[,"T90"]<=-threshold]
matplot(t(tcgamat),type="l",col="grey",ylab="VULCAN NES",xaxt="n",lty=1,main="Repressed TFs",xlim=c(1,3.3),ylim=c(min(tcgamat),0))
axis(1,at=c(1:3),labels=colnames(tcgamat))
abline(h=c(0,threshold,-threshold),lty=2)
matplot(t(tcgamat[tfclass,]),type="l",col="cornflowerblue",lty=1,lwd=2,add=TRUE)
# Repel a bit
plabels<-tfclass
oricoords<-sort(setNames(tcgamat[plabels,3],plabels))
newcoords<-setNames(seq(min(oricoords),max(oricoords),length.out=length(oricoords)),names(oricoords))
text(3,newcoords,label=names(oricoords),pos=4,cex=0.6,font=2)
mtext("TCGA network")
```

```{r, fig.cap=capfig("Global TF activity after Estradiol Treatment in MCF7 cells, inferred using the TCGA network, highlighting TFs significantly upregulated at 45 minutes but not at 90 minutes"), message=FALSE, warning=FALSE,dev.args=list(pointsize=14)}
threshold<-p2z(0.05)
tfclass<-tfs[tcgamat[,"T45"]>=threshold&tcgamat[,"T90"]<=threshold]
matplot(t(tcgamat),type="l",col="grey",ylab="VULCAN NES",xaxt="n",lty=1,main="Candidate Cyclic TFs",xlim=c(1,3.3),ylim=c(0,max(tcgamat)))
axis(1,at=c(1:3),labels=colnames(tcgamat))
abline(h=c(0,threshold,-threshold),lty=2)
matplot(t(tcgamat[tfclass,]),type="l",col="seagreen",lty=1,lwd=2,add=TRUE)
# Repel a bit
plabels<-tfclass
oricoords<-sort(setNames(tcgamat[plabels,3],plabels))
newcoords<-setNames(seq(min(oricoords),max(oricoords),length.out=length(oricoords)),names(oricoords))
text(3,newcoords,label=names(oricoords),pos=4,cex=0.6,font=2)
mtext("TCGA network")
```

```{r, fig.cap=capfig("Global TF activity after Estradiol Treatment in MCF7 cells, inferred using the TCGA network, highlighting TFs significantly upregulated at 90 minutes but not at 45 minutes"), message=FALSE, warning=FALSE,dev.args=list(pointsize=14)}
## Dn-Up TF class
threshold<-p2z(0.05)
tfclass<-tfs[tcgamat[,"T45"]<=threshold&tcgamat[,"T45"]>0&tcgamat[,"T90"]>=threshold]
matplot(t(tcgamat),type="l",col="grey",ylab="VULCAN NES",xaxt="n",lty=1,main="Late responders",xlim=c(1,3.3),ylim=c(0,max(tcgamat)))
axis(1,at=c(1:3),labels=colnames(tcgamat))
abline(h=c(0,threshold,-threshold),lty=2)
matplot(t(tcgamat[tfclass,]),type="l",col="darkorange",lty=1,lwd=2,add=TRUE)
# Repel a bit
plabels<-tfclass
oricoords<-sort(setNames(tcgamat[plabels,3],plabels))
newcoords<-setNames(seq(min(oricoords),min(oricoords)*1.2,length.out=length(oricoords)),names(oricoords))
text(3,newcoords,label=names(oricoords),pos=4,cex=0.6,font=2)
mtext("TCGA network")
```

#### Comparing TCGA and METABRIC results
The following scatterplots compare the activity inferred by VULCAN using two alternative networks, derived from the TCGA and METABRIC breast cancer datasets using ARACNe-AP (Giorgi et al., 2016) with default parameters.
```{r, fig.cap=capfig("Comparing TF activities inferred by VULCAN using two different breast cancer dataset-derived networks on the ER 45 minutes signature. PCC indicates the Pearson Correlation Coefficient"), message=FALSE, warning=FALSE,dev.args=list(pointsize=16)}
common<-intersect(rownames(tcgamat),rownames(metabricmat))
set.seed(1)
x<-tcgamat[common,"T45"]
y<-metabricmat[common,"T45"]
plot(x,y,xlab="TCGA, 45mins",ylab="METABRIC, 45mins",pch=20,col="grey",xlim=c(min(x)*1.2,max(x)*1.2))
grid()
#toshow<-c("ESR1","GATA3","RARA","HSF1")
toshow<-names(which(rank(rank(-abs(x))+rank(-abs(y)))<=20))
#toshow<-unique(c(toshow,names(sort(rank(-abs(x))[1:10])),names(sort(rank(-abs(y)))[1:10])))
textplot2(x[toshow],y[toshow],words=toshow,new=FALSE,cex=0.8)
lm1<-lm(y~x)
abline(lm1$coef)
pcc<-cor.test(x,y)
mtext(paste0("PCC=",signif(pcc$estimate,2)," (p=",signif(pcc$p.value,2),")"))
```
```{r, fig.cap=capfig("Comparing TF activities inferred by VULCAN using two different breast cancer dataset-derived networks on the ER 90 minutes signature. PCC indicates the Pearson Correlation Coefficient"), message=FALSE, warning=FALSE,dev.args=list(pointsize=16)}
x<-tcgamat[common,"T90"]
y<-metabricmat[common,"T90"]
plot(x,y,xlab="TCGA, 90mins",ylab="METABRIC, 90mins",pch=20,col="grey",xlim=c(min(x)*1.2,max(x)*1.2))
grid()
#toshow<-c("ESR1","GATA3","RARA","HSF1")
toshow<-names(which(rank(rank(-abs(x))+rank(-abs(y)))<=20))
#toshow<-unique(c(toshow,names(sort(rank(-abs(x))[1:10])),names(sort(rank(-abs(y)))[1:10])))
textplot2(x[toshow],y[toshow],words=toshow,new=FALSE,cex=0.8)
lm1<-lm(y~x)
abline(lm1$coef)
pcc<-cor.test(x,y)
mtext(paste0("PCC=",signif(pcc$estimate,2)," (p=",signif(pcc$p.value,2),")"))
abline(lm1$coef)
```

#### Testing with a different context network as a negative control
Regulatory networks can be tissue-specific, and so we built a third network using the same parameters as the two breast cancer data-based ones. The third network is built on leukemic samples from the TCGA AML dataset. Our results show a weaker activity for all TFs, which is only weakly correlated to that inferred via the TCGA breast cancer network.
```{r, fig.cap=capfig("Comparison between activities inferred through a breast cancer TCGA dataset and the AML dataset. PCC indicates the Pearson Correlation Coefficient"), message=FALSE, warning=FALSE}
negative_90=vobj_negative_90$mrs[,"NES"]
negative_45=vobj_negative_45$mrs[,"NES"]
tfs<-names(negative_45)
negativemat<-cbind(rep(0,length(negative_45)),negative_45[tfs],negative_90[tfs])
colnames(negativemat)<-c("T0","T45","T90")

common<-intersect(rownames(tcgamat),rownames(negativemat))

par(mfrow=c(1,2),oma=c(0,0,2,0))
# Scatterplot, negative vs tcga 45mins
x<-tcgamat[common,"T45"]
y<-negativemat[common,"T45"]
plot(x,y,xlab="TCGA, 45mins",ylab="negative, 45mins",pch=20,col="grey",xlim=c(min(x)*1.2,max(x)*1.2))
grid()
lm1<-lm(y~x)
abline(lm1$coef)
pcc<-cor.test(x,y)
mtext(paste0("PCC=",signif(pcc$estimate,2)," (p=",signif(p.adjust(pcc$p.value,method="bonferroni",n=1000),2),")"))

abline(lm1$coef)

# Scatterplot, negative vs tcga 90mins
x<-tcgamat[common,"T90"]
y<-negativemat[common,"T90"]
plot(x,y,xlab="TCGA, 90mins",ylab="negative, 90mins",pch=20,col="grey",xlim=c(min(x)*1.2,max(x)*1.2))
grid()
lm1<-lm(y~x)
abline(lm1$coef)
pcc<-cor.test(x,y)
mtext(paste0("PCC=",signif(pcc$estimate,2)," (p=",signif(p.adjust(pcc$p.value,method="bonferroni",n=1000),2),")"))
abline(lm1$coef)

title("Comparison with Negative Network", outer=TRUE)
```
```{r, echo=FALSE}
par<-par.backup
```

## Pathway enrichment analysis
In the previous paragraphs we generated signatures of differential binding centered around the peaks extracted by VULCAN using the \textit{vulcan.annotate} function. We apply two methods that assess the enrichment of gene pathways over continuous signatures: GSEA (Subramaniam et al., 2004) and aREA (Alvarez et al., 2016).
First, we will load the VULCAN object containing our dataset:
```{r, message=FALSE, warning=FALSE}
# Load imported vulcan object
load("results/001_vobj.rda")
```
Then, we need gene pathways. A manually annotated and very exhaustive list is available from the MsiGDB database by the Broad Institute
```{r, message=FALSE, warning=FALSE}
# Load MsigDB
load("msigdb/MSigDB_v5.0_human.rda")
pathways<-msigDBentrez$c2.all.v5.0.entrez.gmt
```
First, we compare aREA and GSEA. Despite being similar in purpose, the two methods differ in their core algorithm and implementations. Notably, aREA processes the entire analysis in one go, by virtue of matrix multiplications. We expect aREA to be faster, while using more RAM than GSEA. Both GSEA and aREA are implemented in the VULCAN package.
Here we calculate GSEA on the 90' vs 00' signature:
```{r, message=FALSE, warning=FALSE}
fname<-"results/003_pathwayComparison_GSEA_90.rda"
if(!file.exists(fname)){
  start<-Sys.time()
  results_gsea_90<-vulcan.pathways(vobj,pathways,contrast=c("t90","t0"),method="GSEA",np=1000)
  end<-Sys.time()
  time_gsea<-end-start
  save(results_gsea_90,time_gsea,file=fname)
} else {
  load(fname)
}
```
Then, we calculate aREA on the same 90' vs 00' signature:
```{r, message=FALSE, warning=FALSE}
fname<-"results/003_pathwayComparison_REA_90.rda"
if(!file.exists(fname)){
  start<-Sys.time()
  results_rea_90<-vulcan.pathways(vobj,pathways,contrast=c("t90","t0"),method="REA")[,1]
  end<-Sys.time()
  time_rea<-end-start
  save(results_rea_90,time_rea,file=fname)
} else {
  load(fname)
}
```
We then compare GSEA and aREA:
```{r, fig.cap=capfig("Comparison of GSEA and aREA on a differential binding signature"), message=FALSE, warning=FALSE}
plot(results_rea_90,results_gsea_90,xlab="aREA enrichment score",ylab="GSEA enrichment score",pch=20,main="Pathway Enrichment Analysis at 90 vs 0",
     ylim=c(-max(abs(results_gsea_90)),max(abs(results_gsea_90))),col="grey")
th<-p2z(0.1)
abline(h=c(th,-th),v=c(th,-th),lty=2)
lm1<-lm(results_gsea_90~results_rea_90)
abline(lm1$coef,lwd=2)
pcc<-cor(results_rea_90,results_gsea_90,method="p")
mtext(paste0("PCC=",signif(pcc,3),", run on ",length(pathways)," pathways"))
# Estrogen
keywords<-c("ESTROGEN","ESTRADIOL","ESR1")
keypaths<-c()
for(keyword in keywords){
  keypaths<-c(keypaths,grep(keyword,names(pathways),value=TRUE))
}
keypaths<-unique(keypaths)
# Breast Cancer
keywords<-c("BREAST_CANCER")
keypaths2<-c()
for(keyword in keywords){
  keypaths2<-c(keypaths2,grep(keyword,names(pathways),value=TRUE))
}
keypaths2<-unique(keypaths2)

points(results_rea_90[keypaths],results_gsea_90[keypaths],pch=16,col="red3")
points(results_rea_90[keypaths2],results_gsea_90[keypaths2],pch=16,col="navy")
legend("topleft",pch=16,col=c("red3","navy"),cex=1,legend=c("Estrogen-Related Pathways","Breast Cancer Pathways"))
legend("bottomright",
       legend=c(
         paste0("REA: ",signif(time_rea,2)," ",units(time_rea)),
         paste0("GSEA: ",signif(time_gsea,2)," ",units(time_gsea))
       ),
       title="Runtime"
)
```
aREA appears to be faster. Therefore, we will use it to calculate the 45' signature as well:
```{r, message=FALSE, warning=FALSE}
fname<-"results/003_pathwayComparison_REA_45.rda"
if(!file.exists(fname)){
  start<-Sys.time()
  results_rea_45<-vulcan.pathways(vobj,pathways,contrast=c("t45","t0"),method="REA")[,1]
  end<-Sys.time()
  time_rea<-end-start
  save(results_rea_45,time_rea,file=fname)
} else {
  load(fname)
}
```
The results of aREA in both 90'vs00' and 45'vs00' signatures show the presence of known Estrogen-related pahtways:
```{r, fig.cap=captab("Caption"), message=FALSE, warning=FALSE}
topdn<-sort(results_rea_90)[1:20]
topup<-sort(results_rea_90,dec=TRUE)[1:20]

topdn<-cbind(topdn,z2p(topdn))
topup<-cbind(topup,z2p(topup))
colnames(topdn)<-colnames(topup)<-c("NES, 90' vs 0'","pvalue")

rownames(topdn)<-tolower(rownames(topdn))
rownames(topup)<-tolower(rownames(topup))
rownames(topdn)<-gsub("_"," ",rownames(topdn))
rownames(topup)<-gsub("_"," ",rownames(topup))
topdn[,1]<-signif(topdn[,1],3)
topup[,1]<-signif(topup[,1],3)
topdn[,2]<-signif(topdn[,2],2)
topup[,2]<-signif(topup[,2],2)

firstup <- function(x) {
  substr(x, 1, 1) <- toupper(substr(x, 1, 1))
  x
}
rownames(topdn)<-firstup(rownames(topdn))
rownames(topup)<-firstup(rownames(topup))
```
```{r, fig.cap=captab("aREA results: upregulated MsigDBpathways at 90mins"), message=FALSE, warning=FALSE}
# Plot your table with table Grob in the library(gridExtra)
grid.newpage()
grid.table(topup)
```
```{r, fig.cap=captab("aREA results: downregulated MsigDBpathways at 90mins"), message=FALSE, warning=FALSE}
grid.newpage()
grid.table(topdn)
```
```{r, fig.cap=captab("aREA results at 45mins"), message=FALSE, warning=FALSE}
topdn<-sort(results_rea_45)[1:20]
topup<-sort(results_rea_45,dec=TRUE)[1:20]

topdn<-cbind(topdn,z2p(topdn))
topup<-cbind(topup,z2p(topup))
colnames(topdn)<-colnames(topup)<-c("NES, 45' vs 0'","pvalue")

rownames(topdn)<-tolower(rownames(topdn))
rownames(topup)<-tolower(rownames(topup))
rownames(topdn)<-gsub("_"," ",rownames(topdn))
rownames(topup)<-gsub("_"," ",rownames(topup))
topdn[,1]<-signif(topdn[,1],3)
topup[,1]<-signif(topup[,1],3)
topdn[,2]<-signif(topdn[,2],2)
topup[,2]<-signif(topup[,2],2)

firstup <- function(x) {
  substr(x, 1, 1) <- toupper(substr(x, 1, 1))
  x
}
rownames(topdn)<-firstup(rownames(topdn))
rownames(topup)<-firstup(rownames(topup))

```
```{r, fig.cap=captab("aREA results: upregulated MsigDBpathways at 45mins"), message=FALSE, warning=FALSE}
# Plot your table with table Grob in the library(gridExtra)
grid.newpage()
grid.table(topup)
```
```{r, fig.cap=captab("aREA results: downregulated MsigDBpathways at 45mins"), message=FALSE, warning=FALSE}
grid.newpage()
grid.table(topdn)
```

## The GRHL2 Transcription Factor
Our analysis shows that the genes repressed by the Transcription Factor GRHL2 are occupied by the ER complex, using both TCGA-derived and METABRIC-derived regulatory models.

### Network similarity between ESR1 and GRHL2
The following analysis highlights, with a Venn Diagram for both the METABRIC and TCGA ARACNe-derived regulatory models, the overlap between the ESR1 (Estrogen Receptor) and GRHL2 networks
```{r, fig.cap=capfig("Overlap between ESR1 and GRH2 networks in TCGA (left) and METABRIC (right) networks"), message=FALSE, warning=FALSE}
# Load networks
load("networks/brca-tf-regulon.rda")
tcga_regulon<-regul
rm(regul)

load("networks/metabric-regulon-tfs.rda")
metabric_regulon<-regulon
rm(regulon)

# Select TFs
tf1<-"ESR1"
tf2<-"GRHL2"

par(mfrow=c(1,2))

# Venn diagram in TCGA
reg1<-tcga_regulon[[list_symbol2eg[[tf1]]]]
reg2<-tcga_regulon[[list_symbol2eg[[tf2]]]]
targets1<-names(reg1$tfmode)
targets2<-names(reg2$tfmode)
vennlist<-list()
vennlist[[tf1]]<-targets1
vennlist[[tf2]]<-targets2
venn(vennlist)
title("Network overlap in TCGA")

# Venn diagram in METABRIC
reg1<-metabric_regulon[[list_symbol2eg[[tf1]]]]
reg2<-metabric_regulon[[list_symbol2eg[[tf2]]]]
targets1<-names(reg1$tfmode)
targets2<-names(reg2$tfmode)
vennlist<-list()
vennlist[[tf1]]<-targets1
vennlist[[tf2]]<-targets2
venn(vennlist)
title("Network overlap in METABRIC")

par<-par.backup
```

### Correlation between ESR1 and GRHL2 expression in breast cancer datasets
The following analysis assesses the inverse correlation between GRHL2 and ESR1 expressions in TCGA and METABRIC, and highlights the phenomenon for specific tumor subtypes, according to the PAM50 nomenclature (Perou et al., 2000)
```{r, fig.cap=capfig("Correlation between GRHL2 and ESR1 expression in the TCGA breast cancer dataset"), message=FALSE, warning=FALSE}
# TCGA Color annotation
load("data/brca-expmat.rda")
load("data/brca-subtypes.rda")
expmat<-expmat[,names(subtypes)]
colors<-setNames(rep("black",length(subtypes)),subtypes)
colors[subtypes=="Basal"]<-"#FF0000AA"
colors[subtypes=="LumA"]<-"#00FFFFAA"
colors[subtypes=="LumB"]<-"#0000FFAA"
colors[subtypes=="Her2"]<-"#FFFF00AA"
colors[subtypes=="Tumor, Normal-Like"]<-"#00FF00AA"
colors[subtypes=="Normal, Tumor-Like"]<-"#00FF00AA"
colors[subtypes=="Normal"]<-"#00FF00AA"
colors_tcga<-colors
tcga_expmat<-expmat

# METABRIC Color annotation
load("data/metabric-expmat.rda")
load("data/metabric-subtypes.rda")
expmat<-expmat[,names(subtypes)]
colors<-setNames(rep("black",length(subtypes)),subtypes)
colors[subtypes=="Basal"]<-"#FF0000AA"
colors[subtypes=="LumA"]<-"#00FFFFAA"
colors[subtypes=="LumB"]<-"#0000FFAA"
colors[subtypes=="Her2"]<-"#FFFF00AA"
colors[subtypes=="Tumor, Normal-Like"]<-"#00FF00AA"
colors[subtypes=="Normal, Tumor-Like"]<-"#00FF00AA"
colors[subtypes=="Normal"]<-"#00FF00AA"
colors_metabric<-colors
metabric_expmat<-expmat

x<-tcga_expmat[list_symbol2eg[[tf1]],]
y<-tcga_expmat[list_symbol2eg[[tf2]],]
plot(x,y,xlab=tf1,ylab=tf2,pch=20,main="Correlation between GRHL2 and ESR1 expression in TCGA",col=colors_tcga)
pcc<-cor(x,y)
mtext(paste0("PCC=",signif(pcc,3)))
grid()
legend("bottomright",col=c("red","cyan","blue","yellow","green"),legend=c("Basal","LumA","LumB","Her2","Normal-Like"),pch=16)

```
```{r, message=FALSE, warning=FALSE}
x<-metabric_expmat[list_symbol2eg[[tf1]],]
y<-metabric_expmat[list_symbol2eg[[tf2]],]
plot(x,y,xlab=tf1,ylab=tf2,pch=20,main="Correlation between GRHL2 and ESR1 expression in METABRIC",col=colors_metabric)
pcc<-cor(x,y)
mtext(paste0("PCC=",signif(pcc,3)))
grid()
legend("bottomright",col=c("red","cyan","blue","yellow","green"),legend=c("Basal","LumA","LumB","Her2","Normal-Like"),pch=16)

```

### Enrichment of GRHL2 network for gene pathways
In order to analyze the GRHL2 network, we perform a simple hypergeometric test to overlap its targets with all pathways contained in hte manually curated Reactome collection of MsigDB pathways (C2 v5.0).
```{r, message=FALSE, warning=FALSE}
tf<-"GRHL2"
regulon<-tcga_regulon

# Pathways
wass<-load("msigdb/MSigDB_v5.0_human.rda") # "msigDBentrez" "msigDBsymbol"
ngenes<-length(unique(unlist(msigDBentrez$c2.all.v5.0.entrez.gmt)))

### Hypergeometric test implementation
# Hypergeometric enrichment function
enrich<-function(list1,list2,pathway){
  l1<-length(list1)
  l2<-length(list2)
  overlap<-length(intersect(list1,list2))
  q<-overlap-1
  m<-l1
  n<-ngenes-l1
  k<-l2
  p<-phyper(q,m,n,k,lower.tail=FALSE,log.p=FALSE)
  return(p)
}

# Fisher P integration function
fisherp<-function (ps) {
  Xsq <- -2 * sum(log(ps))
  p.val <- pchisq(Xsq, df = 2 * length(ps), lower.tail = FALSE)
  return(p.val)
}

# Regulon vs. pathway
networkUP<-names(regulon[[s2e(tf)]]$tfmode)[regulon[[s2e(tf)]]$tfmode>=0]
networkDN<-names(regulon[[s2e(tf)]]$tfmode)[regulon[[s2e(tf)]]$tfmode<0]

### Enrichment Loop C2
fname<-paste0("results/004_results_hypergeom_C2_",tf,".rda")
#pathways<-msigDBentrez$c2.cp.biocarta.v5.0.entrez.gmt
pathways<-msigDBentrez$c2.all.v5.0.entrez.gmt
if(!file.exists(fname)){
  results<-list()
  ntests<-0
  tfs<-names(regulon)
  pb<-txtProgressBar(0,length(tfs),style=3)
  pbi<-1
  for(tf in tfs){
    plimit<-0.05
    setTxtProgressBar(pb,pbi)
    sublistUP<-setNames(rep(NA,length(pathways)),names(pathways))
    sublistDN<-setNames(rep(NA,length(pathways)),names(pathways))
    i<-1
    for(pathway in pathways){
      networkUP<-names(regulon[[tf]]$tfmode)[regulon[[tf]]$tfmode>=0]
      networkDN<-names(regulon[[tf]]$tfmode)[regulon[[tf]]$tfmode<0]
      pUP<-enrich(networkUP,pathway,ngenes)
      pDN<-enrich(networkDN,pathway,ngenes)
      ntests<-ntests+1
      sublistUP[i]<-pUP
      sublistDN[i]<-pDN
      i<-i+1
    }
    sublistUP<-sublistUP[sublistUP<=plimit]
    sublistDN<-sublistDN[sublistDN<=plimit]
    results[[tf]]<-list(up=sublistUP,dn=sublistDN)
    pbi<-pbi+1
  }
  save(results,ntests,file=fname)
} else {
  load(fname)
}

### Now, plotting
# Set parameters
topn<-20
subsetting<-"REACTOME_"

## Start plot
# Prepare input results
res<-results[[s2e(tf)]]

# Correct pvalues and integrate
resUP<-p.adjust(res$up,method="none",n=ntests) # convert to bonferroni when deploying
resDN<-p.adjust(res$dn,method="none",n=ntests)
union<-union(names(resUP),names(resDN))
tmptable<-cbind(resUP[union],resDN[union])
tmptable[is.na(tmptable)]<-1
rownames(tmptable)<-union
integrated<-apply(tmptable,1,fisherp)

# Optional subsetting
toshow<-names(sort(integrated))
toshow<-toshow[grep(subsetting,toshow)]
toshow<-toshow[1:topn]

# Prepare input
resUP<--log10(res$up[toshow])
resDN<--log10(res$dn[toshow])
names(resUP)<-toshow
names(resDN)<-toshow
resUP[is.na(resUP)]<-0
resDN[is.na(resDN)]<-0

```

```{r, fig.cap=capfig("Pathways most significantly overlapping the GRHL2 network"), message=FALSE, warning=FALSE}
# Canvas
upperlim<-ceiling(max(c(resUP,resDN)))
par(mar=c(0,0,0,0))
plot(0,col="white",xlim=c(-10,12),ylim=c(-2,topn+3),xaxt="n",yaxt="n",type="n",frame.plot=FALSE,
     xlab="",ylab="",xaxs="i",yaxs="i")
# text(c(0:10),c(0:10),labels=c(0:10),pos=1,offset=0)
# text(c(1,2,2,1),c(1,2,1,2),labels=c("1_1","2_2","2_1","1_2"),pos=1,offset=0)

# Title
regsize<-length(regulon[[s2e(tf)]]$tfmode)
text(5,topn+2.8,paste0("Pathway enrichment of ",tf," regulon"),pos=1,offset=0,font=2)
text(5,topn+2,paste0("regulon size: ",regsize),pos=1,offset=0,font=3,cex=0.8)

# Axis horizontal
abline(h=0)
segments(0:10,0,0:10,-0.3)
text(5,-1,"Hypergeometric -log10(pvalue)",pos=1,offset=0)
text(1:10,-0.5,cex=0.7,
     labels=round(signif(seq(0,upperlim,length.out=11)),2)[2:11]
)

# Axis vertical
abline(v=0)
#segments(0,topn:0,-0.3,topn:0)

# Significance line
abline(v=(-log10(0.05)/upperlim)*10,lty=2)

# Bars
for(i in topn:1){
  upval<-(resUP[i]/upperlim)*10
  dnval<-(resDN[i]/upperlim)*10
  rect(0,i+0.25,upval,i,col="salmon")
  rect(0,i,dnval,i-0.25,col="cornflowerblue")
}

# Pathway labels
pathlabels<-gsub(subsetting,"",toshow)
pathlabels<-tolower(pathlabels)
pathlabels<-gsub("_"," ",pathlabels)
for(i in topn:1){
  text(0,i,pathlabels[i],pos=2,offset=0.1,cex=0.8)
}

# Percentage of pathway overlap
r<-names(regulon[[s2e(tf)]]$tfmode)
percs<-c()
for(i in topn:1){
  p<-pathways[[toshow[i]]]
  intrsct<-intersect(p,r)
  perc<-100*length(intrsct)/length(p)
  percs<-c(percs,perc)
}
names(percs)<-rev(toshow)
spercs<-(percs/max(percs))*4
lines(spercs,topn:1,lty=1,lwd=2)

# Ruler of pathway overlap
segments(0,topn+1,4,topn+1)
text(2,topn+1.5,pos=1,offset=0,font=3,cex=0.7,labels="% pathway")
segments(0:4,topn+1,0:4,topn+0.8)
text(1:4,topn+0.65,cex=0.5,
     labels=round(seq(0,max(percs),length.out=5),1)[2:5]
)
par<-par.backup
```

## Comparing VULCAN results with other tools and approaches
Our VULCAN results were obtained using independent gene regulatory models and multiple replicates to generate ER-response signatures.

### VULCAN and qPLEX-RIME
We set to testing the performance of VULCAN against a complementary experimental approach called qPLEX-RIME (Holding et al., submitted), which aims at identifyin interactors of ER within the ER-chromatin complex at estrogen-responding promoters. We have QRIME results for MCF7 cells treated with estradiol at both 45 and 90 minutes, the same experimental setup for the VULCAN input dataset.
```{r qrime, message=FALSE, warning=FALSE}
# Load imported vulcan object
load("results/001_vobj.rda")

#### Vulcan Analysis (multiple networks)
## TCGA network, ARACNe with proteomics centroids
load("aracne/regulon-tcga-centroids.rda")
tcga_regulon<-regulon
rm(regulon)

## METABRIC network, ARACNe with proteomics centroids
load("aracne/regulon-metabric-centroids.rda")
metabric_regulon<-regulon
rm(regulon)

## Specify contrast and network
fname<-"results/005_vobj_networks.rda"
list_eg2symbol<-as.list(org.Hs.egSYMBOL[mappedkeys(org.Hs.egSYMBOL)]) 
if(!file.exists(fname)){
  vobj_tcga_90<-vulcan(vobj,network=tcga_regulon,contrast=c("t90","t0"),annotation=list_eg2symbol)
  vobj_tcga_45<-vulcan(vobj,network=tcga_regulon,contrast=c("t45","t0"),annotation=list_eg2symbol)
  vobj_metabric_90<-vulcan(vobj,network=metabric_regulon,contrast=c("t90","t0"),annotation=list_eg2symbol)
  vobj_metabric_45<-vulcan(vobj,network=metabric_regulon,contrast=c("t45","t0"),annotation=list_eg2symbol)
  save(
    vobj_tcga_90,
    vobj_tcga_45,
    vobj_metabric_90,
    vobj_metabric_45,
    file=fname
  )
} else {
  load(fname)
}

# Merge proteomics and TF aracne results
vobj2_tcga_90<-vobj_tcga_90
vobj2_tcga_45<-vobj_tcga_45
vobj2_metabric_90<-vobj_metabric_90
vobj2_metabric_45<-vobj_metabric_45

fname<-"results/002_vobj_networks.rda"
load(fname)
vobj_tcga_90$mrs<-rbind(vobj_tcga_90$mrs,vobj2_tcga_90$mrs[setdiff(rownames(vobj2_tcga_90),rownames(vobj_tcga_90)),])
vobj_tcga_45$mrs<-rbind(vobj_tcga_45$mrs,vobj2_tcga_45$mrs[setdiff(rownames(vobj2_tcga_45),rownames(vobj_tcga_45)),])
vobj_metabric_90$mrs<-rbind(vobj_metabric_90$mrs,vobj2_metabric_90$mrs[setdiff(rownames(vobj2_metabric_90),rownames(vobj_metabric_90)),])
vobj_metabric_45$mrs<-rbind(vobj_metabric_45$mrs,vobj2_metabric_45$mrs[setdiff(rownames(vobj2_metabric_45),rownames(vobj_metabric_45)),])

# Metabric Results
metabric_90=vobj_metabric_90$mrs[,"pvalue"]
metabric_45=vobj_metabric_45$mrs[,"pvalue"]
tfs<-names(metabric_45)
metabricmat<-cbind(metabric_45[tfs],metabric_90[tfs])
colnames(metabricmat)<-c("T45","T90")

# TCGA Results
tcga_90=vobj_tcga_90$mrs[,"pvalue"]
tcga_45=vobj_tcga_45$mrs[,"pvalue"]
tfs<-names(tcga_45)
tcgamat<-cbind(tcga_45[tfs],tcga_90[tfs])
colnames(tcgamat)<-c("T45","T90")

## QRIME results
raw45<-read.delim("qrime/Results/ER_45min_vs_ER_0minExcludeMissingValues_QuantileNormalization.txt",as.is = TRUE)
qrime45<-setNames(raw45$P.Value,raw45$Gene)
qrime45<-qrime45[!is.na(qrime45)]

raw90<-read.delim("qrime/Results/ER_90min_vs_ER_0minExcludeMissingValues_QuantileNormalization.txt",as.is = TRUE)
qrime90<-setNames(raw90$P.Value,raw90$Gene)
qrime90<-qrime90[!is.na(qrime90)]
```
Here, we will compare ER-binding pvalues obtained with QRIME, with the VULCAN results aimed at identifying TFs upstream of the observed differential promoter binding.
```{r, fig.cap=capfig("Comparison of significance between the QRIME method (x-axis) and the VULCAN method (y-axis) at two time points using two regulatory networks for VULCAN"), message=FALSE, warning=FALSE}
### Comparisons, pvalues
par(mfrow=c(2,2))
# TCGA 45
main<-"TCGA network, 45 mins"
common<-intersect(rownames(tcgamat),names(qrime45))
x<-qrime45[common]
x<--log10(x)
y<-tcgamat[common,"T45"]
y<--log10(y)
plot(x,y,pch=20,xlab="-log10(p) QRIME",ylab="-log10(p) VULCAN",col="grey",main=main,
     xlim=c(min(x)-max(x)*0.3,max(x)*1.1),
     ylim=c(min(y)-max(y)*0.2,max(y)*1.1)
)
topx<-names(sort(x,dec=TRUE))[1:10]
topy<-names(sort(y,dec=TRUE))[1:10]
top<-union(topx,topy)
textplot2(x[top],y[top],words=top,new=FALSE)
grid()

# TCGA 90
main<-"TCGA network, 90 mins"
common<-intersect(rownames(tcgamat),names(qrime90))
x<-qrime90[common]
x<--log10(x)
y<-tcgamat[common,"T90"]
y<--log10(y)
plot(x,y,pch=20,xlab="-log10(p) QRIME",ylab="-log10(p) VULCAN",col="grey",main=main,
     xlim=c(min(x)-max(x)*0.3,max(x)*1.1),
     ylim=c(min(y)-max(y)*0.2,max(y)*1.1)
)
topx<-names(sort(x,dec=TRUE))[1:10]
topy<-names(sort(y,dec=TRUE))[1:10]
top<-union(topx,topy)
textplot2(x[top],y[top],words=top,new=FALSE)
grid()

# METABRIC 45
main<-"METABRIC network, 45 mins"
common<-intersect(rownames(metabricmat),names(qrime45))
x<-qrime45[common]
x<--log10(x)
y<-metabricmat[common,"T45"]
y<--log10(y)
plot(x,y,pch=20,xlab="-log10(p) QRIME",ylab="-log10(p) VULCAN",col="grey",main=main,
     xlim=c(min(x)-max(x)*0.3,max(x)*1.1),
     ylim=c(min(y)-max(y)*0.2,max(y)*1.1)
)
topx<-names(sort(x,dec=TRUE))[1:10]
topy<-names(sort(y,dec=TRUE))[1:10]
top<-union(topx,topy)
textplot2(x[top],y[top],words=top,new=FALSE)
grid()

# METABRIC 90
main<-"METABRIC network, 90 mins"
common<-intersect(rownames(metabricmat),names(qrime90))
x<-qrime90[common]
x<--log10(x)
y<-metabricmat[common,"T90"]
y<--log10(y)
plot(x,y,pch=20,xlab="-log10(p) QRIME",ylab="-log10(p) VULCAN",col="grey",main=main,
     xlim=c(min(x)-max(x)*0.3,max(x)*1.1),
     ylim=c(min(y)-max(y)*0.2,max(y)*1.1)
)
topx<-names(sort(x,dec=TRUE))[1:10]
topy<-names(sort(y,dec=TRUE))[1:10]
top<-union(topx,topy)
textplot2(x[top],y[top],words=top,new=FALSE)
grid()

```
```{r, fig.cap=capfig("Comparison of Normalized Enrichment Score between the QRIME method (x-axis) and the VULCAN method (y-axis) at two time points using two regulatory networks for VULCAN"), message=FALSE, warning=FALSE}

# TCGA Results
tcganes_90=vobj_tcga_90$mrs[,"NES"]
tcganes_45=vobj_tcga_45$mrs[,"NES"]
tfs<-names(tcganes_45)
tcganesmat<-cbind(tcganes_45[tfs],tcganes_90[tfs])
colnames(tcganesmat)<-c("T45","T90")

# METABRIC Results
metabricnes_90=vobj_metabric_90$mrs[,"NES"]
metabricnes_45=vobj_metabric_45$mrs[,"NES"]
tfs<-names(metabricnes_45)
metabricnesmat<-cbind(metabricnes_45[tfs],metabricnes_90[tfs])
colnames(metabricnesmat)<-c("T45","T90")

par(mfrow=c(2,2))
# TCGA 45
main<-"TCGA network, 45 mins"
common<-intersect(rownames(tcganesmat),names(qrime45))
x<-qrime45[common]
x<--log10(x)
y<-tcganesmat[common,"T45"]
plot(x,y,pch=20,xlab="-log10(p) QRIME",ylab="NES VULCAN",col="grey",main=main,
     xlim=c(min(x)-max(x)*0.3,max(x)*1.1),
     ylim=c(min(y)-max(y)*0.2,max(y)*1.1)
)
topx<-names(sort(x,dec=TRUE))[1:10]
topy<-names(sort(y,dec=TRUE))[1:10]
top<-union(topx,topy)
textplot2(x[top],y[top],words=top,new=FALSE)
abline(h=c(-p2z(0.05),p2z(0.05)),lty=3)
grid()

# TCGA 90
main<-"TCGA network, 90 mins"
common<-intersect(rownames(tcganesmat),names(qrime90))
x<-qrime90[common]
x<--log10(x)
y<-tcganesmat[common,"T90"]
plot(x,y,pch=20,xlab="-log10(p) QRIME",ylab="NES VULCAN",col="grey",main=main,
     xlim=c(min(x)-max(x)*0.3,max(x)*1.1),
     ylim=c(min(y)-max(y)*0.2,max(y)*1.1)
)
topx<-names(sort(x,dec=TRUE))[1:10]
topy<-names(sort(y,dec=TRUE))[1:10]
top<-union(topx,topy)
textplot2(x[top],y[top],words=top,new=FALSE)
abline(h=c(-p2z(0.05),p2z(0.05)),lty=3)
grid()

# METABRIC 45
main<-"METABRIC network, 45 mins"
common<-intersect(rownames(metabricnesmat),names(qrime45))
x<-qrime45[common]
x<--log10(x)
y<-metabricnesmat[common,"T45"]
plot(x,y,pch=20,xlab="-log10(p) QRIME",ylab="NES VULCAN",col="grey",main=main,
     xlim=c(min(x)-max(x)*0.3,max(x)*1.1),
     ylim=c(min(y)-max(y)*0.2,max(y)*1.1)
)
topx<-names(sort(x,dec=TRUE))[1:10]
topy<-names(sort(y,dec=TRUE))[1:10]
top<-union(topx,topy)
textplot2(x[top],y[top],words=top,new=FALSE)
abline(h=c(-p2z(0.05),p2z(0.05)),lty=3)
grid()

# METABRIC 90
main<-"METABRIC network, 90 mins"
common<-intersect(rownames(metabricnesmat),names(qrime90))
x<-qrime90[common]
x<--log10(x)
y<-metabricnesmat[common,"T90"]
plot(x,y,pch=20,xlab="-log10(p) QRIME",ylab="NES VULCAN",col="grey",main=main,
     xlim=c(min(x)-max(x)*0.3,max(x)*1.1),
     ylim=c(min(y)-max(y)*0.2,max(y)*1.1)
)
topx<-names(sort(x,dec=TRUE))[1:10]
topy<-names(sort(y,dec=TRUE))[1:10]
top<-union(topx,topy)
textplot2(x[top],y[top],words=top,new=FALSE)
abline(h=c(-p2z(0.05),p2z(0.05)),lty=3)
grid()
```

### Comparing Mutual Information and Partial Correlation networks
As shown before in literature, partial correlation and mutual information networks are highly similar. While requiring a mutual information network due to the VIPER engine used by the VULCAN pipeline, we set out here to compare Mutual Information networks with partial correlation ones, using different correlation thresholds.
```{r pcor, message=FALSE, warning=FALSE}
filename<-"results/006_comparison1_aracne.rda"
if(!file.exists(filename)){
  # Load aracne network
  load("aracne/regulon-tcga-centroids.rda")

  # Load expression matrix
  expmat<-as.matrix(read.delim("aracne/tcga-expmat.dat",as.is=TRUE,row.names=1))

  ### Better pipeline
  tfs<-names(regulon)
  genes<-rownames(expmat)

  # Remove zero-SD genes
  sds<-apply(expmat,1,sd)
  expmat<-expmat[sds>=0.5,]
  dim(expmat)

  # Build Covariance Matrix
  pcormat<-cov(t(expmat))
  # Build pcormat
  pcormat<-cor2pcor(pcormat)
  rownames(pcormat)<-colnames(pcormat)<-rownames(expmat)

  # Keep only TFs
  diag(pcormat)<-0
  tfs<-intersect(tfs,rownames(pcormat))
  pcormat<-pcormat[tfs,]

  # Transform my network into a vector of edges
  myedges<-c()
  for(centroid in names(regulon)){
    targets<-names(regulon[[centroid]]$tfmode)
    newedges<-paste0(centroid,"___",targets)
    myedges<-c(myedges,newedges)
  }
  length(myedges) # 268394

  ### Loop over rs
  jis<-c()
  nedges<-c()
  rs<-seq(0.05,0.4,by=0.025)
  for(r in rs){
    # Transform networks into vectors of edges
    matches<-which(abs(pcormat)>=r,arr.ind=TRUE)
    pcoredges<-paste0(rownames(pcormat)[matches[,1]],"___",colnames(pcormat)[matches[,2]])
    nedges<-c(nedges,nrow(matches))
    # Calculate Jaccard
    ji<-length(intersect(myedges,pcoredges))/length(union(myedges,pcoredges))
    jis<-c(jis,ji)
  }
  names(jis)<-names(nedges)<-rs

  # Generate random network vectors
  random<-list()
  for(ii in 1:length(rs)){
    n<-nedges[ii]
    message("Doing ",rs[ii])
    pb<-txtProgressBar(0,1000,style=3)
    randomjis<-c()
    for(i in 1:1000){
      x<-sample(1:nrow(pcormat),n,replace=TRUE)
      y<-sample(1:ncol(pcormat),n,replace=TRUE)
      matches<-cbind(x,y)
      pcoredges<-paste0(rownames(pcormat)[matches[,1]],"___",colnames(pcormat)[matches[,2]])
      ji<-length(intersect(myedges,pcoredges))/length(union(myedges,pcoredges))
      randomjis<-c(randomjis,ji)
      setTxtProgressBar(pb,i)
    }
    random[[ii]]<-randomjis
  }
  names(random)<-as.character(rs)

  save(jis,rs,nedges,random,file=filename)

} else {
  load(filename)
}
```
We generate several partial correlation networks using the same input as the ARACNe network used by VULCAN (the TCGA breast cancer dataset). We tested the overlap of every partial correlation network with the ARACNe network using the Jaccard Index (JI) criterion.
```{r, fig.cap=capfig("Sheer number of overlapping edges at different Pcor thresholds"), message=FALSE, warning=FALSE}
par(mfrow=c(1,1))
bp<-barplot(jis,xlab="pcor r",ylab="JI",ylim=c(0,max(jis)*1.1))
text(bp[,1],jis,labels=kmgformat(nedges),pos=3,cex=0.7)
mtext("Number of overlapping edges between Pcor and Aracne",cex=0.8)
```
Finally, we show how the Jaccard Index between partial correlation networks and the ARACNe network is always significantly higher than expected by selecting random network edges.
```{r, fig.cap=capfig("Comparing the Jaccard Index between the ARACNe network and Partial correlation networks (indicated as arrows) and random networks (indicated as distributions)"), message=FALSE, warning=FALSE}
par(mfrow=c(5,3))
for(i in 1:length(random)){
  dd<-density(random[[i]])
  plot(dd,xlim=c(min(random[[i]]),jis[i]*1.1),main=paste0("r=",rs[i]),xlab="Jaccard Index",lwd=2)
  arrows(jis[i],max(dd$y),jis[i],0,lwd=2)
}
```

### Comparing alternative methods for target enrichment analysis
We developed three independent methods to compare VULCAN with. The first is common, the second is simple but crude, and the third is also common but too stringent.

```{r vipercomparison, message=FALSE, warning=FALSE}
### Classic vulcan result
# vobj_tcga_90<-vulcan(vobj,network=tcga_regulon,contrast=c("t90","t0"),annotation=list_eg2symbol)
wass<-load("results/002_vobj_networks.rda")

### Prepare the workspace
# Load imported vulcan object
load("results/001_vobj.rda")

# Network
load("networks/brca-tf-regulon.rda")
tcga_regulon<-regul
rm(regul)

# Entrez to symbol
library(org.Hs.eg.db)
list_eg2symbol<-as.list(org.Hs.egSYMBOL[mappedkeys(org.Hs.egSYMBOL)]) 
```

1. A T-test based method, that takes the targets of a TF and integrates their pvalue in a specific contrast. Good but hard to keep control over the p-value (the classic integration step using the Fisher's Integration method vastly underestimates them). Also, GRHL2 doesn't shine with this method. Unless differently specified, the p-values are always Bonferroni-corrected
```{r, fig.cap=capfig("Comparison between VULCAN/VIPER and T-test integration"), message=FALSE, warning=FALSE}
#######################################################
### TTEST INTEGRATION
#######################################################
if(TRUE){
  ### TTest functions
  msvipertt<-function(signature,network,minsize=10){
    # First, convert tscores into pvalues
    psignature<-2*pt(signature[,1], nrow(signature)-2, lower=FALSE)

    # For each regulon, take the targets and integrate their pvalue
    tfpvalues<-c()
    for(tf in names(network)){
      targets<-names(network[[tf]]$tfmode)
      ptargets<-psignature[intersect(targets,names(psignature))]
      if(length(ptargets)>=minsize){
        pt<-fisherp(ptargets)
        tfpvalues<-c(tfpvalues,pt)
        names(tfpvalues)[length(tfpvalues)]<-tf
      }
    }

    return(tfpvalues)
  }
  vulcantt<-function(vobj, network, contrast, annotation = NULL, minsize = 10) {
    tfs <- names(network)
    samples <- vobj$samples
    normalized <- vobj$normalized

    # Prepare output objects
    msvipers <- matrix(NA, ncol = 3, nrow = length(tfs))
    rownames(msvipers) <- tfs
    # Define contrast
    a <- samples[[contrast[1]]]
    b <- samples[[contrast[2]]]
    # Vulcan msviper implementation
    signature <- rowTtest(normalized[, a], normalized[, b])$statistic
    tfpvalues <- msvipertt(signature, network, minsize = minsize)
    if(!is.null(annotation)){
      names(tfpvalues)<-annotation[names(tfpvalues)]
    }
    return(tfpvalues)
  }

  ### Repeat the same analysis but with TT
  ttp_90<-p.adjust(vulcantt(vobj,network=tcga_regulon,contrast=c("t90","t0"),annotation=list_eg2symbol),method="bonferroni")
  ttp_45<-p.adjust(vulcantt(vobj,network=tcga_regulon,contrast=c("t45","t0"),annotation=list_eg2symbol),method="bonferroni")

  ### Compare vulcan and TT
  vulcanp_90<-vobj_tcga_90$msviper$es$p.value
  vulcanp_45<-vobj_tcga_45$msviper$es$p.value

  toshow<-c("ESR1","GATA3","GRHL2")

  par(mfrow=c(1,2))

  common<-intersect(names(ttp_45),names(vulcanp_45))
  x<--log10(vulcanp_45[common]+1e-320)
  y<--log10(ttp_45[common]+1e-320)
  plot(x,y,pch=20,main="Method Comparison, 45' vs 00'",xlab="VULCAN pvalue",ylab="Ttest integration pvalue",col="darkgrey")
  grid()
  pcc<-cor.test(x,y)
  mtext(paste0("PCC=",signif(pcc$estimate,3)," (p=",signif(pcc$p.value,3),")"))
  textplot2(x[toshow],y[toshow],words=toshow,new=FALSE)

  common<-intersect(names(ttp_90),names(vulcanp_90))
  x<--log10(vulcanp_90[common]+1e-320)
  y<--log10(ttp_90[common]+1e-320)
  plot(x,y,pch=20,main="Method Comparison, 90' vs 00'",xlab="VULCAN pvalue",ylab="Ttest integration pvalue",col="darkgrey")
  grid()
  pcc<-cor.test(x,y)
  mtext(paste0("PCC=",signif(pcc$estimate,3)," (p=",signif(pcc$p.value,3),")"))
  textplot2(x[toshow],y[toshow],words=toshow,new=FALSE)

  par(mfrow=c(1,1))

}
```

2. A fraction of targets method, defining for every TF the fraction of their targets that are also differentially bound. A crude alternative to VULCAN, which is ignoring the MI strength of interaction and the individual strengths of differential bindings.
```{r, fig.cap=capfig("Comparison between VULCAN/VIPER and a fraction of targets found method"), message=FALSE, warning=FALSE}
#######################################################
### FRACTION TEST
#######################################################
if(TRUE){
  ### TTest functions
  msviperfrac<-function(signature,network,minsize=10){
    # First, convert tscores into pvalues
    psignature<-2*pt(signature[,1], nrow(signature)-2, lower=FALSE)

    # For each regulon, take the targets and integrate their pvalue
    tfpvalues<-c()
    for(tf in names(network)){
      targets<-names(network[[tf]]$tfmode)
      ptargets<-psignature[intersect(targets,names(psignature))]
      ptargets<-ptargets
        frac<-length(ptargets)/length(targets)
        tfpvalues<-c(tfpvalues,frac)
        names(tfpvalues)[length(tfpvalues)]<-tf
    }
    return(tfpvalues)
  }
  vulcanfrac<-function(vobj, network, contrast, annotation = NULL, minsize = 10) {
    tfs <- names(network)
    samples <- vobj$samples
    normalized <- vobj$normalized

    # Prepare output objects
    msvipers <- matrix(NA, ncol = 3, nrow = length(tfs))
    rownames(msvipers) <- tfs
    # Define contrast
    a <- samples[[contrast[1]]]
    b <- samples[[contrast[2]]]
    # Vulcan msviper implementation
    signature <- rowTtest(normalized[, a], normalized[, b])$statistic
    tfpvalues <- msviperfrac(signature, network, minsize = minsize)
    if(!is.null(annotation)){
      names(tfpvalues)<-annotation[names(tfpvalues)]
    }
    return(tfpvalues)
  }

  ### Repeat the same analysis but with TT
  ttp_90<-vulcanfrac(vobj,network=tcga_regulon,contrast=c("t90","t0"),annotation=list_eg2symbol)
  ttp_45<-vulcanfrac(vobj,network=tcga_regulon,contrast=c("t45","t0"),annotation=list_eg2symbol)

  ### Compare vulcan and TT
  vulcanp_90<-vobj_tcga_90$msviper$es$p.value
  vulcanp_45<-vobj_tcga_45$msviper$es$p.value

  toshow<-c("ESR1","GATA3","GRHL2")

  par(mfrow=c(1,2))

  common<-intersect(names(ttp_45),names(vulcanp_45))
  x<--log10(vulcanp_45[common])
  y<-ttp_45[common]
  plot(x,y,pch=20,main="Method Comparison, 45' vs 00'",xlab="VULCAN pvalue",ylab="Fraction of Differentially Bound Targets",col="darkgrey")
  grid()
  pcc<-cor.test(x,y)
  mtext(paste0("PCC=",signif(pcc$estimate,3)," (p=",signif(pcc$p.value,3),")"))
  textplot2(x[toshow],y[toshow],words=toshow,new=FALSE)

  common<-intersect(names(ttp_90),names(vulcanp_90))
  x<--log10(vulcanp_90[common])
  y<-ttp_90[common]
  plot(x,y,pch=20,main="Method Comparison, 90' vs 00'",xlab="VULCAN pvalue",ylab="Fraction of Differentially Bound Targets",col="darkgrey")
  grid()
  pcc<-cor.test(x,y)
  mtext(paste0("PCC=",signif(pcc$estimate,3)," (p=",signif(pcc$p.value,3),")"))
  textplot2(x[toshow],y[toshow],words=toshow,new=FALSE)

  par(mfrow=c(1,1))

}
```
3. A Fisher's Exact Method (different from the Fisher's Integration!) which assesses the overlap between networks and significant differential binding. Here the problem is that the test is too stringent (as observed in the original VIPER paper) and even without p-value correction we get nothing significant (I tried different threshold of differential binding significance). The plot is uninspiring highlights the non-significance of the results of this method:
```{r, fig.cap=capfig("Comparison between VULCAN/VIPER and Fisher's Exact Test method"), message=FALSE, warning=FALSE}
#######################################################
### FISHER'S EXACT TEST
#######################################################
if(TRUE){
  ### TTest functions
  msviperfet<-function(signature,network,minsize=10){
    # First, convert tscores into pvalues
    psignature<-2*pt(signature[,1], nrow(signature)-2, lower=FALSE)

    # For each regulon, take the targets and calculate how many are significant
    tfpvalues<-c()
    for(tf in names(network)){
      targets<-names(network[[tf]]$tfmode)
      ptargets<-psignature[intersect(targets,names(psignature))]
      if(length(ptargets)>=minsize){
        # Prepare the contingency table
        ul<-names(ptargets)[ptargets<=0.05]
        ur<-setdiff(targets,ul)
        dl<-setdiff(names(psignature)[psignature<=0.05],ul)
        ctable<-rbind(
          c(length(ul),length(ur)),
          c(length(dl),0)
        )
        #ctable[2,2]<-5000-sum(ctable)
        fet<-fisher.test(ctable,alternative="greater")
        pt<-fet$p.value
        tfpvalues<-c(tfpvalues,pt)
        names(tfpvalues)[length(tfpvalues)]<-tf
      }
    }

    return(tfpvalues)
  }
  vulcanfet<-function(vobj, network, contrast, annotation = NULL, minsize = 10) {
    tfs <- names(network)
    samples <- vobj$samples
    normalized <- vobj$normalized

    # Prepare output objects
    msvipers <- matrix(NA, ncol = 3, nrow = length(tfs))
    rownames(msvipers) <- tfs
    # Define contrast
    a <- samples[[contrast[1]]]
    b <- samples[[contrast[2]]]
    # Vulcan msviper implementation
    signature <- rowTtest(normalized[, a], normalized[, b])$statistic
    tfpvalues <- msviperfet(signature, network, minsize = minsize)
    if(!is.null(annotation)){
      names(tfpvalues)<-annotation[names(tfpvalues)]
    }
    return(tfpvalues)
  }

  ### Repeat the same analysis but with FET
  ttp_90<-vulcanfet(vobj,network=tcga_regulon,contrast=c("t90","t0"),annotation=list_eg2symbol)
  ttp_45<-vulcanfet(vobj,network=tcga_regulon,contrast=c("t45","t0"),annotation=list_eg2symbol)

  ### Compare vulcan and TT
  vulcanp_90<-vobj_tcga_90$msviper$es$p.value
  vulcanp_45<-vobj_tcga_45$msviper$es$p.value

  toshow<-c("ESR1","GATA3","GRHL2")

  par(mfrow=c(1,2))

  common<-intersect(names(ttp_45),names(vulcanp_45))
  x<--log10(vulcanp_45[common]+1e-320)
  y<--log10(ttp_45[common]+1e-320)
  plot(x,y,pch=20,main="Method Comparison, 45' vs 00'",xlab="VULCAN pvalue",ylab="Classic Fisher's Exact Test",col="darkgrey")
  grid()
  textplot2(x[toshow],y[toshow],words=toshow,new=FALSE)

  common<-intersect(names(ttp_90),names(vulcanp_90))
  x<--log10(vulcanp_90[common]+1e-320)
  y<--log10(ttp_90[common]+1e-320)
  plot(x,y,pch=20,main="Method Comparison, 90' vs 00'",xlab="VULCAN pvalue",ylab="Classic Fisher's Exact Test",col="darkgrey")
  grid()
  textplot2(x[toshow],y[toshow],words=toshow,new=FALSE)

  par(mfrow=c(1,1))
}
```

### Comparing VULCAN with online tools
We will generate BED files for the peaks, as reported by DiffBind, we will then test the pathway enrichment for each of these peaks using the Great software v3.0.0 (http://bejerano.stanford.edu/great/public/html/), the ChIP-Enrich pipeline (http://chip-enrich.med.umich.edu/) and the ISMARA tool (http://ismara.unibas.ch).
Parameters for GREAT: defaults (Basal plus extension, 5.0kb upstream, 1.0 kb downstream)
Parameters for ChIP-Enrich: defaults (Nearest TSS, pathways size<=2000, Biocarta, KEGG, Reactome, TFs).
Parameters for ISMARA: defaults (ChIP-Seq mode, hg19)

The VULCAN analysis shows a significant overlap in terms of significant pathways with the GREAT method. ChIP-enrich computes enrichment for a number of TFs which are amongst the most significant in VULCAN, but it surprisingly fails at identifing ESR1 as the top Transcription Factor affected by our experiment. ISMARA succeeds at identifying ESR1 using a motif-based analysis, but doesn't identify other candidate binding TFs, as expected, being the experiment targeted at the estrogen receptor.

```{r, message=FALSE, warning=FALSE}
### Number of tested pathways (for statistical comparison later)
load("msigdb/MSigDB_v5.0_human.rda")
raw<-msigDBentrez$c2.all.v5.0.entrez.gmt
universe<-names(raw)[grep("BIOCARTA_|REACTOME_|KEGG_|PID_|ST_",names(raw))]

### DiffBind object (containing peak location and intensity)
wass<-load("results/001_diffbind.rda")

### Generate input BEDs for chip enrich and GREAT (ismara takes BAMs as input)
fname<-paste0("results/008_comparison_contrast90vs00_UP_p1.bed")
if(!file.exists(fname)){
  contrasts<-c("contrast45vs00","contrast90vs00","contrast90vs45")
  for (contrast in contrasts) {
    c<-switch(contrast,contrast45vs00=3,contrast90vs00=2,contrast90vs45=1)
    for(pgreat in c(1,0.05)){
      bed<-as.data.frame(dba.report(dbaobj,contrast=c,method=DBA_DESEQ2,bNormalized=TRUE, bCounts=TRUE, th=pgreat))
      if(nrow(bed)>0) { # Nothing individually significant if no rows were produced
        bedup<-bed[bed$Fold>0,1:3]
        beddn<-bed[bed$Fold<0,1:3]
        if(nrow(bedup)>0){
          write.table(bedup,
                      file=paste0("results/008_comparison_",contrast,"_UP_p",pgreat,".bed"),
                      row.names=FALSE,col.names=FALSE,sep="\t",quote=FALSE
          )
        }
        if(nrow(beddn)>0){
          write.table(beddn,
                      file=paste0("results/008_comparison_",contrast,"_DN_p",pgreat,".bed"),
                      row.names=FALSE,col.names=FALSE,sep="\t",quote=FALSE
          )
        }
      }
    }
  }
}
```

```{r, fig.cap=capfig("Comparison of results from the VULCAN and GREAT methods"), message=FALSE, warning=FALSE}
### Pathway Comparison: VULCAN vs. GREAT
load_obj <- function(f){
  env <- new.env()
  nm <- load(f, env)[1]
  env[[nm]]
}
contrasts<-c("45","90")
par(mfrow=c(2,1))
for(c in contrasts){
  # Our REA pathways
  nes.tpathways<-load_obj(paste0("results/003_pathwayComparison_REA_",c,".rda")) # results_rea_45

  # GREAT pathways
  rawgreat<-read.delim(paste0("results/great/great_",c,"_p1_UP.tsv"),as.is=TRUE,skip=3)
  table(rawgreat[,1])
  rawgreat<-rawgreat[rawgreat[,1]=="MSigDB Pathway",]
  great<-setNames(rawgreat$BinomBonfP,rawgreat[,2])
  great<-great[great<0.1]

  ### Comparison GREAT/VULCAN
  vsig<-nes.tpathways[z2p(nes.tpathways)<0.1]
  venn(list(VULCAN=names(vsig),GREAT=names(great)))
  title(paste0("Comparison VULCAN/GREAT for pathways at ",c," minutes"))
  ctable<-rbind(c(0,0),c(0,0))
  ctable[1,1]<-length(intersect(names(vsig),names(great)))
  ctable[1,2]<-length(setdiff(names(vsig),names(great)))
  ctable[2,1]<-length(setdiff(names(great),names(vsig)))
  ctable[2,2]<-length(universe)-ctable[1,1]-ctable[1,2]-ctable[2,1]
  fp<-signif(fisher.test(ctable)$p.value,4)
  mtext(paste0("FET p-value: ",fp))
  text(100,3,ctable[2,2])

  common<-intersect(names(nes.tpathways),names(great))
}
par(mfrow=c(1,1))
```

```{r, fig.cap=capfig("Comparison of results from the VULCAN and ChIP-enrich methods"), message=FALSE, warning=FALSE}
### TF and Pathway Comparison: VULCAN vs. ChIP enrich
load("results/002_vobj_networks.rda")
contrasts<-c("45","90")
par(mfrow=c(2,1))
for(c in contrasts){
  # Our VULCAN TF enrichment
  if(c=="45"){
    vulcannes<-vobj_tcga_45$msviper$es$nes
  }
  if(c=="90"){
    vulcannes<-vobj_tcga_90$msviper$es$nes
  }
  # ChIPEnrich pathways + TFs
  rawchipenrich<-read.delim(paste0("results/chipenrichr/",c,"_p1_UP_results.tab"),as.is=TRUE)
  chipenrich_tfs<-rawchipenrich[rawchipenrich[,1]=="Transcription Factors",]
  chipenrich_pathways<-rawchipenrich[rawchipenrich[,1]!="Transcription Factors",]
  t1<-gsub(" ","_",paste0(toupper(chipenrich_pathways[,1]),"_",toupper(chipenrich_pathways[,3])))
  t2<-chipenrich_pathways[,"FDR"]
  chipenrich_pathways<-setNames(t2,t1)

  # TF comparison
  cetfs<-setNames(chipenrich_tfs[,"FDR"],gsub("_.+","",chipenrich_tfs[,"Description"]))
  vutfs<-vulcannes[!is.na(vulcannes)]
  vutfs<-vutfs[vutfs>0]
  common<-intersect(names(vutfs),names(cetfs))

  plot(vutfs,-log10(z2p(vutfs)),xlab="VULCAN NES",ylab="VULCAN -log10(p)",pch=20,col="grey",
       main=paste0("Comparison VULCAN/ChIP-enrich at ",c)
  )
  mtext("Labels indicate significant TFs by ChIP-enrich binding site analysis",cex=0.8)
  set.seed(1)
  textplot2(vutfs[common],-log10(z2p(vutfs))[common],common,new=FALSE)
}
par(mfrow=c(1,1))
```

```{r, message=FALSE, warning=FALSE}
### Comparison: VULCAN vs. ISMARA
ismara<-t(read.delim("results/ismara/ismara_report/activity_table",as.is=TRUE))
vulcan45<-vobj_tcga_45$msviper$es$nes
vulcan90<-vobj_tcga_90$msviper$es$nes

# Compare VULCAN (x axis) at 45 and 90 minutes with ISMARA (yaxis) average + sd

# Prepare ismara values
i45<-ismara[,grep("45",colnames(ismara))]
i90<-ismara[,grep("90",colnames(ismara))]
i45_mean<-apply(i45,1,mean)
i45_sd<-apply(i45,1,sd)
i90_mean<-apply(i90,1,mean)
i90_sd<-apply(i90,1,sd)

common<-intersect(names(vulcan45),rownames(ismara))
length(common) # 148
```

```{r, fig.cap=capfig("Comparison of results from the VULCAN and ISMARA methods"), message=FALSE, warning=FALSE}
# Plots
par(mfrow=c(2,1))

# 45 minutes
x<-vulcan45[common]
y<-i45_mean[common]
uiw<-i45_sd[common]
top<-intersect(names(sort(x,dec=TRUE))[1:20],names(sort(y,dec=TRUE))[1:20])
plotCI(x,y,uiw=uiw,xlab="VULCAN NES",ylab="ISMARA Activity",main="Comparison VULCAN/ISMARA",pch=20,col="grey",xlim=c(min(x),max(x)*1.1))
mtext("45 minutes",cex=0.8)
textplot2(x[top],y[top],words=top,new=FALSE,font=2)
# 90 minutes
x<-vulcan90[common]
y<-i90_mean[common]
uiw<-i90_sd[common]
top<-intersect(names(sort(x,dec=TRUE))[1:25],names(sort(y,dec=TRUE))[1:25])
plotCI(x,y,uiw=uiw,xlab="VULCAN NES",ylab="ISMARA Activity",main="Comparison VULCAN/ISMARA",pch=20,col="grey",xlim=c(min(x),max(x)*1.1))
mtext("90 minutes",cex=0.8)
textplot2(x[top],y[top],words=top,new=FALSE,font=2)
```

### 2018-01-23 ChIP-seq_GRHL2/Manuscript_Figures.Rmd
first commit

---
title: "GRHL2"
author: "Andrew Holding"
date: "1/18/2018"
output:
  pdf_document: default
  html_document: default
---

```{r setup_grhl2, include=FALSE ,warning=FALSE,echo=TRUE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(cache = TRUE)
```

## Pre-processing
### Download files

```{bash Download, eval=FALSE ,warning=FALSE,echo=TRUE}
#!/bin/sh
java -jar ../java/clarity-tools.jar -l SLX-14333
```

### Removal of reads in blacklisted sites

```{bash Blacklist, eval=FALSE ,warning=FALSE,echo=TRUE}
#downloaded from http://mitra.stanford.edu/kundaje/akundaje/release/blacklists/hg38-human/hg38.blacklist.bed.gz

bl=../blacklists/hg38.blacklist.bed

mkdir ./blacklist_filtered
cd SLX-14333
for f in *.bam
do
	echo $f
	bedtools intersect -v -abam $f -b $bl  > ../blacklist_filtered/$f
done

### Re-index

cd ../blacklist_filtered
for f in *.bam
do
samtools index $f
done

```

### Peak Calling

```{bash MACS, eval=FALSE ,warning=FALSE,echo=TRUE}
### MACS peak caller

### Run macs on the blacklisted data
mkdir ./peaks
cd peaks
control=../GRHL2_filtered/SLX-14333.D701_D503.bam
for bam in ../GRHL2_filtered/*.bam
do
root=`basename $bam .bam`
macs2 callpeak -t $bam -c $control -f BAM -n $root -g hs &
done

```

## Differntial binding analysis
### MA plot

```{r DiffBind MA Plot, fig.cap="MA plot showing changes in GRHL2 binding before and after treatment with 100nM E2." ,warning=FALSE,echo=TRUE}

suppressMessages(library(DiffBind))

if(!file.exists("rdata/003_diffbind.rda")) {
    GRHL2 <- dba(sampleSheet="samplesheet/samplesheet.csv")
    GRHL2 <- dba.count(GRHL2, summits=250)
    GRHL2 <- dba.contrast(GRHL2)
    GRHL2 <- dba.analyze(GRHL2)
    save(GRHL2,file="rdata/003_diffbind.rda")
} else {
    load("rdata/003_diffbind.rda")  
}

dba.plotMA(GRHL2, bFlip=1)
```

### PCA

```{r DiffBind PCA, fig.cap="PCA Plot showing clustering of samples by condition" ,warning=FALSE,echo=TRUE}

dba.plotPCA(GRHL2,components = 2:3)

```

### Table of differntiall bound sites

```{r DiffBind Table, results='asis' ,warning=FALSE,echo=TRUE}

library(knitr)
kable(head(as.data.frame(dba.report(GRHL2))))

```

### Number of sites with increased or decreased binding

```{r DiffBind Up and Down ,warning=FALSE,echo=TRUE}

r<-dba.report(GRHL2,th=1)
length(r[r$Fold>0])
length(r[r$Fold<0])

```

## Quality Control 
### Reproducability of peaks
```{r Overlap, fig.cap="Peak overlap +E2" ,warning=FALSE,echo=TRUE}

dba.plotVenn(GRHL2,GRHL2$masks$ER, label1="Rep1", label2="Rep2", main="GRHL2 +E2")
```

```{r Overlap2, fig.cap="Peak overlap -E2", warning=FALSE,echo=TRUE}
dba.plotVenn(GRHL2,GRHL2$masks$none, label1="Rep1", label2="Rep2", main="GRHL2 -E2")
```

```{r Called, warning=FALSE,echo=TRUE}

called_none<-rowSums(GRHL2$called[,c(1:3)])
called_ER<-rowSums(GRHL2$called[,c(4:6)])
print(paste("Peaks called in -E2 samples:", length(called_none[called_none>0])))
print(paste("Peaks called in +E2 samples:", length(called_ER[called_ER>0])))
called_both<-called_none*called_none
print(paste("Peaks called in both settings:", length(called_both[called_both>0])))
```

## VULCAN Analysis

```{r VULCAN ,warning=FALSE,echo=TRUE,results='hide', fig.cap="VULCAN analysis of GRHL2 ChIP-seq data showing change in TF activity on addition of E2 using both METABRIC and TCGA networks"}

suppressMessages(library(vulcan))

dist_calc<-function(method,dfanno,genematrix,genesmore,allsamples){
    # This function structure was strongly suggested
    # by the Bioconductor reviewer
    supportedMethods<-c(
        "closest",
        "strongest",
        "sum",
        "topvar",
        "farthest",
        "lowvar"
    )
    if(!method%in%supportedMethods){
        stop("unsupported method ", method)
    }

    for (gene in genesmore) {
        subanno <- dfanno[dfanno$feature == gene, ]

        if (method == "closest") {
            closest <- which.min(subanno$distanceToStart)
            genematrix[gene, allsamples] <- as.numeric(subanno[closest,
                                                               allsamples])
        }

        if (method == "farthest") {
            farthest <- which.max(subanno$distanceToStart)
            genematrix[gene, allsamples] <- as.numeric(subanno[farthest,
                                                               allsamples])
        }

        if (method == "sum") {
            sums <- apply(subanno[, allsamples], 2, sum)
            genematrix[gene, allsamples] <- as.numeric(sums)
        }

        if (method == "strongest") {
            sums <- apply(subanno[, allsamples], 1, sum)
            top <- which.max(sums)
            genematrix[gene, allsamples] <- as.numeric(subanno[top,
                                                               allsamples])
        }

        if (method == "topvar") {
            vars <- apply(subanno[, allsamples], 1, var)
            top <- which.max(vars)
            genematrix[gene, allsamples] <- as.numeric(subanno[top,
                                                               allsamples])
        }

        if (method == "lowvar") {
            vars <- apply(subanno[, allsamples], 1, var)
            top <- which.min(vars)
            genematrix[gene, allsamples] <- as.numeric(subanno[top,
                                                               allsamples])
        }

    }
    return(genematrix)
}

vulcan.import.dba<- function (dbaobj, samples,intervals = NULL)  
{
    dbcounts <- dbaobj
    listcounts <- dbcounts$peaks
    names(listcounts) <- dbcounts$samples[, 1]
    first <- listcounts[[1]]
    rawmat <- matrix(NA, nrow = nrow(first), ncol = length(listcounts) + 
                         3)
    colnames(rawmat) <- c("Chr", "Start", "End", names(listcounts))
    rownames(rawmat) <- 1:nrow(rawmat)
    rawmat <- as.data.frame(rawmat)
    rawmat[, 1] <- as.character(first[, 1])
    rawmat[, 2] <- as.integer(first[, 2])
    rawmat[, 3] <- as.integer(first[, 3])
    for (i in 1:length(listcounts)) {
        rawmat[, names(listcounts)[i]] <- as.numeric(listcounts[[i]]$RPKM)
    }
    peakrpkms <- rawmat
    rm(rawmat)
    first <- listcounts[[1]]
    rawmat <- matrix(NA, nrow = nrow(first), ncol = length(listcounts) + 
                         3)
    colnames(rawmat) <- c("Chr", "Start", "End", names(listcounts))
    rownames(rawmat) <- 1:nrow(rawmat)
    rawmat <- as.data.frame(rawmat)
    rawmat[, 1] <- as.character(first[, 1])
    rawmat[, 2] <- as.integer(first[, 2])
    rawmat[, 3] <- as.integer(first[, 3])
    for (i in 1:length(listcounts)) {
        rawmat[, names(listcounts)[i]] <- as.integer(listcounts[[i]]$Reads)
    }
    peakcounts <- rawmat
    rm(rawmat)
    vobj <- list(peakcounts = peakcounts, samples = samples, 
                 peakrpkms = peakrpkms)
    return(vobj)
}

prependSampleNames <- function(vobj,prependString){
    colnames(vobj$peakcounts)[0:-3]<-paste0(prependString,colnames(vobj$peakcounts)[0:-3])
    vobj$samples[[2]]<-paste0(prependString,vobj$samples[[2]])
    vobj$samples[[1]]<-paste0(prependString,vobj$samples[[1]])
    colnames(vobj$peakrpkms)[0:-3]<-paste0(prependString,colnames(vobj$peakrpkms)[0:-3])
    return(vobj)
}

loadVulcanNetworks<-function(){
    regulons<-list()
    load("networks/laml-tf-regulon.rda")
    regulons$laml<-regul
    rm(regul)
    load("networks/brca-tf-regulon.rda")
    regulons$tcga<-regul
    rm(regul)
    load("networks/metabric-regulon-tfs.rda")
    regulons$metabric<-regulon
    rm(regulon)
    return(regulons)
}

#Slow so just load the file below
#vobj<-vulcan.import("samplesheet/samplesheet.csv")
#load(file="003_vobj.Rda")
samples <- list()
samples[['ER']]<-c('1a','2a','3a')
samples[['none']]<-c('1b','2b','3b')

vobj <-vulcan.import.dba(GRHL2,samples)

#vobj<-vulcan.annotate(vobj,lborder=-10000,rborder=10000,method='sum')

vobj <-prependSampleNames(vobj,"X")

lborder=-10000
rborder=10000
method='sum'
#DEBUG
#source("https://bioconductor.org/biocLite.R")
#biocLite("TxDb.Hsapiens.UCSC.hg38.knownGene")
suppressMessages(library("TxDb.Hsapiens.UCSC.hg38.knownGene"))

annotation <- toGRanges(TxDb.Hsapiens.UCSC.hg38.knownGene, 
                        feature = "gene")
gr <- GRanges(vobj$peakcounts)
seqlevels(annotation)<- sub('chr','',seqlevels(annotation))
anno <- annotatePeakInBatch(gr, AnnotationData = annotation, 
                            output = "overlapping", FeatureLocForDistance = "TSS", 
                            bindingRegion = c(lborder, rborder))
dfanno <- anno
names(dfanno) <- seq_len(length(dfanno))
dfanno <- as.data.frame(dfanno)
allsamples <- unique(unlist(vobj$samples))
genes <- unique(dfanno$feature)
peakspergene <- table(dfanno$feature)
rawcounts <- matrix(NA, nrow = length(genes), ncol = length(allsamples))
colnames(rawcounts) <- allsamples
rownames(rawcounts) <- genes
genesone <- names(peakspergene)[peakspergene == 1]
for (gene in genesone) {
    rawcounts[gene, allsamples] <- as.numeric(dfanno[dfanno$feature == 
                                                         gene, allsamples])
}

genesmore <- names(peakspergene)[peakspergene > 1]
rawcounts <- dist_calc(method, dfanno, rawcounts, genesmore, 
                       allsamples)

gr <- GRanges(vobj$peakrpkms)
anno <- annotatePeakInBatch(gr, AnnotationData = annotation, 
                            output = "overlapping", FeatureLocForDistance = "TSS", 
                            bindingRegion = c(lborder, rborder))
dfanno <- anno
names(dfanno) <- seq_len(length(dfanno))
dfanno <- as.data.frame(dfanno)
allsamples <- unique(unlist(vobj$samples))
genes <- unique(dfanno$feature)
peakspergene <- table(dfanno$feature)
rpkms <- matrix(NA, nrow = length(genes), ncol = length(allsamples))
rownames(rpkms) <- genes
genesone <- names(peakspergene)[peakspergene == 1]
colnames(rpkms)<-allsamples
for (gene in genesone) {
    rpkms[gene, allsamples] <- as.numeric(dfanno[dfanno$feature == 
                                                     gene, allsamples])
}
genesmore <- names(peakspergene)[peakspergene > 1]

rpkms <- dist_calc(method, dfanno, rpkms, genesmore, allsamples)
rawcounts <- matrix(as.numeric(rawcounts), nrow = nrow(rawcounts), 
                    dimnames = dimnames(rawcounts))
rpkms <- matrix(as.numeric(rpkms), nrow = nrow(rpkms), dimnames = dimnames(rpkms))
vobj$rawcounts <- rawcounts
colnames(rpkms)<-allsamples
vobj$rpkms <- rpkms

#DEBUG ENDS

vobj<-vulcan.normalize(vobj)

load(file="networks/metabric-regulon-tfs.rda")

regulons <- loadVulcanNetworks()
suppressMessages(library("org.Hs.eg.db"))
list_eg2symbol<-as.list(org.Hs.egSYMBOL[mappedkeys(org.Hs.egSYMBOL)])

vobj_results<-list()

#test<-vulcan(vobj,       network=regulon, contrast=c("none","ER")    )

networks<-c("tcga","metabric") #,"laml")

for (network in networks) {
    vobj_results[[network]]<-vulcan(vobj,
                                       network=regulons[[network]],
                                       contrast=c("none","ER"),
                                       annotation=list_eg2symbol)
}

vobj_objects<-list(vobj_results)
names(vobj_objects)<-c("GRHL2")
networks<-c("tcga","metabric")
```

### Transcription factor activity

```{r Vulcan plots ,warning=FALSE,echo=TRUE}

plotVulcan <-function(vobj,threshold,network_title,title,plotTF,xlim,ylim) {
    threshold<-sign(threshold)*p2z(abs(threshold))
    network=vobj$mrs[,"NES"]
    tfs<-names(network)
    networkmat<-cbind(rep(0,length(network)),network[tfs])
    colnames(networkmat)<-c("0h","45min")
    matplot(t(networkmat),type="l",col="grey",ylab="VULCAN NES",xaxt="n",lty=1,main=title,xlim=xlim,ylim=ylim)
    axis(1,at=c(1:2),labels=colnames(networkmat))
    abline(h=c(0,threshold,-threshold),lty=2)
    text(2,networkmat[plotTF,2],label=names(networkmat[,2][plotTF]),pos=4,cex=0.6,font=2,col="red3")
    mtext(network_title)    
}

par(mfrow=c(2,3))
for(network in networks){
    for (vobj_name in names(vobj_objects))
    {

        vobj<-vobj_objects[[vobj_name]]
        vobj<-vobj[[network]]
        #TFs<-getTFs(vobj)
        TFs=c("ESR1","PGR","FOXA1","GRHL2") #Overide TFS
        plotVulcan(vobj,0.05, paste0(network," Network"),vobj_name,TFs,xlim=c(1,2.3), ylim=c(min(vobj$mrs[,"NES"]),max(vobj$mrs[,"NES"])))

        #TFs<-getTFs(vobj,0.05)
        plotVulcan(vobj,0.05, paste0(network," Network"),vobj_name,TFs,xlim=c(1,2.3), ylim=c(0,max(vobj$mrs[,"NES"])))

        #TFs<-getTFs(vobj,-0.05)
        plotVulcan(vobj,-0.05, paste0(network," Network"),vobj_name,TFs,xlim=c(1,2.3), ylim=c(min(vobj$mrs[,"NES"]),0))
    }
}

```

### METABRIC TGCA comparison

```{r vulcan scatter ,warning=FALSE,echo=TRUE, fig.cap="Comparision of results of VULCAN analysis for METABRIC and TGCA"}

interect<-intersect(rownames(vobj_results$metabric$mrs), 
                       rownames(vobj_results$tcga$mrs))
plot(-log10(vobj_results$metabric$mrs[interect,"pvalue"]),
     -log10(vobj_results$tcga$mrs[interect,"pvalue"]),
     pch=20, xlab="Metabric Network Enrichment Score", ylab="TCGA Network Enrichment Score",main="VULCAN analysis of GRHL2 ChIP-Seq")

filtered<-(log10(vobj_results$metabric$mrs[interect,"pvalue"])^2+log10(vobj_results$tcga$mrs[interect,"pvalue"])^2)>2

text(-log10(vobj_results$metabric$mrs[interect,"pvalue"][filtered]),
     -log10(vobj_results$tcga$mrs[interect,"pvalue"][filtered])-0.1,
     labels=interect[filtered])

points(-log10(vobj_results$metabric$mrs[interect,"pvalue"][c("GATA3","PGR","FOXA1","ESR1","GRHL2")]),
       -log10(vobj_results$tcga$mrs[interect,"pvalue"][c("GATA3","PGR","FOXA1","ESR1","GRHL2")]),pch=20,col="red")
text(-log10(vobj_results$metabric$mrs[interect,"pvalue"][c("GATA3","PGR","FOXA1","ESR1","GRHL2")]),
     -log10(vobj_results$tcga$mrs[interect,"pvalue"][c("GATA3","PGR","FOXA1","ESR1","GRHL2")])-0.1,col="red",
     labels=c("GATA3","PGR","FOXA1","ESR1","GRHL2"))
```

## Motif Analysis 
### Export bed files

```{r bed files up ,warning=FALSE,echo=TRUE}

#Note homer needs this as Hg format so you need to change from "1" to "chr1" etc.
df<-as.data.frame(dba.report(GRHL2))
df$seqnames<-paste0("chr",df$seqnames)

kable(head(df))
#write.bed(df,"bed/up.bed")
```

```{r bed file down ,warning=FALSE,echo=TRUE}
df_all<-as.data.frame(dba.report(GRHL2, th=1))
df_all$seqnames<-paste0("chr",df_all$seqnames)

kable(head(df_all))
#write.bed(df_all,"bed/all.bed")
```
### Homer

```{bash Motif, eval=FALSE ,warning=FALSE,echo=TRUE}

#Homer

mkdir motifAnalysis
cd motifAnalysis
findMotifsGenome.pl ../bed/up.bed hg38 grhl2UpSites
findMotifsGenome.pl ../bed/all.bed hg38 grhl2AllSites

```

### Generate promoter locations

Generate Bed file of Promoters

```{r, eval=FALSE ,warning=FALSE,echo=TRUE}

require(TxDb.Hsapiens.UCSC.hg38.knownGene)
txdb <- TxDb.Hsapiens.UCSC.hg38.knownGene
p<-promoters(genes(txdb), upstream = 1500, downstream = 500)

write.bed<-function(df, filename){
    write.table(as.data.frame(df)[,1:3],
                filename, quote=FALSE, sep="\t",
                row.names=FALSE, col.names=FALSE)
}

write.bed(p,"bed/promoter.bed")

```

## Ovlerap of GRHL2 binding sites with published data

### Extract ChIA-PET Data

```{bash Merge ChIApet, eval=FALSE}
#ChIA-pet all 3 https://www.encodeproject.org/experiments/ENCSR000BZZ/
cd bed
multiIntersectBed -i *bed6_sorted.bed | awk '{if ($4 > 1) {print} }' > hglft_ChIA_combined.bed
```

### Find overlapping sites

```{bash, eval=FALSE ,warning=FALSE,echo=TRUE}
#p300 from Zwart, EMBO, 2011
#ER from Hurtado 2011
#gro-seq GSE43835 
#Rest from Caroll MLL3 paper

cd bed
mkdir overlaps

#generate gro-seq interect
#file1=hglft_E2.40m.rep1.bed 
#file2=hglft_E2.40m.rep2.bed 
#output=gro-seq.bed
#bedtools intersect -sorted -a $file1 -b $file2 > overlaps/$output

file1=hglft_foxa1.bed 
file2=up.bed
output=foxa1-up.bed
bedtools intersect -a $file1 -b $file2 > overlaps/$output

file1=hglft_ER_Hurtado_2011.bed 
output=er-up.bed
bedtools intersect -a $file1 -b $file2 > overlaps/$output

#file1=gro-seq.bed 
#output=gro-up.bed
#bedtools intersect -a $file1 -b $file2 > overlaps/$output

file1=hglft_ChIA_combined.bed 
output=ChIAPet_up.bed
bedtools intersect -a $file1 -b $file2 > overlaps/$output

file1=hglft_h3k4me1.bed 
output=h3k4me1-up.bed
bedtools intersect -a $file1 -b $file2 > overlaps/$output

file1=hglft_h3k4me3.bed 
output=h3k4me3-up.bed
bedtools intersect -a $file1 -b $file2 > overlaps/$output

file1=hglft_p300_ctrl.bed  
output=p300_ctrl_up.bed
bedtools intersect -a $file1 -b $file2 > overlaps/$output

file1=hglft_p300_e2.bed  
output=p300_e2_up.bed
bedtools intersect -a $file1 -b $file2 > overlaps/$output

#all peaks

file1=hglft_foxa1.bed 
file2=all.bed
output=foxa1-all.bed
bedtools intersect -a $file1 -b $file2 > overlaps/$output

file1=hglft_ER_Hurtado_2011.bed 
output=er-all.bed
bedtools intersect -a $file1 -b $file2 > overlaps/$output

#file1=gro-seq.bed 
#output=gro-all.bed
#bedtools intersect -a $file1 -b $file2 > overlaps/$output

file1=hglft_ChIA_combined.bed  
output=ChIAPet_all.bed
bedtools intersect -a $file1 -b $file2 > overlaps/$output

file1=hglft_h3k4me1.bed 
output=h3k4me1-all.bed
bedtools intersect -a $file1 -b $file2 > overlaps/$output

file1=hglft_h3k4me3.bed 
output=h3k4me3-all.bed
bedtools intersect -a $file1 -b $file2 > overlaps/$output

file1=hglft_p300_ctrl.bed  
output=p300_ctrl_all.bed
bedtools intersect -a $file1 -b $file2 > overlaps/$output

file1=hglft_p300_e2.bed  
output=p300_e2_all.bed
bedtools intersect -a $file1 -b $file2 > overlaps/$output

##Overlap with promoters

file1=overlaps/ChIAPet_up.bed   
file2=promoter.bed
output=ChIAPet_up_prom.bed
bedtools intersect -a $file1 -b $file2 > overlaps/$output

file1=overlaps/ChIAPet_all.bed   
file2=promoter.bed
output=ChIAPet_all_prom.bed
bedtools intersect -a $file1 -b $file2 > overlaps/$output

```

### Count number of overlaping sites

```{bash, eval=FALSE ,warning=FALSE,echo=TRUE}
cd bed/overlaps
wc -l *.bed > overlaps.txt
```

### Table of number of overlaping sites for each factor

```{r, warning=FALSE,echo=TRUE ,warning=FALSE,echo=TRUE}
overlaps<-read.table("bed/overlaps/overlaps.txt")
kable(overlaps)
```

### GRHL2 binding overlap with known factors

```{r, warning=FALSE,echo=TRUE ,warning=FALSE,echo=TRUE,fig.cap="GRHL2 binding overlap with known factors"}

overlapsDF<-cbind(overlaps[1:16,],c('all','all','up','up','all','up','all','up','all','up','all','up','all','up','all','up'))
overlapsDF<-cbind(overlapsDF,c('all','promoter','all','promoter',rep("all",12)))
overlapsDF<-cbind(overlapsDF,c(rep("None",14),"E2","E2"))
colnames(overlapsDF)<-c("Number","Factor","GRHL2","Feature","Treatment")
overlapsDF$Factor<-c(rep("ChIAPet",4),rep("ER",2),rep("FOXA1",2),rep("H3K4Me1",2),rep("H3K4Me3",2),rep("P300",4))
overlapsDF$Number<-as.numeric(overlapsDF$Number)
ChIAPet<-overlapsDF[overlapsDF$Factor=='ChIAPet' & overlapsDF$Feature=='all',]
ChIAPet$Number<-overlapsDF[overlapsDF$Factor=='ChIAPet' & overlapsDF$Feature=='all',]$Number - overlapsDF[overlapsDF$Factor=='ChIAPet' & overlapsDF$Feature=='promoter',]$Number
ChIAPet$Feature<-c('enchancer','enchancer')
overlapsDF<-rbind(overlapsDF,ChIAPet)
overlapsDF<-data.frame(overlapsDF)
overlapsDF[overlapsDF$GRHL2=='up',]$Number<-(overlapsDF[overlapsDF$GRHL2=='up',]$Number/355)*100 #355 is the total number of up sites established by DiffBind
overlapsDF[overlapsDF$GRHL2=='all',]$Number<-(overlapsDF[overlapsDF$GRHL2=='all',]$Number/42721)*100 #Total number of GRHL2 sites

library(lattice)

barchart(Number ~ Factor ,data=overlapsDF[overlapsDF$Factor != 'ChIAPet',], 
         par.settings = simpleTheme(col=c("lightblue","pink")),
         group=GRHL2,auto.key=list(space="top", columns=2, text=c('All GRHL2 Sites','GRHL2 Sites Responsive to E2')),
         main="GRHL2 Binding Overlap with Known Factors", ylab="Percentage of GRHL2 sites that overlap", 
         xlab="Factor")
```

### GRHL2 binding overlap with P300 binding sites

```{r P300, fig.cap="GRHL2 binding overlap with P300 binding sites"}

barchart(Number ~ Treatment ,data=overlapsDF[overlapsDF$Factor=='P300',],
         group=GRHL2,
         auto.key=list(space="top",columns=2, text=c('All GRHL2 Sites','GRHL2 Responsive to E2')),
         par.settings = simpleTheme(col=c("lightblue","pink")),
         main="GRHL2 overlap with P300 binding sites",
         xlab="P300 ChIP-Seq Condition",
         ylab="Percentage of GRHL2 sites that overlap")
```

### GRHL2 overlap with ChIA-PET sites

```{r ChIAPet, fig.cap="GRHL2 overlap with ChIA-PET sites"}

barchart(Number ~ Feature ,data=overlapsDF[overlapsDF$Factor=='ChIAPet',],
         group=GRHL2,
         auto.key=list(space="top",columns=2, text=c('All GRHL2 Sites','GRHL2 Sites Responsive to E2')),
         par.settings = simpleTheme(col=c("lightblue","pink")),
         main="GRHL2 overlap with ChIA-PET sites",
         xlab="Genomic Feature ChIP-Seq Condition",
         ylab="Percentage of GRHL2 sites that overlap")
```

---
#Gviz for chr2:11,517,385-11,630,055 #GREB1 and chr17:40,296,682-40,349,154 #RARA
---

### 2018-01-23 Manuscript_Figures_Combined.Rmd
first commit

---
title: "VULCAN Analysis Combined"
author: "Andrew Holding"
date: "1/22/2018"
output:
  pdf_document:
    fig_caption: yes
    fig_height: 8
    fig_width: 8
    highlight: tango
    number_sections: yes
    toc: yes
    toc_depth: 3
  html_document:
    fig_caption: yes
    fig_height: 8
    fig_width: 8
    highlight: tango
    number_sections: yes
    theme: united
    toc: yes
    toc_depth: 3
always_allow_html: yes
---

```{r setup_master, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(cache = TRUE)
```

#Estrogen Receptor-alpha qPCR Analysis

```{r ER qRPCR, child = 'qPCR_ER/Manuscript_Figures.Rmd'}
```

#Estrogen Receptor-alpha ChIP-Seq and VULCAN Analysis

```{r ER ChIP-seq, child = 'ChIP-seq_ER/rmarkdown.Rmd'}
```

#Estrogen Receptor-alpha qPLEX-RIME Analysis

```{r ER qRIME, child = 'qPLEX-RIME_ER/02.Supplementary.Rmd'}
```

#GRHL2 ChIP-Seq Analysis

```{r GRHL2 Chip, child = 'ChIP-seq_GRHL2/Manuscript_Figures.Rmd'}
```

#GRHL2 qPLEX-RIME Analysis

```{r GRHL2 qRIME, child = 'qPLEX-RIME_GRHL2/rRunTMTanalysis.Rmd'}
```

# Technical Session Info
The following code describes the R environment used to generate this document and will help making it fully reproducible should there be future updates in any of the packages.
```{r sessionInfo}
sessionInfo()
```

### 2018-01-23 qPLEX-RIME_GRHL2/rRunTMTanalysis.Rmd
first commit

---
title: |
  | Proteomics RIME-TMT analysis report
  | Holding: Changes in GRHL2 interactome on Activation
always_allow_html: yes
output:
  pdf_document:
    toc: yes
    toc_depth: '3'
  html_notebook:
    number_sections: yes
    toc: yes
    toc_depth: 3
  html_document:
    toc: yes
    toc_depth: '3'

---

```{r setup_qplex, include=FALSE}

library(UniProt.ws)
library(plyr)
library(dplyr)
library(magrittr)
library(qPLEXanalyzer)
library(knitr)
library(readxl)
library(grid)
library(ggplot2)
library(DT)
library(rmarkdown)
options(stringsAsFactors=F)

# knitr options
opts_knit$set(eval.after = 'fig.cap')
```

```{r numberingFunctions, include=F}
# This deals with captioning and links to figure and tables
# Autonumbering is handled automatically in PDFs, but with HTML we need a function to deal with it

# reflink generates links to figures/tables.
# refcaption creates an anchor and a caption for a table/figure
# the "refName" argument matches these together
# e.g. reflink("myFig") will create a link in the text to the figure with a caption created
#   by refcaption("myFig", "This is the caption")
# add refType="Table" to link/caption a table

numberingList <- list(Figure=list(startFig=0), Table=list(startTab=0))

getNumber <- function(refName, refType, numList=numberingList){
    if(!refName%in%names(numList[[refType]])){
        numList[[refType]][[refName]] <- max(unlist(numList[[refType]]))+1 
        numberingList <<- numList
    }
    refNum <- numList[[refType]][[refName]]
    return(refNum)
}
reflink <- function(refName, refType="Figure"){
    refNum <- getNumber(refName, refType)
    paste0("[", refType, " ", refNum, "](#", refName, ")")
}
refcaption <- function(refName, refCaption, refType="Figure"){
    refNum <- getNumber(refName, refType)
    if(refType=="Table"){
        cap <- htmltools::tags$caption(style = 'caption-side: top; text-align: center;',
                                       htmltools::strong(paste0('Table ', refNum, ': ', refCaption)))
    }else{
        cap <- paste0("**", refType, " ", refNum, ". ", refCaption, "  **")
    }
    cat(paste0("<a id=\"", refName, "\"></a>"))
    return(cap)
}

```

```{r myCovPlotFunct, include=F}
coveragePlot2 <- function(data, ProteinID, ProteinName, fastaFile, myCol="brown")
{
    suppressPackageStartupMessages(library(GenomicRanges))
    suppressPackageStartupMessages(library(Biostrings))

    if(!is(data,"MSnSet"))
    stop('data has to be of class MSnSet..')
    if(!is(ProteinID,"character"))
    stop('ProteinID has to be of class character..')
    if(!is(ProteinName,"character"))
    stop('ProteinName has to be of class character..')
    if(!is(fastaFile,"character"))
    stop('fastaFile has to be of class character..')
    if(!is(myCol,"character"))
    stop('myCol has to be of class character..')
    ProteinName <- as.character(ProteinName)

    # extract peptide sequence from MsnSet object
    peptide_seq <- fData(data) %>% 
        filter(Master.Protein.Accessions==ProteinID) %>% 
        use_series(Annotated.Sequence) %>% 
        gsub("^.{4}([A-Z]+).{4}$", "\\1", .)

    ## read protein sequence from fastafile
    Protein_seq <- readAAStringSet(fastaFile)

    ## match peptide sequence with protein sequence and store co-ordinates
    getPosition <- function(peptideSeq, ProteinSeq=Protein_seq){
        vmatchPattern(peptideSeq, ProteinSeq) %>% 
            as.data.frame() %>% 
            dplyr::select(start, end) %>% 
            return()
    }

    features <- lapply(peptide_seq, getPosition) %>% 
      do.call(rbind, .)

    # get percent coverage 
    protWidth <- width(Protein_seq)
    coverage <- GRanges("feature",IRanges(features$start,features$end)) %>% reduce() %>% width() %>% sum()
    Perct <- round(coverage/protWidth*100,2)
    SubTitle <- paste0("Number of Unique Peptides: ", length(unique(peptide_seq)), "\n% Coverage: ", Perct)

    nTicks <- min(c(7, ceiling(protWidth/50) + 1))
    brkTicks <- round(seq(0,protWidth, length.out = nTicks), 0)

    features %>% 
      distinct() %>% 
      ggplot() +
          geom_rect(aes(xmin=start-1, xmax=end, ymin=0, ymax=10), fill=myCol) +
          geom_rect(xmin=0, xmax=protWidth, ymin=0, ymax=10, colour="black", fill=NA, size=0.15) +
          labs(title=ProteinName,
               subtitle=SubTitle) +
          theme(axis.title.y=element_blank(),
                axis.text.y=element_blank(),
                axis.ticks.y=element_blank(),
                panel.background = element_rect(fill = "white"), 
                panel.border = element_blank(),
                plot.subtitle = element_text(hjust = 0.5),
                plot.title = element_text(hjust = 0.5)) +
          scale_x_continuous(limits=c(0,protWidth), breaks=brkTicks) +
          scale_y_continuous(limits=c(0,10), breaks=c(0,10), expand = c(0, 0)) %>% 
    return()
}

```

<!--
Project Title: Holding: Changes in GRHL2 interactome on Activation
Redmine number: 5045
Proteomics LIMS:
    Project Name: GRHL2 0-45 minutes
    Project ID: PR882
    Quantification method: TMT
Reagent used: 11-plex
-->

```{r getData, include=F}

# constants
baitProtein <- "GRHL2"
UniProtAccession <- "Q6ISB3"
intensitiesFile <- "intensities/Velos_PR882_ANH_07122017_Peptide_Intensities.xlsx"
outputDir <- "results/"

# read the intensities excel file and replace NA in the IgG columns with 0.001
intensities <-  read_xlsx(intensitiesFile, 1) %>% 
    mutate_at(vars(starts_with("IGG")), funs(ifelse(is.na(.), 0.001, .)))

# the sample names used by proteomics start with numbers, so R has adds an X to the beginning of the 
# column names. Change the column names to so the "0min" is "Ctrl" "45min" is "ER"
# also "read_xlsx" leaves spaces in the column names, need to replace these with "."
colnames(intensities) <- gsub(" ", ".", gsub("X*0min", "Ctrl", gsub("X*45min", "ER", colnames(intensities))))

# create a metadata table
metadata <- data.frame(Experiment=colnames(intensities)[7:17],
                       Label=factor(c(rep("Ctrl", 4), rep("ER", 4), rep("IgG", 3))),
                       Bio.Rep=c(rep(1:4, 2), 1:3),
                       Tech.Rep=NA,
                       Analyt.Rep=NA,
                       Grp=c(rep(1, 8), rep(2, 3)))

# create MSnset object
rawMSnset <- convertToMSnset(data = intensities, metadata = metadata, 
                             indExpData = 7:17, indFData = 1:6,
                             rmMissing=TRUE)

# fasta file of protein sequence
fastaFile <- paste0("accessoryFiles/", UniProtAccession, ".fasta")
fastaUrl <- paste0("http://www.uniprot.org/uniprot/", UniProtAccession, ".fasta")
download.file(fastaUrl, fastaFile, "wget")

# annotation table
humanUP <- UniProt.ws::UniProt.ws(taxId=9606)
keys <- unique(intensities$Master.Protein.Accessions)
columns <- c("ENTRY-NAME", "PROTEIN-NAMES", "GENES")
annot <- UniProt.ws::select(humanUP, keys, columns, "UNIPROTKB") %>% 
    mutate(GeneSymbol=gsub(" .*", "", GENES)) %>% 
    dplyr::select(Protein="UNIPROTKB", Gene="ENTRY-NAME", Description="PROTEIN-NAMES", GeneSymbol)
```

```{r dataForIntroduction, include=F}

# count protein peptide representation
numberOfSamples <- nrow(metadata)
numberOfPeptides <- nrow(intensities)
numberOfProteins <- length(unique(intensities$Master.Protein.Accessions))

peptidesPerProtein <- table(intensities$Master.Protein.Accessions)
```

## Introduction

This report provides a summary of the results of your proteomics experiment.

Your experiment has **`r numberOfSamples`** samples. **`r numberOfPeptides`** peptides from **`r numberOfProteins`** 
unique proteins were captured in the experiment. Each protein is represented by **`r min(peptidesPerProtein)`** to **`r max(peptidesPerProtein)`**
peptides, but median number of peptides per protein was **`r median(peptidesPerProtein)`**.

## Raw data QC

### Coverage plot

`r reflink("covFig")` shows the coverage of the bait protein, `r baitProtein`, in terms of peptides detected.

```{r coveragePlot, echo=F, fig.width=10, fig.height=2, fig.cap=covCap, results="asis"}
covCap <- refcaption("covFig", paste0("Plot of peptide coverage for ", baitProtein, " (UniprotID: ", UniProtAccession, ")"))
coveragePlot2(data=rawMSnset, ProteinID=UniProtAccession, ProteinName=baitProtein, fastaFile=fastaFile)
```

### Intensity Plot

`r reflink("intFig")` shows the distribution of raw peptide intensities for each sample.

```{r rawIntensityPlot, echo=F, fig.width=10, fig.cap=intCap, results="asis"}
intCap <- refcaption("intFig", "Raw intensities plot.")
sampleColours <- assignColours(rawMSnset, groupColourSchemes = 
                                 c("Blues","Reds","Greens"))
rawIP <- intensityPlot(rawMSnset, sampleColours, title = "Peptide intensity distribution")
print(rawIP)
```

### Peptide intensities for GHRL2

`r reflink("baitIntFig")` shows the raw intensities for each peptide detected for the bait protein `r baitProtein`
in each sample.

```{r baitPeptideIntensities, echo=F, fig.width=10, fig.cap=baitIntCap, results="asis"}
baitIntCap <- refcaption("baitIntFig", "Peptide intensities for GRHL2.")
peptideIntensityPlot(rawMSnset, combinedIntensities=NULL, UniProtAccession,
                                 title = "", ylab = "log2(peptide intensity)",
                                 minIntensity = NA, maxIntensity = NA,
                                 selectedSequence = NULL, selectedModifications = NULL)
```

### Correlation Plot

`r reflink("corFig")` shows a correlation matrix to visualize the level of linear association of samples
within and between groups based on the raw peptide intensities.

```{r rawCorrPlot, echo=F, fig.width=10, fig.cap=corCap, results="asis"}
corCap <- refcaption("corFig", "Correlation plot based on raw intensities.")
corrPlot(rawMSnset, method="shade")
```

### Hierichical clustering dendrogram

`r reflink("hcFig")` shows a dendrogram displaying the hierarchical relationship among samples. The vertical axis 
shows the dissimilarity (measured by  means  of  the  Euclidean  distance)  between  samples:  similar  
samples  appear  on  the  same  branches. Colors  correspond  to sample groups.

```{r rawHCPlot, echo=F, fig.width=10, fig.cap=hcCap, results="asis"}
hcCap <- refcaption("hcFig", "Hierachical clustering based on raw intensities.")
hcTmp <- rawMSnset
exprs(hcTmp) <- exprs(hcTmp)+0.01
label_color <- c(rep("red",4),rep("darkgreen",4),rep("orange",2))
hierarchicalPlot(hcTmp,label_color,branchlength=95,title="Clustering plot")
rm(hcTmp)

```

### PCA Plot

`r reflink("pcFig")` shows a visual representation of the scaled loading of the first two dimensions of a principle
component analysis of the raw peptide intensities.

```{r rawPCAPlot, echo=F, fig.width=10, fig.cap=pcCap, results="asis"}
pcCap <- refcaption("pcFig", "Principle Component Analysis based on raw intensities.")
grpcolor <- c("red", "darkgreen", "orange")
rawPCP <- qPLEXanalyzer::pcaPlot(rawMSnset, groupColours=grpcolor, title = "PCA plot", 
        labels = rawMSnset$Bio.Rep, labelsize=2, legend=TRUE)# + 
    theme(aspect.ratio=1)
print(rawPCP)
```

### QC Conclusion

The sample Ctrl rep 1 appears to be an outlier. The differential anlaysis was carried out twice, first
with this sample included and then again with the sample excluded.

## Full data set analysis

The following section shows the results for analysis of all 8 samples.

```{r analysisFull, include=F}
# normalisation
normFullMSnset <- groupScaling(rawMSnset,func=median, Grp="Grp")
pnormFullMSnset <- summarizeIntensities(normFullMSnset, sum, annot)

# differential analysis
contrasts <- c(ER_v_Ctrl = "ER - Ctrl")
contrastName <- gsub(" - ", "_v_", contrasts)
diffstats <- computeDiffStats(data=pnormFullMSnset, contrasts=contrasts, 
                              applyLog2Transform = TRUE)
diffexpFull <- getContrastResults(diffstats=diffstats, contrast=contrasts, 
                              controlGroup = "IgG", ann = annot) %T>% 
    write.csv(paste0(outputDir, "/DifferentialAbundance.", contrastName, ".FullData.csv"), row.names=F)
```

### Effect of within group normalisation

Normalisation was carried out using median scaling. The experimental samples and the IgG control samples
were normalised separately.
`r reflink("fullQCFig")` show the effects of normalisation on the intensity plots and the principle component
analysis.

```{r withinGroupNormaliation, echo=F, fig.width=10, fig.cap=fullQCCap, results="asis"}
fullQCCap <- refcaption("fullQCFig", "Effects of normalisation on peptide intensities.")
normIP <- intensityPlot(normFullMSnset, sampleColours, title = "WithinGrp Scaling") + 
    theme(legend.position = "bottom")
rawPCP <- qPLEXanalyzer::pcaPlot(normFullMSnset, groupColours=grpcolor, title = "PCA plot", 
        labels = rawMSnset$Bio.Rep, labelsize=2, legend=TRUE) +
    theme(legend.position = "bottom")

grid.newpage()
pushViewport(viewport(layout = grid.layout(nrow = 1, ncol = 2)))
 define_region <- function(row, col){
   viewport(layout.pos.row = row, layout.pos.col = col)
}
print(normIP, vp = define_region(row = 1, col = 1)) 
print(rawPCP, vp = define_region(row = 1, col = 2))

```

### Differential analysis results

`r reflink("fullDATab")` shows differential abundancy results. No proteins were statistically differentially abundant.
`r reflink("fullQCFig")` shows an MA plot and a volcano plot of the differental analysis results.

```{r daResTabFull, echo=F, results="asis"}

deFColnames <- c("Protein", "Gene", "Description", "Gene Symbol", "log2(Fold Change)", "Adjusted p-value")
deFCap <- refcaption("fullDATab", "Differential abundancy results using all 8 samples.", refType = "Table")
deFOptions <- list(columnDefs = list(list(className = 'dt-center', targets="_all")), dom='ltip')
diffexpFull %>% 
    dplyr::select(Protein, Gene, Description, GeneSymbol, logFC, adj.P.Val) %>% 
    datatable(colnames=deFColnames, options=deFOptions, rownames = FALSE, caption=deFCap)

```

```{r resPlotFull, echo=F, fig.width=10, fig.cap=fullDACap, results="asis"}
fullDACap <- refcaption("fullDAFig", "MA plot and volcano plot of differential protein abundancy.")

MAPlot <- qPLEXanalyzer::maPlot(diffexpFull, significanceLevel = 0.0025, minLogFoldChangeForLabelling = 1.5,
             title="")  +
          labs(y="log2(Fold Change)")

volPlot <- diffexpFull %>% 
    mutate(phredAdjPval=-log10(adj.P.Val)) %>% 
    mutate(Signif=ifelse(adj.P.Val<=0.01, "<=0.01", ifelse(adj.P.Val<=0.05, "<=0.05", ">0.05"))) %>% 
    ggplot() +
        geom_point(aes(x=logFC, y=phredAdjPval, fill=Signif), colour="black", shape=21) +
        scale_fill_manual(values=c("red", "orange", "black"), name="Adjusted p-value", limits=c("<=0.01", "<=0.05", ">0.05"), drop = TRUE) +
        theme_minimal() +
        labs(x="log2(Fold Change)", y="-log10(Adjusted p-value)") +
    theme(legend.position = "bottom")

grid.newpage()
pushViewport(viewport(layout = grid.layout(nrow = 1, ncol = 2)))
 define_region <- function(row, col){
   viewport(layout.pos.row = row, layout.pos.col = col)
}
print(MAPlot, vp = define_region(row = 1, col = 1)) 
print(volPlot, vp = define_region(row = 1, col = 2))

```

## Remove sample Ctrl 1

The following section shows the same analysis, but with the Ctrl rep 1 sample removed

```{r analysisTrim, include=F}
trimMSnset <- rawMSnset[,-1]
sampleColours <- assignColours(trimMSnset, groupColourSchemes = c("Blues","Reds","Greens"))

# normalisation
normTrimMSnset <- groupScaling(trimMSnset, func=median, Grp="Grp")

# summarise
pnormTrimMSnset <- summarizeIntensities(normTrimMSnset, sum, annot)
```

### Effect of within group normalisation

Normalisation was carried out using median scaling. The experimental samples and the IgG control samples
were normalised separately.
`r reflink("trimQCFig")` show the effects of normalisation on the intensity plots and the principle component
analysis.

```{r withinGroupNormaliationTrim, echo=F, fig.width=10, fig.cap=trimQCCap, results="asis"}
trimQCCap <- refcaption("trimQCFig", "Effects of normalisation on peptide intensities.")
normIP <- intensityPlot(normTrimMSnset, sampleColours, title = "WithinGrp Scaling") +
    theme(legend.position = "bottom")
rawPCP <- qPLEXanalyzer::pcaPlot(normTrimMSnset, groupColours=grpcolor, title = "PCA plot", 
        labels = trimMSnset$Bio.Rep, labelsize=2, legend=TRUE) +
    theme(legend.position = "bottom")

grid.newpage()
pushViewport(viewport(layout = grid.layout(nrow = 1, ncol = 2)))
 define_region <- function(row, col){
   viewport(layout.pos.row = row, layout.pos.col = col)
}
print(normIP, vp = define_region(row = 1, col = 1)) 
print(rawPCP, vp = define_region(row = 1, col = 2))

```

### ER - Ctrl

```{r contrastERCtrl, include=F}
# differential analysis
contrasts <- c(ER_v_Ctrl = "ER - Ctrl")
contrastName <- gsub(" - ", "_v_", contrasts)
diffstats <- computeDiffStats(data=pnormTrimMSnset, contrasts=contrasts, 
                              applyLog2Transform = TRUE)
diffexpTrim <- getContrastResults(diffstats=diffstats, contrast=contrasts, 
                              controlGroup = "IgG", ann = annot) %T>% 
    write.csv(paste0(outputDir, "/DifferentialAbundance.", contrastName, ".SevenSamples.csv"), row.names=F)
```

`r reflink("trimDATab")` shows differential abundancy results for the contrasts `r contrastName`.
`r reflink("trimQCFig")` shows an MA plot and a volcano plot of the differental analysis results

```{r daResTabTrim, echo=F, results="asis"}

deFCap <- refcaption("trimDATab", "Differential abundancy results using all 8 samples.", refType = "Table")
deFColnames <- c("Protein", "Gene", "Description", "Gene Symbol", "log(Fold Change)", "Adjusted p-value")
deFOptions <- list(columnDefs = list(list(className = 'dt-center', targets="_all")), dom='ltip')
diffexpTrim %>% 
    dplyr::select(Protein, Gene, Description, GeneSymbol, logFC, adj.P.Val) %>% 
    datatable(colnames=deFColnames, options=deFOptions, rownames = FALSE, caption=deFCap)

```

```{r resPlotTrim, echo=F, fig.width=10, fig.cap=trimDACap, results="asis"}
trimDACap <- refcaption("trimDAFig", "MA plot and volcano plot of differential protein abundancy.")

MAPlot <- qPLEXanalyzer::maPlot(diffexpTrim, significanceLevel = 0.0025, minLogFoldChangeForLabelling = 1.5,
             title=names(contrasts)) +
          labs(y="log2(Fold Change)")

volPlot <- diffexpTrim %>% 
    mutate(phredAdjPval=-log10(adj.P.Val)) %>% 
    mutate(Signif=ifelse(adj.P.Val<=0.01, "<=0.01", ifelse(adj.P.Val<=0.05, "<=0.05", ">0.05"))) %>% 
    ggplot() +
        geom_point(aes(x=logFC, y=phredAdjPval, fill=Signif), colour="black", shape=21) +
        scale_fill_manual(values=c("red", "orange", "black"), name="Adjusted p-value", limits=c("<=0.01", "<=0.05", ">0.05")) +
        theme_minimal() +
        labs(x="log2(Fold Change)", y="-log10(Adjusted p-value)") +
    theme(legend.position = "bottom")

grid.newpage()
pushViewport(viewport(layout = grid.layout(nrow = 1, ncol = 2)))
 define_region <- function(row, col){
   viewport(layout.pos.row = row, layout.pos.col = col)
}
print(MAPlot, vp = define_region(row = 1, col = 1)) 
print(volPlot, vp = define_region(row = 1, col = 2))

```

### Ctrl - IgG

```{r contrastCtrlIgG, include=F}
# differential analysis
contrasts <- c(Ctrl_v_IgG = "Ctrl - IgG")
contrastName <- gsub(" - ", "_v_", contrasts)
diffstats <- computeDiffStats(data=pnormTrimMSnset, contrasts=contrasts, 
                              applyLog2Transform = TRUE)
diffexpCtrlIgG <- getContrastResults(diffstats=diffstats, contrast=contrasts, 
                              controlGroup = NULL, ann = annot) %T>% 
    write.csv(paste0(outputDir, "/DifferentialAbundance.", contrastName, ".SevenSamples.csv"), row.names=F)
```

`r reflink("CtrlIgGDATab")` shows differential abundancy results for the contrasts `r contrastName`.
`r reflink("CtrlIgGQCFig")` shows an MA plot and a volcano plot of the differental analysis results.

```{r daResTabCtrlIgG, echo=F, results="asis"}

deFCap <- refcaption("CtrlIgGDATab", "Differential abundancy results using all 8 samples.", refType = "Table")
deFColnames <- c("Protein", "Gene", "Description", "Gene Symbol", "log(Fold Change)", "Adjusted p-value")
deFOptions <- list(columnDefs = list(list(className = 'dt-center', targets="_all")), dom='ltip')
diffexpCtrlIgG %>% 
    dplyr::select(Protein, Gene, Description, GeneSymbol, logFC, adj.P.Val) %>% 
    datatable(colnames=deFColnames, options=deFOptions, rownames = FALSE, caption=deFCap)

```

```{r resPlotCtrlIgG, echo=F, fig.width=10, fig.cap=CtrlIgGDACap, results="asis"}
CtrlIgGDACap <- refcaption("CtrlIgGDAFig", "MA plot and volcano plot of differential protein abundancy.")

MAPlot <- qPLEXanalyzer::maPlot(diffexpCtrlIgG, significanceLevel = 0.00, minLogFoldChangeForLabelling = 1000,
                 title=names(contrasts), ylab="log2(Fold Change)")

volPlot <- diffexpCtrlIgG %>% 
    mutate(phredAdjPval=-log10(adj.P.Val)) %>% 
    mutate(Signif=ifelse(adj.P.Val<=0.01, "<=0.01", ifelse(adj.P.Val<=0.05, "<=0.05", ">0.05"))) %>% 
    ggplot() +
        geom_point(aes(x=logFC, y=phredAdjPval, fill=Signif), colour="black", shape=21) +
        scale_fill_manual(values=c("red", "orange", "black"), name="Adjusted p-value", limits=c("<=0.01", "<=0.05", ">0.05")) +
        theme_minimal() +
        labs(x="log2(Fold Change)", y="-log10(Adjusted p-value)") +
    theme(legend.position = "bottom")

grid.newpage()
pushViewport(viewport(layout = grid.layout(nrow = 1, ncol = 2)))
 define_region <- function(row, col){
   viewport(layout.pos.row = row, layout.pos.col = col)
}
print(MAPlot, vp = define_region(row = 1, col = 1)) 
print(volPlot, vp = define_region(row = 1, col = 2))

```

### ER - IgG

```{r contrastERIgG, include=F}
# differential analysis
contrasts <- c(ER_v_IgG = "ER - IgG")
contrastName <- gsub(" - ", "_v_", contrasts)
diffstats <- computeDiffStats(data=pnormTrimMSnset, contrasts=contrasts, 
                              applyLog2Transform = TRUE)
diffexpERIgG <- getContrastResults(diffstats=diffstats, contrast=contrasts, 
                              controlGroup = NULL, ann = annot) %T>% 
    write.csv(paste0(outputDir, "/DifferentialAbundance.", contrastName, ".SevenSamples.csv"), row.names=F)
```

`r reflink("ERIgGDATab")` shows differential abundancy results for the contrasts `r contrastName`.
`r reflink("ERIgGQCFig")` shows an MA plot and a volcano plot of the differental analysis results.

```{r daResTabERIgG, echo=F, results="asis"}

deFCap <- refcaption("ERIgGDATab", "Differential abundancy results using all 8 samples.", refType = "Table")
deFColnames <- c("Protein", "Gene", "Description", "Gene Symbol", "log(Fold Change)", "Adjusted p-value")
deFOptions <- list(columnDefs = list(list(className = 'dt-center', targets="_all")), dom='ltip')
diffexpERIgG %>% 
    dplyr::select(Protein, Gene, Description, GeneSymbol, logFC, adj.P.Val) %>% 
    datatable(colnames=deFColnames, options=deFOptions, rownames = FALSE, caption=deFCap)

```

```{r resPlotERIgG, echo=F, fig.width=10, fig.cap=ERIgGDACap, results="asis"}
ERIgGDACap <- refcaption("ERIgGDAFig", "MA plot and volcano plot of differential protein abundancy.")

MAPlot <- qPLEXanalyzer::maPlot(diffexpERIgG, significanceLevel = 0.00, minLogFoldChangeForLabelling = 1000,
                 title=names(contrasts), ylab="log2(Fold Change)")

volPlot <- diffexpERIgG %>% 
    mutate(phredAdjPval=-log10(adj.P.Val)) %>% 
    mutate(Signif=ifelse(adj.P.Val<=0.01, "<=0.01", ifelse(adj.P.Val<=0.05, "<=0.05", ">0.05"))) %>% 
    ggplot() +
        geom_point(aes(x=logFC, y=phredAdjPval, fill=Signif), colour="black", shape=21) +
        scale_fill_manual(values=c("red", "orange", "black"), name="Adjusted p-value", limits=c("<=0.01", "<=0.05", ">0.05")) +
        theme_minimal() +
        labs(x="log2(Fold Change)", y="-log10(Adjusted p-value)") +
    theme(legend.position = "bottom")

grid.newpage()
pushViewport(viewport(layout = grid.layout(nrow = 1, ncol = 2)))
 define_region <- function(row, col){
   viewport(layout.pos.row = row, layout.pos.col = col)
}
print(MAPlot, vp = define_region(row = 1, col = 1)) 
print(volPlot, vp = define_region(row = 1, col = 2))

```

### 2018-01-23 ChIP-seq_ER/qrime/01.Prepare_data.R

@@ -1,0 +1,34 @@
+suppressMessages(library(dplyr, quietly = TRUE))
+
+metadata <- read.delim("SampleData/metadata.txt", check.names = FALSE)
+metadata <- metadata %>%
+  mutate(Sample = paste(Run, Group, sep = ":"))
+
+peptideData <- read.delim("SampleData/200116_peptide_intensities_new_reprocessing.txt", stringsAsFactors = FALSE, check.names = FALSE)
+peptideData <- peptideData %>%
+  rename(Sequence = `Annotated Sequence`, Protein = `Master Protein Accessions`)
+
+retainedColumns <- c("Sequence", "Modifications", "Protein")
+
+runs <- levels(metadata$Run)
+
+for (run in runs)
+{
+  cat(run, "\n", sep = "")
+
+  runMetadata <- metadata %>% filter(Run == run)
+
+  samples <- runMetadata$Sample
+  groups <- runMetadata$Group
+
+  runData <- peptideData %>%
+    select(one_of(retainedColumns), one_of(samples)) %>%
+    mutate(Missing = apply(select(., one_of(samples)), 1, function(x) sum(is.na(x)))) %>%
+    filter(Missing != length(samples)) %>%
+    select(-Missing)
+
+  colnames(runData) <- c(retainedColumns, as.character(groups))
+
+  write.table(runData, file = paste("ProcessedData/", run, ".txt", sep = ""), quote = FALSE, sep = "\t", row.names = FALSE)
+}
+

### 2018-01-23 ChIP-seq_ER/qrime/Functions/normalization.R

@@ -1,0 +1,41 @@
+library(dplyr)
+library(preprocessCore)
+
+# Functions for normalizing intensity data
+
+# Performs scale normalization for the specified samples in the given MSnSet
+# object such that the intensities have been scaled so that each sample has
+# the same median value.
+
+# Based on the limma implementation (normalizeMedianValues function) but with
+# the option to scale to the median intensities calculated for specified rows
+# (peptide measurements).
+
+normalizeMedianScaling <- function(msnset, samples = NULL, rowsForCalculatingMedian = NULL)
+{
+  intensities <- msnset %>% exprs %>% as.data.frame
+
+  if (is.null(samples))
+    samples <- colnames(intensities)
+
+  if (!all(samples %in% colnames(intensities)))
+    stop("Could not find one or more of the given samples")
+
+  if (length(samples) == 1) return(msnset)
+
+  if (is.null(rowsForCalculatingMedian))
+    rowsForCalculatingMedian <- 1:nrow(intensities)
+
+  medianIntensities <- intensities %>%
+    summarize_each(funs(median(.[rowsForCalculatingMedian], na.rm = TRUE)), one_of(samples)) %>%
+    mutate_each(funs(log)) %>%
+    as.numeric
+
+  scalingFactors <- exp(medianIntensities - mean(medianIntensities))
+  cat(scalingFactors, "\n")
+
+  exprs(msnset)[,samples] <- t(t(intensities[,samples]) / scalingFactors)
+
+  return(msnset)
+}
+

### 2018-01-23 ChIP-seq_ER/qrime/Functions/plots.R

@@ -1,0 +1,307 @@
+library(ggplot2)
+library(dplyr)
+
+# intensity distribution plot
+
+# intensities is a data frame containing columns for each sample
+# sampleColours is a named vector that maps samples to colours
+
+intensityDistributionPlot <- function(
+  intensities,
+  samples,
+  sampleColours,
+  title = "",
+  xlab = "intensity",
+  minIntensity = NA,
+  maxIntensity = NA,
+  showLegend = TRUE
+)
+{
+  if (is.null(samples) || length(samples) == 0) return(ggplot() + theme_bw())
+
+  # subset of samples within the given intensity data frame
+  samples <- samples[which(samples %in% colnames(intensities))]
+  if (length(samples) == 0) return(ggplot() + theme_bw())
+
+  colours <- sampleColours[samples] %>% as.character
+
+  intensities <- intensities %>%
+    select(one_of(samples)) %>%
+    gather(sample, intensity) %>%
+    filter(complete.cases(.))
+
+  intensities$sample <- factor(intensities$sample, levels = samples)
+
+  plot <- ggplot(intensities, aes(x = intensity, colour = sample))
+  plot <- plot + stat_density(geom = "line", position = "identity")
+  plot <- plot + scale_colour_manual(values = colours, drop = FALSE)
+  plot <- plot + ggtitle(title)
+  plot <- plot + xlab(xlab)
+  plot <- plot + theme_bw()
+  plot <- plot + theme(
+    plot.title = element_text(size = 8),
+    axis.text.x = element_text(size = 8),
+    axis.title.x = element_text(size = 8),
+    axis.text.y = element_blank(),
+    axis.title.y = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.title = element_blank(),
+    legend.key = element_blank(),
+    legend.key.width = unit(0.3, "cm"),
+    legend.key.height = unit(0.3, "cm"),
+    legend.text = element_text(size = 6),
+    legend.justification = c(1,1),
+    legend.position = c(1,1),
+    legend.background = element_rect(fill = "transparent")
+  )
+
+  if (!is.na(minIntensity) && !is.na(maxIntensity))
+    plot <- plot + xlim(minIntensity, maxIntensity)
+
+  if (!showLegend) plot <- plot + theme(legend.position = "none")
+
+  return(plot)
+}
+
+
[…]

### 2018-01-23 ChIP-seq_ER/results/ismara/ismara_report/js/armada.js

@@ -1,0 +1,225 @@
+/*global $ */
+
+function armada_get_remote_dir() {
+    /* returns remote results dir from url parsing */
+    "use strict";
+    var remote_dir, url, url_match;
+    remote_dir = null;
+    url = document.URL;
+    url_match = url.match(/^(http|https):\/\/w*\.?(mara|ismara|bc2-web78.bc2)\.unibas\.ch.*\/(\S+)\/i?s?mara_report/);
+    if (url_match) {
+        remote_dir = url_match[3];
+    }
+    return remote_dir;
+}
+
+function armada_send_data(remote_dir) {
+    /* Put entered data into json format and sends it server */
+    "use strict";
+    console.log("send_data");
+    var data = [];
+    var rows =  $("#sample_index_table tbody tr");
+    var rows_length = rows.length;
+    var rpl_flag  = null;
+    var replicates_map = {};
+    var replicate_count = {};
+    var i;
+    for (i = 0; i < rows_length; i += 1) {
+        var smp_name, smp_idx, rpl_idx, smp_new_name;
+        smp_idx = $(rows[i]).find(".condition").val();
+        /* all samples should be assigned to new datapoints */
+        if (smp_idx === "0") {
+            alert("Please set timepoints for all samples.");
+            $(".armada_submit").show();
+            return false;
+        }
+        console.log($(rows[i]).find("td a").html());
+        smp_name = $(rows[i]).find("td a").html().replace(/^\s+|\s+$/g, "");
+        console.log($(rows[i]).find(".timepoint_name input").val());
+        smp_new_name = $(rows[i]).find(".timepoint_name input").val().replace(/\s+/g, '_');
+        console.log($(rows[i]).find(".gap input").val());
+        var gap = $(rows[i]).find(".gap input").val();
+        if (gap == undefined) {
+            gap = 0;
+        }
+        console.log("point1");
+        console.log(gap);
+        console.log($(rows[i]).find(".replicate").val());
+        rpl_idx = $(rows[i]).find(".replicate").val();
+        /* Check that replica indices are given either for all samples or for none */
+        if (rpl_flag === null) {
+            rpl_flag = Boolean(Number(rpl_idx));
+        } else if (rpl_flag !== Boolean(Number(rpl_idx))) {
+            alert("Please specify replicate indices either for all samples or for none.");
+            $(".armada_submit").show();
+            return false;
+        }
+        /* Check for repeating replicate number in one condiiton */
+        if (replicates_map[smp_idx] === undefined) {
[…]

### 2018-01-23 ChIP-seq_ER/results/ismara/ismara_report/js/average.js

@@ -1,0 +1,183 @@
+/*global $ */
+
+function average_get_remote_dir() {
+    /* returns remote results dir from url parsing */
+    "use strict";
+    var remote_dir, url, url_match;
+    remote_dir = null;
+    url = document.URL;
+    url_match = url.match(/^(http|https):\/\/w*\.?(mara|ismara|bc2-web78.bc2)\.unibas\.ch.*\/(\S+)\/i?s?mara_report/);
+    if (url_match) {
+        remote_dir = url_match[3];
+    }
+    return remote_dir;
+}
+
+function average_send_data(remote_dir) {
+    /* Put entered data into json format and sends it server */
+    "use strict";
+    var data = [];
+    var rows =  $("#sample_index_table tbody tr");
+    var rows_length = rows.length;
+    var rpl_flag  = null;
+    var replicates_map = {};
+    var replicate_count = {};
+    var i;
+    for (i = 0; i < rows_length; i += 1) {
+        var smp_name, smp_idx, rpl_idx, smp_new_name;
+        smp_idx = $(rows[i]).find(".condition").val();
+        /* all samples should be assigned to new datapoints */
+        if (smp_idx === "0") {
+            alert("Please asign all samples to new condiitons.");
+            return false;
+        }
+        smp_name = $(rows[i]).find("td a").html().replace(/^\s+|\s+$/g, "");
+        smp_new_name = $(rows[i]).find(".condition_name input").val().replace(/\s+/g, '_');
+        rpl_idx = $(rows[i]).find(".replicate").val();
+        /* Check that replica indices are given either for all samples or for none */
+        if (rpl_flag === null) {
+            rpl_flag = Boolean(Number(rpl_idx));
+        } else if (rpl_flag !== Boolean(Number(rpl_idx))) {
+            alert("Please specify replicate indices either for all samples or for none.");
+            return false;
+        }
+        /* Check for repeating replicate number in one condiiton */
+        if (replicates_map[smp_idx] === undefined) {
+            replicates_map[smp_idx] = [rpl_idx];
+        } else if (replicates_map[smp_idx].indexOf(rpl_idx) > -1 && rpl_idx > 0) {
+            alert("You have given the same batch/replicate number more than once for one condiiton");
+            return false;
+        } else {
+            replicates_map[smp_idx].push(rpl_idx);
+        }
+        
+        if (replicate_count[rpl_idx] === undefined) {
+            replicate_count[rpl_idx] = 1;
[…]

### 2018-01-23 ChIP-seq_ER/results/ismara/ismara_report/js/jquery.fileupload-ui.js

@@ -1,0 +1,724 @@
+/*
+ * jQuery File Upload User Interface Plugin 6.9.2
+ * https://github.com/blueimp/jQuery-File-Upload
+ *
+ * Copyright 2010, Sebastian Tschan
+ * https://blueimp.net
+ *
+ * Licensed under the MIT license:
+ * http://www.opensource.org/licenses/MIT
+ */
+
+/*jslint nomen: true, unparam: true, regexp: true */
+/*global define, window, document, URL, webkitURL, FileReader */
+
+(function (factory) {
+    'use strict';
+    if (typeof define === 'function' && define.amd) {
+        // Register as an anonymous AMD module:
+        define([
+            'jquery',
+            'tmpl',
+            'load-image',
+            './jquery.fileupload-fp'
+        ], factory);
+    } else {
+        // Browser globals:
+        factory(
+            window.jQuery,
+            window.tmpl,
+            window.loadImage
+        );
+    }
+}(function ($, tmpl, loadImage) {
+    'use strict';
+
+    // The UI version extends the FP (file processing) version or the basic
+    // file upload widget and adds complete user interface interaction:
+    var parentWidget = ($.blueimpFP || $.blueimp).fileupload;
+    $.widget('blueimpUI.fileupload', parentWidget, {
+
+        options: {
+            // By default, files added to the widget are uploaded as soon
+            // as the user clicks on the start buttons. To enable automatic
+            // uploads, set the following option to true:
+            autoUpload: false,
+            // The following option limits the number of files that are
+            // allowed to be uploaded using this widget:
+            maxNumberOfFiles: undefined,
+            // The maximum allowed file size:
+            maxFileSize: undefined,
+            // The minimum allowed file size:
+            minFileSize: undefined,
+            // The regular expression for allowed file types, matches
+            // against either file type or file name:
+            acceptFileTypes:  /.+$/i,
+            // The regular expression to define for which files a preview
+            // image is shown, matched against the file type:
+            previewSourceFileTypes: /^image\/(gif|jpeg|png)$/,
+            // The maximum file size of images that are to be displayed as preview:
+            previewSourceMaxFileSize: 5000000, // 5MB
+            // The maximum width of the preview images:
+            previewMaxWidth: 80,
+            // The maximum height of the preview images:
+            previewMaxHeight: 80,
+            // By default, preview images are displayed as canvas elements
+            // if supported by the browser. Set the following option to false
+            // to always display preview images as img elements:
+            previewAsCanvas: true,
+            // The ID of the upload template:
[…]

### 2018-01-23 ChIP-seq_ER/results/ismara/ismara_report/js/mara.js

@@ -1,0 +1,306 @@
+/*global $:false */
+
+function parseURL(url) {
+
+	var parser = document.createElement('a'),
+		searchObject = {},
+		queries, split, i;
+
+	// Let the browser do the work
+	parser.href = url;
+
+	// Convert query string to object
+	queries = parser.search.replace(/^\?/, '').split('&');
+	for( i = 0; i < queries.length; i++ ) {
+		split = queries[i].split('=');
+		searchObject[split[0]] = split[1];
+	}
+
+	return {
+		protocol: parser.protocol,
+		host: parser.host,
+		hostname: parser.hostname,
+		port: parser.port,
+		pathname: parser.pathname,
+		search: parser.search,
+		searchObject: searchObject,
+		hash: parser.hash
+	};
+}
+
+function get_protocol(data_url) {
+    var parser = parseURL(data_url);
+    return parser.protocol;
+}
+
+function get_filename(data_url) {
+    var parser = parseURL(data_url);
+    return parser.pathname.split("/").pop();
+}
+
+function validate_user_data() {
+    var validate_result = true;
+
+    if ($("#type_rnaseq").attr("checked") === "checked" || $("#type_chipseq").attr("checked") === "checked") {
+        if ($("#organism_row input:checked").length === 0) {
+            alert("Please choose one of organisms.");
+            $("#organism_row").css("border", "2px solid red");
+            validate_result = false;
+        }
+    }
+    var data_url = $("#url").val();
+    if (data_url === undefined) {
+        data_url = "";
+    }
+    if (data_url.trim() !== "") {
+        var filename = get_filename(data_url);
+        if (!filename.match(/(tar\.gz|\.tar|\.zip|\.tgz)$/i)) {
+            alert("Could not find valid filename from your URL. It should be .tar, .tgz, .tar.gz or .zip file.");
+            validate_result = false;
+        }
+    }
+    return validate_result;
+}
+
+function submit_user_data() {
+    var email = $("#email").val();
+    var project = $("#project").val();
+    var type = $("#type_row input:checked").val();
+    var organism = $("#organism_row input:checked").val() === undefined ? "" : $("#organism_row input:checked").val();
+    var mirna = $("#mirna_row input:checked").val();
+    var data_url = $("#url").val();
+    var protocol = "";
+    if (data_url === undefined || data_url.trim() === "") {
+        data_url = "";
+    } else {
+        protocol = data_url.split("://")[0];
[…]

### 2018-01-23 ChIP-seq_ER/results/ismara/ismara_report/js/search_gene.js

@@ -1,0 +1,129 @@
+/*global $:false */
+
+function myFormatter(label, series) {
+    if (series.lines.show == 0) {
+        return(null)
+    } 
+    return(label);
+}
+
+function get_expr_mean(data) {
+    var sum = 0;
+    for (var i in data["data"]) {
+        sum = sum + parseFloat(data["data"][i][1])
+    }
+    mean = sum / data["data"].length
+    return(mean);   
+}
+
+
+function get_min_max(data) {
+    var min = data["data"][0][1];
+    var max = data["data"][0][1];
+    for (var i in data["data"]) {
+        var value = data["data"][i][1];
+        if (value > max) {
+            max = value;
+        } 
+        else if (value < min) {
+            min = value;
+        }
+    }
+    return([min, max]);    
+}
+
+function make_expr_plot(data, samples, id, plots) {
+    var factor = 1
+    if (samples.length > 14) {
+        factor = Math.floor(samples.length / 10);
+    }
+    var samples2show = [];
+    for (i = 0; i < samples.length; i++) {
+        if (i % factor == 0) {
+            samples2show[i] = samples[i]
+        } else {
+            samples2show[i] = [i, ""];
+        }
+    }
+    options = {legend: {container: $("#" + id.replace("_div", "_legend"))}, xaxis: {ticks: samples2show}};
+    plots.push($.plot("#" + id, data, options));
+    var i = 0;
+    // fix colors in prom expr plot
+    $.each(data, function(key, val) {
+        val.color = i;
+        ++i;
+    });
+}
+
+
+function sum_datasets(data1, data2) {
+    for (var i in data1["data"]) {
+        data1["data"][i][1] = data1["data"][i][1] + data2["data"][i][1];
+    }
+    return(data1);
+}
+
+function subtract_datasets(data1, data2) {
+    for (var i in data1["data"]) {
+        data1["data"][i][1] = data1["data"][i][1] - data2["data"][i][1];
+    }
+    return(data1);
+}
+
+function changeStatus(event) {
+    plotType = event.data.param1;
+/*    var show = $(this).attr('show');
+    if (show === "1" || show === undefined) {
+        $(this).attr("show", "0");
+        $(this).find("button").text("Off");
+    }
+    else {
+        $(this).attr("show", "1");
+        $(this).find("button").text("On");
+    }*/
+    if (plotType == "expr") {
+        replot($(this));
+    } else if (plotType == "act") {
[…]

### 2018-01-23 ChIP-seq_ER/results/ismara/ismara_report/js/search_gene_submit.js

@@ -1,0 +1,69 @@
+/*global $ */
+
+function search_gene_get_remote_dir() {
+    /* returns remote results dir from url parsing */
+    "use strict";
+    var remote_dir, url, url_match;
+    remote_dir = null;
+    url = document.URL;
+    url_match = url.match(/^(http|https):\/\/w*\.?(mara|ismara|bc2-web78.bc2)\.unibas\.ch.*\/(\S+)\/i?s?mara_report/);
+    if (url_match) {
+        remote_dir = url_match[3];
+    }
+    return remote_dir;
+}
+
+function search_gene_send_data() {
+    /* Put entered data into json format and sends it server */
+    "use strict";
+    var term = $("#search_gene input[name='term']").val().trim();
+    var dir = $("#search_gene input[name='dir']").val()
+
+    if (term == "") {
+        alert("Please enter gene name!");
+        return;
+    }
+    
+    var post_url = "/fcgi/search_gene";
+    var secure = 0;
+    if (document.URL.match(/\/secure\//)) {
+        post_url = "/secure/fcgi/search_gene";
+        secure = 1;
+    }
+
+    $.ajax({
+        type: "POST",
+        url: post_url, 
+        data: {
+            "dir": dir,
+            "term": term,
+            "secure": secure
+        },
+        beforeSend: function() {$("body").addClass("loading");},
+        complete: function(){$("body").removeClass("loading");}
+    })
+        .done(
+            function(x){
+                console.log("x:" + x);
+                if (x == "Fail") {
+                    alert("Term " + term + " not found.");
+                }
+                else {
+                    var scratch_url = "https://" + window.location.hostname;
+                    /*scratch_url += "/ISMARA/scratch/"*/
+                    var data = window.location.pathname.split("/");
+                    scratch_url += data.slice(0,data.findIndex(function(x){return x == "ismara_report"}) - 1).join("/") + "/";
+                    console.log(scratch_url + dir + "/" + term + ".html");
+                    window.location = scratch_url + dir + "/" + term + ".html";
+                }
+            });
+}
+
+$(function () {
+    var remote_dir = search_gene_get_remote_dir();
+    if (remote_dir) {
+        $("#search_gene").show();
+//        $("#search_gene_button").button();
+    }
+    $("#search_gene_button").click(function(){search_gene_send_data();});
+});

### 2018-01-23 ChIP-seq_GRHL2/000_download.sh

@@ -1,0 +1,2 @@
+#!/bin/sh
+java -jar ../java/clarity-tools.jar -l SLX-14333

### 2018-01-23 ChIP-seq_GRHL2/001_mergeReads.sh

@@ -1,0 +1,17 @@
+mkdir ./mergedReads
+
+cd ./mergedReads
+
+
+for bam in $(find ../SLX-14333/SLX-14333.D???_D???.HM3WHBBXX.s_2.*.bam -type f);
+do
+	
+	bamOut=${bam/.HM3WHBBXX.s_2.bwa.homo_sapiens/}	
+	bam2=${bam/HM3WHBBXX/HM3WTBBXX}
+	bam2=${bam2/s_2/s_4}
+	samtools merge `basename $bamOut` $bam  $bam2  
+done
+
+
+
+

### 2018-01-23 ChIP-seq_GRHL2/001a_sortAndIndex.sh

@@ -1,0 +1,15 @@
+cd mergedReads
+mkdir sorted
+
+for bam in *bam
+do
+	samtools sort $bam -o sorted/$bam
+done
+
+cd sorted
+
+for bam in *bam
+do
+	samtools index $bam
+done
+

### 2018-01-23 ChIP-seq_GRHL2/001b_blacklist.sh

@@ -1,0 +1,27 @@
+#mkdir blacklists
+#cd blacklists
+#wget http://mitra.stanford.edu/kundaje/akundaje/release/blacklists/hg38-human/hg38.blacklist.bed.gz
+#gunzip hg38.blacklist.bed.gz
+#cd ..
+
+
+bl=../../blacklists/hg38.blacklist.bed
+
+mkdir ./GRHL2_filtered
+cd mergedReads/sorted/
+for f in *.bam
+do
+	echo $f
+	bedtools intersect -v -abam $f -b $bl  > ../../GRHL2_filtered/$f
+done
+
+### Re-index
+
+cd ../../GRHL2_filtered
+for f in *.bam
+do
+samtools index $f
+done
+
+cd ..
+

### 2018-01-23 ChIP-seq_GRHL2/002_peakCalling.sh

@@ -1,0 +1,15 @@
+### MACS peak caller
+
+### Run macs on the blacklisted data
+mkdir ./peaks
+cd peaks
+control=../GRHL2_filtered/SLX-14333.D701_D503.bam
+for bam in ../GRHL2_filtered/*.bam
+do
+root=`basename $bam .bam`
+macs2 callpeak -t $bam -c $control -f BAM -n $root -g hs &
+done
+
+
+
+

### 2018-01-23 ChIP-seq_GRHL2/003_diffbind.R

@@ -1,0 +1,411 @@
+
+######################
+#
+# Functions
+#
+######################
+
+
+
+write.bed<-function(dba, filename){
+    score <- -10*(log10(dba$FDR)) 
+    write.table(cbind(dba[,1:3],rownames(dba),score),
+                filename, quote=FALSE, sep="\t",
+                row.names=FALSE, col.names=FALSE)
+}
+
+
+getTFs <-function(vobj,threshold=NA){
+    if (is.na(threshold)){
+        TFs<-names(vobj$mrs[,"NES"]) 
+    }else if (threshold>0) {
+        TFs<-names(vobj$mrs[,"NES"][(vobj$mrs[,"NES"]>p2z(threshold))]) 
+    }else if (threshold<0) {
+        TFs<-names(vobj$mrs[,"NES"][(vobj$mrs[,"NES"]<sign(threshold)*p2z(-threshold))])
+    }
+    if (length(TFs) == 0)
+    {   #plot Function doesn't like no labels.
+        TFs<-names(vobj$mrs[,"NES"]) 
+    }
+    return(TFs)
+}
+
+vulcan.import.dba<- function (dbaobj, samples,intervals = NULL)  
+{
+    dbcounts <- dbaobj
+    listcounts <- dbcounts$peaks
+    names(listcounts) <- dbcounts$samples[, 1]
+    first <- listcounts[[1]]
+    rawmat <- matrix(NA, nrow = nrow(first), ncol = length(listcounts) + 
+                         3)
+    colnames(rawmat) <- c("Chr", "Start", "End", names(listcounts))
+    rownames(rawmat) <- 1:nrow(rawmat)
+    rawmat <- as.data.frame(rawmat)
+    rawmat[, 1] <- as.character(first[, 1])
+    rawmat[, 2] <- as.integer(first[, 2])
+    rawmat[, 3] <- as.integer(first[, 3])
+    for (i in 1:length(listcounts)) {
+        rawmat[, names(listcounts)[i]] <- as.numeric(listcounts[[i]]$RPKM)
+    }
+    peakrpkms <- rawmat
+    rm(rawmat)
+    first <- listcounts[[1]]
+    rawmat <- matrix(NA, nrow = nrow(first), ncol = length(listcounts) + 
+                         3)
+    colnames(rawmat) <- c("Chr", "Start", "End", names(listcounts))
+    rownames(rawmat) <- 1:nrow(rawmat)
[…]

### 2018-01-23 ChIP-seq_GRHL2/004_motif.sh

@@ -1,0 +1,5 @@
+mkdir motifAnalysis
+cd motifAnalysis
+findMotifsGenome.pl ../bed/up.bed hg38 grhl2UpSites
+findMotifsGenome.pl ../bed/all.bed hg38 grhl2AllSites
+

### 2018-01-23 ChIP-seq_GRHL2/005_bed_overlap.sh

@@ -1,0 +1,86 @@
+#p300 from Zwart, EMBO, 2011
+#ER from Hurtado 2011
+#ChIA-pet all 3 https://www.encodeproject.org/experiments/ENCSR000BZZ/
+#gro-seq GSE43835 
+#Rest from Caroll MLL3 paper 
+
+cd bed
+mkdir overlaps
+
+#generate gro-seq interect
+#file1=hglft_E2.40m.rep1.bed 
+#file2=hglft_E2.40m.rep2.bed 
+#output=gro-seq.bed
+#bedtools intersect -sorted -a $file1 -b $file2 > overlaps/$output
+
+file1=hglft_foxa1.bed 
+file2=up.bed
+output=foxa1-up.bed
+bedtools intersect -a $file1 -b $file2 > overlaps/$output
+
+file1=hglft_ER_Hurtado_2011.bed 
+output=er-up.bed
+bedtools intersect -a $file1 -b $file2 > overlaps/$output
+
+#file1=gro-seq.bed 
+#output=gro-up.bed
+#bedtools intersect -a $file1 -b $file2 > overlaps/$output
+
+
+file1=hglft_ChIA_combined.bed 
+output=ChIAPet_up.bed
+bedtools intersect -a $file1 -b $file2 > overlaps/$output
+
+file1=hglft_h3k4me1.bed 
+output=h3k4me1-up.bed
+bedtools intersect -a $file1 -b $file2 > overlaps/$output
+
+file1=hglft_h3k4me3.bed 
+output=h3k4me3-up.bed
+bedtools intersect -a $file1 -b $file2 > overlaps/$output
+
+file1=hglft_p300_ctrl.bed  
+output=p300_ctrl_up.bed
+bedtools intersect -a $file1 -b $file2 > overlaps/$output
+
+file1=hglft_p300_e2.bed  
+output=p300_e2_up.bed
+bedtools intersect -a $file1 -b $file2 > overlaps/$output
+
+
+#all peaks
+
+file1=hglft_foxa1.bed 
+file2=all.bed
+output=foxa1-all.bed
+bedtools intersect -a $file1 -b $file2 > overlaps/$output
+
+file1=hglft_ER_Hurtado_2011.bed 
+output=er-all.bed
+bedtools intersect -a $file1 -b $file2 > overlaps/$output
+
+#file1=gro-seq.bed 
+#output=gro-all.bed
[…]

### 2018-01-23 ChIP-seq_GRHL2/006_promoters.R

@@ -1,0 +1,11 @@
+require(TxDb.Hsapiens.UCSC.hg38.knownGene)
+txdb <- TxDb.Hsapiens.UCSC.hg38.knownGene
+p<-promoters(genes(txdb), upstream = 1500, downstream = 500)
+
+write.bed<-function(df, filename){
+    write.table(as.data.frame(df)[,1:3],
+                filename, quote=FALSE, sep="\t",
+                row.names=FALSE, col.names=FALSE)
+}
+
+write.bed(p,"bed/promoter.bed")

### 2018-01-23 ChIP-seq_GRHL2/006_promoters.sh

@@ -1,0 +1,10 @@
+cd bed
+file1=overlaps/ChIAPet_up.bed   
+file2=promoter.bed
+output=ChIAPet_up_prom.bed
+bedtools intersect -a $file1 -b $file2 > overlaps/$output
+
+file1=overlaps/ChIAPet_all.bed   
+file2=promoter.bed
+output=ChIAPet_all_prom.bed
+bedtools intersect -a $file1 -b $file2 > overlaps/$output

### 2018-01-23 ChIP-seq_GRHL2/bed/combinedChIAPET.sh

@@ -1,0 +1,1 @@
+multiIntersectBed -i *bed6_sorted.bed | awk '{if ($4 > 1) {print} }' > hglft_ChIA_combined.bed		

### 2018-01-23 ChIP-seq_GRHL2/bigWig/bam2BigWig.sh

@@ -1,0 +1,6 @@
+for filename in ../GRHL2_filtered/*.bam 
+do
+ echo $filename
+ output=`basename $filename`
+ bamCoverage -b $filename -o $filename.bw
+done

### 2018-01-23 ChIP-seq_GRHL2/peaks/SLX-14333.D701_D501_model.r

@@ -1,0 +1,17 @@
+# R script for Peak Model
+#  -- generated by MACS
+p <- c(0.0307806134618,0.0358863504609,0.0408787890727,0.0458200606709,0.0508527019363,0.0506918913221,0.0506918913221,0.050856356723,0.0508490471496,0.0505712833615,0.05027524564,0.0501217445993,0.0500376845055,0.050333722227,0.0503739248805,0.0503300674403,0.0504104727474,0.0503410318004,0.0503666153072,0.0503629605205,0.0503446865871,0.0506041764417,0.0506918913221,0.0504652945477,0.0506334147352,0.0505420450681,0.0503629605205,0.05027524564,0.0502350429865,0.0506882365355,0.0508600115097,0.0505822477216,0.0505201163479,0.0507211296156,0.0505456998547,0.050658998242,0.0506992008955,0.0508307732162,0.0508051897094,0.0505785929349,0.0504835684811,0.0507613322692,0.0506334147352,0.0506626530287,0.0507796062026,0.0508819402298,0.0509988934037,0.0510390960572,0.0510939178575,0.051134120511,0.0511012274309,0.0509221428833,0.0512254901782,0.0509988934037,0.0509184880966,0.0511962518847,0.0510427508439,0.0509842742569,0.0509806194703,0.0509733098969,0.0510573699906,0.0510866082841,0.0509879290436,0.0508453923629,0.0508600115097,0.0501582924661,0.0502167690531,0.0501144350259,0.0501656020395,0.0501509828927,0.0501217445993,0.0501144350259,0.0501217445993,0.0504433658276,0.0505274259213,0.0510829534974,0.0512656928317,0.0515069087529,0.0515251826863,0.0513863007923,0.0516384810736,0.0517006124472,0.0518029464744,0.0516494454336,0.0516531002203,0.0514265034458,0.0512912763385,0.0511121917909,0.0512656928317,0.0514703608861,0.0513863007923,0.0515324922597,0.0518906613548,0.0519052805016,0.0518687326347,0.0522342113032,0.0521464964228,0.0525558325315,0.0523036522503,0.0520405076089,0.052120912916,0.0521866990763,0.051912590075,0.0520770554758,0.0522671043834,0.052062436329,0.052051471969,0.0518175656211,0.0518285299812,0.0522122825831,0.0521136033426,0.0520331980356,0.0520660911157,0.0519966501687,0.0518468039146,0.0518796969948,0.0520551267557,0.0519052805016,0.0519710666619,0.0518285299812,
[…]

### 2018-01-23 ChIP-seq_GRHL2/peaks/SLX-14333.D701_D502_model.r

@@ -1,0 +1,17 @@
+# R script for Peak Model
+#  -- generated by MACS
+p <- c(0.0354933074034,0.041437790254,0.0470080908616,0.0528831378321,0.059005067932,0.0588507659761,0.0588584810739,0.0582991364838,0.058140976979,0.0583338544239,0.0580869712945,0.0580869712945,0.0584650110864,0.0584071478529,0.0581024014901,0.0579673872787,0.0578863787518,0.0580831137456,0.0579751023765,0.0577899400294,0.0578015126761,0.0576510682691,0.0576279229757,0.0576240654268,0.0576433531713,0.0579596721809,0.0581564071746,0.058368572364,0.0584264355974,0.0584920139287,0.0585653073577,0.0589549197964,0.0588777688184,0.0586848913735,0.0588160480361,0.0586810338246,0.0589857801875,0.0588700537206,0.0588199055849,0.0586424583357,0.0585691649066,0.058329996875,0.0587118942158,0.0589973528342,0.0590320707743,0.0593098142949,0.0590976491055,0.0590552160677,0.0591323670456,0.0592480935125,0.059494976642,0.0595258370331,0.0593676775284,0.0592789539037,0.0593020991971,0.0591670849857,0.0592095180236,0.0593098142949,0.0594872615442,0.0593098142949,0.0594023954684,0.0596261333045,0.0594101105662,0.0596762814401,0.0596724238912,0.0594602587019,0.0593985379195,0.0592249482192,0.0592056604747,0.0595952729133,0.0594525436041,0.0591593698879,0.0594641162508,0.0592789539037,0.0592789539037,0.0593483897839,0.0595219794842,0.0596531361467,0.0596299908534,0.0594255407618,0.0591786576324,0.0592982416482,0.0591477972412,0.0593098142949,0.05911693685,0.0591439396923,0.0592982416482,0.059305956746,0.0593715350773,0.0597109993802,0.0597688626137,0.0597842928093,0.0597302871247,0.0596608512445,0.0598537286894,0.0599848853519,0.0600466061343,0.0603590675949,0.0603320647527,0.0600774665254,0.0601931929924,0.0601739052479,0.0601816203457,0.0601507599545,0.060170047699,0.0600658938788,0.0596994267335,0.0595566974243,0.0593676775284,0.0593908228217,0.0592558086103,0.0593406746861,0.0594911190931,0.0595759851688,0.0593329595883,0.0590937915567,0.0592943840993,0.0592712388059,0.0594023954684,0.0593
[…]

### 2018-01-23 ChIP-seq_GRHL2/peaks/SLX-14333.D701_D503_model.r

@@ -1,0 +1,17 @@
+# R script for Peak Model
+#  -- generated by MACS
+p <- c(0.0395328543787,0.0460981319809,0.0528151875061,0.0595922482567,0.0662669471523,0.0662175310843,0.0663269523777,0.066510497773,0.0663446009734,0.066309303782,0.0663693090074,0.0662175310843,0.0657269001237,0.0653209824225,0.065433933435,0.0655362952901,0.0655433547284,0.065384517367,0.065684543494,0.0659280941147,0.066058693723,0.0663198929394,0.066559913841,0.0667222809215,0.0666410973812,0.0663904873222,0.0663834278839,0.0661328178249,0.0660233965315,0.0659069157999,0.0655786519198,0.0653527498948,0.0656068896729,0.065684543494,0.0654904089413,0.0658469105745,0.0658327916979,0.0659563318679,0.0657410190003,0.0658257322596,0.0662175310843,0.0663834278839,0.0662281202417,0.0660657531612,0.0658045539448,0.0655645330432,0.0653351012991,0.0651974422526,0.0652927446694,0.0653386310182,0.0651727342186,0.0649468321936,0.0649115350021,0.0650456643295,0.0654762900647,0.0651127289932,0.0652539177588,0.0654727603456,0.0653492201756,0.0650244860147,0.0651550856229,0.065508057537,0.0657445487194,0.0656174788303,0.065508057537,0.0659775101827,0.0658222025405,0.0658786780468,0.0661998824886,0.0668458210914,0.0665175572113,0.0665246166496,0.0662951849054,0.066436373671,0.0663693090074,0.0663198929394,0.0665316760878,0.0666305082238,0.0667081620449,0.0663269523777,0.0664151953562,0.0663198929394,0.0661681150163,0.0661787041737,0.0663022443437,0.0666305082238,0.0667434592363,0.066835231934,0.0665634435601,0.066584621875,0.0667540483937,0.0668917074402,0.0670893717121,0.0673682195242,0.0672340901969,0.0668281724957,0.0669234749125,0.0668105239,0.0668881777211,0.0670293664867,0.066785815866,0.0667081620449,0.0666834540109,0.0664222547945,0.0668034644617,0.0667328700789,0.0664787303007,0.0663163632202,0.0664752005816,0.0664434331093,0.0668599399679,0.0669093560359,0.067036425925,0.0670999608695,0.0672093821629,0.0671740849715,0.0670187773293,0.067061133959,0.0671176094652,0.0672587982309,0.067
[…]

### 2018-01-23 ChIP-seq_GRHL2/peaks/SLX-14333.D701_D504_model.r

@@ -1,0 +1,17 @@
+# R script for Peak Model
+#  -- generated by MACS
+p <- c(0.0398576286986,0.0463213922772,0.0532300029842,0.059912820073,0.0666225975938,0.0668888318601,0.0670303741282,0.0670876650463,0.0671281056943,0.0669697131562,0.0671618062344,0.0674010800686,0.0671045153163,0.0671887666664,0.0672696479625,0.0670034136962,0.0670034136962,0.0666124874318,0.0665383462437,0.0664372446236,0.0661440499253,0.0661743804114,0.0660159878732,0.065803674471,0.0657261632289,0.0657430134989,0.0654127482066,0.0656857225809,0.0657025728509,0.0659991376032,0.0659654370631,0.065830634903,0.065837375011,0.0662384114374,0.0662653718695,0.0662384114374,0.0665653066758,0.0663294028955,0.0662013408434,0.0660699087373,0.0662957023555,0.0663260328415,0.0665551965137,0.0662990724095,0.0660732787913,0.0661912306814,0.0659047760911,0.0660429483052,0.0660496884132,0.0661036092773,0.065834004957,0.065790194255,0.0657463835529,0.0654902594487,0.0654868893947,0.0655576605287,0.065813784633,0.0660699087373,0.0662653718695,0.0660867590073,0.0663732135976,0.0662249312214,0.0663024424635,0.0663631034356,0.0664709451637,0.0662518916534,0.0660496884132,0.0659081461451,0.0659452167391,0.0662080809514,0.0660800188993,0.0663597333815,0.0663900638676,0.06678773024,0.0668618714281,0.0671045153163,0.0673437891505,0.0672460575844,0.066831540942,0.0667506596459,0.0668820917521,0.0668551313201,0.0664237644076,0.0661305697093,0.0661137194393,0.0659485867931,0.0661204595473,0.0662721119775,0.0666091173778,0.0665282360817,0.0664675751097,0.0665214959737,0.0669933035342,0.0672662779085,0.0673134586645,0.0673606394206,0.0674988116347,0.067761675847,0.067829076927,0.0680312801672,0.0680177999512,0.0681391218954,0.0680481304373,0.0679739892492,0.0679942095732,0.0682975144335,0.0680919411393,0.067761675847,0.0677010148749,0.0677347154149,0.0677347154149,0.067808856603,0.0676403539029,0.0676673143349,0.0676403539029,0.0675796929308,0.0678863678451,0.0682772941095,0.0684457968097,0.06834469518
[…]

### 2018-01-23 ChIP-seq_GRHL2/peaks/SLX-14333.D702_D501_model.r

@@ -1,0 +1,17 @@
+# R script for Peak Model
+#  -- generated by MACS
+p <- c(0.0307879827374,0.0359820834311,0.041058285244,0.0462098113418,0.0516986592376,0.0515119860096,0.0515873102946,0.0518493078075,0.0519475568748,0.0518853324655,0.0519442819059,0.0519442819059,0.0522554039525,0.0524027775535,0.052127680165,0.0522030044499,0.0522226542634,0.0518984323412,0.0522652288592,0.0523045284862,0.0523340032064,0.0523569279888,0.0520982054448,0.0520621807867,0.0520359810354,0.052163704823,0.052458452025,0.0528186986053,0.0526876998488,0.0528121486675,0.0532477195327,0.0533623434446,0.053457317543,0.0533885431959,0.0533197688488,0.0532116948747,0.0528383484188,0.0527630241338,0.0528645481701,0.0527040746934,0.0522423040769,0.0520949304758,0.0520916555069,0.0522128293567,0.0522848786727,0.0523340032064,0.0522324791701,0.0523078034551,0.0525272263722,0.0526745999732,0.0528219735742,0.0529988218954,0.0529103977348,0.0530282966156,0.0532509945016,0.0532706443151,0.0535064420767,0.0531854951234,0.0530643212737,0.0532280697192,0.0532968440664,0.0531494704654,0.0534802423254,0.0533263187866,0.0530806961182,0.0529595222685,0.0530315715845,0.0533983681026,0.0533557935068,0.0532280697192,0.0533132189109,0.0534049180405,0.0533656184135,0.0535555666104,0.0538437638746,0.054089386543,0.0540664617606,0.0542138353616,0.0541876356103,0.0540402620093,0.0540304371026,0.0541024864186,0.0540271621337,0.0538994383461,0.053830663999,0.0537127651181,0.0535457417037,0.0535195419524,0.0536931153047,0.0540959364808,0.0541155862943,0.0541745357347,0.0542989845533,0.0542334851751,0.0542596849264,0.0540533618849,0.0541647108279,0.0541974605171,0.0540926615119,0.0540042373513,0.0540468119471,0.0541090363564,0.0542171103305,0.0546101065999,0.0546166565377,0.054901578833,0.0550948019988,0.0550358525584,0.0552913001335,0.055274925289,0.0551537514392,0.0552945751024,0.0552290757242,0.0552028759729,0.0552258007553,0.0550162027449,0.0549539783356,0.0554190239211,0.0553567995117,0.0549736281
[…]

### 2018-01-23 ChIP-seq_GRHL2/peaks/SLX-14333.D702_D502_model.r

@@ -1,0 +1,17 @@
+# R script for Peak Model
+#  -- generated by MACS
+p <- c(0.00501421889214,0.00644685286133,0.00644685286133,0.00644685286133,0.00787948683051,0.00787948683051,0.00787948683051,0.00716316984592,0.00644685286133,0.00573053587674,0.00787948683051,0.00644685286133,0.00787948683051,0.0085958038151,0.0085958038151,0.0093121207997,0.0093121207997,0.0093121207997,0.0100284377843,0.0100284377843,0.00716316984592,0.00787948683051,0.0085958038151,0.00787948683051,0.00716316984592,0.00716316984592,0.0093121207997,0.0093121207997,0.00787948683051,0.0093121207997,0.0093121207997,0.0085958038151,0.00716316984592,0.00787948683051,0.00787948683051,0.00716316984592,0.0093121207997,0.0093121207997,0.0093121207997,0.0100284377843,0.0114610717535,0.0128937057227,0.0128937057227,0.0121773887381,0.0114610717535,0.0136100227072,0.0100284377843,0.0114610717535,0.0136100227072,0.0128937057227,0.0121773887381,0.0121773887381,0.0114610717535,0.0128937057227,0.0136100227072,0.0121773887381,0.0128937057227,0.0114610717535,0.0093121207997,0.0085958038151,0.0085958038151,0.00787948683051,0.0085958038151,0.0085958038151,0.00787948683051,0.00787948683051,0.00787948683051,0.00787948683051,0.0100284377843,0.0114610717535,0.0114610717535,0.0107447547689,0.0107447547689,0.0093121207997,0.0093121207997,0.00787948683051,0.00716316984592,0.00644685286133,0.00501421889214,0.00429790190755,0.00573053587674,0.00644685286133,0.00573053587674,0.00716316984592,0.0093121207997,0.0100284377843,0.0100284377843,0.0114610717535,0.0107447547689,0.0100284377843,0.0085958038151,0.0085958038151,0.0085958038151,0.00716316984592,0.00573053587674,0.00573053587674,0.0085958038151,0.00716316984592,0.00787948683051,0.00787948683051,0.0085958038151,0.00787948683051,0.0085958038151,0.0100284377843,0.0100284377843,0.0100284377843,0.00716316984592,0.00787948683051,0.00787948683051,0.0100284377843,0.0107447547689,0.0114610717535,0.0128937057227,0.0121773887381,0.0136100227072,0.0136100227072,0.0128937057227,0.0121773887381,0.01217738
[…]

### 2018-01-23 ChIP-seq_GRHL2/peaks/SLX-14333.D702_D503_model.r

@@ -1,0 +1,17 @@
+# R script for Peak Model
+#  -- generated by MACS
+p <- c(0.0406506212106,0.047443754921,0.0540841624403,0.0606863884118,0.0674127044137,0.0676895206351,0.0674763403266,0.0674317951876,0.0674222498006,0.0673108869529,0.0672122512878,0.0670722522792,0.0670690704836,0.0668908899272,0.0671931605139,0.0667763452839,0.067110433827,0.067345886705,0.067119979214,0.0672090694921,0.0670404343227,0.0669417986576,0.066992707388,0.0671804333313,0.0669163442924,0.067301341566,0.0671358881922,0.0669163442924,0.0670817976662,0.0670499797097,0.0673236141355,0.0675749759918,0.067654520883,0.0677277021829,0.0675049764875,0.0673236141355,0.0674349769832,0.0675049764875,0.0675177036701,0.0673808864572,0.0673522502963,0.0673331595224,0.0673967954354,0.067473158531,0.067718156796,0.067600430357,0.0676513390873,0.0674954311005,0.0674667949397,0.067291796179,0.0672345238573,0.0670499797097,0.0667254365535,0.0669195260881,0.0669258896794,0.0666618006405,0.0664645293103,0.0666586188449,0.0665058926537,0.066820890423,0.0666554370492,0.0668495265838,0.0669577076359,0.0668463447882,0.0670054345706,0.0671581607618,0.06717406974,0.0673808864572,0.0674349769832,0.0674031590267,0.0677467929568,0.067600430357,0.0677308839786,0.0677277021829,0.0675845213787,0.067899519148,0.0676640662699,0.0674604313484,0.0676577026786,0.0673967954354,0.0670086163663,0.0666840732101,0.0666077101145,0.0665090744494,0.0664549839233,0.0664995290624,0.0668304358099,0.0667158911666,0.0662990759365,0.0663245303017,0.0667795270795,0.0671677061487,0.0671358881922,0.0671708879443,0.0673395231137,0.0673713410702,0.0673713410702,0.0676608844743,0.0682017897346,0.068189062552,0.0678486104176,0.0677817927089,0.0679058827392,0.0676354301091,0.0676577026786,0.0671677061487,0.0672217966747,0.0671804333313,0.066747709123,0.0667158911666,0.0670308889358,0.0674317951876,0.0675177036701,0.0677945198915,0.0675527034222,0.0678645193958,0.067889973761,0.0676513390873,0.0678708829871,0.0677881563002,0.06764
[…]

### 2018-01-23 ChIP-seq_GRHL2/peaks/SLX-14333.D702_D504_model.r

@@ -1,0 +1,17 @@
+# R script for Peak Model
+#  -- generated by MACS
+p <- c(0.0321644127167,0.0376190354628,0.042873059992,0.0481902866991,0.0535872031088,0.0535020175647,0.0535899510296,0.0534662945945,0.0532409650907,0.0530293751907,0.0529359458842,0.0528507603401,0.0528397686569,0.0527710706375,0.0525320015297,0.0526006995492,0.0527463393505,0.0529524334089,0.0530046439037,0.0532052421205,0.0532217296452,0.053199746279,0.0533591256842,0.0534058403374,0.0537658179594,0.0538125326127,0.053746582514,0.0535267488517,0.0535762114257,0.0534772862776,0.053298671427,0.0532876797439,0.0532574526153,0.0532876797439,0.0531337961803,0.0529716688543,0.0529469375673,0.053348134001,0.0532162338037,0.0532629484569,0.0535267488517,0.0535817072672,0.0538510035036,0.0539746599386,0.0540928205321,0.0541367872646,0.0541175518191,0.0539883995425,0.0540103829087,0.0541285435022,0.0542137290464,0.0539993912256,0.0537191033062,0.0535624718218,0.0532794359816,0.0534553029114,0.0534690425153,0.0534140840997,0.0533975965751,0.0533756132088,0.0531118128141,0.0532327213283,0.0532849318231,0.053175014992,0.0533371423179,0.0531612753881,0.053150283705,0.0528287769738,0.0531035690517,0.0531255524179,0.0531915025166,0.0533206547932,0.0532684442985,0.0534772862776,0.0532134858829,0.0532876797439,0.0533646215257,0.0535734635049,0.0534937738023,0.0533426381595,0.0533179068725,0.0531008211309,0.0529029708349,0.0528535082608,0.0531365441011,0.0530541064777,0.0531530316257,0.0532766880608,0.0533069151894,0.0532217296452,0.0532904276647,0.0534498070699,0.0535789593465,0.0538152805334,0.0538345159789,0.0539279452854,0.0538345159789,0.0536668928114,0.0538042888503,0.054007634988,0.0542192248879,0.0542274686503,0.0542192248879,0.0539306932062,0.0538070367711,0.0537823054841,0.0535954468711,0.053609186475,0.053510261327,0.0533371423179,0.0530980732102,0.0530843336063,0.0533206547932,0.053436067466,0.0536201781581,0.0537685658802,0.0539554244932,0.0539938953841,0.054081828849,0.0542274686
[…]

### 2018-01-23 ChIP-seq_GRHL2/peaks/SLX-14333.D703_D501_model.r

@@ -1,0 +1,17 @@
+# R script for Peak Model
+#  -- generated by MACS
+p <- c(0.0481963321802,0.0562735321839,0.064220263465,0.0720187348341,0.0796926678772,0.0796808070843,0.0795444079653,0.0797994150138,0.0794732432075,0.079479173604,0.0791530017977,0.0789632291104,0.0788090388019,0.0788861339562,0.078933577128,0.0788386907843,0.0788031084055,0.0787378740442,0.078797178009,0.0788386907843,0.0788742731632,0.0788268299914,0.0788742731632,0.0787675260266,0.0787912476125,0.0788624123702,0.0784354238238,0.0781804167752,0.0782575119294,0.0781922775682,0.0782693727224,0.0783049551013,0.0782990247048,0.0785718229428,0.0787141524583,0.0788090388019,0.0791233498153,0.0792597489343,0.0789157859385,0.0789988114892,0.0785836837358,0.0787556652336,0.0787615956301,0.0785836837358,0.0785006581851,0.0783227462907,0.0780677392421,0.0784354238238,0.0785362405639,0.0783701894625,0.0787022916653,0.0784117022379,0.0784532150132,0.0784176326344,0.0783286766872,0.0783346070837,0.0784117022379,0.0779609921055,0.0779016881407,0.0778661057619,0.0779016881407,0.0779609921055,0.0780796000351,0.077978783295,0.0781744863787,0.0780262264668,0.0782100687576,0.078103321621,0.0780499480527,0.0782219295506,0.0778245929865,0.0779016881407,0.0780380872597,0.0779847136914,0.07781866259,0.0780736696386,0.0781507647928,0.0781270432069,0.0783049551013,0.0782812335154,0.0787082220618,0.0788327603878,0.0783464678766,0.0785303101674,0.0784887973921,0.0786251965111,0.0783939110484,0.0782456511365,0.0786370573041,0.0784354238238,0.0784117022379,0.0782871639118,0.078648918097,0.0782812335154,0.0784828669956,0.0785065885815,0.0784887973921,0.0788624123702,0.0788683427667,0.0793190528991,0.0792182361589,0.0793131225026,0.0794969647934,0.0794613824146,0.0795977815336,0.0795444079653,0.0796333639124,0.0793902176568,0.0793487048815,0.0791352106082,0.0791530017977,0.0790166026787,0.0788742731632,0.0792894009167,0.0788505515773,0.0789750899033,0.0793783568639,0.079751971842
[…]

### 2018-01-23 ChIP-seq_GRHL2/peaks/SLX-14333.D703_D502_model.r

@@ -1,0 +1,17 @@
+# R script for Peak Model
+#  -- generated by MACS
+p <- c(0.0411229480844,0.0478262304382,0.0544761710225,0.0612327951457,0.0680143120946,0.0676835931243,0.0677404910117,0.0674560015749,0.0672426344974,0.0674346648672,0.0672497467333,0.0672888640308,0.0672461906153,0.0670221551839,0.0669119155271,0.0672355222614,0.0668265686961,0.0670363796557,0.0672461906153,0.0670150429479,0.0671075020149,0.0673742108619,0.0673315374463,0.0674168842774,0.0673493180361,0.0672177416716,0.0675022311084,0.067452445457,0.0672604150872,0.0673030885027,0.0674773382827,0.0670363796557,0.0671643999022,0.0673102007386,0.0675875779394,0.0676907053603,0.0676338074729,0.0678222817248,0.0677227104219,0.067712042068,0.0676480319447,0.0676978175962,0.0675769095856,0.0673422058002,0.0672212977896,0.0673244252104,0.0670719408353,0.0669759256504,0.0672390783794,0.0671323948406,0.0671252826047,0.0670826091892,0.0673564302721,0.0674239965133,0.0674239965133,0.0672426344974,0.0675804657035,0.0673102007386,0.0672106294357,0.0672924201488,0.0669723695324,0.0671395070765,0.0671679560202,0.0674595576929,0.0676480319447,0.0679574142072,0.0679040724378,0.0679823070329,0.0679929753868,0.0678293939607,0.0679182969097,0.0678258378427,0.067889847966,0.0677476032476,0.0674666699288,0.0672604150872,0.0670932775431,0.0671643999022,0.0673350935643,0.0674453332211,0.0676480319447,0.0677796083093,0.0675200116982,0.067452445457,0.0675946901754,0.0675840218215,0.0677867205452,0.0678507306684,0.0677404910117,0.0679147407917,0.0674737821647,0.0675520167598,0.0677191543039,0.0677547154835,0.0673350935643,0.0671715121382,0.0670150429479,0.0670577163635,0.0672212977896,0.0671288387227,0.0672817517949,0.0673137568566,0.0675662412317,0.0674026598055,0.0679218530276,0.0679111846738,0.0677831644272,0.0677973888991,0.0674666699288,0.0675057872264,0.067975194797,0.0681209956334,0.0681032150436,0.0683450310648,0.0683983728342,0.0684552707216,0.0685655103783,0.068590403204,0.069116708662,0.06912
[…]

### 2018-01-23 ChIP-seq_GRHL2/peaks/SLX-14333.D703_D503_model.r

@@ -1,0 +1,17 @@
+# R script for Peak Model
+#  -- generated by MACS
+p <- c(0.00626870817596,0.00783588521995,0.00783588521995,0.00940306226394,0.0109702393079,0.0141045933959,0.0141045933959,0.0141045933959,0.0172389474839,0.0203733015719,0.0156717704399,0.0141045933959,0.0156717704399,0.0141045933959,0.0125374163519,0.0109702393079,0.00940306226394,0.00940306226394,0.00783588521995,0.00626870817596,0.00626870817596,0.00626870817596,0.00626870817596,0.00783588521995,0.00783588521995,0.00783588521995,0.00783588521995,0.00783588521995,0.00783588521995,0.00626870817596,0.00940306226394,0.0125374163519,0.0109702393079,0.0125374163519,0.0125374163519,0.0109702393079,0.0109702393079,0.0109702393079,0.00940306226394,0.0109702393079,0.00783588521995,0.00470153113197,0.00470153113197,0.00156717704399,0.00313435408798,0.00470153113197,0.00626870817596,0.00626870817596,0.00783588521995,0.00783588521995,0.00940306226394,0.0109702393079,0.0109702393079,0.0125374163519,0.0125374163519,0.0109702393079,0.00940306226394,0.00940306226394,0.00783588521995,0.00626870817596,0.00626870817596,0.00470153113197,0.00470153113197,0.00313435408798,0.00313435408798,0.00313435408798,0.00626870817596,0.00626870817596,0.00783588521995,0.00940306226394,0.00783588521995,0.00783588521995,0.00783588521995,0.00940306226394,0.00783588521995,0.00940306226394,0.00626870817596,0.00626870817596,0.00470153113197,0.00313435408798,0.00470153113197,0.00470153113197,0.00626870817596,0.00470153113197,0.00470153113197,0.00313435408798,0.00626870817596,0.00626870817596,0.00940306226394,0.00940306226394,0.00783588521995,0.00940306226394,0.00783588521995,0.00783588521995,0.00783588521995,0.00940306226394,0.00626870817596,0.00626870817596,0.00313435408798,0.00313435408798,0.00313435408798,0.00313435408798,0.00313435408798,0.00313435408798,0.00313435408798,0.00156717704399,0.00156717704399,0.00156717704399,0.00313435408798,0.00313435408798,0.00313435408798,0.00156717704399,0.00313435408798,0.00313435408798,0.003134354087
[…]

### 2018-01-23 ChIP-seq_GRHL2/peaks/SLX-14333.D703_D504_model.r

@@ -1,0 +1,17 @@
+# R script for Peak Model
+#  -- generated by MACS
+p <- c(0.0274660786895,0.0320644964335,0.0365493062097,0.0411720685182,0.0458435199557,0.0457596664557,0.0455703198427,0.0455243356652,0.045518925762,0.0455405653749,0.0455919594556,0.0455973693588,0.0459003239396,0.0457894209234,0.0457028624718,0.0457623714073,0.0458299951976,0.0457650763589,0.0458110605363,0.0456974525685,0.0459354883106,0.0459517180203,0.0458489298589,0.0460274566655,0.0461410646332,0.0464872984399,0.0464737736818,0.0466414806819,0.0466306608754,0.0468091876819,0.0466252509722,0.0465116430044,0.0466171361173,0.0464818885366,0.0463980350366,0.0462681973591,0.0463763954237,0.0465197578592,0.0466712351496,0.046584676698,0.0465170529076,0.0467334490367,0.0466063163109,0.0468091876819,0.0468118926335,0.0469363204078,0.0470796828433,0.0471824710047,0.0471446016821,0.0473393581983,0.0475205899564,0.0475314097629,0.0474989503435,0.0476720672468,0.0478776435695,0.0478235445372,0.0479587921179,0.0478695287147,0.048021006005,0.0480372357147,0.0483374853439,0.0484970774891,0.0487810974086,0.0488054419731,0.0488189667312,0.0488920004248,0.0487269983763,0.0490353628603,0.0488541311022,0.0489650341184,0.0488838855699,0.0490597074249,0.0489488044087,0.0489893786829,0.0492598738443,0.0493897115218,0.0495736482315,0.0494059412315,0.0496466819251,0.0495574185219,0.0496034026993,0.04965479678,0.0497711096994,0.049773814651,0.049584468038,0.0496412720219,0.0495952878445,0.0497170106671,0.049892832522,0.0498820127155,0.0499469315543,0.0498847176672,0.0500010305866,0.0499469315543,0.0503932485706,0.0503310346835,0.0504906268287,0.0503932485706,0.050460872361,0.0507205477159,0.0506772684901,0.0505122664416,0.0506962031514,0.0509477636515,0.050698908103,0.0510803062806,0.0513210469742,0.0516483461195,0.0516970352486,0.0517078550551,0.0517592491357,0.0521406473133,0.0524192573296,0.0524760613135,0.0527059822007,0.0527817208459,0.0527925406523,0.0529818872653,0.0530143466847,0.0531955784428,0.0
[…]

### 2018-01-23 ChIP-seq_GRHL2/peaks/SLX-14333.D704_D501_model.r

@@ -1,0 +1,17 @@
+# R script for Peak Model
+#  -- generated by MACS
+p <- c(0.0296331965535,0.0347628180856,0.0397035728456,0.0446216635929,0.0495813050301,0.0494566529605,0.049486871644,0.0493811062516,0.0498570505175,0.0503329947833,0.0501025773213,0.0497890584795,0.0498494958466,0.050023253277,0.050091245315,0.0502423387327,0.0502498934036,0.050287666758,0.0499325972263,0.0496492970681,0.0498268318339,0.0497890584795,0.0497437304542,0.0498948238719,0.0498306091694,0.0499779252517,0.0499703705808,0.0499741479162,0.0500799133086,0.0501743466947,0.049887269201,0.0501025773213,0.0501139093276,0.0497172891061,0.0497663944668,0.049520867663,0.0496946250934,0.0500119212707,0.0499325972263,0.0498494958466,0.0502423387327,0.0500270306124,0.0499288198909,0.0499741479162,0.0501516826821,0.0501516826821,0.0499325972263,0.0498759371947,0.0500308079479,0.050091245315,0.0501630146884,0.0502763347517,0.0504312055049,0.0504387601758,0.0504727561948,0.0504123188277,0.0505822989226,0.0505747442517,0.0506276269479,0.0505105295492,0.0504991975429,0.0508316030619,0.0508164937201,0.051088461872,0.0508958177644,0.0513793167012,0.0513528753531,0.0512282232834,0.0510280245049,0.0515077461062,0.0516172888341,0.0512055592708,0.0513339886759,0.0514850820936,0.0515417421252,0.0512282232834,0.0513528753531,0.051360430024,0.0513868713721,0.0511375672328,0.0508013843783,0.0512773286442,0.0514813047581,0.0511111258847,0.0510431338467,0.0511337898974,0.0510582431885,0.0510997938784,0.051522855448,0.0514699727518,0.0516890582075,0.0516550621885,0.0515492967961,0.0519005889924,0.0520063543848,0.0521007877709,0.0523085412202,0.0524105292772,0.0521121197772,0.0523274278975,0.0521687798088,0.0521574478025,0.0518930343215,0.0517306088974,0.0516966128784,0.0517154995556,0.0516172888341,0.0512811059796,0.0515304101189,0.051522855448,0.0516059568278,0.0518968116569,0.0521347837899,0.052255658524,0.052527626676,0.0523198732266,0.052161225138,0.0525880640431,0.0525351813469,0.0526371694038,
[…]

### 2018-01-23 ChIP-seq_GRHL2/peaks/SLX-14333.D704_D502_model.r

@@ -1,0 +1,17 @@
+# R script for Peak Model
+#  -- generated by MACS
+p <- c(0.0469157745548,0.0549102363225,0.0624395031166,0.0700376876846,0.0777794509482,0.077785194096,0.0776703311396,0.0776818174352,0.077687560583,0.0774118894876,0.0771879067225,0.0771362183922,0.0779517453828,0.0778426255742,0.0775669544788,0.0780149200089,0.0782733616608,0.0780780946349,0.0780034337132,0.0784111972086,0.078945309956,0.0790946317993,0.078589234791,0.0786122073823,0.0789280805125,0.0789051079212,0.0785318033128,0.0789165942169,0.079106118095,0.0784916012781,0.0781355261131,0.0779804611219,0.0778196529829,0.0782101870348,0.0781297829653,0.0778311392786,0.0781010672262,0.0776760742874,0.0775152661484,0.0779000570524,0.07772201947,0.0778598550177,0.0780493788958,0.0782216733304,0.0784341697998,0.0784628855389,0.0780953240784,0.0776645879917,0.0778943139046,0.0775497250353,0.0777047900265,0.0773429717137,0.0770902732096,0.0764470406536,0.0761771127059,0.0756544862542,0.0758727258714,0.0763551502884,0.0760048182713,0.0761885990016,0.0761024517843,0.0760105614191,0.0759933319757,0.0764872426883,0.0765733899056,0.0771132458009,0.0771017595052,0.0771304752443,0.0772223656095,0.0774578346702,0.0774808072615,0.0775784407744,0.0775037798527,0.077365944305,0.0773544580094,0.0770673006183,0.0773774306007,0.0775095230006,0.0776014133657,0.0772166224617,0.0772912833833,0.0775324955918,0.0781240398175,0.0785432896085,0.0787213271909,0.0788361901474,0.078847676443,0.0789051079212,0.0791290906863,0.0794105049295,0.0795023952947,0.0795598267729,0.0794679364077,0.0788821353299,0.0788706490343,0.0795425973294,0.0792784125296,0.0792152379036,0.0792554399383,0.0793875323382,0.0789395668081,0.0790257140255,0.0789567962516,0.0790199708776,0.0787902449648,0.0783595088782,0.0780666083393,0.0779115433481,0.0777622215047,0.0774176326354,0.0777909372438,0.0778139098351,0.0775784407744,0.0777449920612,0.078204443887,0.0779804611219,0.0784743718346,0.07878450181
[…]

### 2018-01-25 ChIP-seq_GRHL2/Manuscript_Figures.Rmd
Added Groseq data

@@
 
----
-#Gviz for chr2:11,517,385-11,630,055 #GREB1 and chr17:40,296,682-40,349,154 #RARA
----
+### GRHL2 overlap with Gro-SEQ data
+
+```{r GroSeq, fig.cap="Gro-Seq data at GRHL2 sites. Blue is control samples, pink is E2 treated samples. Data from GSE43836."}
+allPeaks<-read.csv("txt/All_GRHL2_peaks_GSE43836.csv")
+upPeaks<-read.csv("txt/Up_GRHL2_peaks_GSE43836.csv")
+
+par(mfrow=c(1,2))
+
+plot(allPeaks[1:2],ylim=c(0,6),type="n", lwd=2, ylab="Read Depth", main="All GRHL2 Sites")
+lines(allPeaks[c(1,2)], lwd=2, col="lightblue")
+lines(allPeaks[c(1,3)], lwd=2,  col="lightblue")
+lines(allPeaks[c(1,4)], lwd=2,  col="pink")
+lines(allPeaks[c(1,5)],  lwd=2, col="pink")
+
+plot(upPeaks[c(1,5)],ylim=c(0,6),type="n", ylab="Read Depth", main="E2 responsive GRHL2 Sites")
+lines(upPeaks[c(1,2)], lwd=2,col="lightblue")
+lines(upPeaks[c(1,3)], lwd=2, col="lightblue")
+lines(upPeaks[c(1,4)], lwd=2, col="pink")
+lines(upPeaks[c(1,5)],  lwd=2,col="pink")
+
+```

### 2018-01-29 Manuscript_Figures_Combined.Rmd
Update Rmd with instruction on how to run

@@
 knitr::opts_chunk$set(cache = TRUE)
+```
+
+#Introduction
+
+To reduce the overall size of this project read count and network data is stored within Rdata files.
+These are avalible from https://github.com/andrewholding/VULCANSupplementary.
+The raw data is avalible from the gene expression omnibus if reprocessing is required.
+The following script will download all the required files and rebuild the network file.
+The Rmd files are designed to recreate the processed data from the raw data if the processed data is not found.
+
+```{bash download, eval=FALSE}
+
+git clone https://github.com/andrewholding/VULCANSupplementary.git
+
+#Rebuild network
+cd ChIP-seq_ER/data/
+cat brca-expmat.rda.part1 brca-expmat.rda.part2 > brca-expmat.rda
+
 ```

### 2018-01-31 eRNA/plot_eRNA.R
eRNA

@@ -1,0 +1,31 @@
+library("ggpubr")
+eRNA<-read.csv("txt/eRNA_values.csv")
+
+oe<-eRNA[eRNA$Experiment=="overexpression",]
+
+oe_comparisons <- list( c("Control","GRHL2") )
+oe_p <- ggboxplot(oe, x = "Treatment", y = "eRNA",
+               add = "jitter", color="Treatment",
+               facet.by="Gene", ylab="Relative eRNA levels")
+oe_p <- oe_p + stat_compare_means(comparisons = oe_comparisons,
+                       method = "wilcox.test" ,
+                       paired=TRUE)
+oe_p
+
+si<-eRNA[eRNA$Experiment=="knockdown",]
+
+si_comparisons <- list( c("siCtrl","siGRHL2") )
+p <- ggboxplot(si, x = "Treatment", y = "eRNA",
+               add = "jitter", color="Treatment",
+               facet.by="Gene", ylab="Relative eRNA levels")
+p + stat_compare_means(comparisons = si_comparisons,
+                       method = "wilcox.test" ,
+                       paired = TRUE)
+
+#Figure for manuscript
+png("plot/eRNA.png",point=15)
+oe_p
+dev.off()
+
+
+wilcox.test(si$eRNA[si$Treatment=='siCtrl'],si$eRNA[si$Treatment=='siGRHL2'],paired=TRUE, alternative="less")

### 2018-01-31 eRNA/plot_eRNA.Rmd
Added eRNA to Rmd

---
title: "eRNA"
author: "Andrew Holding"
date: "1/31/2018"
output:
  html_document: default
  pdf_document: default
---

```{r eRNAsetup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Effect of GRHL2 overexpression on eRNA transcription

```{r eRNAlib, message=FALSE, warning=FALSE}
library("ggpubr")
```

```{r oeGRLH2, fig.cap="Effect of Overexpression of GRHL2 on eRNA at E2 responsive binding sites."}

eRNA<-read.csv("txt/eRNA_values.csv")

oe<-eRNA[eRNA$Experiment=="overexpression",]

oe_comparisons <- list( c("Control","GRHL2") )
oe_p <- ggboxplot(oe, x = "Treatment", y = "eRNA",
               add = "jitter", color="Treatment",
               facet.by="Gene", ylab="Relative eRNA levels")
oe_p <- oe_p + stat_compare_means(comparisons = oe_comparisons,
                       method = "wilcox.test" ,
                       paired=TRUE)
oe_p

```

## Effect of GRHL2 knockdown on eRNA transcription

```{r siGRHL2, fig.cap="Effect of knockdown of GRHL2 on eRNA at E2 responsive binding sites."}
si<-eRNA[eRNA$Experiment=="knockdown",]

si_comparisons <- list( c("siCtrl","siGRHL2") )
p <- ggboxplot(si, x = "Treatment", y = "eRNA",
               add = "jitter", color="Treatment",
               facet.by="Gene", ylab="Relative eRNA levels")
p + stat_compare_means(comparisons = si_comparisons,
                       method = "wilcox.test" ,
                       paired = TRUE)

```

###siRNA combined test
```{r wilcoxsiRNA}

wilcox.test(si$eRNA[si$Treatment=='siCtrl'],
            si$eRNA[si$Treatment=='siGRHL2'],
            paired=TRUE, alternative="less")

```

### 2018-02-05 ChIP-seq_GRHL2/Manuscript_Figures.Rmd
Update Readme

@@
      -log10(vobj_results$tcga$mrs[interect,"pvalue"]),
-     pch=20, xlab="Metabric Network Enrichment Score", ylab="TCGA Network Enrichment Score",main="VULCAN analysis of GRHL2 ChIP-Seq")
+     pch=20, xlab="Metabric Network Enrichment Score", ylab="TCGA Network Enrichment Score",main="VULCAN analysis of GRHL2 ChIP-Seq",
+     col="gray"
+     )
 
-filtered<-(log10(vobj_results$metabric$mrs[interect,"pvalue"])^2+log10(vobj_results$tcga$mrs[interect,"pvalue"])^2)>2
+filtered<-(log10(vobj_results$metabric$mrs[interect,"pvalue"])^2+log10(vobj_results$tcga$mrs[interect,"pvalue"])^2)>3
+filtered[c("GATA3","PGR","FOXA1","ESR1","GRHL2")]<-rep(FALSE,5)
+
+points(-log10(vobj_results$metabric$mrs[interect,"pvalue"][filtered]),
+     -log10(vobj_results$tcga$mrs[interect,"pvalue"][filtered]),
+     labels=interect[filtered],cex=1)
 
@@
      -log10(vobj_results$tcga$mrs[interect,"pvalue"][filtered])-0.1,
-     labels=interect[filtered])
+     labels=interect[filtered],cex=1)
 
 points(-log10(vobj_results$metabric$mrs[interect,"pvalue"][c("GATA3","PGR","FOXA1","ESR1","GRHL2")]),
-       -log10(vobj_results$tcga$mrs[interect,"pvalue"][c("GATA3","PGR","FOXA1","ESR1","GRHL2")]),pch=20,col="red")
-text(-log10(vobj_results$metabric$mrs[interect,"pvalue"][c("GATA3","PGR","FOXA1","ESR1","GRHL2")]),
-     -log10(vobj_results$tcga$mrs[interect,"pvalue"][c("GATA3","PGR","FOXA1","ESR1","GRHL2")])-0.1,col="red",
-     labels=c("GATA3","PGR","FOXA1","ESR1","GRHL2"))
+       -log10(vobj_results$tcga$mrs[interect,"pvalue"][c("GATA3","PGR","FOXA1","ESR1","GRHL2")]),pch=20,col="red",cex=1.5)
+text(-log10(vobj_results$metabric$mrs[interect,"pvalue"][c("FOXA1","ESR1","GRHL2")]),
+     -log10(vobj_results$tcga$mrs[interect,"pvalue"][c("FOXA1","ESR1","GRHL2")])+0.2,col="red",
+     labels=c("FOXA1","ESR1","GRHL2"),cex=1.25,font=2)
+text(-log10(vobj_results$metabric$mrs[interect,"pvalue"][c("GATA3","PGR")])+0.2,
+     -log10(vobj_results$tcga$mrs[interect,"pvalue"][c("GATA3","PGR")]),col="red",
+     labels=c("GATA3","PGR"),cex=1.25,font=2)
 ```

### 2018-02-05 Manuscript_Figures_Combined.Rmd
Update Readme

@@
 output:
-  pdf_document:
-    fig_caption: yes
-    fig_height: 8
-    fig_width: 8
-    highlight: tango
-    number_sections: yes
-    toc: yes
-    toc_depth: 3
   html_document:
@@
     theme: united
+    toc: yes
+    toc_depth: 3
+  pdf_document:
+    fig_caption: yes
+    fig_height: 8
+    fig_width: 8
+    highlight: tango
+    number_sections: yes
     toc: yes
@@
 
+#GRHL2 qPLEX-RIME Analysis
+```{r GRHL2 eRNA, child = 'eRNA/plot_eRNA.Rmd'}
+```
+
 # Technical Session Info

### 2018-02-09 ChIP-seq_GRHL2/Manuscript_Figures.Rmd
Add Suppl. Figure 92

@@
 ```
+
+###  GRHL2 overlap with Gro-SEQ data is reproducible between studies
+
+```{r GroSeq2, fig.cap="Analysis of GRHL2 sites of Gro-Seq data from  GSE43836 and GSE45822 both showed that E2 responsive GRHL2 resonsive sites are transcriptionally responsive to E2." }
+GSE43836<-read.csv("txt/averaged_GSE43836.csv")
+
+colnames(GSE43836)[1]<-"Distance from Center/bp"
+
+par(mfrow=c(2,2))
+
+plot(GSE43836[1:2],ylim=c(0,6),type="n", lwd=2, ylab="Read Depth", main="All GRHL2 Sites [GSE43836]")
+lines(GSE43836[c(1,2)], lwd=2, col="lightblue")
+lines(GSE43836[c(1,3)], lwd=2,  col="pink")
+lines(GSE43836[c(1,4)], lwd=2,  col="deeppink")
+legend("topright", legend=c("0 min", "45 min", "90 min"),
+       col=c("lightblue","pink","deeppink"),pch=20, cex=0.5, horiz = TRUE)
+
+plot(GSE43836[1:2],ylim=c(0,6),type="n", lwd=2, ylab="Read Depth", main="All GRHL2 Sites [GSE43836]")
+lines(GSE43836[c(1,5)], lwd=2, col="lightblue")
+lines(GSE43836[c(1,6)], lwd=2,  col="pink")
+lines(GSE43836[c(1,7)], lwd=2,  col="deeppink")
+legend("topright", legend=c("0 min", "45 min", "90 min"),
+       col=c("lightblue","pink","deeppink"),pch=20, cex=0.5,horiz = TRUE)
+
+GSE45822<-read.csv("txt/averaged_GSE45822.csv")
+
+colnames(GSE45822)[1]<-"Distance from Center/bp"
+
+plot(GSE45822[1:2],ylim=c(0,6),type="n", lwd=2, ylab="Read Depth", main="All GRHL2 Sites [GSE45822]")
+lines(GSE45822[c(1,3)], lwd=2, col="lightblue")
+lines(GSE45822[c(1,2)], lwd=2,  col="deeppink")
+legend("topright", legend=c("Veh", "E2"),
+       col=c("lightblue","deeppink"),pch=20, cex=0.5,horiz = TRUE)
+
+plot(GSE45822[1:2],ylim=c(0,6),type="n", lwd=2, ylab="Read Depth", main="All GRHL2 Sites [GSE45822]")
+lines(GSE45822[c(1,5)], lwd=2, col="lightblue")
+lines(GSE45822[c(1,4)], lwd=2,  col="deeppink")
+legend("topright", legend=c("Veh", "E2"),
+       col=c("lightblue","deeppink"),pch=20, cex=0.5,horiz = TRUE)
+
+```

### 2018-02-14 ChIP-seq_ER/rmarkdown.Rmd
Adjusted figures to make them easier it read

@@
 matplot(t(t(metabricmat["ESR1",])),type="l",col="red",lty=1,lwd=2,add=TRUE)
-text(3,metabricmat["ESR1",3],label="ESR1",pos=4,cex=0.6,font=2,col="red3")
+text(3,metabricmat["ESR1",3],label="ESR1",pos=4,cex=1,font=2,col="red3")
 mtext("METABRIC network")
@@
 oricoords<-sort(setNames(metabricmat[plabels,3],plabels))
-newcoords<-setNames(seq(min(oricoords),max(oricoords),length.out=length(oricoords)),names(oricoords))
-text(3,newcoords,label=names(oricoords),pos=4,cex=0.6,font=2)
-text(3,newcoords["ESR1"],label="ESR1",pos=4,cex=0.6,font=2,col="red3")
+newcoords<-setNames(seq(min(oricoords)*0.7,max(oricoords),length.out=length(oricoords)),names(oricoords))
+text(3,newcoords,label=names(oricoords),pos=4,cex=0.8,font=2)
+text(3,newcoords["ESR1"],label="ESR1",pos=4,cex=1,font=2,col="red3")
 mtext("METABRIC network")
@@
 oricoords<-sort(setNames(metabricmat[plabels,3],plabels))
-newcoords<-setNames(seq(min(oricoords),max(oricoords),length.out=length(oricoords)),names(oricoords))
-text(3,newcoords,label=names(oricoords),pos=4,cex=0.6,font=2)
+newcoords<-setNames(seq(min(oricoords),max(oricoords)*0.5,length.out=length(oricoords)),names(oricoords))
+text(3,newcoords,label=names(oricoords),pos=4,cex=0.8,font=2)
 mtext("METABRIC network")
@@
 newcoords<-setNames(seq(min(oricoords),max(oricoords),length.out=length(oricoords)),names(oricoords))
-text(3,newcoords,label=names(oricoords),pos=4,cex=0.6,font=2)
+text(3,newcoords,label=names(oricoords),pos=4,cex=0.8,font=2)
 mtext("METABRIC network")
@@
 oricoords<-sort(setNames(metabricmat[plabels,3],plabels))
-newcoords<-setNames(seq(min(oricoords),min(oricoords)*1.2,length.out=length(oricoords)),names(oricoords))
-text(3,newcoords,label=names(oricoords),pos=4,cex=0.6,font=2)
+newcoords<-setNames(seq(min(oricoords)*0.7,min(oricoords)*1.4,length.out=length(oricoords)),names(oricoords))
+text(3,newcoords,label=names(oricoords),pos=4,cex=0.8,font=2)
 mtext("METABRIC network")

### 2018-03-12 eRNA/plot_eRNA.Rmd
mRNA

@@
 ---
-title: "eRNA"
+title: "eRNA and mRNA"
 author: "Andrew Holding"
@@
 ```
+
+## Effect of GRHL2 overexpression on mRNA transcription
+
+```{r mRNA}
+qpcr<-read.csv("txt/QPCR GREB1 XBP1 TFF1.csv", header=TRUE)
+df<-data.frame(qpcr)
+
+p <- ggboxplot(df, x = "Sample", y = "Expression", ylab="Relative Expression",
+               xlab="Condition",
+               color = "Sample", palette =c("#00AFBB", "#E7B800", "#FC4E07"),
+               add = "jitter", shape = "Sample", facet.by="Target",
+               outlier=FALSE)
+
+my_comparisons <- list( c("Ctrl", "siGRHL2"), c("Ctrl","GRHL2-DDK"))
+p + 
+    stat_compare_means(comparisons = my_comparisons,method="t.test", paired=TRUE)  +
+    theme(axis.text.x = element_text(angle = 90, hjust = 1))
+
+```

### 2018-03-12 eRNA/plot_mRNA.R

@@ -1,0 +1,15 @@
+qpcr<-read.csv("txt/QPCR GREB1 XBP1 TFF1.csv", header=TRUE)
+df<-data.frame(qpcr)
+library("ggpubr")
+p <- ggboxplot(df, x = "Sample", y = "Expression", ylab="Relative Expression",
+               xlab="Condition",
+               color = "Sample", palette =c("#00AFBB", "#E7B800", "#FC4E07"),
+               add = "jitter", shape = "Sample", facet.by="Target",
+               outlier=FALSE)
+p
+
+my_comparisons <- list( c("Ctrl", "siGRHL2"), c("Ctrl","GRHL2-DDK"))
+p + 
+    stat_compare_means(comparisons = my_comparisons,method="t.test", paired=TRUE)  +
+    theme(axis.text.x = element_text(angle = 90, hjust = 1))
+

### 2018-03-12 ChIP-seq_GRHL2/Manuscript_Figures.Rmd
Added mRNA data

@@
 
-plot(GSE43836[1:2],ylim=c(0,6),type="n", lwd=2, ylab="Read Depth", main="All GRHL2 Sites [GSE43836]")
+plot(GSE43836[1:2],ylim=c(0,6),type="n", lwd=2, ylab="Read Depth", main="Resonsive GRHL2 Sites [GSE43836]")
 lines(GSE43836[c(1,5)], lwd=2, col="lightblue")

### 2018-11-26 eRNA/oe_delta.R
Clean up

@@ -1,0 +1,17 @@
+
+library("ggpubr")
+eRNA<-read.csv("txt/oe.csv")
+
+eRNA<-eRNA[1:135,]
+
+p <- ggboxplot(eRNA, x = "Condition", y = "Fold",
+                   add = "jitter", color="Condition",
+                   facet.by=c("Gene","CellTye"),ylab="Relative eRNA levels to empty vector")
+comp.ev<-list(c("delta","OE"),c("delta","EV"),c("EV","OE"))
+p2<-p + stat_compare_means(comparisons=comp.ev,
+                           method = "t.test",paired=TRUE,alternative="greater",label.y=c(1.25,1.35,1.45))
+p2
+
+pdf("plot/oe_delta.pdf")
+p2
+dev.off()

### 2018-11-26 S2_ChIP-seq_ER/qrime/03.QRIMESceneInvestigatonCombined.Rmd
remove big files

--- 2018-01-23 ChIP-seq_ER/qrime/02.Supplementary.Rmd
@@
 ---
-title: "Supplementary"
+title: "QRIME Scene Investigation"
 author: "Andrew Holding & Matt Eldridge"
@@
 
-# Introduction
-
-This report is from an analysis of the RIME-TMT proteomics data for an experiment carried out by Andrew Holding to quantify specific interactors of estrogen receptor (ER) and forkhead box protein A1 (FOXA1) in MCF-7 cells.
-
-The RIME (Rapid Immunoprecipitation of Endogenous Proteins) technique was used to pull down ER and FOXA1 and their interacting proteins, which were subsequently assessed and quantitated using mass spectrometry with tandem mass tags (TMT).
-TMT runs were carried out for 3 biological replicates, estrogen-starved MCF-7 cells from the same cell line at different times/passages, before and after addition of estrogen with samples taken at 3 different time points -- 0, 45 and 90 minutes after estrogen addition.
-A pull down of Immunoglobulin G (IgG) was also run as a control.
-
-The digests from two of the biological replicates (PR622 and PR650) were analyzed with 3 separate runs of the mass spectrometer to increase coverage; data from the repeated runs were combined prior to the analysis carried out here.
-The other sample (PR590) was run once only.
-
-`r figr("samples_table", TRUE, type="Table")` shows the isobaric tags used for each sample within each of the runs.
+\newpage
+
+# SampleSheet
 
@@
 )
+
+print(
+  xtable(
+    contrastsTable %>% select(Contrast),
+    caption = "Configured Constrasts."
+  ),
+  include.rownames = FALSE,
+  size = "\\setlength{\\tabcolsep}{12pt}"
+)
 ```
@@
 
-## Missing intensity values
-
-Missing intensity values can result even though ions are present at detectable concentrations due to the stochastic nature of the mass spectrometry acquisition method.
-It is reasonable to expect that these missing values are randomly distributed in the data.
-Alternatively, missing values may occur when there is a low abundance of ions, below the limit of detection of the instrument.
-These biologically relevant missing values are not randomly distributed, affecting only those proteins that are expressed at low levels in the samples analysed.
-The R/Bioconductor package `MSnBase` provides imputation methods for both types of missing value.
-
-In this analysis, missing values have been handled in 3 different ways.
-The first approach is to exclude all peptide-level mesasurements where there is a missing value for one or more of the samples within a run.
-In addition, two imputation methods have been performed: one in which the missing values are set to the smallest non-missing value in the data for that run and the other in which k-nearest neighbour (KNN) averaging is applied.
+\newpage
+
+# Normalization of intensity data
 
@@
 
-```{r normalized_intensity_plots, echo=FALSE, warning=FALSE, message=FALSE, fig.cap="Density plots of normalized intensities from each TMT run where scale normalization was applied to peptide intensities that include misssing values.", fig.height=8, fig.width=5}
-
-analysis <- "IncludeMissingValues_ScaleNormalization"
+```{r normalized_intensity_plots, echo=FALSE, warning=FALSE, message=FALSE, fig.cap="Density plots of normalized intensities from each TMT run where QUANTILE normalization was applied to peptide intensities that include misssing values.", fig.height=6.5, fig.width=5}
+
+analysis <- "IncludeMissingValues_QuantileNormalization"
 
@@
 ```
-
-The statistical analysis of differential binding between groups that follows requires at least 3 observations per group.
-`r figr("replicate_count_table", TRUE, type="Table")` gives the number of protein identified in just a single run, two runs or all three runs.
-A total of `r (replicateCounts %>% filter(N == "all"))$Count` proteins were identified across all runs of which `r (replicateCounts %>% filter(N == "3"))$Count` were observed in all three runs, allowing for a statistical analysis.
-log~2~ fold changes are still computed for proteins identified only in one or two of the runs, but no measure of the statistical significance is given in the differential binding analysis.
 
@@
 
-Protein-level intensities were scaled to sum to 1.0 for each protein within a run, allowing for comparison across runs in a principal component analysis (PCA).
-
-`r figr("pca_plot", TRUE, type="Figure")` shows a PCA plot for the first two principal components using all proteins that were sampled in all three runs.
-The plot shows a clear separation of the ER, FOXA1 and IgG control pull-downs.
-The 0, 45 and 90 minute timepoints within each of the ER and FOXA1 groups are not completely separated in the first two principal components.
-`r figr("pca_plot_er", TRUE, type="Figure")` shows the PCA plot for just the ER samples.
-
 ```{r pca_plot, echo=FALSE, warning=FALSE, message=FALSE, fig.cap="Principal Componenent Analysis for proteins sampled in all 3 runs.
@@
 
-```{r pca_plot_er, echo=FALSE, warning=FALSE, message=FALSE, fig.cap="Principal Componenent Analysis for ER-interactng proteins sampled in all 3 runs.
-The first two principal components are displayed.", fig.height=5}
-
-metadata <- metadata %>% filter(grepl("^ER", Group))
-
-print(pcaPlot(scaledIntensities, metadata$Sample, metadata$Group, groupColours, labels = metadata$Run, legend = TRUE))
-```
+\newpage
 
@@
 ```
-
-The IgG control was used to assess non-specific binding.
-The binding of a protein is considered to be non-specific if the log~2~ fold change relative to the IgG control is less than `r controlLogFoldChangeThreshold`.
-In the differential binding analysis, protein intensity values are fitted to a linear model and the fitted values for each condition being compared are also compared to the IgG control.
-For each comparison of two groups, the maximum log~2~ fold change from each of the two groups above the IgG control is used to determine if the binding is specific.
-
-\newpage
-
-```{r contrast_1, include = FALSE}
-contrastGroups <- c("ER 45min", "ER 0min")
-contrastName <- paste(contrastGroups, collapse = " vs ")
-processingDescription <- "excluding peptides with missing intensities, no normalization"
-analysis <- "ExcludeMissingValues_NoNormalization"
-```
-
-## `r paste(contrastName, processingDescription, sep = ", ")`
-
-```{r results_1, include = FALSE}
-
-data <- combinedProteinData[[analysis]]
-exprs(data) <- log2(exprs(data))
-
-results <- createContrastResults(
-  data,
-  fittedLinearModels[[analysis]],
-  fittedLinearModelsExcludingControls[[analysis]],
-  fittedContrasts[[analysis]],
-  runs,
-  contrastGroups,
-  controlGroup,
-  proteinInfo)
-```
-
-```{r ma_plot_1, echo=FALSE, warning=FALSE, message=FALSE, fig.cap=paste("MA plot of the average intensity against log~2~ fold change for the ", contrastName, " comparison (", processingDescription, ").
-Top ranking differentially-expressed proteins with false discovery rate below 0.05 are highlighted in pink.
-Open circles indicate that the protein is non-specific from the IgG control comparison.", sep = ""), fig.height=6}
-print(maPlot(results,
-             significanceLevel = significanceLevel,
-             controlLogFoldChangeThreshold = controlLogFoldChangeThreshold,
-             minLogFoldChangeForLabelling = 1.5,
-             maxNumberLabelledProteins = 20,
-             pointSize = 1.5))
-```
-
-```{r volcano_plot_1, echo=FALSE, warning=FALSE, message=FALSE, fig.cap=paste("Volcano plot of the average intensity against log~2~ fold change for the ", contrastName, " comparison (", processingDescription, ").
-Top ranking differentially-expressed proteins with false discovery rate below 0.05 are highlighted in pink.
-Open circles indicate that the protein is non-specific from the IgG control comparison.", sep = ""), fig.height=6}
-print(volcanoPlot(results,
-                  significanceLevel = significanceLevel,
-                  minLogFoldChange = -2.0,
-                  maxLogFoldChange = 2.75,
-                  controlLogFoldChangeThreshold = controlLogFoldChangeThreshold,
-                  minLogFoldChangeForLabelling = 1.5,
-                  maxNumberLabelledProteins = 20,
-                  pointSize = 1.5))
-```
-
-```{r pvalue_histogram_1, echo=FALSE, warning=FALSE, message=FALSE, fig.cap=paste("Histogram of p-values for the ", contrastName, " comparison (", processingDescription, ")", sep = ""), fig.height=3.5, fig.width=4.5}
-print(histogram(results, "P.Value", xlab = "p-value"))
-```
-
-```{r qq_plot_1, echo=FALSE, warning=FALSE, message=FALSE, fig.cap=paste("QQ plot of the adjusted p-values for the ", contrastName, " comparison (", processingDescription, ")", sep = ""), fig.height=5, fig.width=5}
-q <- results %>%
-  select(adj.P.Val) %>%
-  filter(!is.na(adj.P.Val)) %>%
-  arrange(adj.P.Val) %>%
-  unlist %>%
-  as.numeric
-qqplot(-log10(ppoints(length(q))),
-       -log10(q),
-       xlab=expression(Expected~~-log[10](italic(q))),
-       ylab=expression(Observed~~-log[10](italic(q))))
-abline(0, 1, col = "red")
-```
-
-\newpage
 
@@
 ```{r contrast_3, include = FALSE}
-contrastGroups <- c("ER 45min", "ER 0min")
+contrastGroups <- c("ER 90min", "ER 0min")
 contrastName <- paste(contrastGroups, collapse = " vs ")
-processingDescription <- "excluding peptides with missing intensities, quantile normalization excluding IgG control"
-analysis <- "ExcludeMissingValues_QuantileNormalizationExcludingIgG"
+processingDescription <- "excluding peptides with missing intensities, quantile normalization"
+analysis <- "ExcludeMissingValues_QuantileNormalization"
 ```
@@
   proteinInfo)
+
+print(contrastGroups)
+
 ```
@@
 ```{r contrast_4, include = FALSE}
-contrastGroups <- c("ER 45min", "ER 0min")
+contrastGroups <- c("FOXA1 45min", "FOXA1 0min")
 contrastName <- paste(contrastGroups, collapse = " vs ")
-processingDescription <- "excluding peptides with missing intensities, scale normalization"
-analysis <- "ExcludeMissingValues_ScaleNormalization"
+processingDescription <- "excluding peptides with missing intensities, quantile normalization"
+analysis <- "ExcludeMissingValues_QuantileNormalization"
 ```
@@
 
-```{r volcano_plot_4, echo=FALSE, warning=FALSE, message=FALSE, fig.cap=paste("Volcano plot of the average intensity against log~2~ fold change for the ", contrastName, " comparison (", processingDescription, ").
-Top ranking differentially-expressed proteins with false discovery rate below 0.05 are highlighted in pink.
-Open circles indicate that the protein is non-specific from the IgG control comparison.", sep = ""), fig.height=6}
-print(volcanoPlot(results,
-                  significanceLevel = significanceLevel,
-                  minLogFoldChange = -2.0,
-                  maxLogFoldChange = 2.75,
-                  controlLogFoldChangeThreshold = controlLogFoldChangeThreshold,
-                  minLogFoldChangeForLabelling = 1.5,
-                  maxNumberLabelledProteins = 50,
-                  pointSize = 1.5))
-```
-
-```{r pvalue_histogram_4, echo=FALSE, warning=FALSE, message=FALSE, fig.cap=paste("Histogram of p-values for the ", contrastName, " comparison (", processingDescription, ")", sep = ""), fig.height=3.5, fig.width=4.5}
-print(histogram(results, "P.Value", xlab = "p-value"))
-```
-
-```{r qq_plot_4, echo=FALSE, warning=FALSE, message=FALSE, fig.cap=paste("QQ plot of the adjusted p-values for the ", contrastName, " comparison (", processingDescription, ")", sep = ""), fig.height=5, fig.width=5}
-q <- results %>%
-  select(adj.P.Val) %>%
-  filter(!is.na(adj.P.Val)) %>%
-  arrange(adj.P.Val) %>%
-  unlist %>%
-  as.numeric
-qqplot(-log10(ppoints(length(q))),
-       -log10(q),
-       xlab=expression(Expected~~-log[10](italic(q))),
-       ylab=expression(Observed~~-log[10](italic(q))))
-abline(0, 1, col = "red")
-```
-
 \newpage
@@
 ```{r contrast_5, include = FALSE}
-contrastGroups <- c("ER 45min", "ER 0min")
+contrastGroups <- c("FOXA1 90min", "FOXA1 0min")
 contrastName <- paste(contrastGroups, collapse = " vs ")
-processingDescription <- paste("excluding peptides with missing intensities, scale normalization using top", numberOfPeptidesWithHighestIgG, "IgG peptides")
-analysis <- "ExcludeMissingValues_ScaleNormalizationTopIgG"
+processingDescription <- "excluding peptides with missing intensities, quantile normalization"
+analysis <- "ExcludeMissingValues_QuantileNormalization"
 ```
@@
 
-```{r volcano_plot_5, echo=FALSE, warning=FALSE, message=FALSE, fig.cap=paste("Volcano plot of the average intensity against log~2~ fold change for the ", contrastName, " comparison (", processingDescription, ").
-Top ranking differentially-expressed proteins with false discovery rate below 0.05 are highlighted in pink.
-Open circles indicate that the protein is non-specific from the IgG control comparison.", sep = ""), fig.height=6}
-print(volcanoPlot(results,
-                  significanceLevel = significanceLevel,
-                  minLogFoldChange = -2.0,
-                  maxLogFoldChange = 2.75,
-                  controlLogFoldChangeThreshold = controlLogFoldChangeThreshold,
-                  minLogFoldChangeForLabelling = 1.5,
-                  maxNumberLabelledProteins = 50,
-                  pointSize = 1.5))
-```
-
-```{r pvalue_histogram_5, echo=FALSE, warning=FALSE, message=FALSE, fig.cap=paste("Histogram of p-values for the ", contrastName, " comparison (", processingDescription, ")", sep = ""), fig.height=3.5, fig.width=4.5}
-print(histogram(results, "P.Value", xlab = "p-value"))
-```
-
-```{r qq_plot_5, echo=FALSE, warning=FALSE, message=FALSE, fig.cap=paste("QQ plot of the adjusted p-values for the ", contrastName, " comparison (", processingDescription, ")", sep = ""), fig.height=5, fig.width=5}
-q <- results %>%
-  select(adj.P.Val) %>%
-  filter(!is.na(adj.P.Val)) %>%
-  arrange(adj.P.Val) %>%
-  unlist %>%
-  as.numeric
-qqplot(-log10(ppoints(length(q))),
-       -log10(q),
-       xlab=expression(Expected~~-log[10](italic(q))),
-       ylab=expression(Observed~~-log[10](italic(q))))
-abline(0, 1, col = "red")
-```
-
-\newpage
-
-```{r contrast_6, include = FALSE}
+\newpage
+
+# Differential binding results tables
+
+```{r contrast_results_thresholds, include = FALSE}
+significanceLevel <- 0.05
+logFoldChangeThreshold <- 1.0
+controlLogFoldChangeThreshold <- 1.0
+```
+
+The following tables contain the top ranking differentially bound proteins for each comparison.
+Included are all proteins that reach a statistical signficance of `r significanceLevel` in terms of the adjusted p-value and those with an absolute log~2~ fold change of `r logFoldChangeThreshold` or above.
+
+The IgG column gives the larger of the log~2~ fold changes for the two groups against the IgG control and an asterisk indicates specific binding where this log~2~ fold change is above a threshold of `r controlLogFoldChangeThreshold`.
+N is the number of replicates in which the protein was observed.
+
+In all cases, peptide intensities were quantile normalized and measurements with missing values were removed prior to summarization at the protein level.
+
+```{r contrast_results_table, include = FALSE}
+
+contrastResultsTable <- function(results, contrastName,
+                                 significanceLevel = 0.05,
+                                 logFoldChangeThreshold = 1.0,
+                                 controlLogFoldChangeThreshold = 1.0)
+{
+  resultsTable <- results %>%
+    select(Protein, Gene, N, logFC, AveExpr, adj.P.Val, B, logFCcontrol)
+
+  resultsTable$Specific <- ifelse(resultsTable$logFCcontrol >= controlLogFoldChangeThreshold, "*", "")
+
+  resultsTable <- resultsTable %>%
+    filter(adj.P.Val < 0.05 | abs(logFC) >= 1.0) %>%
+    arrange(desc(abs(logFC)))
+
+  colnames(resultsTable) <- c("Protein", "Gene", "N", "log2FC", "Avg Expr", "p-value", "B", "IgG", "")
+
+  print(
+    xtable(
+      resultsTable,
+      align = "lllrrrrrrl",
+      display = c("d", "s", "s", "d", "f", "f", "g", "f", "f", "s"),
+      caption = paste("Top ranking differentially bound proteins from the", contrastName, "comparison,  sorted by log2 fold change.")
+    ),
+    include.rownames = FALSE,
+    tabular.environment = "longtable",
+    floating = FALSE,
+    add.to.row = addtorow
+  )
+}
+```
+
+\newpage
+
+```{r results_table_contrast_1_config, include = FALSE}
 contrastGroups <- c("ER 45min", "ER 0min")
 contrastName <- paste(contrastGroups, collapse = " vs ")
-processingDescription <- "KNN nearest neighbour averaging imputation of missing values, quantile normalization"
-analysis <- "ImputeKNN_QuantileNormalization"
-```
-
-## `r paste(contrastName, processingDescription, sep = ", ")`
-
-```{r results_6, include = FALSE}
+processingDescription <- "excluding peptides with missing intensities, quantile normalization"
+analysis <- "ExcludeMissingValues_QuantileNormalization"
+```
+
+## `r contrastName`
+
+```{r results_table_contrast_1, echo=FALSE, warning=FALSE, message=FALSE, results='asis'}
 
@@
   proteinInfo)
-```
-
-```{r ma_plot_6, echo=FALSE, warning=FALSE, message=FALSE, fig.cap=paste("MA plot of the average intensity against log~2~ fold change for the ", contrastName, " comparison (", processingDescription, ").
-Top ranking differentially-expressed proteins with false discovery rate below 0.05 are highlighted in pink.
-Open circles indicate that the protein is non-specific from the IgG control comparison.", sep = ""), fig.height=6}
-print(maPlot(results,
-             significanceLevel = significanceLevel,
-             controlLogFoldChangeThreshold = controlLogFoldChangeThreshold,
-             minLogFoldChangeForLabelling = 1.5,
-             maxNumberLabelledProteins = 50,
-             pointSize = 1.5))
-```
-
-```{r volcano_plot_6, echo=FALSE, warning=FALSE, message=FALSE, fig.cap=paste("Volcano plot of the average intensity against log~2~ fold change for the ", contrastName, " comparison (", processingDescription, ").
-Top ranking differentially-expressed proteins with false discovery rate below 0.05 are highlighted in pink.
-Open circles indicate that the protein is non-specific from the IgG control comparison.", sep = ""), fig.height=6}
-print(volcanoPlot(results,
-                  significanceLevel = significanceLevel,
-                  minLogFoldChange = -2.0,
-                  maxLogFoldChange = 2.75,
-                  controlLogFoldChangeThreshold = controlLogFoldChangeThreshold,
-                  minLogFoldChangeForLabelling = 1.5,
-                  maxNumberLabelledProteins = 50,
-                  pointSize = 1.5))
-```
-
-```{r pvalue_histogram_6, echo=FALSE, warning=FALSE, message=FALSE, fig.cap=paste("Histogram of p-values for the ", contrastName, " comparison (", processingDescription, ")", sep = ""), fig.height=3.5, fig.width=4.5}
-print(histogram(results, "P.Value", xlab = "p-value"))
-```
-
-```{r qq_plot_6, echo=FALSE, warning=FALSE, message=FALSE, fig.cap=paste("QQ plot of the adjusted p-values for the ", contrastName, " comparison (", processingDescription, ")", sep = ""), fig.height=5, fig.width=5}
-q <- results %>%
-  select(adj.P.Val) %>%
-  filter(!is.na(adj.P.Val)) %>%
-  arrange(adj.P.Val) %>%
-  unlist %>%
-  as.numeric
-qqplot(-log10(ppoints(length(q))),
-       -log10(q),
-       xlab=expression(Expected~~-log[10](italic(q))),
-       ylab=expression(Observed~~-log[10](italic(q))))
-abline(0, 1, col = "red")
-```
-
-\newpage
-
-# Comparison of differential binding analysis approaches
-
-In this section, comparisons of the differential binding results obtained using differing methods for handling missing values and varying normalization techniques are presented.
-The B-statistic (log odds) is used to rank the proteins from most significantly differentially expressed to least significant.
-Scatter plots of both the B-statistic generated for the two methods being compared and the rank are presented.
-The protein with the highest B-statistic, i.e.
-the most significantly differentially expressed protein, is given a rank of 1.
-
-Differing imputation methods have relatively little effect on the results for most proteins.
-This is largely because only a small proportion of the proteins have peptide-level measurements containing missing values.
-There are a few proteins that are only represented in the differential binding results when imputation is carried out, for which all peptide-level measurements contain missing values.
-These proteins are not represented in the comparions with the analysis in which measurements with missing values are removed.
-
-Different normalization approaches have a more substantial impact on the differential binding analysis.
-
-\newpage
-
-## Removal of missing values vs KNN imputation (quantile normalization in both)
-
-```{r exclude_missing_vs_knn_imputation, include = FALSE}
-
-contrastGroups <- c("ER 45min", "ER 0min")
-contrastName <- paste(contrastGroups, collapse = " vs ")
-
-analysis <- "ExcludeMissingValues_QuantileNormalization"
-data <- combinedProteinData[[analysis]]
-exprs(data) <- log2(exprs(data))
-results1 <- createContrastResults(
-  data,
-  fittedLinearModels[[analysis]],
-  fittedLinearModelsExcludingControls[[analysis]],
-  fittedContrasts[[analysis]],
-  runs,
-  contrastGroups,
-  controlGroup,
-  proteinInfo)
-
-analysis <- "ImputeKNN_QuantileNormalization"
-data <- combinedProteinData[[analysis]]
-exprs(data) <- log2(exprs(data))
-results2 <- createContrastResults(
-  data,
-  fittedLinearModels[[analysis]],
-  fittedLinearModelsExcludingControls[[analysis]],
-  fittedContrasts[[analysis]],
-  runs,
-  contrastGroups,
-  controlGroup,
-  proteinInfo)
-
-results <- inner_join(
-  results1 %>% select(Protein, B1 = B),
-  results2 %>% select(Protein, B2 = B),
-  by = "Protein"
-  ) %>%
-  filter(complete.cases(.)) %>%
-  mutate(rank1 = rank(-B1), rank2 = rank(-B2))
-```
-
-```{r exclude_missing_vs_knn_imputation_B_plot, echo=FALSE, warning=FALSE, message=FALSE, fig.cap=paste("Plot of B-statistic (log odds) for differential expression of proteins for the ", contrastName, " contrast, comparing results following exclusion of missing values and KNN imputation.", sep = ""), fig.height=2.9, fig.width=3.5}
-print(scatterPlot(results, "B1", "B2", "B-statistic (excluding missing values)", "B-statistic (KNN imputation)"))
-```
-
-```{r exclude_missing_vs_knn_imputation_rank_plot, echo=FALSE, warning=FALSE, message=FALSE, fig.cap=paste("Plot of the rank of differentially-expressed proteins for the ", contrastName, " contrast, comparing results following exclusion of missing values and KNN imputation.", sep = ""), fig.height=2.9, fig.width=3.5}
-print(scatterPlot(results, "rank1", "rank2", "Rank (excluding missing values)", "Rank (KNN imputation)"))
-```
-
-\newpage
-
-## Removal of missing values vs lowest value imputation (quantile normalization in both)
-
-```{r exclude_missing_vs_min_imputation, include = FALSE}
-
-contrastGroups <- c("ER 45min", "ER 0min")
-contrastName <- paste(contrastGroups, collapse = " vs ")
-
-analysis <- "ExcludeMissingValues_QuantileNormalization"
-data <- combinedProteinData[[analysis]]
-exprs(data) <- log2(exprs(data))
-results1 <- createContrastResults(
-  data,
-  fittedLinearModels[[analysis]],
-  fittedLinearModelsExcludingControls[[analysis]],
-  fittedContrasts[[analysis]],
-  runs,
-  contrastGroups,
-  controlGroup,
-  proteinInfo)
-
-analysis <- "ImputeMin_QuantileNormalization"
-data <- combinedProteinData[[analysis]]
-exprs(data) <- log2(exprs(data))
-results2 <- createContrastResults(
-  data,
-  fittedLinearModels[[analysis]],
-  fittedLinearModelsExcludingControls[[analysis]],
-  fittedContrasts[[analysis]],
-  runs,
-  contrastGroups,
-  controlGroup,
-  proteinInfo)
-
-results <- inner_join(
-  results1 %>% select(Protein, B1 = B),
-  results2 %>% select(Protein, B2 = B),
-  by = "Protein"
-  ) %>%
-  filter(complete.cases(.)) %>%
-  mutate(rank1 = rank(-B1), rank2 = rank(-B2))
-```
-
-```{r exclude_missing_vs_min_imputation_B_plot, echo=FALSE, warning=FALSE, message=FALSE, fig.cap=paste("Plot of B-statistic (log odds) for differential expression of proteins for the ", contrastName, " contrast, comparing results following exclusion of missing values and lowest value imputation.", sep = ""), fig.height=2.9, fig.width=3.5}
-print(scatterPlot(results, "B1", "B2", "B-statistic (excluding missing values)", "B-statistic (lowest value imputation)"))
-```
-
-```{r exclude_missing_vs_min_imputation_rank_plot, echo=FALSE, warning=FALSE, message=FALSE, fig.cap=paste("Plot of the rank of differentially-expressed proteins for the ", contrastName, " contrast, comparing results following exclusion of missing values and lowest value imputation.", sep = ""), fig.height=2.9, fig.width=3.5}
-print(scatterPlot(results, "rank1", "rank2", "Rank (excluding missing values)", "Rank (lowest value imputation)"))
-```
-
-\newpage
-
-## Quantile vs Scale normalization (missing values excluded in both)
-
-```{r quantile_vs_scale_normalization, include = FALSE}
-
-contrastGroups <- c("ER 45min", "ER 0min")
-contrastName <- paste(contrastGroups, collapse = " vs ")
-
-analysis <- "ExcludeMissingValues_QuantileNormalization"
-data <- combinedProteinData[[analysis]]
-exprs(data) <- log2(exprs(data))
-results1 <- createContrastResults(
-  data,
-  fittedLinearModels[[analysis]],
-  fittedLinearModelsExcludingControls[[analysis]],
-  fittedContrasts[[analysis]],
-  runs,
-  contrastGroups,
-  controlGroup,
-  proteinInfo)
-
-analysis <- "ExcludeMissingValues_ScaleNormalization"
-data <- combinedProteinData[[analysis]]
-exprs(data) <- log2(exprs(data))
-results2 <- createContrastResults(
-  data,
-  fittedLinearModels[[analysis]],
-  fittedLinearModelsExcludingControls[[analysis]],
-  fittedContrasts[[analysis]],
-  runs,
-  contrastGroups,
-  controlGroup,
-  proteinInfo)
-
-results <- inner_join(
-  results1 %>% select(Protein, B1 = B),
-  results2 %>% select(Protein, B2 = B),
-  by = "Protein"
-  ) %>%
-  filter(complete.cases(.)) %>%
-  mutate(rank1 = rank(-B1), rank2 = rank(-B2))
-```
-
-```{r scale_vs_quantile_normalization_B_plot, echo=FALSE, warning=FALSE, message=FALSE, fig.cap=paste("Plot of B-statistic (log odds) for differential expression of proteins for the ", contrastName, " contrast, comparing results following quantile and scale normalization.", sep = ""), fig.height=2.9, fig.width=3.5}
-print(scatterPlot(results, "B1", "B2", "B-statistic (quantile normalization)", "B-statistic (scale normalization)"))
-```
-
-```{r scale_vs_quantile_normalization_rank_plot, echo=FALSE, warning=FALSE, message=FALSE, fig.cap=paste("Plot of the rank of differentially-expressed proteins for the ", contrastName, " contrast, comparing results following quantile and scale normalization.", sep = ""), fig.height=2.9, fig.width=3.5}
-print(scatterPlot(results, "rank1", "rank2", "Rank (quantile normalization)", "Rank (scale normalization)"))
-```
-
-\newpage
-
-## Scale normalization based on all peptides vs peptides with highest IgG intensities (missing values excluded in both)
-
-```{r scale_normalization_all_vs_topIgG, include = FALSE}
-
-contrastGroups <- c("ER 45min", "ER 0min")
-contrastName <- paste(contrastGroups, collapse = " vs ")
-
-analysis <- "ExcludeMissingValues_ScaleNormalization"
-data <- combinedProteinData[[analysis]]
-exprs(data) <- log2(exprs(data))
-results1 <- createContrastResults(
-  data,
-  fittedLinearModels[[analysis]],
-  fittedLinearModelsExcludingControls[[analysis]],
-  fittedContrasts[[analysis]],
-  runs,
-  contrastGroups,
-  controlGroup,
-  proteinInfo)
-
-analysis <- "ExcludeMissingValues_ScaleNormalizationTopIgG"
-data <- combinedProteinData[[analysis]]
-exprs(data) <- log2(exprs(data))
-results2 <- createContrastResults(
-  data,
-  fittedLinearModels[[analysis]],
-  fittedLinearModelsExcludingControls[[analysis]],
-  fittedContrasts[[analysis]],
-  runs,
-  contrastGroups,
-  controlGroup,
-  proteinInfo)
-
-results <- inner_join(
-  results1 %>% select(Protein, B1 = B),
-  results2 %>% select(Protein, B2 = B),
-  by = "Protein"
-  ) %>%
-  filter(complete.cases(.)) %>%
-  mutate(rank1 = rank(-B1), rank2 = rank(-B2))
-```
-
-```{r scale_normalization_all_vs_topIgG_B_plot, echo=FALSE, warning=FALSE, message=FALSE, fig.cap=paste("Plot of B-statistic (log odds) for differential expression of proteins for the ", contrastName, " contrast, comparing results following scale normalization based on all peptides and those with the highest IgG intensities.", sep = ""), fig.height=2.5, fig.width=3.5}
-print(scatterPlot(results, "B1", "B2", "B-statistic (scale normalization based on all peptides)", "B-statistic (scale normalization, peptides with highest IgG)"))
-```
-
-```{r scale_normalization_all_vs_topIgG_rank_plot, echo=FALSE, warning=FALSE, message=FALSE, fig.cap=paste("Plot of the rank of differentially-expressed proteins for the ", contrastName, " contrast, comparing results following scale normalization based on all peptides and those with the highest IgG intensities.", sep = ""), fig.height=2.5, fig.width=3.5}
-print(scatterPlot(results, "rank1", "rank2", "Rank (scale normalization based on all peptides)", "Rank (scale normalization, peptides with highest IgG)"))
-```
-
-\newpage
-
-# Comparisoin of variance between specfic and non-specific binding
-
-```{r  echo=FALSE, warning=FALSE, message=FALSE}
-
-maPlot <- function(differentialExpressionResults, selectedGenes = NULL,
-                   xlab = expression(average~log[2](intensity)),
-                   ylab = expression(log[2](fold~change)),
-                   significanceLevel = 0.05,
-                   minLogFoldChangeForLabelling = 1.5,
-                   maxNumberLabelledProteins = 50,
-                   controlLogFoldChangeThreshold = -Inf,
-                   pointSize = 2.5) {
-  if (!"logFCcontrol" %in% colnames(differentialExpressionResults))
-    differentialExpressionResults$logFCcontrol <- Inf
-
-  differentialExpressionResults$group <-
-    ifelse(differentialExpressionResults$Gene %in% selectedGenes, 0,
-    ifelse(!is.na(differentialExpressionResults$adj.P.Val) & differentialExpressionResults$adj.P.Val <= significanceLevel & differentialExpressionResults$logFCcontrol >= controlLogFoldChangeThreshold, 1,
-    ifelse(!is.na(differentialExpressionResults$adj.P.Val) & differentialExpressionResults$adj.P.Val <= significanceLevel, 2,
-    ifelse(abs(differentialExpressionResults$logFC) >= minLogFoldChangeForLabelling & differentialExpressionResults$logFCcontrol >= controlLogFoldChangeThreshold, 3,
-    ifelse(abs(differentialExpressionResults$logFC) >= minLogFoldChangeForLabelling, 4,
-    ifelse(differentialExpressionResults$logFCcontrol >= controlLogFoldChangeThreshold, 5, 6))))))
-
-  # 0 - selected
-  # 1 - significant, specific
-  # 2 - significant, non-specific
-  # 3 - non-significant, large logfc, specific
-  # 4 - non-significant, large logfc, non-specific
-  # 5 - specific
-  # 6 - non-specific
-
-  differentialExpressionResults <- differentialExpressionResults %>% arrange(desc(group))
-
-  differentialExpressionResults$group <- factor(differentialExpressionResults$group, 0:6)
-
-  differentialExpressionResults <- differentialExpressionResults %>%
-    mutate(label = paste(" ", Gene)) %>%
-    mutate(logFCcontrolAlpha = pmin(logFCcontrol, 2.0))
-
-  labelSubset <- bind_rows(
-    differentialExpressionResults %>%
-      filter(group != 0 & adj.P.Val <= significanceLevel) %>%
-      arrange(desc(abs(logFC))) %>%
-      slice(1:(maxNumberLabelledProteins / 2)),
-    differentialExpressionResults %>%
-      filter(group != 0 & abs(logFC) >= minLogFoldChangeForLabelling) %>%
-      arrange(desc(abs(logFC))) %>%
-      slice(1:maxNumberLabelledProteins)
-  ) %>%
-    unique %>%
-    slice(1:maxNumberLabelledProteins)
-
-    if (maxNumberLabelledProteins == 0) labelSubset <- labelSubset %>% slice(0)
-
-  # plot <- ggplot(differentialExpressionResults, aes(x = AveExpr, y = logFC, colour = group, size = group, shape = group, alpha = logFCcontrolAlpha))
-  plot <- ggplot(differentialExpressionResults, aes(x = AveExpr, y = logFC, colour = group))+ coord_cartesian(ylim = c(-3.5,3.5))
-  plot <- plot + geom_hline(yintercept = 0, color="gray50", size = 0.5)
-  plot <- plot + geom_point(alpha=0.75)
-  plot <- plot + geom_text(data = labelSubset, aes(label = label), hjust = 0, size = 3.5)
-  plot <- plot + geom_point(data = subset(differentialExpressionResults, group == 0), alpha = 0.7)
-  plot <- plot + geom_text(data = subset(differentialExpressionResults, group == 0), aes(label = label), hjust = 0, size = 4.0, alpha = 1)
-  plot <- plot + scale_colour_manual(values = c("green", "red", "blue", "red", "blue", "red", "blue"), drop = FALSE)
-  # plot <- plot + scale_colour_manual(values = c("blue", "deeppink3", "deeppink3", "gray50", "gray50", "gray50", "gray50"), drop = FALSE)
-  # plot <- plot + scale_size_manual(values = c(1.2 * pointSize, pointSize, pointSize, 0.8 * pointSize, 0.8 * pointSize, 0.6 * pointSize, 0.6 * pointSize), drop = FALSE)
-  # plot <- plot + scale_shape_manual(values = c(16, 16, 1, 16, 1, 16, 1), drop = FALSE)
-  #plot <- plot + scale_alpha(range = c(0.2, 1.0))
-  plot <- plot + xlab(xlab)
-  plot <- plot + ylab(ylab)
-  plot <- plot + theme_bw()
-  plot <- plot + theme(
-    text = element_text(size = 14),
-    legend.position = "none"
-  )
-
-  save(differentialExpressionResults,file="Results/QRIME.RData") 
-  return(plot)
-}
-
-contrastGroups <- c("ER 45min", "ER 0min")
-contrastName <- paste(contrastGroups, collapse = " vs ")
-
-analysis <- "ExcludeMissingValues_QuantileNormalization"
-
-data <- combinedProteinData[[analysis]]
-exprs(data) <- log2(exprs(data))
-
-results <- createContrastResults(
-  data,
-  fittedLinearModels[[analysis]],
-  fittedLinearModelsExcludingControls[[analysis]],
-  fittedContrasts[[analysis]],
-  runs,
-  c("ER 45min", "ER 0min"),
-  controlGroup,
-  proteinInfo)
-
-```
-
-```{r MA_specificVsNonSpecfic, echo=FALSE, warning=FALSE, message=FALSE, fig.cap="MA plot showing the distribution of non-specific binding.
-Non-specific binding was identified as proteins that showed less than a log2 enrichment over the IgG channel at either 0 or 45 minutes.
-The majority of non-specific binding (blue) is found to show less change between time points whereas the specific interactions (red) shows a broader distribution.", fig.height=2.5, fig.width=3.5}
-
-print(maPlot(results,
-             significanceLevel = 0.5,
-             controlLogFoldChangeThreshold = 1.0,
-             minLogFoldChangeForLabelling = 1.5,
-             maxNumberLabelledProteins = 0,
-             pointSize = 1.5))
-```
-
-```{r  echo=FALSE, warning=FALSE, message=FALSE}
-
-load("Results/QRIME.RData")
-as.numeric(differentialExpressionResults$group)%%2->differentialExpressionResults$group
-
-tmp<-differentialExpressionResults
-
-```
-
-```{r density_specificVsNonSpecfici, echo=FALSE, warning=FALSE, message=FALSE, fig.cap="Density plot right showing the distribution of non-specific binding.
-Non-specific binding was identified as proteins that showed less than a log2 enrichment over the IgG channel at either 0 or 45 minutes.
-The majority of non-specific binding (blue) is found to show less change between time points whereas the specific interactions (red) shows a broader distribution.", fig.height=3.5, fig.width=4.5}
-
-plot(density(tmp$logFC[tmp$group==1]),col="blue",ylab="Density", xlab="Fold Change", main="", xlim=c(3,-3), lwd=2)
-lines(density(tmp$logFC[tmp$group==0]),col="red",lwd=2)
-
-```
-
-```{r box_specificVsNonSpecfici, echo=FALSE, warning=FALSE, message=FALSE, fig.cap=" Box plot showing the variance of non-specific vs non-specific binding.
-Non-specific binding was identified as proteins that showed less than a log2 enrichment over the IgG channel at either 0 or 45 minutes.
-The the variance of the log fold changes for the non-specific binding interactions was significantly less than that of the specific interactions (F-test, one-sided, p-value < 2.2e*-16).", fig.height=3.5, fig.width=4.5}
-
-boxplot(tmp$logFC[tmp$group==1],tmp$logFC[tmp$group==0],ylab="Fold Change", names=c("Non-specific","Specific"), col=c("#0000FFAA","#FF0000AA"))
-
-```
-
-```{r  echo=FALSE, warning=FALSE, message=FALSE}
-
-var.test(tmp$logFC[tmp$group==1],tmp$logFC[tmp$group==0], alternative="less")
-
-```
-
-\newpage
-
-# Differential binding results tables
-
-```{r contrast_results_thresholds, include = FALSE}
-significanceLevel <- 0.05
-logFoldChangeThreshold <- 1.0
-controlLogFoldChangeThreshold <- 1.0
-```
-
-The following tables contain the top ranking differentially expressed proteins for each comparison.
-Included are all proteins that reach a statistical signficance of `r significanceLevel` in terms of the adjusted p-value and those with an absolute log~2~ fold change of `r logFoldChangeThreshold` or above.
-
-The IgG column gives the larger of the log~2~ fold changes for the two groups against the IgG control and an asterisk indicates specific binding where this log~2~ fold change is above a threshold of `r controlLogFoldChangeThreshold`.
-N is the number of replicates in which the protein was observed.
-
-In all cases, peptide intensities were quantile normalized and measurements with missing values were removed prior to summarization at the protein level.
-
-```{r contrast_results_table, include = FALSE}
-
-contrastResultsTable <- function(results, contrastName,
+
+contrastResultsTable(results, contrastName,
                      significanceLevel = 0.05,
@@
                      controlLogFoldChangeThreshold = 1.0)
-{
-  resultsTable <- results %>%
-    select(Protein, Gene, N, logFC, AveExpr, adj.P.Val, B, logFCcontrol)
-
-  resultsTable$Specific <- ifelse(resultsTable$logFCcontrol >= controlLogFoldChangeThreshold, "*", "")
-
-  resultsTable <- resultsTable %>%
-    filter(adj.P.Val < 0.05 | abs(logFC) >= 1.0) %>%
-    arrange(desc(abs(logFC)))
-
-  colnames(resultsTable) <- c("Protein", "Gene", "N", "log2FC", "Avg Expr", "p-value", "B", "IgG", "")
-
-  print(
-    xtable(
-      resultsTable,
-      align = "lllrrrrrrl",
-      display = c("d", "s", "s", "d", "f", "f", "g", "f", "f", "s"),
-      caption = paste("Top ranking differentially expressed proteins from the", contrastName, "comparison,  sorted by log2 fold change.")
-    ),
-    include.rownames = FALSE,
-    tabular.environment = "longtable",
-    floating = FALSE,
-    add.to.row = addtorow
-  )
-}
-```
-
-\newpage
-
-```{r results_table_contrast_1_config, include = FALSE}
-contrastGroups <- c("ER 45min", "ER 0min")
+
+resultsFile <- paste("Results/", paste(gsub(" ", "_", contrastGroups), collapse = "_vs_"), analysis, ".txt", sep = "")
+write.table(results, resultsFile, sep = "\t", row.names = FALSE, quote = FALSE)
+```
+
+\newpage
+
+```{r results_table_contrast_2_config, include = FALSE}
+contrastGroups <- c("ER 90min", "ER 0min")
 contrastName <- paste(contrastGroups, collapse = " vs ")
@@
 
-```{r results_table_contrast_1, echo=FALSE, warning=FALSE, message=FALSE, results='asis'}
+```{r results_table_contrast_2, echo=FALSE, warning=FALSE, message=FALSE, results='asis'}
 
@@
 
-resultsFile <- paste("Results/", paste(gsub(" ", "_", contrastGroups), collapse = "_vs_"), analysis, ".txt", sep = "")
+resultsFile <- paste("Results/",paste(gsub(" ", "_", contrastGroups), collapse = "_vs_"), analysis, ".txt", sep = "")
 write.table(results, resultsFile, sep = "\t", row.names = FALSE, quote = FALSE)
@@
 
-```{r results_table_contrast_2_config, include = FALSE}
-contrastGroups <- c("ER 90min", "ER 0min")
+```{r results_table_contrast_3_config, include = FALSE}
+contrastGroups <- c("ER 90min", "ER 45min")
 contrastName <- paste(contrastGroups, collapse = " vs ")
@@
 
-```{r results_table_contrast_2, echo=FALSE, warning=FALSE, message=FALSE, results='asis'}
+```{r results_table_contrast_3, echo=FALSE, warning=FALSE, message=FALSE, results='asis'}
 
@@
 
-```{r results_table_contrast_3_config, include = FALSE}
-contrastGroups <- c("ER 90min", "ER 45min")
+```{r results_table_contrast_4_config, include = FALSE}
+contrastGroups <- c("FOXA1 45min", "FOXA1 0min")
 contrastName <- paste(contrastGroups, collapse = " vs ")
@@
 
-```{r results_table_contrast_3, echo=FALSE, warning=FALSE, message=FALSE, results='asis'}
+```{r results_table_contrast_4, echo=FALSE, warning=FALSE, message=FALSE, results='asis'}
 
@@
 
-```{r results_table_contrast_4_config, include = FALSE}
-contrastGroups <- c("FOXA1 45min", "FOXA1 0min")
+```{r results_table_contrast_5_config, include = FALSE}
+contrastGroups <- c("FOXA1 90min", "FOXA1 0min")
 contrastName <- paste(contrastGroups, collapse = " vs ")
@@
 
-```{r results_table_contrast_4, echo=FALSE, warning=FALSE, message=FALSE, results='asis'}
+```{r results_table_contrast_5, echo=FALSE, warning=FALSE, message=FALSE, results='asis'}
 
@@
 
-```{r results_table_contrast_5_config, include = FALSE}
-contrastGroups <- c("FOXA1 90min", "FOXA1 0min")
+```{r results_table_contrast_6_config, include = FALSE}
+contrastGroups <- c("FOXA1 90min", "FOXA1 45min")
 contrastName <- paste(contrastGroups, collapse = " vs ")
@@
 
-```{r results_table_contrast_5, echo=FALSE, warning=FALSE, message=FALSE, results='asis'}
+```{r results_table_contrast_6, echo=FALSE, warning=FALSE, message=FALSE, results='asis'}
 
@@
 
-resultsFile <- paste("Results/",paste(gsub(" ", "_", contrastGroups), collapse = "_vs_"), analysis, ".txt", sep = "")
-write.table(results, resultsFile, sep = "\t", row.names = FALSE, quote = FALSE)
-```
-
-\newpage
-
-```{r results_table_contrast_6_config, include = FALSE}
-contrastGroups <- c("FOXA1 90min", "FOXA1 45min")
-contrastName <- paste(contrastGroups, collapse = " vs ")
-processingDescription <- "excluding peptides with missing intensities, quantile normalization"
-analysis <- "ExcludeMissingValues_QuantileNormalization"
-```
-
-## `r contrastName`
-
-```{r results_table_contrast_6, echo=FALSE, warning=FALSE, message=FALSE, results='asis'}
-
-data <- combinedProteinData[[analysis]]
-exprs(data) <- log2(exprs(data))
-
-results <- createContrastResults(
-  data,
-  fittedLinearModels[[analysis]],
-  fittedLinearModelsExcludingControls[[analysis]],
-  fittedContrasts[[analysis]],
-  runs,
-  contrastGroups,
-  controlGroup,
-  proteinInfo)
-
-contrastResultsTable(results, contrastName,
-                     significanceLevel = 0.05,
-                     logFoldChangeThreshold = 1.0,
-                     controlLogFoldChangeThreshold = 1.0)
-
 resultsFile <- paste("Results/", paste(gsub(" ", "_", contrastGroups), collapse = "_vs_"), analysis, ".txt", sep = "")

### 2018-11-26 S2_ChIP-seq_ER/rmarkdown.Rmd
remove big files

--- 2018-02-14 ChIP-seq_ER/rmarkdown.Rmd
@@
 ### VULCAN and qPLEX-RIME
-We set to testing the performance of VULCAN against a complementary experimental approach called qPLEX-RIME (Holding et al., submitted), which aims at identifyin interactors of ER within the ER-chromatin complex at estrogen-responding promoters.
+We set to testing the performance of VULCAN against a complementary experimental approach called qPLEX-RIME (Papachristou et al., under review), which aims at identifyin interactors of ER within the ER-chromatin complex at estrogen-responding promoters.
 We have QRIME results for MCF7 cells treated with estradiol at both 45 and 90 minutes, the same experimental setup for the VULCAN input dataset.

### 2018-11-26 S6_eRNA/plot_eRNA.Rmd
remove big files

--- 2018-01-31 eRNA/plot_eRNA.Rmd
@@
 output:
-  html_document: default
   pdf_document: default
@@
 ```
+
+## Effect of GRHL2  and GRLH2 delta 425-437 overexpression on eRNA transcription
+
+```{r delta, fig.cap="Overexpression of GRHL2 delta 425-437."}
+eRNA<-read.csv("txt/oe.csv")
+
+eRNA<-eRNA[1:135,]
+
+p <- ggboxplot(eRNA, x = "Condition", y = "Fold",
+                   add = "jitter", color="Condition",
+                   facet.by=c("Gene","CellTye"),ylab="Relative eRNA levels to empty vector")
+comp.ev<-list(c("delta","OE"),c("delta","EV"),c("EV","OE"))
+p2<-p + stat_compare_means(comparisons=comp.ev,
+                           method = "t.test",paired=TRUE,alternative="greater",label.y=c(1.25,1.35,1.45))
+p2
+
+```

### 2018-11-26 eRNA/oe_delta.R

@@ -1,17 +1,0 @@
-
-library("ggpubr")
-eRNA<-read.csv("txt/oe.csv")
-
-eRNA<-eRNA[1:135,]
-
-p <- ggboxplot(eRNA, x = "Condition", y = "Fold",
-                   add = "jitter", color="Condition",
-                   facet.by=c("Gene","CellTye"),ylab="Relative eRNA levels to empty vector")
-comp.ev<-list(c("delta","OE"),c("delta","EV"),c("EV","OE"))
-p2<-p + stat_compare_means(comparisons=comp.ev,
-                           method = "t.test",paired=TRUE,alternative="greater",label.y=c(1.25,1.35,1.45))
-p2
-
-pdf("plot/oe_delta.pdf")
-p2
-dev.off()

### 2018-11-26 eRNA/plot_eRNA.R

@@ -1,50 +1,0 @@
-library("ggpubr")
-eRNA<-read.csv("txt/eRNA_values.csv")
-
-oe<-eRNA[eRNA$Experiment=="overexpression",]
-
-oe_comparisons <- list( c("Control","GRHL2") )
-oe_p <- ggboxplot(oe, x = "Treatment", y = "eRNA",
-               add = "jitter", color="Treatment",
-               facet.by="Gene", ylab="Relative eRNA levels")
-oe_p <- oe_p + stat_compare_means(comparisons = oe_comparisons,
-                       method = "wilcox.test" ,
-                       paired=TRUE)
-oe_p
-
-si<-eRNA[eRNA$Experiment=="knockdown"& eRNA$Tissue=="MCF7",]
-si_comparisons <- list( c("siCtrl","siGRHL2") )
-MCF7p <- ggboxplot(si, x = "Treatment", y = "eRNA",
-               add = "jitter", color="Treatment",
-                facet.by="Gene",ylab="Relative eRNA levels")
-MCF7p<-MCF7p + stat_compare_means(comparisons = si_comparisons,
-                       method = "t.test" ,label.y=0.0014,
-                       paired = TRUE)+theme(axis.title.x=element_blank())
-
-si<-eRNA[eRNA$Experiment=="knockdown"& eRNA$Tissue=="T47D",]
-si_comparisons <- list( c("siCtrl","siGRHL2") )
-T47Dp <- ggboxplot(si, x = "Treatment", y = "eRNA",
-               add = "jitter", color="Treatment",
-               facet.by="Gene",ylab="Relative eRNA levels")
-T47Dp<- T47Dp + stat_compare_means(comparisons = si_comparisons,
-                          method = "t.test" ,
-                          paired = TRUE,label.y=0.0005)+ guides(fill=FALSE, color=FALSE)+theme(axis.title.x=element_blank())
-
-si<-eRNA[eRNA$Experiment=="knockdown"& eRNA$Tissue=="ZR75",]
-si_comparisons <- list( c("siCtrl","siGRHL2") )
-ZR75p <- ggboxplot(si, x = "Treatment", y = "eRNA",
-                   add = "jitter", color="Treatment",
-                   facet.by="Gene",ylab="Relative eRNA levels")
-ZR75p<-ZR75p + stat_compare_means(comparisons = si_comparisons,
[…]

### 2018-11-26 eRNA/plot_mRNA.R

@@ -1,15 +1,0 @@
-qpcr<-read.csv("txt/QPCR GREB1 XBP1 TFF1.csv", header=TRUE)
-df<-data.frame(qpcr)
-library("ggpubr")
-p <- ggboxplot(df, x = "Sample", y = "Expression", ylab="Relative Expression",
-               xlab="Condition",
-               color = "Sample", palette =c("#00AFBB", "#E7B800", "#FC4E07"),
-               add = "jitter", shape = "Sample", facet.by="Target",
-               outlier=FALSE)
-p
-
-my_comparisons <- list( c("Ctrl", "siGRHL2"), c("Ctrl","GRHL2-DDK"))
-p + 
-    stat_compare_means(comparisons = my_comparisons,method="t.test", paired=TRUE)  +
-    theme(axis.text.x = element_text(angle = 90, hjust = 1))
-

### 2018-11-26 S2_ChIP-seq_ER/rmarkdown.Rmd
updated ferderico's code

--- 2018-01-23 ChIP-seq_ER/rmarkdown.Rmd
@@
 ---
-title: "Analysis of the time-dependent response to Estradiol through ChIP-Seq data and the VULCAN package"
+title: Analysis of the time-dependent response to Estradiol through ChIP-Seq data
+  and the VULCAN package
 author: "Federico M. Giorgi"
@@
 output:
+  pdf_document:
+    toc: yes
+    toc_depth: '3'
   html_document:
-    fig_caption: yes
     fig_height: 8
     fig_width: 8
-    toc: true # turns on table of contents
-    toc_depth: 3  # up to three depths of headings (specified by #, ## and ###)
-    number_sections: true  # if you want number sections at each table header
+    highlight: tango
+    number_sections: yes
     theme: united
-    highlight: tango
+    toc: yes
+    toc_depth: 3
+fig_caption: yes
 ---
 
-```{r setup_vulcan, include=FALSE}
+```{r setup, include=FALSE, message=FALSE}
 knitr::opts_chunk$set(echo=TRUE)
@@
 
-```{r captions, include=FALSE}
+```{r captions, include=FALSE, message=FALSE}
 set.seed(1)
@@
 capfig <- function(x){
-  #x <- paste0("Figure S",capfigNo,". ",x)
-  #capfigNo <<- capfigNo + 1
+  x <- paste0("Figure S",capfigNo,". ",x)
+  capfigNo <<- capfigNo + 1
   x
@@
 
-## Introduction
+# Introduction
 Vulcan (VirtUaL ChIP-Seq Analysis through Networks) is a pipeline that combines
@@
 
-## Software setup
+# Software setup
 VULCAN is conveniently provided as an R package available from the Bioconductor repository.
@@
 
-## Import ChIP-Seq data using Vulcan
+# Import ChIP-Seq data using Vulcan
 ChIP-Seq data was generated using a cell line model of ER+ breast cancer, MCF7, at 0’, 45’ and 90’ after stimulation with estradiol.
 The binding profile of ER at each timepoint was then compared between time points using differential binding analysis.
-From the temporal comparison ER binding we established four classes of binding pattern: early responders, repressed transcription factors, late responders and candidate cyclic genes.
+From the temporal comparison ER binding we established four classes of binding pattern: early coactivators, early corepressors, late coactivators and transient coactivators.
 
@@
 
-### Initial import
+## Initial import
 VULCAN requires an input sheet file in CSV format describing the samples and the location of the individual input files
@@
 
-### Annotation
+## Annotation
 VULCAN identifies the location of each peak in each sample, and according to the selected method, assigns a score to each gene promoter. There are a few methods available.
@@
 
-### Normalization
+## Normalization
 At this step, genes are quantified according to the number of reads that could be associated to their promoters.
@@
 
-### Dataset Sample Clustering
+## Dataset Sample Clustering
 Here, we show how the samples cluster together using peak raw counts, VST-normalized peak raw counts and peak RPKMs.
@@
 
-### MA plots
+## MA plots
 Without changing the format of the input data, we can use the Bioconductor DiffBind package to visualize the amplitude of changes in ER binding between time points.
@@
 
-## Apply a coregulatory network over a ChIP-Seq profile
+# Apply a coregulatory network over a ChIP-Seq profile
 Once the data has been loaded, VULCAN applies a regulatory network over differential binding signatures to define Transcription Factors whose networks are most affected by the treatment.
@@
 A third network is used as a negative control, highlighting regulatory mechanisms derived from the Amyloid Leukemia dataset (TCGA).
-### Loading the input networks and processed data
+## Loading the input networks and processed data
 First, we will load the output from the previous paragraphs (chipseq data annotated and normalized) and the three networks.
@@
 ```
-### Running the core Vulcan function
+## Running the core Vulcan function
 The final Vulcan pipeline step requires three input objects:
@@
 ```
-### Visualize the relative TF activity
+## Visualize the relative TF activity
 Vulcan has now generated activity scores for each Transcription Factor, which specify the global ER binding strength on the promoters of their targets.
 
-#### METABRIC results
+### METABRIC results
 The following line plot shows the relative network activity for every TF at two time points.
@@
 These TFs can be grouped in classes.
-For example, TFs whose activity is already significant after 45mins and remains significant at 90 minutes after estradiol treatment can be dubbed **early responders**.
+For example, TFs whose activity is already significant after 45mins and remains significant at 90 minutes after estradiol treatment can be dubbed **early coactivators**.
 ```{r, fig.cap=capfig("Global TF activity after Estradiol Treatment in MCF7 cells, inferred using the METABRIC network, highlighting TFs significantly upregulated at 45 minutes and 90 minutes"), message=FALSE, warning=FALSE,dev.args=list(pointsize=14)}
@@
 tfclass<-tfs[metabricmat[,"T45"]>=threshold&metabricmat[,"T90"]>=threshold]
-matplot(t(metabricmat),type="l",col="grey",ylab="VULCAN NES",xaxt="n",lty=1,main="Early responders",xlim=c(1,3.3),ylim=c(0,max(metabricmat)))
+matplot(t(metabricmat),type="l",col="grey",ylab="VULCAN NES",xaxt="n",lty=1,main="Early coactivators",xlim=c(1,3.3),ylim=c(0,max(metabricmat)))
 axis(1,at=c(1:3),labels=colnames(metabricmat))
@@
 ```
-TFs whose *repressed* targets are bound will yield a negative activity score. These **repressed TFs** are symmetrically opposite to early responder TFs.
+TFs whose *repressed* targets are bound will yield a negative activity score. These **early corepressors** are symmetrically opposite to early responder TFs.
 ```{r, fig.cap=capfig("Global TF activity after Estradiol Treatment in MCF7 cells, inferred using the METABRIC network, highlighting TFs significantly downregulated at 45 minutes and 90 minutes"), message=FALSE, warning=FALSE,dev.args=list(pointsize=14)}
@@
 tfclass<-tfs[metabricmat[,"T45"]<=-threshold&metabricmat[,"T90"]<=-threshold]
-matplot(t(metabricmat),type="l",col="grey",ylab="VULCAN NES",xaxt="n",lty=1,main="Repressed TFs",xlim=c(1,3.3),ylim=c(min(metabricmat),0))
+matplot(t(metabricmat),type="l",col="grey",ylab="VULCAN NES",xaxt="n",lty=1,main="Early corepressors",xlim=c(1,3.3),ylim=c(min(metabricmat),0))
 axis(1,at=c(1:3),labels=colnames(metabricmat))
@@
 Some TFs appear to have their targets bound at 45 minutes, but then unoccupied at 90 minutes.
-This "updown" behavior is consistent to what observed in previous literature about the cyclic properties of certain components of the ER DNA-binding complex, and therefore we dubbed them **candidate cyclic TFs**.
+This "updown" behavior is consistent to what observed in previous literature about the cyclic properties of certain components of the ER DNA-binding complex, and therefore we dubbed them **transient coactivators**.
 ```{r, fig.cap=capfig("Global TF activity after Estradiol Treatment in MCF7 cells, inferred using the METABRIC network, highlighting TFs significantly upregulated at 45 minutes but not at 90 minutes"), message=FALSE, warning=FALSE,dev.args=list(pointsize=14)}
@@
 tfclass<-tfs[metabricmat[,"T45"]>=threshold&metabricmat[,"T90"]<=threshold]
-matplot(t(metabricmat),type="l",col="grey",ylab="VULCAN NES",xaxt="n",lty=1,main="Candidate Cyclic TFs",xlim=c(1,3.3),ylim=c(min(metabricmat),max(metabricmat)))
+matplot(t(metabricmat),type="l",col="grey",ylab="VULCAN NES",xaxt="n",lty=1,main="Transient coactivators",xlim=c(1,3.3),ylim=c(min(metabricmat),max(metabricmat)))
 axis(1,at=c(1:3),labels=colnames(metabricmat))
@@
 ```
-Finally, a category of TFs appear to be activated but at 90 minutes only. We call this category of TFs **late responders**.
+Finally, a category of TFs appear to be activated but at 90 minutes only. We call this category of TFs **late coactivators**.
 ```{r, fig.cap=capfig("Global TF activity after Estradiol Treatment in MCF7 cells, inferred using the METABRIC network, highlighting TFs significantly upregulated at 90 minutes but not at 45 minutes"), message=FALSE, warning=FALSE,dev.args=list(pointsize=14)}
@@
 tfclass<-tfs[metabricmat[,"T45"]<=threshold&metabricmat[,"T45"]>0&metabricmat[,"T90"]>=threshold]
-matplot(t(metabricmat),type="l",col="grey",ylab="VULCAN NES",xaxt="n",lty=1,main="Late responders",xlim=c(1,3.3),ylim=c(0,max(metabricmat)))
+matplot(t(metabricmat),type="l",col="grey",ylab="VULCAN NES",xaxt="n",lty=1,main="Late coactivators",xlim=c(1,3.3),ylim=c(0,max(metabricmat)))
 axis(1,at=c(1:3),labels=colnames(metabricmat))
@@
 
-#### TCGA results
+### TCGA results
 All the TFs and the four categories of TFs were inferred also using the TCGA network.
@@
 tfclass<-tfs[tcgamat[,"T45"]>=threshold&tcgamat[,"T90"]>=threshold]
-matplot(t(tcgamat),type="l",col="grey",ylab="VULCAN NES",xaxt="n",lty=1,main="Early responders",xlim=c(1,3.3),ylim=c(0,max(tcgamat)))
+matplot(t(tcgamat),type="l",col="grey",ylab="VULCAN NES",xaxt="n",lty=1,main="Early coactivators",xlim=c(1,3.3),ylim=c(0,max(tcgamat)))
 axis(1,at=c(1:3),labels=colnames(tcgamat))
@@
 tfclass<-tfs[tcgamat[,"T45"]<=-threshold&tcgamat[,"T90"]<=-threshold]
-matplot(t(tcgamat),type="l",col="grey",ylab="VULCAN NES",xaxt="n",lty=1,main="Repressed TFs",xlim=c(1,3.3),ylim=c(min(tcgamat),0))
+matplot(t(tcgamat),type="l",col="grey",ylab="VULCAN NES",xaxt="n",lty=1,main="Early corepressors",xlim=c(1,3.3),ylim=c(min(tcgamat),0))
 axis(1,at=c(1:3),labels=colnames(tcgamat))
@@
 tfclass<-tfs[tcgamat[,"T45"]>=threshold&tcgamat[,"T90"]<=threshold]
-matplot(t(tcgamat),type="l",col="grey",ylab="VULCAN NES",xaxt="n",lty=1,main="Candidate Cyclic TFs",xlim=c(1,3.3),ylim=c(0,max(tcgamat)))
+matplot(t(tcgamat),type="l",col="grey",ylab="VULCAN NES",xaxt="n",lty=1,main="Transient coactivators",xlim=c(1,3.3),ylim=c(0,max(tcgamat)))
 axis(1,at=c(1:3),labels=colnames(tcgamat))
@@
 tfclass<-tfs[tcgamat[,"T45"]<=threshold&tcgamat[,"T45"]>0&tcgamat[,"T90"]>=threshold]
-matplot(t(tcgamat),type="l",col="grey",ylab="VULCAN NES",xaxt="n",lty=1,main="Late responders",xlim=c(1,3.3),ylim=c(0,max(tcgamat)))
+matplot(t(tcgamat),type="l",col="grey",ylab="VULCAN NES",xaxt="n",lty=1,main="Late coactivators",xlim=c(1,3.3),ylim=c(0,max(tcgamat)))
 axis(1,at=c(1:3),labels=colnames(tcgamat))
@@
 
-#### Comparing TCGA and METABRIC results
+### Comparing TCGA and METABRIC results
 The following scatterplots compare the activity inferred by VULCAN using two alternative networks, derived from the TCGA and METABRIC breast cancer datasets using ARACNe-AP (Giorgi et al., 2016) with default parameters.
@@
 
-#### Testing with a different context network as a negative control
+### Testing with a different context network as a negative control
 Regulatory networks can be tissue-specific, and so we built a third network using the same parameters as the two breast cancer data-based ones.
@@
 
-## Pathway enrichment analysis
+# Pathway enrichment analysis
 In the previous paragraphs we generated signatures of differential binding centered around the peaks extracted by VULCAN using the \textit{vulcan.annotate} function.
@@
 
-## The GRHL2 Transcription Factor
+# The GRHL2 Transcription Factor
 Our analysis shows that the genes repressed by the Transcription Factor GRHL2 are occupied by the ER complex, using both TCGA-derived and METABRIC-derived regulatory models.
 
-### Network similarity between ESR1 and GRHL2
+## Network similarity between ESR1 and GRHL2
 The following analysis highlights, with a Venn Diagram for both the METABRIC and TCGA ARACNe-derived regulatory models, the overlap between the ESR1 (Estrogen Receptor) and GRHL2 networks
@@
 
-### Correlation between ESR1 and GRHL2 expression in breast cancer datasets
+## Correlation between ESR1 and GRHL2 expression in breast cancer datasets
 The following analysis assesses the inverse correlation between GRHL2 and ESR1 expressions in TCGA and METABRIC, and highlights the phenomenon for specific tumor subtypes, according to the PAM50 nomenclature (Perou et al., 2000)
@@
 
-### Enrichment of GRHL2 network for gene pathways
+## Enrichment of GRHL2 network for gene pathways
 In order to analyze the GRHL2 network, we perform a simple hypergeometric test to overlap its targets with all pathways contained in hte manually curated Reactome collection of MsigDB pathways (C2 v5.0).
@@
 
-## Comparing VULCAN results with other tools and approaches
+# Comparing VULCAN results with other tools and approaches
 Our VULCAN results were obtained using independent gene regulatory models and multiple replicates to generate ER-response signatures.
 
-### VULCAN and qPLEX-RIME
-We set to testing the performance of VULCAN against a complementary experimental approach called qPLEX-RIME (Holding et al., submitted), which aims at identifyin interactors of ER within the ER-chromatin complex at estrogen-responding promoters.
+## VULCAN and QRIME
+We set to testing the performance of VULCAN against a complementary experimental approach called QRIME (Holding et al., submitted), which aims at identifyin interactors of ER within the ER-chromatin complex at estrogen-responding promoters.
 We have QRIME results for MCF7 cells treated with estradiol at both 45 and 90 minutes, the same experimental setup for the VULCAN input dataset.
@@
 
-### Comparing Mutual Information and Partial Correlation networks
+## Comparing Mutual Information and Partial Correlation networks
 As shown before in literature, partial correlation and mutual information networks are highly similar.
@@
 ```{r, fig.cap=capfig("Comparing the Jaccard Index between the ARACNe network and Partial correlation networks (indicated as arrows) and random networks (indicated as distributions)"), message=FALSE, warning=FALSE}
-par(mfrow=c(5,3))
+par(mfrow=c(3,5))
 for(i in 1:length(random)){
@@
 
-### Comparing alternative methods for target enrichment analysis
+## Comparing alternative methods for target enrichment analysis
 We developed three independent methods to compare VULCAN with. The first is common, the second is simple but crude, and the third is also common but too stringent.
@@
 
-### Comparing VULCAN with online tools
+## Comparing VULCAN with online tools
 We will generate BED files for the peaks, as reported by DiffBind, we will then test the pathway enrichment for each of these peaks using the Great software v3.0.0 (http://bejerano.stanford.edu/great/public/html/), the ChIP-Enrich pipeline (http://chip-enrich.med.umich.edu/) and the ISMARA tool (http://ismara.unibas.ch).
@@
 contrasts<-c("45","90")
-par(mfrow=c(2,1))
+par(mfrow=c(1,2))
 for(c in contrasts){
@@
 # Plots
-par(mfrow=c(2,1))
+par(mfrow=c(1,2))
 
@@
 ```
+
+# Testing VULCAN on different datasets
+We will show how VULCAN can be used on more independent ChIP-Seq datasets, arising from patient data and different sample types.
+
+## Breast Cancer Xenografts
+We tested  VULCAN on another ChIP-Seq dataset (available on GEO series GSE110824). We used a TCGA-derived breast cancer network.
+```{r xenogratfs, message=FALSE, warning=FALSE}
+
+list_eg2symbol<-as.list(org.Hs.egSYMBOL[mappedkeys(org.Hs.egSYMBOL)]) 
+fname<-"results/009_vobj_xeno.rda"
+
+if(!file.exists(fname)){
+  ctcf<-readRDS("chipseq/holdingNAR/ctcf.rds")
+  pdx<-readRDS("chipseq/holdingNAR/pdx.rds")
+
+  # Build input object
+  peakcounts<-matrix(NA,nrow=nrow(pdx$peaks[[1]]),ncol=nrow(pdx$samples)+3)
+  colnames(peakcounts)<-c("Chr","Start","End",pdx$samples$SampleID)
+  peakcounts[,"Chr"]<-pdx$peaks[[1]]$Chr
+  peakcounts[,"Start"]<-pdx$peaks[[1]]$Start
+  peakcounts[,"End"]<-pdx$peaks[[1]]$End
+  peakcounts<-as.data.frame(peakcounts)
+  peakrpkms<-peakcounts
+  for(i in 1:nrow(pdx$samples)){
+    peakcounts[,i+3]<-pdx$peaks[[i]]$Reads
+    peakrpkms[,i+3]<-pdx$peaks[[i]]$RPKM
+  }
+  samples<-list(A=c("PDX01","PDX04"),B=c("PDX03","PDX05"),X="PDX02")
+  vobj<-list(peakcounts=peakcounts,samples=samples,peakrpkms=peakrpkms)
+
+  # And then standard VULCAN pipeline
+  vobj<-vulcan.annotate(vobj,lborder=-10000,rborder=10000,method='sum')
+  vobj<-vulcan.normalize(vobj)
+  names(vobj$samples)
+  vobj_xeno<-viper(vobj$normalized,regulon=tcga_regulon)
+  rownames(vobj_xeno)<-list_eg2symbol[rownames(vobj_xeno)]
+  save(vobj_xeno,file=fname)  
+  rm(vobj,ctcf,pdx)
+} else {
+  load(fname)
+}
+```
+We show again a co-activity pattern of ESR1 with FOXA1 and GATA3, and an inverse correlation between ESR1 and FOXC1.
+```{r, fig.cap=capfig("VULCAN Activity scores for a few TFs extracted from the ER-targeted ChIP-Seq Xenograft dataset GSE110824"), message=FALSE, warning=FALSE}
+sometfs<-c("ESR1","FOXA1","GATA3","FOXC1")
+vobj_xeno<-vobj_xeno[,c("PDX01","PDX02","PDX03","PDX04","PDX05")]
+xenoidconverter<-setNames(c("AB555B","VHI0244o2","AB580","STG195","V0980U"),c("PDX01","PDX02","PDX03","PDX04","PDX05"))
+plot(vobj_xeno[sometfs[1],],type="l",lwd=4,ylab=c("Relative NES in sample"),xaxt="n",main="Xenograft dataset",xlab="")
+axis(1,at=c(1:ncol(vobj_xeno)),labels=xenoidconverter[colnames(vobj_xeno)],font=2)
+lines(vobj_xeno[sometfs[2],],type="l",lwd=4,col=2,lty=2)
+lines(vobj_xeno[sometfs[3],],type="l",lwd=4,col=3,lty=3)
+lines(vobj_xeno[sometfs[4],],type="l",lwd=4,col=4,lty=4)
+grid()
+legend("topright",legend=sometfs,bg="white",col=1:4,lty=1:4,lwd=4)
+mtext("GEO series GSE110824",cex=0.8)
+```
+
+## Prostate Cancer Data
+We tested VULCAN on patient ChIP-Seq samples from human prostate cancer cell lines.
+We used a TCGA-derived prostate cancer network, available on the _aracne.networks_ Bioconductor package.
+We obtained the raw data from the Sahu et al., 2013 Cancer Research study available on GEO series [GSE39880](https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE39880).
+This dataset focuses on Androgen Receptor (AR) binding events in LNCaP-1F5 prostate-derived cells treated with CPA, RU486 and DHT.
+Reads were aligned using Bowtie2 and peaks were called using Macs 1.4.
+Only samples with biological replicates in the dataset are shown.
+```{r prostate, message=FALSE, warning=FALSE}
+# Import
+sheetfile<-"chipseq/prostateAR/sheetmini.csv"
+fname<-"results/010_input_prostateAR.rda"
+if(!file.exists(fname)){
+  vobj<-vulcan.import(sheetfile)
+  save(vobj,file=fname)
+} else {
+  load(fname)
+}
+
+fnamm<-"results/010_vulcan_prostateAR.rda"
+if(!file.exists(fnamm)){
+  samples<-list(AR_CPA=c("AR_CPA_1","AR_CPA_2"),AR_RU486=c("AR_RU486_1","AR_RU486_2"),AR_DHT=c("AR_DHT_1","AR_DHT_2"))
+  vobj<-list(peakcounts=vobj$peakcounts,samples=samples,peakrpkms=vobj$peakrpkms)
+
+  # Annotation
+  vobj<-vulcan.annotate(vobj,lborder=-10000,rborder=10000,method="sum")
+
+  # Normalization
+  vobj_prostateAR<-vulcan.normalize(vobj)
+  save(vobj_prostateAR,file="results/010_vobj_prostateAR.rda")
+
+  # Obtain prostate carcinoma (prad) regulon from bioconductor
+  library(aracne.networks)
+  data(regulonprad)
+
+  # Vulcan inference
+  names(vobj_prostateAR$samples)
+  vobj_prostateAR<-viper(vobj_prostateAR$normalized,regulon=regulonprad)
+  rownames(vobj_prostateAR)<-list_eg2symbol[rownames(vobj_prostateAR)]
+  save(vobj_prostateAR,file=fnamm)  
+} else {
+  load(fnamm)
+}
+```
+The bar plots show the relative VULCAN Normalized Enrichment Score calculated on absolute peak intensities after treating cells with Dihydrotestosterone (DHT) and partial AR modulators cyproterone acetate (CPA) and Mifepristone (RU486).
+FOXA1 network binding is higher in presence of the strong AR recruiter DHT, partially confirming the results of the original study.
+Two replicates for each treatment were produced and are reported in matching colors.
+```{r, fig.cap=capfig("VULCAN Activity scores for FOXA1 in Prostate cell lines (dataset GSE39880)"), message=FALSE, warning=FALSE}
+par(las=2)
+toplot<-vobj_prostateAR["FOXA1",]
+names(toplot)<-gsub("AR_","",names(toplot))
+names(toplot)<-gsub("_"," ",names(toplot))
+barplot(toplot,cex.names=0.8,col=c(2,2,3,3,4,4),ylab="FOXA1 Vulcan Score")
+par(las=1)
+title("AR ChIP-Seq in Prostate Cells")
+mtext("GEO series GSE39880",cex=0.8)
+```
+
+# Characterize GHRL2
+We will perform a Fisher's Exact one tail test analysis using the MSIGDB database, to collect clues on the function of this transcriptional regulator.
+Note: the hypergeometric R implementation for the one-tailed Fisher test, as described here: https://mengnote.blogspot.it/2012/12/calculate-correct-hypergeometric-p.html (Results are identical to a Fisher test, but faster to compute).
+```{r grhl2, message=FALSE, warning=FALSE}
+fnamm<-"results/011_grhl2_fet.rda"
+if(!file.exists(fnamm)){
+  load("data/msigdb.v6.1.entrez.rda")
+  targets_grhl2<-names(tcga_regulon[["79977"]]$tfmode)
+
+  ### Calculate Contigency Tables
+  contingencyFunction<-function(set,list){
+    if(!exists("universegenelist")){
+      universegenelist<<-unique(unlist(list))
+    }
+    out<-t(sapply(list,function(x){
+      ul<-intersect(x,set)
+      ur<-setdiff(set,ul)
+      dl<-setdiff(x,ul)
+      dr<-setdiff(universegenelist,c(ul,ur,dl))
+      return(c(length(ul),length(ur),length(dl),length(dr)))
+    }))
+    return(out)
+  }
+  grhl2_ctables<-contingencyFunction(targets_grhl2,msigdb)
+
+  grhl2_fet<-apply(grhl2_ctables,1,function(f){
+    w<-f[1]
+    x<-f[2]
+    y<-f[3]
+    z<-f[4]
+    p<-phyper(w-1,x+w,z+y,y+w,lower.tail=FALSE)
+  })
+
+  save(grhl2_fet,file=fnamm)
+
+  rm(msigdb)
+} else {
+  load(fnamm)
+}
+```
+```{r, fig.cap=captab("10 gene sets from the MsigDB chemical/genetic perturbetion set, which are most overlapping with GRHL2 targets identified by our network analysis"), message=FALSE, warning=FALSE}
+# Write the output NICELY
+classnames<-setNames(
+  c(
+    "C1 positional gene sets",
+    "C2 chemical and genetic perturbations",
+    "C2 BioCarta gene sets",
+    "C2 KEGG gene sets",
+    "C2 Reactome gene sets",
+    "C2 Canonical pathways",
+    "C3 microRNA targets",
+    "C3 transcription factor targets",
+    "C4 cancer gene neighborhoods",
+    "C4 cancer modules",
+    "C5 GO biological process",
+    "C5 GO cellular component",
+    "C5 GO molecular function",
+    "C6 oncogenic signatures",
+    "C7 immunologic signatures",
+    "H hallmark gene sets"
+  ),
+  c("c1_all","c2_cgp","c2_cpbiocarta","c2_cpkegg","c2_cpreactome","c2_cp","c3_mir","c3_tft",      
+    "c4_cgn","c4_cm","c5_bp","c5_cc","c5_mf","c6_all","c7_all","h_all")
+)
+
+# Print Enrichment
+fet<-grhl2_fet
+classes<-unique(gsub(";_;.+","",names(fet)))
+append<-FALSE
+class<-classes[2]
+subfet<-fet[grep(paste0(class,";_;"),names(fet))]
+names(subfet)<-gsub(".+;_;","",names(subfet))
+subfet<-sort(subfet,dec=FALSE)
+subfet<-cbind(names(subfet),subfet,p.adjust(subfet,method="bonferroni"))
+colnames(subfet)<-c("Gene Set","P-value","Q-value")
+subfet[,2]<-signif(as.numeric(subfet[,2]),3)
+subfet[,3]<-signif(as.numeric(subfet[,3]),3)
+rownames(subfet)<-NULL
+
+grid.newpage()
+grid.table(subfet[1:10,])
+
+```
+
+# Overlap with Motif analysis
+We compare our analysis with a direct binding strategy.
+In order to do so, we directly compare the gene networks used by VULCAN with transcription factor targets	gene sets that share upstream cis-regulatory motifs which can function as potential transcription factor binding sites.
+This motif analysis is based on work by Xie et al.
+2005 and processed by the MsigDB project (C3 collection).
+```{r motif, fig.cap=capfig("Example: VULCAN scores of GATA3 and ESR1 in our dataset. Individual samples are indicated."), message=FALSE, warning=FALSE}
+load("data/msigdb.v6.1.entrez.rda")
+
+# VULCAN score correlation example: GATA3 vs ESR1
+load("results/001_vobj.rda") # vobj
+fname<-"results/012_vulcanscores.rda"
+if(!file.exists(fname)){
+  vulcanscores<-viper(vobj$normalized,regulon=tcga_regulon)
+  save(vulcanscores,file=fname)
+} else {
+  load(fname)
+}
+X<-vulcanscores[list_symbol2eg["ESR1"],]
+Y<-vulcanscores[list_symbol2eg["GATA3"],]
+plot(X,Y,xlab="ESR1 VULCAN score",ylab="GATA3 VULCAN score",main="VULCAN score in our dataset",pch=NA)
+text(X,Y,labels=colnames(vulcanscores))
+grid()
+abline(lm(Y~X)$coef)
+scc<-signif(cor(X,Y,method="s"),3)
+mtext(paste0("Spearman correlation: ",scc),cex=0.8)
+```
+
+```{r, fig.cap=capfig("Example of target intersection between GATA3 and ER according to the MSigDB database of canonical TF-specific motifs in putative target gene promoters"), message=FALSE, warning=FALSE}
+# Motif-based targets overlap example: GATA3 vs. ESR1
+motifdb<-msigdb[grep("c3_tft;_;",names(msigdb))]
+grep("GATA3",names(motifdb),value=TRUE)
+grep("ER",names(motifdb),value=TRUE)
+ER_Targets<-msigdb[["c3_tft;_;ER_Q6"]]
+GATA3_Targets<-msigdb[["c3_tft;_;GATA3_01"]]
+# Significance of overlap (Fisher Test)
+ul<-intersect(GATA3_Targets,ER_Targets)
+ur<-setdiff(GATA3_Targets,ul)
+dl<-setdiff(ER_Targets,ul)
+setp<-fet<-fisher.test(rbind(c(length(ul),length(ur)),c(length(dl),20000-length(ul)-length(ur)-length(dl))),alternative="greater")$p
+# Plot it
+library(VennDiagram)
+v<-venn.diagram(list(GATA3_Targets=GATA3_Targets, ER_Targets=ER_Targets),
+                fill = c("orange", "blue"),
+                alpha = c(0.5, 0.5), cat.cex=1,
+                filename=NULL,lty=2,
+                #main=paste0("FET p=",signif(setp,4)),
+                cex=0.6,rotation.degree = 0,main.pos=c(0.5,0)
+)
+grid.newpage()
+grid.draw(v)
+
+```
+
+We then decided to compare TF overlap according to VULCAN score and a classic approach: overlap of putative target genes with promoters carrying canonical and TF specific binding motifs.
+There is no correlation between the two approaches.
+
+```{r motif2, message=FALSE, warning=FALSE}
+# Enrichment of target overlaps with targets defined by motif analysis
+
+### any2entrez function
+any2entrez<-function(x){
+  tab<-AnnotationDbi::select(org.Hs.eg.db, keys=x, columns=c("ENTREZID"), keytype="ALIAS")
+  symbols<-tab[,1]
+  entrez<-tab[,2]
+  dups<-which(duplicated(symbols))
+  symbols<-symbols[-dups]
+  entrez<-entrez[-dups]
+  out<-setNames(entrez,symbols)
+  return(out)
+}
+
+# Get TFs from the msigdb motif database
+submotifdb<-motifdb[grep("UNKNOWN",names(motifdb),invert=TRUE)]
+names(submotifdb)<-gsub("c3_tft;_;","",names(submotifdb))
+names(submotifdb)<-gsub("ER_","ESR1_",names(submotifdb))
+names(submotifdb)[1:441]<-sapply(strsplit(names(submotifdb)[1:441],"_"),function(x){x[1]})
+names(submotifdb)[441:500]<-sapply(strsplit(names(submotifdb)[441:500],"_"),function(x){x[2]})
+names(submotifdb)<-any2entrez(names(submotifdb))
+submotifdb<-submotifdb[!is.na(names(submotifdb))]
+tfs_with_motifs<-names(submotifdb)
+tfs_with_vulcan<-rownames(vulcanscores)
+tfs_with_both<-intersect(tfs_with_motifs,tfs_with_vulcan)
+
+# Compare VULCAN correlation with MOTIF overlap
+corvulcan<-cor(t(vulcanscores[tfs_with_both,]),method="s")
+fetmotif<-matrix(NA,nrow=nrow(corvulcan),ncol=ncol(corvulcan),dimnames=dimnames(corvulcan))
+for(i in 1:nrow(corvulcan)){
+  for(j in (i+1):nrow(corvulcan)){
+    if(j>nrow(corvulcan)|j==i){
+      next
+    }
+    targets1<-submotifdb[[rownames(corvulcan)[i]]]
+    targets2<-submotifdb[[rownames(corvulcan)[j]]]
+    JI<-length(intersect(targets1,targets2))/length(union(targets1,targets2))
+    fetmotif[i,j]<-JI
+
+  }
+}
+```
+
+```{r, fig.cap=capfig("TF pairs compared in terms of VULCAN score Spearman Correlation Coefficient in our ER dataset and in terms of Jaccard Index of motif-based target intersection according to the MSigDB C3 collection"), message=FALSE, warning=FALSE}
+# Plot the Motif overlap JI with the vulcan-score correlation between TF pairs
+x<-corvulcan[upper.tri(corvulcan,diag=FALSE)]
+y<-fetmotif[upper.tri(fetmotif,diag=FALSE)]
+
+plot(x,y,xlab="SCC VULCAN score",ylab="Jaccard Index of Motif Overlap",main="TF pairs",pch=20)
+minpair<-which(fetmotif==max(fetmotif,na.rm=TRUE),arr.ind=TRUE)
+tf1<-list_eg2symbol[[rownames(corvulcan)[minpair[1,1]]]]
+tf2<-list_eg2symbol[[rownames(corvulcan)[minpair[1,2]]]]
+text(corvulcan[minpair[1,1],minpair[1,2]],fetmotif[minpair[1,1],minpair[1,2]],labels=paste0(tf1,"-",tf2),font=2,col="red3")
+
+tf1<-list_symbol2eg[["ESR1"]]
+tf2<-list_symbol2eg[["GATA3"]]
+text(corvulcan[tf2,tf1],fetmotif[tf2,tf1],labels=paste0(list_eg2symbol[[tf1]],"-",list_eg2symbol[[tf2]]),font=2,col="red3")
+
+rm(msigdb)
+```
+
+# Technical Session Info
+The following code describes the R environment used to generate this document and will help making it fully reproducible should there be future updates in any of the packages.
+```{r sessionInfo}
+sessionInfo()
+```

### 2018-11-26 Manuscript_Figures_Combined.Rmd
Add Suppl. Figure 92
Added mRNA data
code working for markdown

@@
 
-```{r ER qRPCR, child = 'qPCR_ER/Manuscript_Figures.Rmd'}
+```{r ER qRPCR, child = 'S1_qPCR_ER/Manuscript_Figures.Rmd'}
 ```
@@
 
-```{r ER ChIP-seq, child = 'ChIP-seq_ER/rmarkdown.Rmd'}
+```{r ER ChIP-seq, child = 'S2_ChIP-seq_ER/rmarkdown.Rmd'}
 ```
@@
 
-```{r ER qRIME, child = 'qPLEX-RIME_ER/02.Supplementary.Rmd'}
-```
-
-#GRHL2 ChIP-Seq Analysis
-
-```{r GRHL2 Chip, child = 'ChIP-seq_GRHL2/Manuscript_Figures.Rmd'}
+```{r ER qRIME, child = 'S3_qPLEX-RIME_ER/02.Supplementary.Rmd'} 
 ```
@@
 
-```{r GRHL2 qRIME, child = 'qPLEX-RIME_GRHL2/rRunTMTanalysis.Rmd'}
+<!-- ```{r GRHL2 qRIME, child = 'qPLEX-RIME_GRHL2/rRunTMTanalysis.Rmd'} -->
+<!-- ``` -->
+
+#GRHL2 ChIP-Seq Analysis
+
+```{r GRHL2 Chip, child = 'S5_ChIP-seq_GRHL2/Manuscript_Figures.Rmd'}
 ```
 
-#GRHL2 qPLEX-RIME Analysis
-```{r GRHL2 eRNA, child = 'eRNA/plot_eRNA.Rmd'}
+#Effect of GRHL2 expression on eRNA and mRNA 
+```{r GRHL2 eRNA, child = 'S6_eRNA/plot_eRNA.Rmd'}
 ```
+
+#Effect of GRHL2 expression on H3K27ac
+
+<!-- ```{r GRHL2 H3K27ac, child = 'S7_ChIP-seq_H3K27ac_siGRHL2/plot_eRNA.Rmd'} -->
+<!-- ``` -->
 

### 2018-11-26 S2_ChIP-seq_ER/rmarkdown.Rmd
code working for markdown

@@
 
-# Technical Session Info
-The following code describes the R environment used to generate this document and will help making it fully reproducible should there be future updates in any of the packages.
-```{r sessionInfo}
-sessionInfo()
-```
+<!-- # Technical Session Info -->
+<!-- The following code describes the R environment used to generate this document and will help making it fully reproducible should there be future updates in any of the packages. -->
+<!-- ```{r sessionInfo} -->
+<!-- sessionInfo() -->
+<!-- ``` -->

### 2018-11-27 S4_qPLEX-RIME_GRHL2/rRunTMTanalysis.Rmd
removed master document to reduce complexity

--- 2018-01-23 qPLEX-RIME_GRHL2/rRunTMTanalysis.Rmd
@@
 fastaUrl <- paste0("http://www.uniprot.org/uniprot/", UniProtAccession, ".fasta")
-download.file(fastaUrl, fastaFile, "wget")
+download.file(fastaUrl, fastaFile)
 
@@
 covCap <- refcaption("covFig", paste0("Plot of peptide coverage for ", baitProtein, " (UniprotID: ", UniProtAccession, ")"))
+#colnames(fData(rawMSnset))[6]<-"Master.Protein.Accessions"
 coveragePlot2(data=rawMSnset, ProteinID=UniProtAccession, ProteinName=baitProtein, fastaFile=fastaFile)
